### 2016-01-07 Rmd/vegetation-types.Rmd

---
title: "Vegetation types"
author: "Lyndon Estes"
date: "January 4, 2016"
output: html_document
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Preparing a habitat rarity layer based on vegetation types

A more straightforward way for calculating biodiversity impacts, which can be based on hectares

## Data
```{r, eval = FALSE}
library(agroEcoTradeoff)
library(gdalUtils)
library(RPostgreSQL)
library(rmapaccuracy)
library(gisutils)

# p_dat <- "/Users/lestes/Dropbox/data/zari/landcover/Spatial Data_FD_ILUAII/"
p_proj <- "/Users/lestes/Dropbox/publications/zambiaToff/private/zambiaToff/"
p_dat <- paste0("/Users/lestes/Dropbox/data/distributions/vegetation/",
                "potential/vecea/Zambia/")
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_devel <- full_path(p_proj, "external/input_devel/biodiversity")

# Load in Vegetation Map of Zambia
veg <- readOGR(dsn = full_path(p_dat, "pnv_vecea_v2_0_Zambia.shp"), 
               layer = "pnv_vecea_v2_0_Zambia", stringsAsFactors = FALSE)
mgrid <- raster(full_path(p_basedat, "ZA-grid.tif"))
msk <- raster(full_path(p_basedat, "ZA-mask.tif"))
cntry <- readOGR(full_path(p_basedat, "africa_countries_alb.sqlite"), 
                 layer = "africa_countries_alb")
zam <- cntry[cntry@data$fips_cntry == "ZA", ]
```

## Intersect grid with vegetation types

To 1 km resolution, in order to turn it into percent coverage

```{r, eval=FALSE}
# transform vegetation map to albers
vegalb <- spTransform(veg, mgrid@crs)

# Test intersects
vtypes <- unique(vegalb$CAT)
vegss <- vegalb[vegalb$CAT == 75, ]
tcrp <- crop(mgrid, extent(vegss))
tint <- intersect(vegss, tcrp)
# plot(tint[1:100, ])
# plot(tcrp)
# tcrp <- rasterToPolygons(tcrp)
# plot(tcrp[1:10, ])
# plot(vegss)
# plot(zam, add = TRUE)
# rm(tint)

```

Initial tests shows this will be tedious in R, so will run this through postgis, for speed (although given the install hassles I could have already run this and been done in R)

Notes on fixing a broken postgis install are appended to postgis

```{r, eval = FALSE}
# Set up new postigs database
Sys.setenv(PATH=paste(Sys.getenv("PATH"),"/usr/local/pgsql/bin/",
                      sep=":"))
# Sys.getenv("PATH")
dbname <- "zambiatoff"
# system(paste("createdb", dbname))
# system("psql -U lestes -h localhost zambiatoff")  # just hangs
drv <- dbDriver("PostgreSQL")
con <- dbConnect(drv, host = "localhost", dbname = dbname, port = 5432)
# dbSendQuery(con, "CREATE EXTENSION postgis;")
# dbSendQuery(con, "CREATE EXTENSION postgis_topology;")
# dbSendQuery(con, "CREATE EXTENSION fuzzystrmatch;")
# dbSendQuery(con, "CREATE EXTENSION postgis_tiger_geocoder;")

# Need to add back in Africa Albers projection
# p_sql <- "~/Dropbox/projects/mappingafrica/external/postgis/af-alb-ref.sql"
# system(paste("psql -f", p_sql, dbname))  # add in prj
srid <- 102022
sql <- paste0("select proj4text from spatial_ref_sys WHERE srid=", srid)
rs <- dbGetQuery(con, sql)

# First convert zambia grid to a set of polygons, for loading into postgis
# Resurrect old gdal_polygonize wrapper
gdal_polygonize <- function(inraster, outshape, attname, 
                            gdalformat = "SQLite") {
  gdal_setInstallation()
  ext_tab <- cbind(c("SQLite", "ESRI Shapefile"), 
                   c(".sqlite", ".shp"))
  ext <- ext_tab[which(ext_tab[, 1] == gdalformat), 2]
  gdal_opts <- getOption("gdalUtils_gdalPath")[[1]]

  pyv <- system("which python", intern = TRUE)
  py_c <- paste(pyv, paste0(gdal_opts$path, "gdal_polygonize.py"))
  rnm <- inraster@file@name
  # rast_nm <- unlist(strsplit(rnm, "/"))[length(unlist(strsplit(rnm, "/")))]
  fullc <- paste(py_c, rnm, "-f", paste0("'", gdalformat, "'"), 
                 paste0(outshape, ext), outshape, attname)
  system(fullc)
  # shp <- readOGR(dsn = paste0(outshape, ext), layer = outshape)
  # return(shp)
}

# Convert to shapefile using gdal_polygonize
setwd(p_devel)
gdal_polygonize(mgrid, "zamgrid", "ID", "ESRI Shapefile")

r <- crop(mgrid, extent(-200000, -100000, -2000000, -1900000))

# Small test grid
# plot(zam)
# plot(as(extent(-200000, -100000, -2000000, -1900000), "SpatialPolygons"), 
#      add = TRUE)
r <- writeRaster(r, file = "testraster.tif", overwrite = TRUE)
gdal_polygonize(r, "testgrid", "ID", "ESRI Shapefile")
polygonizer(r@file@name, outshape = "testgrid2.shp")  # doesn't work with asc

# Send to postgis
postgis_vectab("testgrid.shp", "testgrid", "zambiatoff", "geom", "lestes", con)
# dbRemoveTable(con, "testgrid")

# system("shp2pgsql zamgrid.shp zamgrid zambiatoff > zamgrid.sql")
# system("psql -U lestes -d zambiatoff -f zamgrid.sql")

# Index it
grname_out <- "zamgrid"
dbSendQuery(con, paste0("CREATE INDEX gix_",  grname_out, " ON ", grname_out, 
                        " USING GIST(geom)"))
dbSendQuery(con, paste("VACUUM ANALYZE",  grname_out))
dbSendQuery(con, paste0("CLUSTER ", grname_out, " USING gix_", grname_out))
dbSendQuery(con, paste("ANALYZE",  grname_out))

# A little test, using rmapaccuracy functions
# sql <- paste0("SELECT ID, ST_ASeWKT(geom) FROM ", grname_out, 
#              " WHERE gid in (", paste0(1:10, collapse = ","), ")") 
# qgrid <- dbGetQuery(con, sql) 
# plot(polyFromWkt(qgrid, rs$proj4text))

# Vacuum analyze
create_idx <- function(tabname, idxvarname, idxvar, con) {
  # this first one should be revised to check if there is an index variable or 
  # not first
  dbSendQuery(con, paste0("CREATE INDEX gix_", idxvarname, " ON ",
                          tabname, " USING GIST(", idxvar, ")"))
  dbSendQuery(con, paste("VACUUM ANALYZE", tabname))
  dbSendQuery(con, paste0("CLUSTER ", tabname, " USING gix_", idxvarname))
  dbSendQuery(con, paste("ANALYZE",  tabname))
}

# A function to create tables from shapefiles
postgis_vectab <- function(shp, tabname, dbname, idxvarname, idxvar, user, 
                           con) {
  pgcalls <- list(paste("shp2pgsql", shp, tabname, dbname, ">",
                        paste0(tabname, ".sql")), 
                  paste("psql -U", user, "-d", dbname, "-f", 
                        paste0(tabname, ".sql")))
  for(i in pgcalls) system(i)  # create table in db
  create_idx(tabname, idxvarname, idxvar, con)
}

# send vegetation types to postgis
writeOGR(vegalb, dsn = full_path(p_devel, "vecea.shp"), 
         layer = "vecea", driver = "ESRI Shapefile")
create_pg_vtable("vecea.shp", "vtypes", "zambiatoff", "geom", "lestes", con)

# Update srids for the two fields
sql <- paste0("SELECT UpdateGeometrySRID('vtypes', 'geom', ", srid, ")")  
dbSendQuery(con, sql)
dbClearResult(dbListResults(con)[[1]])
sql <- paste0("SELECT UpdateGeometrySRID('zamgrid', 'geom', ", srid, ")")  
dbSendQuery(con, sql)
dbClearResult(dbListResults(con)[[1]])

# Intersection
isectname <- "vtypegrids"
sql <- paste("CREATE TABLE",  isectname, 
             "(gid serial primary key,", 
             "grid_id integer,", 
             "v_cat integer,",  
             "geom geometry(multipolygon,", srid, ")",
             ");")
d <- dbSendQuery(con, sql)

a <- "zamgrid"
b <- "vtypes"
tick <- Sys.time()
print(paste("Processing intersection, starting at time", tick))
sql <- paste("INSERT INTO", isectname, "(grid_id, v_cat, geom)", 
             "SELECT", 
             "a.id AS grid_id,", 
             "b.gid AS v_cat,",  
             "CASE", 
             "WHEN ST_Within(a.geom,b.geom)", 
             "THEN a.geom", 
             "ELSE ST_Multi(ST_Intersection(a.geom,b.geom))", 
             "END AS geom", 
             "FROM", a, "a",
             "JOIN", b, "b", 
             "ON ST_Intersects(a.geom, b.geom)")
system.time(d <- dbSendQuery(con, sql))  # 6.211
tock <- Sys.time()
print(paste0("Finished intersections at ", tock, ", total time = ", tock - tick))

create_idx()

for(i in vtypes) {
  v <- vegalb[vegalb@data$CAT == i, ]

}
vtypes

dbDisconnect(con)

```

```{r, eval=FALSE, echo=FALSE}
# For cleaning up zari-sourced vegetation map

# trim down some names
# unique(veg@data[, c(1:2, 4)])
ftype <- unique(veg@data$FEATURE_TY)
ftyper <- ftype
ftyper[1] <- "treeless-grassy"
ftyper[2] <- "termitary"
ftyper[3:4] <- c("miombo hilly", "miombo plateau")
ftyper[10] <- "Parinari-copperbelt chipya"
ftyper[15] <- "Munga heavy soils"
ftyper[16] <- "Kalahari woodland"
ftyper[17] <- "Mopane"
ftyper[18] <- "Baikiaea"

# Replace in table
ftype_tab <- cbind(ftype, ftyper)
for(i in 1:nrow(ftype_tab)) {  # i <- 1
  ind <- which(veg@data$FEATURE_TY == ftype_tab[i, 1])
  veg@data$FEATURE_TY[ind] <- ftype_tab[i, 2]
}
colnames(veg@data) <- c("code", "vtype1", "trapnell", "vtype2")
unique(veg@data[, c(1:2, 4)])
# plot(veg[veg$FORESTCLAS == "Inland water", ])
# plot(veg[veg@data$vtype1 == "Itigi forest", ])

```