### 2016-01-15 Rmd/cropland.Rmd
Various scripts for creating terrain, cropland, and actual vegetation layers

---
title: "cropland.Rmd"
author: "Lyndon Estes"
date: "January 14, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# New cropland map of Zambia

Cropland class from the SERVIR landcover map (2010)

## Inputs
```{r, eval = FALSE}
library(agroEcoTradeoff)
library(rworldxtra)
data("countriesHigh")

p_proj <- "/Users/lestes/Dropbox/publications/zambiaToff/private/zambiaToff/"
p_devel <- full_path(p_proj, "external/input_devel/cropland")
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_lc <- "~/Dropbox/data/landcover/Zambia_LandCover_2010_Scheme_II/"

# Load in data
mgrid <- raster(full_path(p_basedat, "ZA-grid.tif"))
lc <- raster(full_path(p_lc, fname = "Zambia_LandCover_2010_Scheme_II.tif"))

```

## Extract cropland class and resample
```{r, eval = FALSE}

# pullout cropland class 8
system.time(cropland <- rast_math(lc == 8, tmpdir = p_devel))

onm <- full_path(p_devel, "cropland-1k.tif")
ext <- bbox(mgrid)[1:4]  # extent   
gdalwarp(srcfile = cropland@file@name, t_srs = projection(mgrid), 
         dstfile = onm, r = "average", ot = "Float32", te = ext, 
         srcnodata = 255, dstnodata = 255, tr = c(1000, 1000), 
         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
r <- raster(onm)
par(mar = rep(0, 4)); plot(r)

```