### 2016-01-15 Rmd/terrain.Rmd

---
title: "terrain"
author: "Lyndon Estes"
date: "January 8, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Filter terrain by workability

The first step entailed getting a percent slope map from EarthEngine, using script XX, from which I exported a map showing areas less than 20% slope built from the new SRTM 30 m dataset for Africa.

## Background data for Earth Engine
```{r, eval = FALSE}
library(agroEcoTradeoff)
library(rworldxtra)
data("countriesHigh")

p_proj <- "/Users/lestes/Dropbox/publications/zambiaToff/private/zambiaToff/"
# p_dat <- paste0("/Users/lestes/Dropbox/data/distributions/vegetation/",
#                 "potential/vecea/Zambia/")

p_devel <- full_path(p_proj, "external/input_devel/terrain")
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")

# Load in Vegetation Map of Zambia
cntry <- readOGR(full_path(p_basedat, "africa_countries_alb.sqlite"), 
                 layer = "africa_countries_alb")
cntrygcs <- spTransform(cntry, raster()@crs)
zambia <- countriesHigh[countriesHigh@data$SOVEREIGNT == "Zambia", ]
zambia@data <- zambia@data[, c("OID_", "NAME")]
mgrid <- raster(full_path(p_basedat, "ZA-grid.tif"))

writeOGR(cntrygcs, dsn = full_path(p_devel, "africa.kml"), 
         layer = "africa", driver = "KML")
writeOGR(zambia, dsn = full_path(p_devel, "zambia.kml"), 
         layer = "zambia", driver = "KML")
```

## Process slope filter from Earth Engine
```{r, eval = FALSE}

farmable <- raster(full_path(p_devel, "farmable.tif"))

onm <- full_path(p_devel, "farmable-1k.tif")
ext <- bbox(mgrid)[1:4]  # extent   
gdalwarp(srcfile = farmable@file@name, t_srs = projection(mgrid), 
         dstfile = onm, r = "average", ot = "Float32", te = ext, 
         srcnodata = 255, dstnodata = 255, tr = c(1000, 1000), 
         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
r <- raster(onm)
par(mar = rep(0, 4)); plot(r)

```