### 2016-01-15 Rmd/vegetation-types.Rmd

---
title: "Vegetation types"
author: "Lyndon Estes"
date: "January 4, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Preparing a habitat rarity layer based on vegetation types

A more straightforward way for calculating biodiversity impacts, which can be based on hectares

## Data
```{r, eval = FALSE}
library(agroEcoTradeoff)
library(gdalUtils)
library(RPostgreSQL)
library(rmapaccuracy)
library(gisutils)
library(rworldxtra)
data("countriesHigh")

# p_dat <- "/Users/lestes/Dropbox/data/zari/landcover/Spatial Data_FD_ILUAII/"
p_proj <- "/Users/lestes/Dropbox/publications/zambiaToff/private/zambiaToff/"
p_dat <- paste0("/Users/lestes/Dropbox/data/distributions/vegetation/",
                "potential/vecea/Zambia/")
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_devel <- full_path(p_proj, "external/input_devel/biodiversity")

# Load in Vegetation Map of Zambia
veg <- readOGR(dsn = full_path(p_dat, "pnv_vecea_v2_0_Zambia.shp"), 
               layer = "pnv_vecea_v2_0_Zambia", stringsAsFactors = FALSE)
mgrid <- raster(full_path(p_basedat, "ZA-grid.tif"))
# length(which(!is.na(values(mgrid)))) # diff(cellStats(mgrid, range))
# plot(mgrid)
msk <- raster(full_path(p_basedat, "ZA-mask.tif"))
cntry <- readOGR(full_path(p_basedat, "africa_countries_alb.sqlite"), 
                 layer = "africa_countries_alb")
zam <- cntry[cntry@data$fips_cntry == "ZA", ]
```

## Intersect grid with vegetation types

To 1 km resolution, in order to turn it into percent coverage

```{r, eval=FALSE}
# transform vegetation map to albers
vegalb <- spTransform(veg, mgrid@crs)

# Test intersects
vtypes <- unique(vegalb$CAT)
vegss <- vegalb[vegalb$CAT == 75, ]
tcrp <- crop(mgrid, extent(vegss))
tint <- intersect(vegss, tcrp)
# plot(tint[1:100, ])
# plot(tcrp)
# tcrp <- rasterToPolygons(tcrp)
# plot(tcrp[1:10, ])
# plot(vegss)
# plot(zam, add = TRUE)
# rm(tint)

```

Initial tests shows this will be tedious in R, so will run this through postgis, for speed (although given the install hassles I could have already run this and been done in R)

Notes on fixing a broken postgis install are in evernote.

```{r, eval = FALSE}
# Set up new postigs database
Sys.setenv(PATH=paste(Sys.getenv("PATH"),"/usr/local/pgsql/bin/",
                      sep=":"))
# Sys.getenv("PATH")
dbname <- "zambiatoff"
# system(paste("createdb", dbname))
# system("psql -U lestes -h localhost zambiatoff")  # just hangs
drv <- dbDriver("PostgreSQL")
con <- dbConnect(drv, host = "localhost", dbname = dbname, port = 5432)
# dbSendQuery(con, "CREATE EXTENSION postgis;")
# dbSendQuery(con, "CREATE EXTENSION postgis_topology;")
# dbSendQuery(con, "CREATE EXTENSION fuzzystrmatch;")
# dbSendQuery(con, "CREATE EXTENSION postgis_tiger_geocoder;")

# Need to add back in Africa Albers projection
# p_sql <- "~/Dropbox/projects/mappingafrica/external/postgis/af-alb-ref.sql"
# system(paste("psql -f", p_sql, dbname))  # add in prj
srid <- 102022
sql <- paste0("select proj4text from spatial_ref_sys WHERE srid=", srid)
rs <- dbGetQuery(con, sql)

# First convert zambia grid to a set of polygons, for loading into postgis
# Resurrect old gdal_polygonize wrapper
gdal_polygonize <- function(inraster, outshape, attname, 
                            gdalformat = "SQLite") {
  gdal_setInstallation()
  ext_tab <- cbind(c("SQLite", "ESRI Shapefile"), 
                   c(".sqlite", ".shp"))
  ext <- ext_tab[which(ext_tab[, 1] == gdalformat), 2]
  gdal_opts <- getOption("gdalUtils_gdalPath")[[1]]

  pyv <- system("which python", intern = TRUE)
  py_c <- paste(pyv, paste0(gdal_opts$path, "gdal_polygonize.py"))
  rnm <- inraster@file@name
  # rast_nm <- unlist(strsplit(rnm, "/"))[length(unlist(strsplit(rnm, "/")))]
  fullc <- paste(py_c, rnm, "-f", paste0("'", gdalformat, "'"), 
                 paste0(outshape, ext), outshape, attname)
  system(fullc)
  # shp <- readOGR(dsn = paste0(outshape, ext), layer = outshape)
  # return(shp)
}

# Convert to shapefile using gdal_polygonize
setwd(p_devel)

# Small test grid
# plot(zam)
# plot(as(extent(-200000, -100000, -2000000, -1900000), "SpatialPolygons"), 
#      add = TRUE)
r <- crop(mgrid, extent(-200000, -100000, -2000000, -1900000))
r <- writeRaster(r, file = "testraster.tif", overwrite = TRUE)
# gdal_polygonize(r, "testgrid", "ID", "ESRI Shapefile")
polygonizer(r@file@name, outshape = "testgrid.shp")  # doesn't work with asc
# gdal_polygonize(mgrid, "zamgrid", "ID", "ESRI Shapefile")
polygonizer(mgrid@file@name, outshape = "zamgrid.shp")  # doesn't work with asc

# Send to postgis
postgis_vectab("testgrid.shp", "testgrid", "zambiatoff", "geom", "lestes", con)
# dbRemoveTable(con, "testgrid")

# system("shp2pgsql zamgrid2.shp zamgrid zambiatoff > zamgrid.sql")
# system("psql -U lestes -d zambiatoff -f zamgrid.sql")

# Index it
postgis_vectab("zamgrid.shp", "zamgrid", "zambiatoff", "geom", "lestes", con)
# create_idx("zamgrid", idxvar = "geom", con)
# dbRemoveTable(con, "zamgrid")

# A little test, using rmapaccuracy functions
# sql <- paste0("SELECT ID, ST_ASeWKT(geom) FROM ", grname_out, 
#              " WHERE gid in (", paste0(1:10, collapse = ","), ")") 
# qgrid <- dbGetQuery(con, sql) 
# plot(polyFromWkt(qgrid, rs$proj4text))

# send vegetation types to postgis
writeOGR(vegalb, dsn = full_path(p_devel, "vecea.shp"), 
         layer = "vecea", driver = "ESRI Shapefile")
postgis_vectab("vecea.shp", "vtypes", "zambiatoff", "geom", "lestes", con)

# Update srids for the two fields
sql <- paste0("SELECT UpdateGeometrySRID('vtypes', 'geom', ", srid, ")")  
dbSendQuery(con, sql)
dbClearResult(dbListResults(con)[[1]])
# sql <- paste0("SELECT UpdateGeometrySRID('testgrid', 'geom', ", srid, ")")  
sql <- paste0("SELECT UpdateGeometrySRID('zamgrid', 'geom', ", srid, ")")  
dbSendQuery(con, sql)
dbClearResult(dbListResults(con)[[1]])

# Intersection
isectname <- "gridveg"
# dbRemoveTable(con, isectname)
sql <- paste0("CREATE TABLE ",  isectname, 
             "(gid serial primary key, ", 
             "grid_id integer, ", 
             "vgid integer, ", 
             "veg_cat integer, ",
             "geom geometry(multipolygon,", srid, ")",
             ");")
d <- dbSendQuery(con, sql)

a <- "zamgrid"
b <- "vtypes"
tick <- Sys.time()
print(paste("Processing intersection, starting at time", tick))
sql <- paste("INSERT INTO", isectname, "(grid_id, vgid, veg_cat, geom)", 
             "SELECT", 
             "a.id AS grid_id,", 
             "b.gid AS vgid,", "b.cat AS veg_cat,",  
             "CASE", 
             "WHEN ST_CoveredBy(a.geom,b.geom)", 
             "THEN a.geom", 
             "ELSE ST_Multi(ST_Intersection(a.geom,b.geom))", 
             "END AS geom", 
             "FROM", a, "AS a",
             "INNER JOIN", b, "AS b", 
             "ON ST_Intersects(a.geom, b.geom)")
system.time(d <- dbSendQuery(con, sql))  # 6.211
tock <- Sys.time()
print(paste0("Finished at ", tock, ", total time = ", tock - tick))

dbDisconnect(con)
```
Intersection in postgis taking far too long (it ran for more than 15 hours before I killed it). Switching to Grass, which is much faster. Table updates of area calculations took unacceptably long there also.  So I ended up doing the full workflow in ArcGIS (!), in script `vegridToRaster.py`.

# Convert vegetation types

## Come up with percent remaining first

First I sent both Zambia landcover and a rasterized version of vecea (in EPSG:4326) created by ArcGIS (`veceaToRaster.py`).  I then 
```{r, eval = FALSE}
# Read in extent of Zambia landcover grid
p_lc <- "~/Dropbox/data/landcover/Zambia_LandCover_2010_Scheme_II/"
lcnm <- full_path(p_lc, fname = "Zambia_LandCover_2010_Scheme_II.tif")
lc <- raster(full_path(p_lc, fname = "Zambia_LandCover_2010_Scheme_II.tif"))
vegrem <- raster(full_path(p_devel, fname = "vegremain.tif"))

lt <- 8
fnm <- full_path(p_devel, "cropland.tif")
pcalc <- paste0("gdal_calc.py -A ", lcnm, " --outfile=", fnm, " --type='Byte'",
                " --overwrite --calc=", "'", lt, "'", " --co 'COMPRESS=LZW'")
system.time(system(pcalc))  # 49 seconds

# Read in unique vegetation types from "actual" vegetation map
v <- gdalinfo(vegrem@file@name, hist=TRUE)[31]  # gdalinfo histogram
v <- as.numeric(strsplit(v, " ")[[1]])  # convert to numeric, get numbers
v <- v[!is.na(v)]  # remove NAs
lcvals <- c(0:(length(v) - 1))[which(v != 0)]  # what are the actuals, less 0

# ran raster calculator - this proves much faster than gdal_calc.py, which is 
# a dog by comparison and always converts to float, despite passing byte type
p_temp <- full_path(p_devel, "temp")
dir.create(p_temp)

# Processed converted areas also
vtr <- lapply(lcvals, function(x) {
  print(paste("Processing veg type", x))
  dang <- Sys.time()
  r <- rast_math(vegrem == x, tmpdir = p_temp)
  print(Sys.time() - dang)
  r
})
# length(vtr) * 4  # rough time to process, about 88 minutes for all types
names(vtr) <- paste0("veg", lcvals)

# aggregate and reproject these types to 1 km resolution
ext <- sapply(c("xmin", "ymin", "xmax", "ymax"), function(x) {
  slot(extent(sa_r), x)
})

p_aveg <- full_path(p_devel, "act_veg")
dir.create(p_aveg)

atypes <- names(vtr)[names(vtr) != "veg0"]  # drop type 0, which are the converted areas
avr <- lapply(1:length(atypes), function(x) {  # x <- 1
  ri <- vtr[atypes[x]]
  nm <- paste0(names(ri), ".tif")
  onm <- full_path(p_aveg, nm)
  ext <- bbox(mgrid)[1:4]  # extent   
  print(paste("warping and masking", nm))
  dang <- Sys.time()
  gdalwarp(srcfile = ri[[1]]@file@name, t_srs = projection(mgrid), 
           dstfile = onm, r = "average", ot = "Float32", te = ext, 
           srcnodata = 255, dstnodata = 255, tr = c(1000, 1000), 
           of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
  # file.remove(lcnm[j])
  print(ptime <- Sys.time() - dang)
  l <- list(raster(onm), as.numeric(gsub("[[:alpha:]]", "", ptime)))
  names(l) <- c(names(ri), "minutes")
  return(l)
})
sum(sapply(avr, function(x) x$minutes)) / 60  # actually in seconds, not mins

# getOption("gdalUtils_gdalPath")
# getOption("")

# stack the rasters
s <- stack(lapply(avr, function(x) x[[1]]))

# Sum for total - the inverse of this will be the farmable areas
remfrac <- calc(s, sum)

veg_dt <- dtraster::as.data.table(s)  # stack and convert to dt
veg_dt[, lapply(.SD, function(x) max(x, na.rm = TRUE))]
veg_dt <- na.omit(veg_dt)

```

# Various rubbish I tried
```{r, eval = FALSE}
# Convert to kmls for creating fusion tables
# http://gis.stackexchange.com/questions/43848/how-to-re-project-shapefiles
system("ogr2ogr -f KML -t_srs EPSG:4326 zamgrid.kml zamgrid.shp")
system("ogr2ogr -f KML -t_srs EPSG:4326 vecea.kml vecea.shp")
system("ogr2ogr -t_srs EPSG:4326 veceagcs.shp vecea.shp")

options("gdal")

fnm <- full_path(p_proj, "external/input_devel/roads/tmdpfr_links.sqlite")
onm <- full_path(p_proj, "external/input_devel/roads/tmdpfr_links.kml")
system(paste("ogr2ogr -f KML", onm, fnm))

zambia <- countriesHigh[countriesHigh@data$SOVEREIGNT == "Zambia", ]
zambia@data <- zambia@data[, c("OID_", "NAME")]
r <- raster(extent(zambia))
# res(r) <- c(0.008333333, 0.008333333)
res(r) <- c(0.00899, 0.00899)
projection(r) <- raster()@crs
r[] <- 1
dataType(r) <- "INT4S"

zambiar <- rasterize(zambia, r, dataType = "INT4S")
dataType(zambiar) <- "INT4S"
vi <- which(!is.na(values(zambiar)))
zambiar[vi] <- as.integer(1:length(vi))
# plot(zambiar)
length(which(!is.na(values(zambiar))))
dataType(zambiar)

zambiawr <- writeRaster(zambiar, format = "GTiff",
                        filename = full_path(p_devel, "zamgridgcs.tif"),
                        datatype = "INT4S", overwrite = TRUE)
dataType(zambiawr)

round(res(r)[1] / res(lc)[1])
res(lc)[1] * 31

# First rasterize landcover types at 30 m resolution
# inm <- full_path(p_dat, "pnv_vecea_v2_0_Zambia.shp")
# fnm <- full_path(p_devel, "vecea.tif")
# gdal_rasterize(src_datasource = inm, l = "pnv_vecea_v2_0_Zambia", ot = "Byte", 
#                a = "CAT", tr = res(lc), at = TRUE, co = "COMPRESS=LZW",
#                dst_filename = fnm, te = bbox(lc)[1:4], of = "GTiff", 
#                verbose = TRUE)
# 
# veg@data <- veg@data[, "CAT", drop = FALSE]
# writeOGR(veg, dsn = full_path(p_devel, "vecea.kml"), layer = "vecea", 
#          driver = "KML")

# a <- rep(0, 40)
# a[sample(1:length(a), size = 40, replace = FALSE)] <- 1
# mean(a)

```

```{r, eval=FALSE, echo=FALSE}
# For cleaning up zari-sourced vegetation map

# trim down some names
# unique(veg@data[, c(1:2, 4)])
ftype <- unique(veg@data$FEATURE_TY)
ftyper <- ftype
ftyper[1] <- "treeless-grassy"
ftyper[2] <- "termitary"
ftyper[3:4] <- c("miombo hilly", "miombo plateau")
ftyper[10] <- "Parinari-copperbelt chipya"
ftyper[15] <- "Munga heavy soils"
ftyper[16] <- "Kalahari woodland"
ftyper[17] <- "Mopane"
ftyper[18] <- "Baikiaea"

# Replace in table
ftype_tab <- cbind(ftype, ftyper)
for(i in 1:nrow(ftype_tab)) {  # i <- 1
  ind <- which(veg@data$FEATURE_TY == ftype_tab[i, 1])
  veg@data$FEATURE_TY[ind] <- ftype_tab[i, 2]
}
colnames(veg@data) <- c("code", "vtype1", "trapnell", "vtype2")
unique(veg@data[, c(1:2, 4)])
# plot(veg[veg$FORESTCLAS == "Inland water", ])
# plot(veg[veg@data$vtype1 == "Itigi forest", ])

```