### 2016-01-18 Rmd/biodiversity-layer.Rmd
Created new biodiversity priority raster

---
title: "Biodiversity layer prep"
author: "Lyndon Estes"
date: "January 15, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Assembling new biodiversity layer

## Inputs
```{r, eval = FALSE}
library(agroEcoTradeoff)

p_proj <- "/Users/lestes/Dropbox/publications/zambiaToff/private/zambiaToff/"
p_bd <- full_path(p_proj, "external/input_devel/biodiversity")
p_dat <- "/Users/lestes/Dropbox/data/zari/"
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_findat <- "~/Dropbox/projects/agroEcoTradeoff/external/data/dt/"
p_pas <- full_path(p_dat, "landcover/Spatial Data_FD_ILUAII")

# Load in data
mgrid <- raster(full_path(p_basedat, "ZA-grid.tif"))
vegrem <- raster(full_path(p_bd, fname = "vegremain.tif"))
msk <- raster(full_path(p_basedat, "ZA-mask.tif"))
mskdt <- fread(full_path(p_findat, "ZA-mask.csv"))
nps <- spTransform(readOGR(full_path(p_pas, "Nat_Parks_GMAs.shp"), 
                           layer = "Nat_Parks_GMAs"), mgrid@crs)
freserves <- spTransform(readOGR(full_path(p_pas, "Forest_Reserves.shp"), 
                                 layer = "Forest_Reserves"), mgrid@crs)
vecea <- readOGR(full_path(p_bd, "veceagcs.shp"), layer = "veceagcs",
                 stringsAsFactors = FALSE)

# assign threat matrix rank from van Breugel et al (2014) 
veceatypes <- unique(vecea@data)
veceatypes <- veceatypes[, c(1, 3)]
veceatypes[order(veceatypes$CODE), ]
vecea@data[vecea@data$CODE == "wd", ][1, ]
endang <- c("VU", "EN", "", "VU", "NA", "VU", "", "", "VU", "EN", "NA", "drop",
            "VU", "VU", "", "VU", "VU", "VU", "", "", "VU")
veceatypes$THRT <- endang
thrtmat <- cbind(unique(endang), c(3, 2, 4, 5, 6))
veceatypes$erank <- 0
for(i in 1:nrow(thrtmat)) {
  veceatypes[veceatypes$THRT == thrtmat[i, 1], "erank"] <- thrtmat[i, 2]
}

```

## Prepare impacts layer
```{r, eval = FALSE}
fnms <- full_path(full_path(p_bd, "act_veg"), 
                  dir(full_path(p_bd, "act_veg"), pattern = ".tif$"))
s <- stack(lapply(fnms, function(x) raster(x)))
sm <- mask(s, mgrid)

# Sum for total - the inverse of this will be the farmable areas
remfrac <- calc(s, sum)
remfrac <- mask(remfrac, mgrid)

# Calculate intactness
w <- matrix(nrow = 11, ncol = 11)
w[] <- 1
intact <- focal(remfrac, w = w, fun = mean, na.rm = TRUE)
# intact <- mask(intact, mgrid)
# plot(intact)

# Calculate a veg type weight based on relative remaining area
veg_dt <- dtraster::as.data.table(sm)  # stack and convert to dt
veg_dt <- na.omit(veg_dt)
vnms <- names(veg_dt)[-c(1:2)]
vtots <- veg_dt[, lapply(.SD, function(x) sum(x * 100)), 
                .SDcols = vnms]  
wgti <- round(1 / (vtots / nrow(veg_dt)), 2)  # bd weights (1 = rarest)
# wgt <- round(vtots / nrow(veg_dt), 4)  # biodiversity weights (1 = most common)
# plot(unlist(wgti), unlist(wgt))

# Check the relative areas against histograms from 30 m layer
v <- gdalinfo(vegrem@file@name, hist=TRUE)[31]  # gdalinfo histogram
v <- as.numeric(strsplit(v, " ")[[1]])  # convert to numeric, get numbers
v <- v[!is.na(v)]  # remove NAs
lcvals <- c(0:(length(v) - 1))[which(v != 0)]  # what are the actuals, less 0
hist30 <- cbind(lcvals, v[v != 0])

# check to see if similar to data.table derived versions (checks out)
plot(round(hist30[-1, 2] / sum(hist30[-1, 2]) * 100, 2), vtots / nrow(veg_dt))

```

## New impacts layer

Will be intactness, PAs, and rarity weights combined

### Weighting in relation to rarity and protectedness
```{r, eval = TRUE}
# Combine protected areas (NPs and GMAs, forest reserves)
nps$code <- ifelse(nps$TYPE == "gma", 2, 1) 
npsr <- rasterize(nps, mgrid, field = "code")  # rasterize NPs/GMAs
npsr[is.na(npsr)] <- 0  
freserves$code <- 3  
freservesr <- rasterize(freserves, mgrid, field = "code")  # forest reserves
freservesr[is.na(freservesr)] <- 0
pas <- npsr + freservesr  # sum them
reclmat <- cbind(0:5, c(0:3, 1, 2))
pas <- reclassify(pas, reclmat)
pas <- mask(pas, mgrid)
# plot(pas)
# plot(freservesr)

# Let's look at how much veg is protected, and maybe use that as basis for 
# weights - rarity + proportion protected
smpa <- mask(stack(s, pas), mgrid)
vegpa_dt <- as.data.table(smpa)
vegpa_dt <- na.omit(vegpa_dt)
setnames(vegpa_dt, "layer", "PA")
protected <- vegpa_dt[, lapply(.SD, sum(x * 100)), by = PA, .SDcols = vnms]
protected <- cbind("PA" = protected[, PA], 
                   protected[, lapply(.SD, function(x) {
                       round(x / sum(x), 3)
                   }), .SDcols = vnms])
# proportion protected in both NPs and GMA, and just in NPs
prop_prot2 <- protected[PA %in% c(1:2), lapply(.SD, sum), .SDcols = vnms]
prop_prot1 <- protected[PA == 1, lapply(.SD, sum), .SDcols = vnms]  # nps only

# stack PAS with intact area
pasconvs <- mask(stack(pas, remfrac), mgrid)
pa_dt <- as.data.table(pasconvs)
pa_dt <- na.omit(pa_dt)
setnames(pa_dt, c("layer.1", "layer.2"), c("pas", "nveg"))
pa_dt[, (1 - sum(nveg) / .N) * 100 , by = "pas"]
pa_dt[, .N, by = "pas"]

# how much of PAs converted
# 15.8 % outside of PAs converted (472,429 km2)
# 0.9 % NPs converted (60,201 km2)
# 3.8 % in GMAs converted (165228 km2)
# 8.3 % forest reserves converted (54091 km2)

# Weighting scheme, factoring in protection status
wgts <- unlist(log(wgti))  # log transform weights
wgtsn <- (wgts - min(wgts)) / diff(range(wgts))
normalize <- function(x) (x - min(x)) / diff(range(x))
wgts_prop <- normalize(unlist(wgtsn + (1 - prop_prot2)))
wgts_propl <- normalize(log(unlist(wgtsn + (1 - prop_prot2))))
wgts_prop_np <- normalize(unlist(wgtsn + (1 - prop_prot1)))

# plots to examine them
labs <- gsub("veg", "", names(wgtsn))
# par(mfrow = c(2, 1), mar = c(2, 2, 2, 2))
# weights against protected proportion (NPs + GMAs)
plot(wgtsn, prop_prot2, pch = "", ylab = "")
text(wgtsn, prop_prot2, labels = labs, cex = 0.75)
# weights against protected proportion (NPs only)
plot(wgtsn, prop_prot1, pch = "", ylab = "")
text(wgtsn, prop_prot1, labels = labs, cex = 0.75)
# additive weights factoring in protected proportion versus protected proportion
plot(wgts_prop2, prop_prot, pch = "", ylab = "")
text(wgts_prop2,  prop_prot, labels = labs, cex = 0.75)
# BD + NP protected proportion versus NP protected proportion
plot(wgts_prop_np, prop_prot1, pch = "", ylab = "")
text(wgts_prop_np,  prop_prot1, labels = labs, cex = 0.75)
# protected proportion weighted versus same but log-transformed
plot(wgts_prop, wgts_propl, pch = "", ylab = "")
text(wgts_prop,  wgts_propl, labels = labs, cex = 0.75)
# Bd weights against BD + protected proportion weights
plot(wgtsn, wgts_prop, pch = "", ylab = "")
lines(c(0, 1), c(0, 1))
text(wgtsn, wgts_prop, labels = labs, cex = 0.75)
# Bd weights against BD + NP protected proportion weights
plot(wgtsn, wgts_prop_np, pch = "", ylab = "")
lines(c(0, 1), c(0, 1))
text(wgtsn, wgts_prop_np, labels = labs, cex = 0.75)
# Bd weights versus log-transformed BD + protected proportion weights
plot(wgtsn, wgts_propl, pch = "", ylab = "")
lines(c(0, 1), c(0, 1))
text(wgtsn,  wgts_propl, labels = labs, cex = 0.75)
# The same but against multiplicative weights
plot(wgtsn, wgtsn * unlist(1 - prop_prot), pch = "", ylab = "")
lines(c(0, 1), c(0, 1))
text(wgtsn,  wgtsn * unlist(1 - prop_prot), labels = labs, cex = 0.75)
# sort(wgts_prop); sort(wgts_prop_np)
# sort(wgtsn)

# check against veceatypes von Breugel categories
veceatypes$wgtsn <- wgtsn[match(veceatypes$CAT, labs)]
veceatypes$wgtsnp <- wgts_prop[match(veceatypes$CAT, labs)]
plot(veceatypes$erank, veceatypes$wgtsn)  # not much connection
plot(veceatypes$erank, veceatypes$wgtsnp)  # not much connection
```

We will use the weights for rarity adjusted by how much the vegetation type falls in parks -- this corresponds to von Breugel risk class A C2 in Table 2.

### Reweight with intactness
```{r, eval = FALSE}
# assign weights to veg types
bdw <- wgts_prop_np
bd1 <- copy(veg_dt)
for(i in names(bdw)) {
  wi <- bdw[i]
  di <- eval(parse(text = paste0("quote(", i, ")")))
  bd1[eval(di) > 0, (i) := wi]
}

# Reduce to single column that takes the maximum weight across vegetation types
bd1[, TID := 1:.N ]
setkey(bd1, "TID")
bd1[, MWGT := apply(.SD, 1, max), .SDcols = vnms]
bd1r <- dt_to_raster(bd1[, .(x, y, MWGT)], mgrid@crs)
# plot(bd1r)
# plot(remfrac == 0)
# plot(smpa[["veg68"]])

# Stack vprior
bdstack <- stack(list("veg" = bd1r, "intact" = crop(intact, bd1r)))
bdr <- calc(bdstack, sum)
bdr <- (bdr - minValue(bdr)) / diff(cellStats(bdr, range, na.rm = TRUE))
# plot(bdr)

# Filter out PAs
# 0.9 % NPs converted (60,201 km2) - set NPs to 0
# 3.8 % in GMAs converted (165228 km2)  - set them to 4 %
# 8.3 % forest reserves converted (54091 km2) - set them to 8.3
pamat <- cbind(1:3, c(1, 1 - 0.038, 1 - 0.083))
pafix <- lapply(1:3, function(x) {  # x <- 2
  pamsk <- crop((pas == x) * pamat[x, 2], bdr)
  bdmsk <- mask(bdr, pamsk, maskvalue = 0)
  bdcalc <- calc(stack(bdmsk, pamsk), max)
})
pafix <- calc(stack(pafix), sum)  # sum those

# final biodiversity layer
bdf <- calc(stack(bdr, pafix), max)
plot(remfrac)
plot(bdf)

writeRaster(bdf, file = full_path(p_bd, "bdprior.tif"))

```