### 2016-01-20 Rmd/all-inputs.Rmd
Many updates as getting ready to assemble and mask inputs

---
title: "Assemble Inputs"
author: "Lyndon Estes"
date: "January 20, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Assemble Inputs

Bring all datasets in, create farmable area mask, mask them, and convert to data.tables for model input

1. For running model, non-grid: 
    + `currprod.rda` - For input, not data.table
    + `vprops.rda` - table of veg type area, rarity value, and prioritization value.

2. For running model:
    + Prioritization (P) and impact (I)
        + `ZA-carbon.tif` - Carbon impacts, to data.table (P|I)
        + `bdprior.tif` - biodiversity prioritization grid (P)
        + `ZA-cost.tif` - cost distance
        + `maize|soybean_S|R|I_10km,tif` - variants thereof (P|I). Separate variants of these must be created prior to model runtime. A function will be creatd here to facilitate that.  
    + Impact only:
        + `ZA-vegremain.tif` - vegetation remaining grid, for assessing conversion impacts. 
    + Masking/allocation
        + `farmable-1k.tif` - slope-based farmable area mask
        + `water-1k.tif` - water-based farmable area mask
        + `ZA-remfrac.tif` - cropland and urban-based farmable area mask
3. For displaying model outputs: 
    + `cropland-1k.tif` - cropland mask

## Inputs
```{r, eval = FALSE}
library(agroEcoTradeoff)

p_proj <- "/Users/lestes/Dropbox/publications/zambiaToff/private/zambiaToff/"
p_devel <- full_path(p_proj, "external/input_devel/cropland")
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_crop <- full_path(p_proj, "external/input_devel/cropland")
p_bd <- full_path(p_proj, "external/input_devel/biodiversity")
p_cost <- full_path(p_proj, "external/input_devel/roads/cost")

# Load in data
mgrid <- raster(full_path(p_basedat, "ZA-grid.tif"))
cropland <- raster(full_path(p_crop, fname = "cropland-1k.tif"))
bd <- raster(full_path(p_bd, fname = "bdprior.tif"))
vegr <- raster(full_path(p_bd, fname = "ZA-vegremain.tif"))
cost <- raster(full_path(p_cost, fname = "zamcost.tif"))

```