### 2016-01-21 Rmd/to-do.Rmd
Updated to-do

---
title: "To-Do List"
author: "Lyndon Estes"
date: "January 4, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Data and Improvements Needed

For the next iteration of the model and the ProcRoyalB paper.

## Improvements

A table of things to do, stating their progress and likelihood of achieving ("D" = doing, "L" is likely to get done, "U" is unlikely to get done, "F" is finished, "A" is abandoned) during the month of January.  __Table last updated: `r Sys.Date()`__

```{r, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
# http://stackoverflow.com/questions/19997242/simple-manual-rmarkdown-tables-that-look-good-in-html-pdf-and-docx
# tabl <- "  
# | Item        | Description           | Doing or Done?  |
# |---------------|:-------------|------|
# | Plausible scenario | A single production target to follow (Alastair)  | Doing |
# | Gridded yield estimates | Maize and soybean, exact method for handling variability & production zones (e.g. irrigation regions) |  Doing |
# | Gridded yield estimates | Run with improved soil profiles (Di) | Unlikely |
# | Upgrade soils | With ILUA database (with Nate?) | Unlikely |
# | Filter terrain | Get rid of unfarmable bits, probably needs new model logic (cell fraction multiplier) | Doing | 
# 
# "
# cat(tabl) # output the table in a format good for HTML/PDF/docx conversion

# make an html table
# http://r-norberg.blogspot.com/2013/06/generating-tables-using-pander-knitr.html
tabl <- cbind(
  "Item" = c("Plausible Scenario", 
             "Gridded yields",
             "Gridded yields",
             "Upgrade soils", 
             "Filter terrain", 
             "Biodiversity impacts", 
             "Impacts weighting (per Tim)", 
             "Work in economics", 
             "Explore econ versus enviro goals", 
             "Better cropland map, yield estimates", 
             "Updated carbon layers"), 
  "Notes" = c("Single production to follow (Alastair)", 
              "Exact method for variability and production zones, including inability to reach potential yields", 
              "Run with improved soils maps", 
              "With ILUA database (and Nate's help)", 
              "Removable unfarmable areas, add new supporting logic", 
              "Habitat rarity/threatenedness", 
              "Weight carbon and biodiversity equally", 
              "Estimate profitability per pixel. Either input modification or code modification", 
              "See where map overlaps are greatest, most divergent", 
              "Upgrade all maps to SERVIR landcover, and rebase yield estimates on that", "Baccini vegetation carbon + ISRIC grids"),
  "Doing" = c("D", "F*", "U", "U", "D", "D", "D", "L", "D", "F**", "F")
)
print(xtable::xtable(tabl), align = rep("c", dim(table)[2] + 1), type = "html", 
      include.rownames = FALSE)
```
*Temporary soybean grid with bias adjustment from FAO GAEZ yields. Developed weighted merged method for different types of management.

**Just using total production estimates, which will not be mapped for now. These will be basis for estimating production targets.