### 2016-02-08 Rmd/empirical-yields.Rmd
Many changes, including development of yield-predictors, empirical-yields (GAM fit to DSSAT yields), paper to-do list, etc.

---
title: "Stage 2 Yield Prediction"
author: "Lyndon Estes"
date: "February 5, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Empirical prediction of yield

A yield transfer function for DSSAT yields

```{r, eval = FALSE}
library(agroEcoTradeoff)
library(matrixStats)
library(doMC)
library(mgcv)

p_root <- lmisc::set_base_path()
p_clim <- fp(p_root, "external/input_devel/climate")
p_yield <- fp(p_root, "external/input_devel/yield")

# Zambia provinces
p_prov <- paste0("/Users/lestes/Dropbox/data/zari/infrastructure/zambia/", 
                 "rda_dump/Final Shape files RDA/")
prov <- readOGR(fp(p_prov, "prov.shp"), layer = "prov")

# Regression variables
load(fp(p_yield, "empirical/regdat.rda"))

```

# Fit model
## Exploratory plots
```{r, eval = FALSE}
plot(prov)
regdt[CROP == "maize", points(x, y, pch = 16, cex = 0.5)]
regdt[CROP == "maize" & YLDS > 8000, 
      points(x, y, pch = 16, cex = 0.5, col = "red")]
regdt[CROP == "maize" & YLDS > 8000, 
      points(x, y, pch = 16, cex = 0.5, col = "red")]
regdt[CROP == "soybean" & YLDS > 1700, 
      points(x, y, pch = 16, cex = 0.5, col = "blue")]

# Prune out yields > 8000 maize > 1700 soybean in subsistence case
mzfilt <- regdt[CROP == "soybean" & YLDS > 1700, unique(.SD), 
                .SDcols = c("x", "y", "SOL", "WTH", "Zone")]
soyfilt <- regdt[CROP == "maize" & YLDS > 8000, unique(.SD), 
                 .SDcols = c("x", "y", "SOL", "WTH", "Zone")]
regnew <- regdt[!x %in% soyfilt$x | !y %in% soyfilt$y]
# regnew[CROP == "maize", points(x, y, pch = 16, cex = 0.5, col = "green")]

# regnew[CROP == "maize", hist(YLDS)]
# regnew[CROP == "maize", hist(YLDR)]

# summary(lm(YLDS ~ poly(MAP, 2), data = regnew[CROP == "maize", ]))
# summary(lm(YLDR ~ poly(MSP, 2), data = regnew[CROP == "maize", ]))
regnew[CROP == "maize", plot(MAP, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "soybean", plot(MAP, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "soybean", plot(MAP, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "soybean", plot(MSP, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "soybean", plot(MSP, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(FP, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(MSP, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(MSP, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(sloc, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(sloc, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(slcl, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(slcl, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(slsi, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(slsi, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(sbdm, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(sbdm, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(slhw, YLDS, pch = 16, cex = 0.5)]  # pH looks pred
regnew[CROP == "maize", plot(slhw, YLDR, pch = 16, cex = 0.5)]  # pH looks pred
regnew[CROP == "maize", plot(GDD, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(KDD, YLDS, pch = 16, cex = 0.5)]
# regnew[CROP == "maize", plot(log(KDD), YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(NKJ, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(NKF, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(ATMX, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(STMX, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(STMX, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SSRT, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SSRM, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(ASRT, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(ASRM, YLDS, pch = 16, cex = 0.5)]

regnew[, round(cor(.SD), 2), .SDcols = names(regnew)[-c(1:13)]]

summary(lm(YLDS ~ poly(sbdm, 2) + poly(slhw, 2), 
           data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ poly(slhw, 2), data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ poly(sbdm, 2), data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ poly(KDD, 2), data = regnew[CROP == "maize", ]))
summary(lm(YLDR ~ poly(STMX, 2), data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ poly(STMX, 2), data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ KDD, data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ poly(GDD, 2), data = regnew[CROP == "maize", ]))

# p <- "sbdm"; d <- "YLDS"
# regnew[CROP == "maize", {
#   lm1 <- lm(get(d) ~ get(p))
#   lm2 <- lm(get(d) ~ poly(get(p), 2))
#   plot(get(p), get(d), pch = 16, cex = 0.1)
#   abline(lm1)
#   dum <- data.frame(floor(min(get(p))):ceiling(max(get(p))))
#   colnames(dum) <- p
#   lines(dum[, 1], predict(lm2, newdata = dum), col = "red")
#   summary(lm1)
#   # summary(lm2)
# }]

reg_plotter <- function(dep, pred) {
  lm1 <- lm(dep ~ pred)
  lm2 <- lm(dep ~ poly(pred, 2))
  plot(pred, dep, pch = 16, cex = 0.1)
  abline(lm1)
  prng <- c(floor(min(pred)), ceiling(max(pred)))
  dum <- data.frame(seq(prng[1], prng[2], by = diff(prng) / 100))
  # dum <- data.frame(floor(min(pred)):ceiling(max(pred)))
  colnames(dum) <- "pred"
  preds <- predict(lm2, newdata = dum)
  lines(dum[, 1], preds, col = "red")
  mtext(round(summary(lm1)$adj.r.squared, 4), side = 3, adj = 0)
  mtext(round(summary(lm2)$adj.r.squared, 4), side = 3, adj = 0.5)
}
# regnew[CROP == "maize", reg_plotter(YLDR, GDD)]  # 
regnew[CROP == "maize", reg_plotter(YLDS, MAP)]  # P 0.045
regnew[CROP == "maize", reg_plotter(YLDS, MSP)]  # P 0.039
regnew[CROP == "maize", reg_plotter(YLDS, JP)]  # P 0.012
# regnew[CROP == "maize", reg_plotter(YLDS, JP - mean(JP))]  # P 0.012
regnew[CROP == "maize", reg_plotter(YLDS, FP)]  # P 0.005
regnew[CROP == "maize", reg_plotter(YLDS, OP)]  # P 0.003

regnew[CROP == "maize", reg_plotter(YLDS, slhw)]  # P 0.17
regnew[CROP == "maize", reg_plotter(YLDS, sloc)]  # P 0.266
# regnew[CROP == "maize", reg_plotter(YLDR, sloc)]  # P 
regnew[CROP == "maize", reg_plotter(YLDS, slcl)]  # P 0.185
regnew[CROP == "maize", reg_plotter(YLDS, sbdm)]  # 0.181
# mtax pairings
regnew[CROP == "maize", reg_plotter(YLDS, STMX)]  # P 0.07
regnew[CROP == "maize", reg_plotter(YLDS, ATMX)]  # P 0.056 
# NK
regnew[CROP == "maize", reg_plotter(YLDS, NKJ)]  # L 0.054
regnew[CROP == "maize", reg_plotter(YLDS, NKF)]  # P 0.027 
# GDD/KDD
regnew[CROP == "maize", reg_plotter(YLDS, GDD)]  # P 0.04
regnew[CROP == "maize", reg_plotter(YLDS, KDD)]  # L 0.027
# srad
regnew[CROP == "maize", reg_plotter(YLDS, SSRT)]  # P 0.012 
regnew[CROP == "maize", reg_plotter(YLDS, SSRM)]  # Same, picked up wrong one
regnew[CROP == "maize", reg_plotter(YLDS, ASRT)]  # P 0.0288 
regnew[CROP == "maize", reg_plotter(YLDS, ASRM)]  # P 0.0299

# best predictors look to be
# MAP, SLOC (but polynomial relationship odd), SLHW, slcl, sbdm, all good
# STMX (linear KDD as a possible alternate), GDD, ASRM

#
regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4279

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + poly(STMX, 2) + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4199

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(slhw, 2) + poly(sloc, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + poly(STMX, 2) + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4044 # SLOC doesn't give as much value, NS on linear

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4293  # Doesn't add much

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4293  # Oct precip add a tiny bit, significant

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + JP + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4312  # Jan precip adds more R2, but makes MAP linear NS

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + FP + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4374  # Feb precip even more R2, but makes MAP linear NS

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + x + y)
  summary(lm1)
  plot(YLDS, predict(lm1))
}]  # 0.4301  # Oct precip add a tiny bit, significant

# regnew[CROP == "maize", {
#   gam1 <- bam(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
#               poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + s(x, y), 
#              family = gaussian(link = "log"), method = "ML")
#   summary(gam1)
#   # plot(YLDS, predict(lm1))
# }]  # 0.4301  # Oct precip add a tiny bit, significant

```

## Reduce variables and fit models

Reduce down to means across years, and select 10% and 90% percentile years

### Maize
```{r, eval = FALSE}
regred <- copy(regnew[CROP == "maize"])
regred[, mid2 := 1:.N]

# regred[, c("Y10S", "Y90S") := as.integer(0)]

regmu <- regnew[CROP == "maize", lapply(.SD, mean), 
                by = list(CROP, WTH, Zone, SOL), 
                .SDcols = names(regnew)[c(3:4, 11:ncol(regnew))]]
regmu[, YST := "mu"]
# regnew[CROP == "maize" & Zone == "Zone2a" & WTH == "AABS" & 
#          SOL == "WI_CMZR003", mean(YLDS)]
qs <- function(x, y, ptile) {
  DF <- cbind(1:length(x), x, y)
  DF <- DF[order(DF[, 2]), ]
  DF[round(nrow(DF) * ptile), 3]
}

y <- 30:60
DF <- cbind(1:31, sample(2000:3000, size = 31), y)
DF <- DF[order(DF[, 2]), ]
DF[round(nrow(DF) * 0.9), 3]
sort(DF[, 2])

a <- regnew[CROP == "maize" & Zone == "Zone3" & WTH == "AEQA" & 
            SOL == "WI_RGTZ054", YLDS]
b <- regnew[CROP == "maize" & Zone == "Zone3" & WTH == "AEQA" & 
            SOL == "WI_RGTZ054", mid]
qs(a, b, 0.9)
sort(a)

# Select out quantile values
m10 <- regred[,
  qs(YLDS, mid2, 0.1), by = list(CROP, WTH, Zone, SOL)
][, V1]
m90 <- regred[,
  qs(YLDS, mid2, 0.9), by = list(CROP, WTH, Zone, SOL) 
][, V1]

# 
y10 <- regred[mid2 %in% m10]
y90 <- regred[mid2 %in% m90]
y10[, YST := "10"]
y90[, YST := "90"]

y1090 <- rbind(y10, y90)
cnames <- colnames(y1090)[colnames(y1090) %in% colnames(regmu)]
setcolorder(regmu, cnames)
regred2 <- rbind(regmu, y1090[, cnames, with = FALSE])

par(mfrow = c(2, 2))
for(i in c("10", "mu", "90")) {
  regred2[CROP == "maize" & YST == i, hist(YLDS)]
}
regred2[CROP == "maize", mean(YLDS), by = YST]

regred2[CROP == "maize", reg_plotter(YLDS, MAP)]  # P 0.071
regred2[CROP == "maize", reg_plotter(YLDS, MSP)]  # P 0.039
regred2[CROP == "maize", reg_plotter(YLDS, JP)]  # P 0.020
regred2[CROP == "maize", reg_plotter(YLDS, FP)]  # P 0.017
regred2[CROP == "maize",  summary(lm(YLDS ~ poly(FP, 2)))]
regred2[CROP == "maize", reg_plotter(YLDS, 100 - slcl + slsi)]  # P 0.102
regred2[CROP == "maize", reg_plotter(YLDS, slcl)]  # P 0.1879
regred2[CROP == "maize", reg_plotter(YLDS, slsi)]  # P 0.03

regred2[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + x + y)
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4301  # Oct precip add a tiny bit, significant

regred2[CROP == "maize" & YST == "mu", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + x + y)
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4301  # Oct precip add a tiny bit, significant

regred2[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4301  # Oct precip add a tiny bit, significant

regred2[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + poly(sloc, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4301  # Oct precip add a tiny bit, significant

regred2[CROP == "maize", {
  gam1 <- gam(YLDS ~ s(MAP) + s(JP) + s(slhw) + s(slcl) + (sbdm) + s(GDD) +
                KDD + s(ASRM), 
            family = gaussian(link = "log"), method = "ML")
}]

gam1 <- gam(YLDS ~ s(MAP, k = 3) + s(JP, k = 3) + s(slhw, k = 3) + 
              s(slcl, k = 3) + s(sbdm, k = 3) + s(GDD, k = 3) + KDD + 
              s(ASRM, k = 3), family = gaussian(link = "log"), method = "ML", 
            data = regred2[CROP == "maize"])
gam2 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + 
              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD + 
              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam3 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + 
              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD + 
              poly(ASRM, 2) + s(x, y), 
            family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam4 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + 
              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD + 
              s(x, y), family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam5 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + 
              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD, 
            family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam6 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + 
              poly(slcl, 2) + poly(GDD, 2) + KDD, 
            family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam7 <- gam(YLDS ~ poly(MAP, 2) + poly(slhw, 2) + 
              poly(slcl, 2) + poly(GDD, 2) + KDD, 
            family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam8 <- gam(YLDS ~ poly(MAP, 2) + slhw +  poly(slcl, 2) + poly(GDD, 2) + KDD, 
            family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam9 <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
              poly(100 - (slcl + slsi), 2) + poly(GDD, 2) + KDD, 
            family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam10 <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
               poly(sloc, 2) + poly(GDD, 2) + KDD, 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
gam11 <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
               sloc + poly(GDD, 2) + KDD, 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
gam12 <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
               sloc + poly(GDD, 2) + KDD + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])

summary(gam3)
plot.gam(gam3, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam4)
plot.gam(gam4, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam5)
plot.gam(gam5, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam6)
plot.gam(gam6, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam7)
plot.gam(gam7, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam8)
plot.gam(gam8, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam9)
plot.gam(gam9, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam10)
plot.gam(gam10, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam11)  # This one or
plot.gam(gam11, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam12)  # of this one, with smooth x, y terms
plot.gam(gam12, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)

par(mfrow = c(2, 2))
plot(regred2[CROP == "maize", YLDS], predict(gam11, type = "response"))
plot(regred2[CROP == "maize", YLDS], predict(gam12, type = "response"))
plot(regred2[CROP == "maize", YLDS], predict(gam4, type = "response"))

# Reduce down to model for mean, 10th, 90th)
gam12mu <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
               poly(GDD, 2) + KDD + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[YST == "mu"])
summary(gam12mu)
plot.gam(gam12mu, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
gam1210 <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
               poly(GDD, 2) + KDD + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[YST == "10"])
summary(gam1210)
plot.gam(gam1210, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
gam1290 <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
               poly(GDD, 2) + KDD + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[YST == "90"])
summary(gam1290)
plot.gam(gam1290, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)

```

### Soybean

```{r, eval = FALSE}
regreds <- copy(regnew[CROP == "soybean"])
regreds[, mid2 := 1:.N]

# regred[, c("Y10S", "Y90S") := as.integer(0)]

regmus <- regreds[, lapply(.SD, mean), by = list(CROP, WTH, Zone, SOL), 
                .SDcols = names(regnew)[c(3:4, 11:ncol(regnew))]]
regmus[, YST := "mu"]

# Select out quantile values
m10 <- regreds[,
  qs(YLDR, mid2, 0.1), by = list(CROP, WTH, Zone, SOL)
][, V1]
m90 <- regreds[,
  qs(YLDR, mid2, 0.9), by = list(CROP, WTH, Zone, SOL) 
][, V1]

# Select out 10th and 90th percentile years
y10s <- regreds[mid2 %in% m10]
y90s <- regreds[mid2 %in% m90]
y10s[, YST := "10"]
y90s[, YST := "90"]

y1090s <- rbind(y10s, y90s)
cnames <- colnames(y1090s)[colnames(y1090s) %in% colnames(regmus)]
setcolorder(regmus, cnames)
regreds2 <- rbind(regmus, y1090s[, cnames, with = FALSE])
regreds2[, slcf := 100 - (slcl + slsi)]

par(mfrow = c(2, 2))
for(i in c("10", "mu", "90")) {
  regreds2[CROP == "soybean" & YST == i, hist(YLDR)]
}
regreds2[CROP == "soybean", mean(YLDR), by = YST]

regreds2[, reg_plotter(YLDR, MAP)]  # L 0.065
regreds2[, reg_plotter(YLDR, MSP)]  # L 0.1105
regreds2[, reg_plotter(YLDR, JP)]  # L 0.08
regreds2[, reg_plotter(YLDR, FP)]  # P 0.016
regreds2[, reg_plotter(YLDR, 100 - slcl + slsi)]  # P 0.12
regreds2[, reg_plotter(YLDR, slcl)]  # P 0.31
regreds2[, reg_plotter(YLDR, slsi)]  # P 0.003
regreds2[, reg_plotter(YLDR, sloc)]  # P 0.31
regreds2[, reg_plotter(YLDR, slhw)]  # P 0.27
regreds2[, reg_plotter(YLDR, GDD)]  # P 0.006
regreds2[, reg_plotter(YLDR, KDD)]  # P 0.006
regreds2[, reg_plotter(YLDR, STMX)]  # L 0.029
regreds2[, reg_plotter(YLDR, NKF)]  # L 0.032
regreds2[, reg_plotter(YLDR, NKJ)]  # P 0.029
regreds2[, reg_plotter(YLDR, SSRT)]  # L 0.023
regreds2[, reg_plotter(YLDR, ASRM)]  # L 0.055

sgam1 <- gam(YLDR ~ s(MAP, k = 3) + s(JP, k = 3) + s(slhw, k = 3) + 
               s(slcl, k = 3) + s(sbdm, k = 3) + s(GDD, k = 3) + KDD + 
               s(ASRM, k = 3), family = gaussian(link = "log"), method = "ML", 
            data = regreds2)

sgam2 <- gam(YLDR ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + 
              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD + 
              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam3 <- gam(YLDR ~ poly(MAP, 2) + poly(FP, 2) + poly(slhw, 2) + 
              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD + 
              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam4 <- gam(YLDR ~ poly(MAP, 2) + poly(OP, 2) + poly(slhw, 2) + 
              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD + 
              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam5 <- gam(YLDR ~ poly(MAP, 2) + poly(OP, 2) + poly(slhw, 2) + 
              poly(slcl, 2) + poly(sloc, 2) + poly(GDD, 2) + KDD + 
              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam6 <- gam(YLDR ~ poly(MAP, 2) + poly(OP, 2) + poly(slhw, 2) + 
              poly(slcl, 2) + sloc + poly(GDD, 2) + KDD + 
              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam7 <- gam(YLDR ~ poly(MAP, 2) + poly(OP, 2) + poly(slhw, 2) + 
              poly(slcl, 2) + poly(GDD, 2) + KDD + 
              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam8 <- gam(YLDR ~ poly(MAP, 2) + poly(slhw, 2) + poly(slcl, 2) + 
               poly(GDD, 2) + KDD + poly(ASRM, 2), 
             family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam9 <- gam(YLDR ~ poly(MAP, 2) + poly(slhw, 2) + poly(slcl, 2) + 
               poly(GDD, 2) + poly(ASRM, 2), 
             family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam10 <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + poly(slcl, 2) + 
               poly(GDD, 2) + poly(ASRM, 2), 
              family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam11 <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + poly(sbdm, 2) + 
               poly(GDD, 2) + poly(ASRM, 2), 
              family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam12 <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + poly(slcl, 2) + 
               poly(GDD, 2) + poly(SSRM, 2), 
              family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam13 <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + poly(slcl, 2) + sloc +  
               poly(GDD, 2) + poly(SSRM, 2), 
              family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam14 <- gam(YLDR ~ poly(MSP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) + 
               poly(GDD, 2) + poly(SSRM, 2), 
              family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam15 <- gam(YLDR ~ poly(MSP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) + 
               poly(GDD, 2) + KDD + poly(SSRM, 2), 
              family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam16 <- gam(YLDR ~ poly(MSP, 2) + poly(OP, 2) + poly(slhw, 2) + poly(slcl, 2) + 
               poly(GDD, 2) + poly(SSRM, 2), 
              family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam17 <- gam(YLDR ~ poly(MSP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) + 
               poly(GDD, 2) + poly(SSRM, 2) + s(x, y), 
              family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam18 <- gam(YLDR ~ poly(MSP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) + 
               poly(GDD, 2) + NKJ + poly(SSRM, 2) + s(x, y), 
              family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam19 <- gam(YLDR ~ poly(MSP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) + 
               poly(GDD, 2) + NKF + poly(SSRM, 2) + s(x, y), 
              family = gaussian(link = "log"), method = "REML", data = regreds2)

summary(sgam1)
plot.gam(sgam1, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam2)
plot.gam(sgam2, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam3)
plot.gam(sgam3, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam4)
plot.gam(sgam4, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam5)
plot.gam(sgam5, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam6)
plot.gam(sgam6, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam7)
plot.gam(sgam7, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam8)
plot.gam(sgam8, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam9)  # MSP stronger than MAP
plot.gam(sgam9, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam10)  # MSP stronger than MAP
plot.gam(sgam10, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam11)  # SBDM much weaker SLCL
plot.gam(sgam11, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam12)  # SSRM a tad better than 
plot.gam(sgam12, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam13)  # SLOC goes in the wrong direction
plot.gam(sgam13, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam14)  # OP adds a bit
plot.gam(sgam14, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam15)  # KDD NS
plot.gam(sgam15, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam16)  # Polynomial term on October precip doesn't add much
plot.gam(sgam16, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam17)  # xy term added to model 15, all remain signif, seems way to go
plot.gam(sgam17, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam18)  # NKJ NS
plot.gam(sgam18, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam19)  # NKF NS
plot.gam(sgam19, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)

# Now see about taking same model and fitting separately to 10th/90th/mean years
sgam17mu <- gam(YLDR ~ poly(MSP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) + 
                  poly(GDD, 2) + poly(SSRM, 2) + s(x, y), 
                family = gaussian(link = "log"), method = "REML", 
                data = regreds2[YST == "mu"])
summary(sgam17mu)  # OP drops out, SSRM drops out
plot.gam(sgam17mu, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
sgam17mu <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + 
                  poly(slcl, 2) + poly(GDD, 2) + s(x, y), 
                family = gaussian(link = "log"), method = "REML", 
                data = regreds2[YST == "mu"])
summary(sgam17mu)  # OP drops out, SSRM drops out
plot.gam(sgam17mu, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
sgam1710 <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + poly(slcl, 2)
                + poly(GDD, 2) + s(x, y), 
                family = gaussian(link = "log"), method = "REML", 
                data = regreds2[YST == "10"])
summary(sgam1710)  # OP drops out, SSRM drops out
plot.gam(sgam1710, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
sgam1790 <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + poly(slcl, 2)
                + poly(GDD, 2) + s(x, y), 
                family = gaussian(link = "log"), method = "REML", 
                data = regreds2[YST == "90"])
summary(sgam1790)  # OP drops out, SSRM drops out
plot.gam(sgam1790, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)

```

### Save GAMs outputs
```{r, eval = FALSE}
save(gam12, gam12mu, gam1210, gam1290, sgam17, sgam17mu, sgam1710, sgam1790, 
     file = fp(p_yield, "gams.rda"))

```