### 2016-02-10 Rmd/DSSAT-yields.Rmd
Updated on DSSAT and empirical modeling side: 1. Created DSSAT-yields.Rmd, for running DSSAT for both maize and soybean. 2. Results of these go into yield-predictors and empirical-yields to create modeled crop yields (using a GAM). 3. zmd-gsod-stations.Rmd creates the station points used in the analysis.

---
title: "DSSAT soybean yields"
author: "Lyndon Estes"
date: "February 9, 2016"
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# DSSAT soybean runs for Zambia

```{r, eval = FALSE}
library(agroEcoTradeoff)
library(gdalUtils)
library(rCropMod)
library(doMC)
library(dismo)

setwd("~/Dropbox/publications/zambiaToff/private/zambiaToff/")
p_root <- lmisc::set_base_path("zambiaToff")
p_cropmod <- "~/Dropbox/projects/rCropMod/external/ext_data/"
p_dssat <- "~/DSSAT45"  # DSSAT main folder
p_wth <- "~/DSSAT45/Weather/"
p_soil <- "~/DSSAT45/SOIL/"
p_clim <- fp(p_root, "external/input_devel/climate")
p_yield <- fp(p_root, "external/input_devel/yield")
p_pgf <- "~/Dropbox/data/climate/afdm/pgf_bc/zambia/"
p_sgrids <- "~/Dropbox/data/soil/"

load(fp(p_yield, "empirical/stations.rda"))

# Read in original soil-climate data from rCropMod
soilgrid <- raster(fp(p_cropmod, "af_soil_grid.tif"))  # soil grid
soilshare <- fread(fp(p_cropmod, "af_soil_share.csv"))  # share of profiles

```

## Weather data
```{r, eval = FALSE}
sdate <- "1948/01/01"
edate <- "2010/12/31"
dts <- seq.Date(as.Date(sdate), as.Date(edate), "days")
yrs <- as.numeric(substr(dts, 1, 4))
ano <- unique(yrs)
mos <- as.numeric(substr(dts, 6, 7))
dys <- as.numeric(substr(dts, 9, 10))

# set up indices for selecting dates
ystrt <- which(mos == 1 & dys == 1) 
ystrt <- ystrt[-length(ystrt)]
yend <- which(mos == 12 & dys == 31)[-1]
# set up seasonal index
sind <- rep(0, length(dts))  # whole year spanning growing season
for(i in 1:length(ystrt)) sind[ystrt[i]:yend[i]] <- i
sindu <- unique(sind)[-1]
sindu2 <- sindu[sindu %in% which(ano %in% 1979:2010)]
ind <- which(sind %in% sindu2)

prec <- brick(fp(p_pgf, "prec.Zambia.nc"))
tmax <- brick(fp(p_pgf, "Tmax.Zambia.nc")) 
tmin <- brick(fp(p_pgf, "Tmin.Zambia.nc"))
srad <- brick(fp(p_pgf, "Rs.Zambia.nc")) 
wind <- brick(fp(p_pgf, "wnd10m.Zambia.nc"))

# extract weather station data
registerDoMC(cores = 7)
wdat <- foreach(i = 1:nrow(zmetf)) %dopar% {  # i = 1 
  print(i)
  rain <- unname(extract(prec, zmetf[i, ])[1, ])
  tmx <- unname(extract(tmax, zmetf[i, ])[1, ])
  tmn <- unname(extract(tmin, zmetf[i, ])[1, ])
  sr <- unname(extract(srad, zmetf[i, ])[1, ])
  wnd <- unname(extract(wind, zmetf[i, ])[1, ])
  cbind(rain, tmx, tmn, sr, wnd)
}

# reduce down and modify
wdat2 <- lapply(wdat, function(x) {
  y <- x[ind, ]
  y[, 2] <- y[, 2] - 273.15
  y[, 3] <- y[, 3] - 273.15
  y[, 4] <- y[, 4] / 1000000 * (3600 * 24)
  y
})
# tstwsta <- cbind.data.frame(dts[ind], wdat2[[1]])
# head(tstwsta); tail(tstwsta)

# get elevations for the weather stations
# raster::getData('ISO3')  # raster is a marvelous package
elev <- raster::getData("alt", country = "ZMB")
zmet_elev <- extract(elev, zmetf)
# plot(elev)

# Write WTH files to climate folder
sdate <- "19790101" 
edate <- "20101231"
zmetxy <- as.data.frame(zmetf)
zmetxy[1:6, ]
wthnms <- substr(zmetxy$nms, 1, 4)
ginds <- grep("g", wthnms)
combis <- expand.grid(data.frame(cbind(LETTERS, LETTERS, LETTERS)))[, 3:1]
combis <- paste0(combis[, 1], combis[, 2], combis[, 3])
wthnms[ginds] <- paste0("G", combis[1:length(ginds)])
for(i in 1:length(wdat2)) {  # i <- 1
  xy <- c("x" = zmetxy[i, "x"], "y" = zmetxy[i, "y"]) 
  elev <- zmet_elev[i]
  name <- wthnms[i]
  d <- wdat2[[i]]  # head(d)
  weather(xy, elev = elev, srad = d[, "sr"], tmax = d[, "tmx"], 
          tmin = d[, "tmn"], prec = d[, "rain"], 
          sdate = sdate, edate = edate, name = name, 
          outdir = fp(p_clim, "DSSAT"))
}
file.copy(dir(fp(p_clim, "DSSAT"), full.names = TRUE), 
          fp(p_wth, dir(fp(p_clim, "DSSAT"))), overwrite = TRUE)

```

## Pull the soil profiles
```{r, eval = FALSE}
soilid <- extract(soilgrid, zmetf)  # grid id corresponding to point
wthsoil <- cbind(wthnms, zmetxy, soilid)
profs <- soilshare[CELL5M %in% soilid]  # profiles at that point
profdat <- read_fullSOL_prof(solfile = full_path(p_soil, "WI.SOL"), 
                             profiles = profs$SoilProfile)

# checks
set.seed(234)
for(i in sample(1:nrow(profdat), 4)) {
  print(profdat[prof == profdat$prof[i], c(1:5, 11:16), with = FALSE])
}

# uniprofs <- unique(profs$SoilProfile)
# profdat2 <- do.call(rbind, lapply(1:length(uniprofs), function(i) {
#   print(i)
#   profdat <- read_SOL_prof(solfile = full_path(p_soil, "WI.SOL"), 
#                            profiles = uniprofs[i]) 
# }))
# profdat$slb3 == profdat2$SLB3

# profdat[pro == "WI_VRBW509"]
profdat[, prof := as.character(prof)]
setkey(profdat, "prof")
wthprofs <- data.table(merge(wthsoil, profs, by.x = "soilid", by.y = "CELL5M"))
setnames(wthprofs, "SoilProfile", "prof")
wthsoilm <- merge(wthprofs, profdat, by.x = "prof", by.y = "prof")
wthsoilmr <- wthsoilm[, 1:37, with = FALSE]
names(wthsoilm)
# profdat[, 1:4, with = FALSE]      
# head(wthprofs)
# wthsoilm[1:10, 1:7, with = FALSE]
# wthsoilm[prof == "WI_VRBW509", 1:20, with = FALSE]
```

## Soybean X-files and runs
```{r, eval = FALSE}

# read in field parameters
field <- readxl::read_excel(fp(p_yield, "DSSAT/fields.xls"))  
field <- cbind(wthsoilmr, field)  # join the soils data
field[, WTH := paste0(wthnms, "7932")]
field[, c("ID_FIELD", "XNAME") := list(fid(field$WTH), xname(field$WTH))]

# Create X table
xtab <- x_tab(fids = field)  # table created, with all fixed field parameters
cult <- c("990008", "990009", "990010")
CLNUM <- 1:length(cult)
pdates <- strftime(as.Date("1979-12-31"), "%y%j")
tcomb <- expand.grid(list("PDATE" = pdates, "INGENO" = cult), 
                     stringsAsFactors = FALSE)
ttab <- cbind("N" = 1:nrow(tcomb), tcomb, 
              t_tab(tvars = c("PDATE", "INGENO"), topts = c("MP", "CU"), 
                    ttab = tcomb))

# Speed this up
xtabl <- lapply(1:nrow(xtab), function(x) { # x <- 1
# xtabl <- lapply(1:10, function(x) { # x <- 1
  print(x)
  d <- xtab[x, ]
  d2 <- do.call(rbind, lapply(1:nrow(ttab), function(x) d))
  upd_col <- colnames(d2)[which(colnames(d2) %in% colnames(ttab))]
  d2[, c(upd_col) := ttab[, upd_col]]
  xt <- cbind(data.table(ttab), d2[, !colnames(d2) %in% upd_col, with = FALSE])
})

xrun <- lapply(xtabl, function(x) {  # x <- xtabl[[1]]
  print(x$XNAME[1])
  xf <- copy(x)
  xf <- as.data.frame(xf)
  xfnm <- x_file(xf, outdir = fp(p_dssat, "Soybean"), z = "01", xtype = ".SNX")
  bname <- batch_file(xl = list(xf), xfiles = xfnm, 
                    outdir = full_path(p_dssat, "Soybean"), btype = "SOY")
  exec_csm(projdir = fp(p_yield, "DSSAT"), csmdir = p_dssat, 
         rundir = full_path(p_dssat, "Soybean"), bname = bname)
  sdat <- read_CSM_outfile(rundir = full_path(p_dssat, "Soybean"), 
                           type = "summary", 
                           vars = c("RUNNO", "TRNO", "FNAM", "SOIL_ID...", 
                                    "PDAT", "MDAT", "HWAH", "PRCM"))
})
# save(xrun, file = fp(p_yield, "DSSAT/soyyields.rda"))

soy_ylds <- rbindlist(lapply(1:length(xrun), function(i) { # i <- 1
  xbinder <- xtabl[[i]][1, list(XNAME, wthnms, x, y, prof)]
  setnames(xbinder, c("wthnms", "prof"), c("WTH", "PROF"))
  cbind(xbinder, xrun[[i]])
}))  
# soy_ylds[, lapply(.(HWAH, PRCM), mean), by = XNAME][, plot(V2, V1)]
# soy_ylds[, lapply(.(HWAH, PRCM), mean), by = XNAME][, mean(V1)]
# 3074 mean with current settings

write.table(soy_ylds, file = fp(p_yield, "DSSAT/soy_ylds.csv"), 
            row.names = FALSE, sep = ",")
# soy_ylds[, unique(XNAME)]

# check on soil configuration
tst <- unique(soy_ylds[, .(x, y, WTH, PROF)])
tst[WTH == "LUSA"]
unixy <- unique(wthprofs[, .(x, y)])
for(i in 1:nrow(unixy)) {  # i = 17
  p <- wthsoilmr[x == unixy$x[i] & y == unixy$y[i], .(wthnms, prof)]
  # p <- wthprofs[x == unixy$x[i] & y == unixy$y[i], .(wthnms, prof)]
  tstp <- tst[x == unixy$x[i] & y == unixy$y[i], .(WTH, PROF)]
  if(!all(tstp == p)) print(paste(i, "has a problem"))
#   if(!all(tstp[order(WTH, PROF)] == p[order(wthnms, prof)])) {
#     print(paste(i, "has a problem"))
#   }
}

```

## Maize yields
```{r, eval = FALSE}
# read in field parameters
field <- readxl::read_excel(fp(p_yield, "DSSAT/mzfields.xls"))  
field <- cbind(wthsoilmr, field)  # join the soils data
field[, WTH := paste0(wthnms, "7932")]
field[, c("ID_FIELD", "XNAME") := list(fid(field$WTH), xname(field$WTH))]

# 
# gdd <- brick("external/input_devel/climate/gdd.tif")
# gddmu <- calc(gdd, mean)
# plot(gddmu)

# Create X table
xtab <- x_tab(fids = field)  # table created, with all fixed field parameters
cult <- c("OP9991", "OP9996", "HY9991", "HY9996")
CLNUM <- 1:length(cult)
pdates <- strftime(as.Date("1979-12-01"), "%y%j")
tcomb <- expand.grid(list("PDATE" = pdates, "INGENO" = cult), 
                     stringsAsFactors = FALSE)
ttab <- cbind("N" = 1:nrow(tcomb), tcomb, 
              t_tab(tvars = c("PDATE", "INGENO"), topts = c("MP", "CU"), 
                    ttab = tcomb))

# Speed this up
xtabl <- lapply(1:nrow(xtab), function(x) { # x <- 1
# xtabl <- lapply(1:10, function(x) { # x <- 1
  print(x)
  d <- xtab[x, ]
  d2 <- do.call(rbind, lapply(1:nrow(ttab), function(x) d))
  upd_col <- colnames(d2)[which(colnames(d2) %in% colnames(ttab))]
  d2[, c(upd_col) := ttab[, upd_col]]
  xt <- cbind(data.table(ttab), d2[, !colnames(d2) %in% upd_col, with = FALSE])
})

cdt <- c("Maize", ".MZX")
mxrun <- lapply(xtabl, function(x) {  # x <- xtabl[[1]]
  print(x$XNAME[1])
  xf <- copy(x)
  xf <- as.data.frame(xf)
  xfnm <- x_file(xf, outdir = fp(p_dssat, cdt[1]), z = "01", xtype = cdt[2])
  bname <- batch_file(xl = list(xf), xfiles = xfnm, 
                    outdir = full_path(p_dssat, cdt[1]), btype = cdt[1])
  exec_csm(projdir = fp(p_yield, "DSSAT"), csmdir = p_dssat, 
         rundir = full_path(p_dssat, cdt[1]), bname = bname)
  sdat <- read_CSM_outfile(rundir = full_path(p_dssat, cdt[1]), 
                           type = "summary", 
                           vars = c("RUNNO", "TRNO", "FNAM", "SOIL_ID...", 
                                    "PDAT", "MDAT", "HWAH", "PRCM"))
})
# save(xrun, file = fp(p_yield, "DSSAT/soyyields.rda"))

mz_ylds <- rbindlist(lapply(1:length(mxrun), function(i) { # i <- 1
  xbinder <- xtabl[[i]][1, list(XNAME, wthnms, x, y, prof)]
  setnames(xbinder, c("wthnms", "prof"), c("WTH", "PROF"))
  cbind(xbinder, mxrun[[i]])
}))  
# mz_ylds[, lapply(.(HWAH, PRCM), mean), by = XNAME][, plot(V2, V1)]
# mz_ylds[, lapply(.(HWAH, PRCM), mean), by = XNAME][, mean(V1)]
# 9024 mean with 1/11 start date (SYMBI no difference of course for maize)
# 8904 with 15/11 start date
# 8518 over whole country

write.table(mz_ylds, file = fp(p_yield, "DSSAT/maize_ylds.csv"), 
            row.names = FALSE, sep = ",")
# mz_ylds[, unique(XNAME)]

# check on soil configuration
tst <- unique(mz_ylds[, .(x, y, WTH, PROF)])
tst[WTH == "LUSA"]
unixy <- unique(wthprofs[, .(x, y)])
for(i in 1:nrow(unixy)) {  # i = 17
  p <- wthsoilmr[x == unixy$x[i] & y == unixy$y[i], .(wthnms, prof)]
  tstp <- tst[x == unixy$x[i] & y == unixy$y[i], .(WTH, PROF)]
  if(!all(tstp == p)) print(paste(i, "has a problem"))
}

```