### 2016-02-10 Rmd/empirical-yields.Rmd

---
title: "Stage 2 Yield Prediction"
author: "Lyndon Estes"
date: "February 5, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Empirical prediction of yield

A yield transfer function for DSSAT yields

```{r, eval = FALSE}
library(agroEcoTradeoff)
library(matrixStats)
library(doMC)
library(mgcv)

p_root <- lmisc::set_base_path()
p_clim <- fp(p_root, "external/input_devel/climate")
p_yield <- fp(p_root, "external/input_devel/yield")
p_sgrids <- fp(p_root, "external/input_devel/soil")

# Zambia provinces
p_prov <- paste0("/Users/lestes/Dropbox/data/zari/infrastructure/zambia/", 
                 "rda_dump/Final Shape files RDA/")
prov <- readOGR(fp(p_prov, "prov.shp"), layer = "prov")

# Regression variables
load(fp(p_yield, "empirical/regdat.rda"))

```

# Fit model
## Exploratory plots
```{r, eval = FALSE}
plot(prov)
regdt[CROP == "maize", points(x, y, pch = 16, cex = 0.5)]
# regdt[CROP == "maize", plot(MAP, YLDS)]
regdt[CROP == "maize" & YLDS > 9000, 
      points(x, y, pch = 16, cex = 0.5, col = "red")]
# regdt[CROP == "soybean", plot(MAP, YLDR)]
# regdt[CROP == "soybean" & YLDS > 1700, 
#       points(x, y, pch = 16, cex = 0.5, col = "blue")]

# Prune out yields > 8000 maize > 1700 soybean in subsistence case
# soyfilt <- regdt[CROP == "soybean" & YLDS > 1700, unique(.SD), 
#                 .SDcols = c("x", "y", "SOL", "WTH", "Zone")]
mzfilt <- regdt[CROP == "maize" & YLDS > 9000, unique(.SD), 
                .SDcols = c("x", "y", "SOL", "WTH")]
# regnew <- regdt[!x %in% soyfilt$x | !y %in% soyfilt$y]
regnew <- regdt[!x %in% mzfilt$x | !y %in% mzfilt$y]
regnew[CROP == "maize", points(x, y, pch = 16, cex = 0.5, col = "green")]

# regnew[CROP == "maize", hist(YLDS)]
# regnew[CROP == "maize", hist(YLDR)]

# summary(lm(YLDS ~ poly(MAP, 2), data = regnew[CROP == "maize", ]))
# summary(lm(YLDR ~ poly(MSP, 2), data = regnew[CROP == "maize", ]))
regnew[CROP == "maize", plot(MAP, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "soybean", plot(MAP, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "soybean", plot(MAP, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "soybean", plot(MSP, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "soybean", plot(MSP, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(FP, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(MSP, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(MSP, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SLOC, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SLOC, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SLCL, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SLCL, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SLSI, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SLSI, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SBDM, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SBDM, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SLHW, YLDS, pch = 16, cex = 0.5)]  # pH looks pred
regnew[CROP == "maize", plot(SLHW, YLDR, pch = 16, cex = 0.5)]  # pH looks pred
regnew[CROP == "maize", plot(GDD, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(KDD, YLDS, pch = 16, cex = 0.5)]
# regnew[CROP == "maize", plot(log(KDD), YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(NKJ, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(NKF, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(ATMX, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(STMX, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(STMX, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SSRT, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SSRM, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(ASRT, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(ASRM, YLDS, pch = 16, cex = 0.5)]

regnew[, round(cor(.SD), 2), .SDcols = names(regnew)[-c(1:13)]]

summary(lm(YLDS ~ poly(SBDM, 2) + poly(SLHW, 2), 
           data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ poly(SLHW, 2), data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ poly(SBDM, 2), data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ poly(KDD, 2), data = regnew[CROP == "maize", ]))
summary(lm(YLDR ~ poly(STMX, 2), data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ poly(STMX, 2), data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ KDD, data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ poly(GDD, 2), data = regnew[CROP == "maize", ]))

# p <- "SBDM"; d <- "YLDS"
# regnew[CROP == "maize", {
#   lm1 <- lm(get(d) ~ get(p))
#   lm2 <- lm(get(d) ~ poly(get(p), 2))
#   plot(get(p), get(d), pch = 16, cex = 0.1)
#   abline(lm1)
#   dum <- data.frame(floor(min(get(p))):ceiling(max(get(p))))
#   colnames(dum) <- p
#   lines(dum[, 1], predict(lm2, newdata = dum), col = "red")
#   summary(lm1)
#   # summary(lm2)
# }]

reg_plotter <- function(dep, pred) {
  lm1 <- lm(dep ~ pred)
  lm2 <- lm(dep ~ poly(pred, 2))
  plot(pred, dep, pch = 16, cex = 0.1)
  abline(lm1)
  prng <- c(floor(min(pred)), ceiling(max(pred)))
  dum <- data.frame(seq(prng[1], prng[2], by = diff(prng) / 100))
  # dum <- data.frame(floor(min(pred)):ceiling(max(pred)))
  colnames(dum) <- "pred"
  preds <- predict(lm2, newdata = dum)
  lines(dum[, 1], preds, col = "red")
  mtext(round(summary(lm1)$adj.r.squared, 4), side = 3, adj = 0)
  mtext(round(summary(lm2)$adj.r.squared, 4), side = 3, adj = 0.5)
}
# regnew[CROP == "maize", reg_plotter(YLDR, GDD)]  # 
regnew[CROP == "maize", reg_plotter(YLDS, MAP)]  # P 0.058
regnew[CROP == "maize", reg_plotter(YLDS, MSP)]  # P 0.050
regnew[CROP == "maize", reg_plotter(YLDS, JP)]  # P 0.0186
# regnew[CROP == "maize", reg_plotter(YLDS, JP - mean(JP))]  # P 0.012
regnew[CROP == "maize", reg_plotter(YLDS, FP)]  # P 0.009
regnew[CROP == "maize", reg_plotter(YLDS, OP)]  # P 0.007

regnew[CROP == "maize", reg_plotter(YLDS, SLHW)]  # P 0.16
regnew[CROP == "maize", reg_plotter(YLDS, SLOC)]  # P 0.43
# regnew[CROP == "maize", reg_plotter(YLDR, SLOC)]  # P 
regnew[CROP == "maize", reg_plotter(YLDS, SLCL)]  # P 0.191
regnew[CROP == "maize", reg_plotter(YLDS, SBDM)]  # 0.20
# mtax pairings
regnew[CROP == "maize", reg_plotter(YLDS, STMX)]  # P 0.07
regnew[CROP == "maize", reg_plotter(YLDS, ATMX)]  # P 0.044 
# NK
regnew[CROP == "maize", reg_plotter(YLDS, NKJ)]  # L 0.056
regnew[CROP == "maize", reg_plotter(YLDS, NKF)]  # P 0.024 
# GDD/KDD
regnew[CROP == "maize", reg_plotter(YLDS, GDD)]  # P 0.03
regnew[CROP == "maize", reg_plotter(YLDS, KDD)]  # L 0.026
# srad
regnew[CROP == "maize", reg_plotter(YLDS, SSRT)]  # P 0.015 
regnew[CROP == "maize", reg_plotter(YLDS, SSRM)]  # 0.0155
regnew[CROP == "maize", reg_plotter(YLDS, ASRT)]  # P 0.038 
regnew[CROP == "maize", reg_plotter(YLDS, ASRM)]  # P 0.039

# best predictors look to be
# MAP, SLOC (but polynomial relationship odd), SLHW, SLCL, SBDM, all good
# STMX (linear KDD as a possible alternate), GDD, ASRM

#
regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(SLHW, 2) + poly(SLCL, 2) +
              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.2941

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(SLHW, 2) + poly(SLCL, 2) +
              poly(SBDM, 2) + poly(GDD, 2) + poly(STMX, 2) + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.2926

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(SLHW, 2) + poly(SLOC, 2) +
              poly(SBDM, 2) + poly(GDD, 2) + poly(STMX, 2) + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4244 # SLOC doesn't give as much value, NS on linear

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(SLHW, 2) + poly(SLCL, 2) +
              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.2961  # Doesn't add much

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + JP + poly(SLHW, 2) + poly(SLCL, 2) +
              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.2972  # Jan precip adds more R2, but makes MAP linear NS

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + FP + poly(SLHW, 2) + poly(SLCL, 2) +
              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.303  # Feb precip even more R2, but makes MAP linear NS

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(SLHW, 2) + poly(SLCL, 2) +
              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + x + y)
  summary(lm1)
  plot(YLDS, predict(lm1))
}]  # 0.4301  # Oct precip add a tiny bit, significant

# regnew[CROP == "maize", {
#   gam1 <- bam(YLDS ~ poly(MAP, 2) + OP + poly(SLHW, 2) + poly(SLCL, 2) +
#               poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + s(x, y), 
#              family = gaussian(link = "log"), method = "ML")
#   summary(gam1)
#   # plot(YLDS, predict(lm1))
# }]  # 0.4301  # Oct precip add a tiny bit, significant

```

## Reduce variables and fit models

Reduce down to means across years, and select 10% and 90% percentile years

### Maize
```{r, eval = FALSE}
regred <- copy(regnew[CROP == "maize"])
regred[, mid2 := 1:.N]

# regred[, c("Y10S", "Y90S") := as.integer(0)]
regmu <- regred[CROP == "maize", lapply(.SD, mean), 
                by = list(CROP, x, y, WTH, SOL), 
                .SDcols = names(regred)[c(9:ncol(regred))]]
regmu[, YST := "mu"]
# regnew[CROP == "maize" & Zone == "Zone2a" & WTH == "AABS" & 
#          SOL == "WI_CMZR003", mean(YLDS)]
qs <- function(x, y, ptile) {
  DF <- cbind(1:length(x), x, y)
  DF <- DF[order(DF[, 2]), ]
  DF[round(nrow(DF) * ptile), 3]
}

y <- 30:60
DF <- cbind(1:31, sample(2000:3000, size = 31), y)
DF <- DF[order(DF[, 2]), ]
DF[round(nrow(DF) * 0.9), 3]
sort(DF[, 2])

a <- regnew[CROP == "maize" & WTH == "AEQA" & SOL == "WI_RGTZ054", YLDS]
b <- regnew[CROP == "maize" & WTH == "AEQA" & SOL == "WI_RGTZ054", mid]
qs(a, b, 0.9)
sort(a)

# Select out quantile values
m10 <- regred[,
  qs(YLDS, mid2, 0.1), by = list(CROP, x, y, WTH, SOL)
][, V1]
m90 <- regred[,
  qs(YLDS, mid2, 0.9), by = list(CROP, x, y, WTH, SOL)
][, V1]

# 
y10 <- regred[mid2 %in% m10]
y90 <- regred[mid2 %in% m90]
y10[, YST := "10"]
y90[, YST := "90"]

y1090 <- rbind(y10, y90)
cnames <- colnames(y1090)[colnames(y1090) %in% colnames(regmu)]
setcolorder(regmu, cnames)
regred2 <- rbind(regmu, y1090[, cnames, with = FALSE])

par(mfrow = c(2, 2))
for(i in c("10", "mu", "90")) {
  regred2[CROP == "maize" & YST == i, hist(YLDS)]
}
regred2[CROP == "maize", mean(YLDS), by = YST]

regred2[CROP == "maize", reg_plotter(YLDS, MAP)]  # P 0.1235
regred2[CROP == "maize", reg_plotter(YLDS, MSP)]  # P 0.1097
regred2[CROP == "maize", reg_plotter(YLDS, JP)]  # P 0.036
regred2[CROP == "maize", reg_plotter(YLDS, FP)]  # P 0.026
regred2[CROP == "maize",  summary(lm(YLDS ~ poly(FP, 2)))]
regred2[CROP == "maize", reg_plotter(YLDS, 100 - SLCL + SLSI)]  # P 0.1243
regred2[CROP == "maize", reg_plotter(YLDS, SLCL)]  # P 0.19
regred2[CROP == "maize", reg_plotter(YLDS, SLSI)]  # P 0.049

regred2[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(SLHW, 2) + poly(SLCL, 2) +
              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + x + y)
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.3324  # Oct precip add a tiny bit, significant

regred2[CROP == "maize" & YST == "mu", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(SLHW, 2) + poly(SLCL, 2) +
              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + x + y)
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.3271  # Oct precip add a tiny bit, significant

regred2[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(SLHW, 2) + poly(SLCL, 2) +
              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.34  # Oct precip add a tiny bit, significant

regred2[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(SLHW, 2) + poly(SLOC, 2) +
              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.466  # Oct precip add a tiny bit, significant

# Fit gams
gam1 <- gam(YLDS ~ s(MAP, k = 3) + s(JP, k = 3) + s(SLHW, k = 3) + 
              s(SLCL, k = 3) + s(SBDM, k = 3) + s(GDD, k = 3) + KDD + 
              s(ASRM, k = 3), family = gaussian(link = "log"), method = "ML", 
            data = regred2[CROP == "maize"])
gam2 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(SLHW, 2) + 
              poly(SLCL, 2) + poly(SBDM, 2) + poly(GDD, 2) + KDD + 
              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam3 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(SLHW, 2) + 
              poly(SLCL, 2) + poly(SBDM, 2) + poly(GDD, 2) + KDD + 
              poly(ASRM, 2) + s(x, y), 
            family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam4 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(SLHW, 2) + 
              poly(SLCL, 2) + poly(SBDM, 2) + poly(GDD, 2) + KDD + 
              s(x, y), family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam5 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(SLHW, 2) + 
              poly(SLCL, 2) + poly(SBDM, 2) + poly(GDD, 2) + KDD, 
            family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam6 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(SLHW, 2) + 
              poly(SLCL, 2) + poly(GDD, 2) + KDD, 
            family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam7 <- gam(YLDS ~ poly(MAP, 2) + poly(SLHW, 2) + 
              poly(SLCL, 2) + poly(GDD, 2) + KDD, 
            family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam8 <- gam(YLDS ~ poly(MAP, 2) + SLHW +  poly(SLCL, 2) + poly(GDD, 2) + KDD, 
            family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam9 <- gam(YLDS ~ poly(MAP, 2) + SLHW + poly(SLCL, 2) + 
              poly(100 - (SLCL + SLSI), 2) + poly(GDD, 2) + KDD, 
            family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam10 <- gam(YLDS ~ poly(MAP, 2) + SLHW + poly(SLCL, 2) + 
               poly(SLOC, 2) + poly(GDD, 2) + KDD, 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
gam11 <- gam(YLDS ~ poly(MAP, 2) + SLHW + poly(SLCL, 2) + 
               SLOC + poly(GDD, 2) + KDD, 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
gam12 <- gam(YLDS ~ poly(MAP, 2) + SLHW + poly(SLCL, 2) + 
               SLOC + poly(GDD, 2) + KDD + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
gam13 <- gam(YLDS ~ poly(MAP, 2) + poly(SLCL, 2) + 
               poly(SLOC, 2) + poly(GDD, 2) + KDD, 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
gam14 <- gam(YLDS ~ poly(MAP, 2) + poly(SLCL, 2) + 
               poly(SLOC, 2) + poly(GDD, 2) + KDD + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
gam15 <- gam(YLDS ~ poly(MSP, 2) + poly(SLCL, 2) + poly(SLOC, 2) + 
               poly(GDD, 2) + KDD + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
gam16 <- gam(YLDS ~ poly(JFMP, 2) + poly(SLCL, 2) + poly(SLOC, 2) + 
               poly(GDD, 2) + KDD + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
gam17 <- gam(YLDS ~ poly(MSP, 2) + poly(SLCL, 2) + SLOC + 
               poly(GDD, 2) + KDD + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
gam18 <- gam(YLDS ~ poly(MSP, 2) + poly(SLCL, 2) + poly(SLHW, 2) + SLOC + 
               poly(GDD, 2) + s(x, y), family = gaussian(link = "log"), 
             method = "REML", data = regred2[CROP == "maize"])
gam19 <- gam(YLDS ~ poly(MSP, 2) + poly(SLCL, 2) + SLHW + SLOC + 
               poly(GDD, 2) + KDD + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
gam20 <- gam(YLDS ~ poly(MAP, 2) + poly(SLCL, 2) + 
               poly(SLOC, 2) + poly(GDD, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
gam21 <- gam(YLDS ~ poly(MSP, 2) + poly(SLCL, 2) + poly(SLOC, 2) + 
               poly(GDD, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
gam22 <- gam(YLDS ~ poly(MAP, 2) + poly(SLCL, 2) + #(SLCL, 2) + 
               s(SLOC, k = 3) + poly(GDD, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
gam23 <- gam(YLDS ~ poly(MAP, 2) + poly(SLHW, 2) + 
               s(SLOC, k = 3) + poly(GDD, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
gam24 <- gam(YLDS ~ poly(MAP, 2) + poly(SLHW, 2) + s(SLCL, k = 3) + 
               s(SLOC, k = 3) + poly(GDD, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])

summary(gam1)
plot.gam(gam1, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam2)
plot.gam(gam2, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam3)
plot.gam(gam3, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam4)
plot.gam(gam4, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam5)
plot.gam(gam5, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam6)
plot.gam(gam6, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam7)
plot.gam(gam7, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam8)
plot.gam(gam8, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam9)
plot.gam(gam9, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam10)
plot.gam(gam10, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam11)  
plot.gam(gam11, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam12)  
plot.gam(gam12, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam13)  # this one, without smooth x, y terms
plot.gam(gam13, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam14)  # of this one, with smooth x, y terms
plot.gam(gam14, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam15)  # this one, with MSP, is best, with smooth x, y terms
plot.gam(gam15, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam16)  # JFMP doesn't add as much
plot.gam(gam16, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam17)  # Turn SLOC into linear predictor, which makes more sense
plot.gam(gam17, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam18)  # add polynomial on SLHW, drop KDD. More R2
plot.gam(gam18, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam19)  # SLHW linear, keep KDD as NS linear 
plot.gam(gam19, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam20)  # SLHW linear, keep KDD as NS linear 
plot.gam(gam20, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam21)  # SLHW linear, keep KDD as NS linear 
plot.gam(gam21, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam22)  ### SLOC smooth, SLCL poly
plot.gam(gam22, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam23)  ### SLOC smooth and SLHW poly
plot.gam(gam23, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam24)
plot.gam(gam24, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)

plot(predst[["SLCL"]])

par(mfrow = c(2, 2))
regred2[CROP == "maize", plot(YLDS, predict(gam15, type = "response"))]
lines(c(2000, 8000), c(2000, 8000), col = "red")
regred2[CROP == "maize", plot(YLDS, predict(gam17, type = "response"))]
lines(c(2000, 8000), c(2000, 8000), col = "red")
regred2[CROP == "maize", plot(YLDS, predict(gam18, type = "response"))]
lines(c(2000, 8000), c(2000, 8000), col = "red")
regred2[CROP == "maize", plot(YLDS, predict(gam19, type = "response"))]
lines(c(2000, 8000), c(2000, 8000), col = "red")

# Reduce down to model for mean, 10th, 90th)
gam18mu <- gam(YLDS ~ poly(MSP, 2) + poly(SLCL, 2) + poly(SLHW, 2) + SLOC + 
               poly(GDD, 2) + s(x, y), family = gaussian(link = "log"), 
               method = "REML", data = regred2[YST == "mu"])
summary(gam18mu)  # Rainfall drops out, but still retains shape
plot.gam(gam18mu, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
gam19mu <- gam(YLDS ~ poly(MSP, 2) + poly(SLCL, 2) + SLHW + SLOC + 
               poly(GDD, 2) + KDD + s(x, y), family = gaussian(link = "log"), 
               method = "REML", data = regred2[YST == "mu"])
summary(gam19mu)  # Rainfall drops out, but still retains shape
plot.gam(gam19mu, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)

# gam1510 <- gam(YLDS ~ poly(MSP, 2) + poly(SLCL, 2) + poly(SLOC, 2) + 
#                poly(GDD, 2) + s(x, y),
#              family = gaussian(link = "log"), method = "REML", 
#              data = regred2[YST == "10"])
# summary(gam1510)
# plot.gam(gam1510, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
# gam1590 <- gam(YLDS ~ poly(MSP, 2) + poly(SLCL, 2) + poly(SLOC, 2) + 
#                poly(GDD, 2) + s(x, y),
#              family = gaussian(link = "log"), method = "REML", 
#              data = regred2[YST == "90"])
# summary(gam1590)
# plot.gam(gam1590, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
plot(regred2[YST == "mu", YLDS], predict(gam18mu, type = "response"))
plot(regred2[YST == "mu", YLDS], predict(gam19mu, type = "response"))
# plot(regred2[YST == "10", YLDS], predict(gam1510, type = "response"))
# plot(regred2[YST == "90", YLDS], predict(gam1590, type = "response"))

```

### Soybean

```{r, eval = FALSE}
regreds <- copy(regnew[CROP == "soybean"])
regreds[, mid2 := 1:.N]

regmus <- regreds[, lapply(.SD, mean), by = list(CROP, x, y, WTH, SOL), 
                  .SDcols = names(regred)[c(9:ncol(regred))]]
regmus[, YST := "mu"]

# Select out quantile values
m10 <- regreds[,
  qs(YLDR, mid2, 0.1), by = list(CROP, x, y, WTH, SOL)
][, V1]
m90 <- regreds[,
  qs(YLDR, mid2, 0.9), by = list(CROP, x, y, WTH, SOL)
][, V1]

# Select out 10th and 90th percentile years
y10s <- regreds[mid2 %in% m10]
y90s <- regreds[mid2 %in% m90]
y10s[, YST := "10"]
y90s[, YST := "90"]

y1090s <- rbind(y10s, y90s)
cnames <- colnames(y1090s)[colnames(y1090s) %in% colnames(regmus)]
setcolorder(regmus, cnames)
regreds2 <- rbind(regmus, y1090s[, cnames, with = FALSE])
regreds2[, slcf := 100 - (SLCL + SLSI)]
regreds2[, ONDP := OP + (MSP - (JP + FP))]
regreds2[, NDP := (MSP - (JP + FP))]

par(mfrow = c(2, 2))
for(i in c("10", "mu", "90")) {
  regreds2[CROP == "soybean" & YST == i, hist(YLDR)]
}
regreds2[CROP == "soybean", mean(YLDR), by = YST]

regreds2[, reg_plotter(YLDR, MAP)]  # L 0.19
regreds2[, reg_plotter(YLDR, MSP)]  # L 0.15
regreds2[, reg_plotter(YLDR, JFMP)]  # L 0.23
regreds2[, reg_plotter(YLDR, JP)]  # L 0.04
regreds2[, reg_plotter(YLDR, FP)]  # P 0.11
regreds2[, reg_plotter(YLDR, NDP)]  # 
regreds2[, reg_plotter(YLDR, ONDP)] 
regreds2[, reg_plotter(YLDR, OP)]

regreds2[, reg_plotter(YLDR, 100 - SLCL + SLSI)]  # P 0.16
regreds2[, reg_plotter(YLDR, SLCL)]  # P 0.15
regreds2[, reg_plotter(YLDR, SLSI)]  # P 0.01
regreds2[, reg_plotter(YLDR, SLOC)]  # L 0.16 (decline)
regreds2[, reg_plotter(YLDR, SLHW)]  # P 0.18
regreds2[, reg_plotter(YLDR, GDD)]  # NS
regreds2[, reg_plotter(YLDR, KDD)]  # NS
regreds2[, reg_plotter(YLDR, STMX)]  # L 0.04
regreds2[, reg_plotter(YLDR, ATMX)]  # L 0.04
regreds2[, reg_plotter(YLDR, NKF)]  # L 0.06
regreds2[, reg_plotter(YLDR, NKJ)]  # P 0.06
regreds2[, reg_plotter(YLDR, SSRT)]  # P 0.11
regreds2[, reg_plotter(YLDR, SSRM)]  # P 0.11
regreds2[, reg_plotter(YLDR, ASRM)]  # L 0.13

sgam1 <- gam(YLDR ~ s(MAP, k = 3) + s(JP, k = 3) + s(SLHW, k = 3) + 
               s(SLCL, k = 3) + s(SBDM, k = 3) + s(GDD, k = 3) + KDD + 
               s(ASRM, k = 3), family = gaussian(link = "log"), method = "ML", 
            data = regreds2)
sgam2 <- gam(YLDR ~ poly(MAP, 2) + poly(JP, 2) + poly(SLHW, 2) + 
              poly(SLCL, 2) + poly(SBDM, 2) + poly(GDD, 2) + KDD + 
              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam2a <- gam(YLDR ~ poly(JFMP, 2) + poly(SLHW, 2) + 
              poly(SLCL, 2) + poly(SBDM, 2) + poly(GDD, 2) + KDD + 
              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
summary(sgam2); summary(sgam2a)  # MAP better
sgam3 <- gam(YLDR ~ poly(MAP, 2) + poly(FP, 2) + poly(SLHW, 2) + 
              poly(SLCL, 2) + poly(SBDM, 2) + poly(GDD, 2) + KDD + 
              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam4 <- gam(YLDR ~ poly(MAP, 2) + poly(OP, 2) + poly(SLHW, 2) + 
              poly(SLCL, 2) + poly(SBDM, 2) + poly(GDD, 2) + KDD + 
              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam5 <- gam(YLDR ~ poly(MAP, 2) + poly(OP, 2) + poly(SLHW, 2) + 
              poly(SLCL, 2) + poly(SLOC, 2) + poly(GDD, 2) + KDD + 
              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam6 <- gam(YLDR ~ poly(MAP, 2) + poly(OP, 2) + poly(SLHW, 2) + 
              poly(SLCL, 2) + SLOC + poly(GDD, 2) + KDD + 
              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam7 <- gam(YLDR ~ poly(MAP, 2) + poly(OP, 2) + poly(SLHW, 2) + 
              poly(SLCL, 2) + poly(GDD, 2) + KDD + 
              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam8 <- gam(YLDR ~ poly(MAP, 2) + poly(SLHW, 2) + poly(SLCL, 2) + 
               poly(GDD, 2) + KDD + poly(ASRM, 2), 
             family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam9 <- gam(YLDR ~ poly(MAP, 2) + poly(SLHW, 2) + poly(SLCL, 2) + 
               poly(GDD, 2) + poly(ASRM, 2),
             family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam9 <- gam(YLDR ~ poly(MAP, 2) + poly(SLHW, 2) + poly(SLCL, 2) + 
               poly(, 2) + poly(ASRM, 2),
             family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam10 <- gam(YLDR ~ poly(MAP, 2) + poly(SLHW, 2) + poly(SLCL, 2) + 
               poly(GDD, 2) + poly(ASRM, 2) + s(x, y),
             family = gaussian(link = "log"), method = "REML", data = regreds2)
summary(sgam10)
sgam11 <- gam(YLDR ~ poly(MSP, 2) + poly(SLHW, 2) + poly(SLCL, 2) + 
               poly(GDD, 2) + poly(ASRM, 2) + s(x, y),
             family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam12 <- gam(YLDR ~ poly(JFMP, 2) + poly(SLHW, 2) + poly(SLCL, 2) + 
               poly(GDD, 2) + poly(ASRM, 2) + s(x, y),
             family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam13 <- gam(YLDR ~ poly(JFMP, 2) + poly(SLCL, 2) + poly(SLOC, 2) + 
                poly(GDD, 2) + poly(ASRM, 2) + s(x, y),
             family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam14 <- gam(YLDR ~ poly(MAP, 2) + poly(SLCL, 2) + poly(SLOC, 2) + 
                poly(GDD, 2) + poly(ASRM, 2) + s(x, y),
             family = gaussian(link = "log"), method = "REML", data = regreds2)
summary(sgam14)

summary(sgam1)
plot.gam(sgam1, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam2)
plot.gam(sgam2, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam2a)
plot.gam(sgam2a, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam3)
plot.gam(sgam3, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam4)
plot.gam(sgam4, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam5)
plot.gam(sgam5, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam6)
plot.gam(sgam6, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam7)
plot.gam(sgam7, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam8)
plot.gam(sgam8, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam9)  
plot.gam(sgam9, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam10)  
plot.gam(sgam10, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam11)  
plot.gam(sgam11, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam12)  
plot.gam(sgam12, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam13)  
plot.gam(sgam13, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam14)  
plot.gam(sgam14, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)

regreds2[, plot(MAP, YLDR)]
regreds2[, plot(JFMP, YLDR)]

plot(regreds2[, YLDR], predict(sgam12, type = "response"))
lines(c(0, 5000), c(0, 5000), col = "red")

# Now see about taking same model and fitting separately to 10th/90th/mean years
sgam12mu <- gam(YLDR ~ poly(MSP, 2) + poly(SLHW, 2) + poly(SLCL, 2) + 
               poly(GDD, 2) + poly(ASRM, 2) + s(x, y),
                family = gaussian(link = "log"), method = "REML", 
                data = regreds2[YST == "mu"])
summary(sgam12mu)  # OP drops out, SSRM drops out
# plot.gam(sgam12mu, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
# sgam1810 <- gam(YLDR ~ poly(MSP, 2) + poly(SLHW, 2) + poly(SLCL, 2) + 
#                   SLOC + poly(GDD, 2) + s(x, y),
#                 family = gaussian(link = "log"), method = "REML", 
#                 data = regreds2[YST == "10"])
# summary(sgam1810)  # OP drops out, SSRM drops out
# plot.gam(sgam1810, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
# sgam1890 <- gam(YLDR ~ poly(MSP, 2) + poly(SLHW, 2) + poly(SLCL, 2) + 
#                   SLOC + poly(GDD, 2) + s(x, y),
#                 family = gaussian(link = "log"), method = "REML", 
#                 data = regreds2[YST == "90"])
# summary(sgam1890)  # OP drops out, SSRM drops out
# plot.gam(sgam1890, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)

par(mfrow = c(2, 2))
plot(regreds2[, YLDR], predict(sgam12, type = "response"))
lines(c(0, 5000), c(0, 5000), col = "red")
plot(regreds2[YST == "mu", YLDR], predict(sgam12mu, type = "response"))
lines(c(0, 5000), c(0, 5000), col = "red")
# plot(regreds2[YST == "10", YLDR], predict(sgam1810, type = "response"))
# plot(regreds2[YST == "90", YLDR], predict(sgam1890, type = "response"))

```

## Load in grid predictors
```{r, eval = FALSE}
soilgrids <- brick(fp(p_sgrids, "soilgrids.tif"))
climvars <- c("rftot", "ndjftot", "jfmtot", "gdd", "kdd", "srmu_ano")
climgrids <- sapply(climvars, function(x) {  # x <- climvars[3]
  print(x)
  brick(dir(p_clim, full.names = TRUE)[grep(x, dir(p_clim))])
})
climgrids_mu <- sapply(climgrids, function(x) calc(x, mean))
climst <- stack(climgrids_mu)
names(climst) <- c("MAP", "MSP", "JFMP", "GDD", "KDD", "ASRM")

# disaggregate down to 1 km
climstbase <- disaggregate(climst[[1]], fact = 30)
climststep <- disaggregate(climst, fact = 20)
climst1km <- resample(climst, climstbase)

# need to warp soilgrids back to that also
ext <- bbox(climst1km)[1:4]  # extent   
onm <- fp(p_sgrids, "soilgridsgcs.tif")
gdalwarp(srcfile = soilgrids@file@name, t_srs = projection(climst1km), 
         dstfile = onm, r = "average", ot = "Float32", te = ext, 
         srcnodata = -32768, dstnodata = -99, tr = res(climst1km), 
         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
soilgcs <- brick(onm)
names(soilgcs) <- c("SLOC", "SLCL", "SLHW")
plot(soilgcs)

# need also a grid of x,y values
x <- soilgcs[[1]]
x[] <- xFromCell(x, 1:ncell(x))
y <- soilgcs[[1]]
y[] <- yFromCell(y, 1:ncell(y))
plot(y)
xy <- stack(x, y)
names(xy) <- c("x", "y")
predst <- stack(xy, climst1km, soilgcs)

```

## Predicted yields surfaces
```{r, eval = FALSE}
mzfull1 <- raster::predict(predst, gam18, type = "response")
mzfull2 <- raster::predict(predst, gam19, type = "response")
mzfull3 <- raster::predict(predst, gam22, type = "response")
mzfull4 <- raster::predict(predst, gam23, type = "response")
soyfull <- raster::predict(predst, sgam10, type = "response")
soyfull2 <- raster::predict(predst, sgam13, type = "response")
soyfull3 <- raster::predict(predst, sgam14, type = "response")
summary(sgam10)
predict(gam15)
# mzfull1[mzfull1 > 8000] <- 8000
# mzfull2[mzfull2 > 8000] <- 8000
# soyfull[soyfull > 5000] <- 5000
par(mfrow = c(2, 2))
plot(mzfull1, zlim = c(0, 9000))
plot(mzfull2, zlim = c(0, 9000))
plot(mzfull3, zlim = c(0, 9000))
plot(mzfull4, zlim = c(0, 9000))

prov@proj4string <- soyfull@crs
soymsk <- crop(soyfull, prov)
soymsk2 <- crop(soyfull2, prov)
soymsk3 <- crop(soyfull3, prov)
plot(soymsk > 5000)
plot(prov, add = TRUE)
plot(soymsk2)
plot(prov, add = TRUE)
plot(soymsk3 > 5000)
plot(prov, add = TRUE)

# Save bricks and models
save(gam18, gam19, gam22, gam23, sgam10, sgam13, sgam14, 
     file = fp(p_yield, "empirical/gams1.rda"))
yldrs <- brick(stack(list(mzfull1, mzfull2, mzfull3, mzfull4, soyfull,
                          soyfull2, soyfull3)))
writeRaster(yldrs, filename = fp(p_yield, "empirical/yield_surfaces1.tif"))

```