### 2016-02-10 Rmd/yield-predictors.rmd

---
title: "Yield Mapping"
author: "Lyndon Estes"
date: "January 31, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Predict DSSAT yields onto finer soil map

```{r, eval = FALSE}
library(agroEcoTradeoff)
library(gdalUtils)
library(rCropMod)
library(matrixStats)
library(doMC)

p_root <- lmisc::set_base_path()
p_crop <- "~/Dropbox/data/agricultural/zambia/dssat/"
p_cropmod <- "~/Dropbox/projects/rCropMod/external/ext_data/"
p_soil <- "~/DSSAT45/SOIL/"
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_clim <- fp(p_root, "external/input_devel/climate")
p_climo <- "~/Dropbox/data/climate/rainfall/zambia/precip/"
p_yield <- fp(p_root, "external/input_devel/yield")
p_pgf <- "~/Dropbox/data/climate/afdm/pgf_bc/zambia/"
p_sgrids <- "~/Dropbox/data/soil/"

# Read in Zambia grid
mgrid <- raster(full_path(p_basedat, "ZA-grid.tif"))

# Zambia provinces
p_prov <- paste0("/Users/lestes/Dropbox/data/zari/infrastructure/zambia/", 
                 "rda_dump/Final Shape files RDA/")
prov <- readOGR(fp(p_prov, "prov.shp"), layer = "prov")

# Read in original soil-climate data from rCropMod
soilgrid <- raster(fp(p_cropmod, "af_soil_grid.tif"))  # soil grid
soilshare <- fread(fp(p_cropmod, "af_soil_share.csv"))  # share of profiles

# Read in soil-climate grid from Di's work
ID <- fread(full_path(p_crop, "Soil_clim_match_ID_master.csv"))
ID[, V1 := NULL]
setkeyv(ID, c("WTH", "SOL", "Zone"))
ID[, SharePct := as.numeric(SharePct)]
zone <- sort(unique(ID$Zone))
for(i in 1:length(zone)) ID[Zone == zone[i], zoneid := i]
setkeyv(ID, c("WTH", "Zone"))
# ID[, FID := 1:.N]
IDr <- subset(unique(ID[, .(x, y, ID, WTH, Zone, zoneid)]))
setkeyv(IDr, c("WTH", "Zone"))
# ID[SOL == "WI_ARBW438"]
gcs <- "+proj=longlat +datum=WGS84"
idgr <- dt_to_raster(IDr[, .(x, y, ID, zoneid)], CRS(gcs))
# plot(idgr)

# Yield data selection points
load(fp(p_yield, "empirical/stations.rda"))
zmetsamps <- extract(idgr, zmetf, cellnumbers = TRUE)
zmetxy <- data.frame(zmetf)[, c("x", "y")]
# xFromCell(idgr, zmetsamps[, "cells"]) - zmetxy$x
# yFromCell(idgr, zmetsamps[, "cells"]) - zmetxy$y
# zmetsamps <- cbind(xyFromCell(idgr, zmetsamps[, "cells"]), zmetsamps)
zmetsamps <- cbind(zmetxy, zmetsamps)  # harmonize with DSSAT zmetfxy

# zmet_dt <- data.table(data.frame(zmetsamps))[, .(x, y)]
zmet_dt <- data.table(data.frame(zmetsamps))
# setkeyv(zmet_dt, c("x", "y"))
setkeyv(zmet_dt, c("ID", "zoneid"))

# setkeyv(IDr, c("x", "y"))
setkeyv(IDr, c("ID", "zoneid"))
IDr_ss <- merge(IDr, zmet_dt)
# IDr_ss <- IDr[zmet_dt]  # pick up IDs, soils, wth variables
IDr_ss[, c("x.x", "y.x", "cells") := NULL]
setnames(IDr_ss, c("x.y", "y.y"), c("x", "y"))
setkeyv(IDr_ss, c("WTH", "Zone"))

# Collect soils up here also 
pts <- IDr_ss
# pts <- cbind.data.frame("x" = 27.4738917, "y" = -16.2783358)  # in Zambia
coordinates(pts) <- ~x+y
soilid <- extract(soilgrid, pts)  # grid id corresponding to point
IDr_ss[, soilid := soilid]
profs <- soilshare[CELL5M %in% soilid]  # profiles at that point
profs[, SharePct := NULL]
IDr_ss <- merge(IDr_ss, profs, by.x = "soilid", by.y = "CELL5M")
IDr_ss[, soilid := NULL]
setnames(IDr_ss, "SoilProfile", "SOL")
setkeyv(IDr_ss, c("Zone", "WTH", "SOL"))

# IDr_ss[, plot(x, y)]
# points(zmet_dt, pch = 16, cex = 0.5)
# tst <- extract(idgr, zmet_dt)
# all(tst[order(tst[, 1]), 2] == zmetsamps[order(zmetsamps[, 4]), 5])

# alternative for yield data selection points:
# Creatre stratified, representative random sample of 50 pts from ID/zone grid
# ztab <- freq(idgr$zoneid)[-5, ]
# ztabs <- round((ztab[, 2] / sum(ztab[, 2])) * 50)
# set.seed(34)
# gsamps <- do.call(rbind, lapply(1:4, function(i) {
#   gsamp <- sampleRandom(idgr$zoneid == i, size = ztabs[i], cells = TRUE)
#   sampxy <- xyFromCell(idgr, gsamp[, "cell"])
# }))
# plot(idgr$ID)  # unique(values(idgr$ID))
# length(which(!is.na(values(idgr$ID))))
# length(which(!is.na(unique(values(idgr$ID)))))
# points(gsamps, pch = 16, cex = 0.5)
# 
# # merge those points with the main ID grid
# gsamps_dt <- data.table(gsamps)
# setkeyv(gsamps_dt, c("x", "y"))
# setkeyv(IDr, c("x", "y"))
# IDr_ss <- IDr[gsamps_dt]  # pick up IDs, soils, wth variables
# setkeyv(IDr_ss, c("WTH", "SOL", "Zone"))
# IDr_ss[, plot(x, y)]
# points(gsamps, pch = 16, cex = 0.5)

# sgrid <- fread(fp(p_cropmod, "Soil_clim_match_ID.csv"), 
#                stringsAsFactors = FALSE)
# setnames(sgrid, old = c("V1", "X", "Y", "WeatherGrid", "SoilType"), 
#          new = c("ID", "x", "y", "WTH", "SOL"))
# wgs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
# sr <- dt_to_raster(sgrid[, list(ID, x, y)], CRSobj = wgs)
# setkeyv(sgrid, c("x", "y"))
# sgrid[SOL == "WI_ARBW438"]
```

# Full DSSAT yield grids
```{r, eval = FALSE}
# crops <- dir(p_crop, pattern = "maize|soybean", full.names = FALSE)
crops <- dir(p_crop, pattern = "maize", full.names = FALSE)  # only maize now
fylds <- lapply(crops, function(X) {  # X <- "maize"
  fnms <- dir(full_path(p_crop, X), pattern = X)
  fl <- unlist(lapply(c("S", "R"), function(x) {
    fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
  }))  # reorder to read subsistence first
  # idt <- copy(IDr)
  # key(idt)
  idt <- copy(ID)
  idt <- subset(unique(idt[, .(x, y, WTH, Zone, SharePct)]))
  setkeyv(idt, c("WTH", "Zone"))
  for(y in 1:length(fl)) { # y <- 1
    # setcolorder(idt, c(4, 7:8, 5:6, 1:3))
    print(paste("Processing", fl[y]))
    cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
    cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
    setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
    cr[, WTH := substr(WTH, 1, 4)]
    cr[, c("V1", "RUNNO", "TRNO", "TNAM") := NULL]
    setkeyv(cr, c("WTH", "Zone"))

    crm <- merge(cr, idt, by = c("WTH", "Zone"))

    # nms <- names(crm)[c(6:7, 8, 1:2, 4, 9, 5, 3)]

    # Soil averaged yields
    solmu <- crm[, round(weighted.mean(HWAH, w = SharePct)), 
                 by = .(WTH, Zone, YR)]
    snm <- gsub("*.*_|\\.csv", "", fl[y])

    # Yield means
    v <- solmu[, round(mean(V1)), by = .(WTH, Zone)]
    setnames(v, "V1", paste0("Y", snm))
    idt <- idt[v]
    setkeyv(idt, c("WTH", "Zone"))
    idt
  }
  idt
})
names(fylds) <- crops

rdt <- fylds$maize[, c(1:2, grep("YS", names(fylds$maize))), with = FALSE]
mzs <- dt_to_raster(rdt, raster()@crs)
rdt <- fylds$maize[, c(1:2, grep("YR", names(fylds$maize))), with = FALSE]
mzr <- dt_to_raster(rdt, raster()@crs)

# par(mar = rep(0, 4))
plot(calc(mzs, max))
plot(calc(mzr, mean))
# plot(calc(soyr, mean))
# plot(zmetf, pch = 16, cex = 0.5, add = TRUE)

```

# Extract subset of DSSAT yields, soils, climate data
## Yield
```{r, eval=FALSE}
# merge function for lists of data.tables
mymerge <- function(x, y) merge(x, y, all = TRUE)

crops <- dir(p_crop, pattern = "maize", full.names = FALSE)  # only maize now
# Yield data for those points
mannms <- c("S", "R")
ylds <- lapply(crops, function(X) {
  # X <- crops[1]
  fnms <- dir(full_path(p_crop, X), pattern = X)
  fl <- unlist(lapply(mannms, function(x) {
    fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
  }))  # reorder to read subsistence first
  # fl <- fl[1:3]
  crml <- lapply(1:length(fl), function(y) { # y <- 1
    idt <- copy(IDr_ss)
    print(paste("Processing", fl[y]))
    cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
    cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
    setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
    cr[, WTH := substr(WTH, 1, 4)]
    cr[, c("V1", "RUNNO", "TRNO", "TNAM") := NULL]
    setkeyv(cr, c("Zone", "WTH", "SOL"))

    # merge IDs with subset location
    # crm <- cr[idt]
    crm <- merge(cr, idt)
    # crm[, SharePct := NULL]
    nms <- names(crm)[c(7:8, 6, 1:2, 9, 3, 5, 4)]
    setcolorder(crm, nms)
    mid <- 1:nrow(crm)
    crm <- cbind(mid, crm)
    keys <- c("mid", "x", "y", "ID", "WTH", "Zone", "zoneid", "SOL", "YR")
    setkeyv(crm, keys)
    # setcolorder(crm, c(nms[grep("HWAH", nms, invert = TRUE)], "HWAH"))
    newnm <- gsub(".*._||\\.csv", "", fl[y])
    setnames(crm, "HWAH", newnm)
    crm
  })
  yldmerge <- Reduce(mymerge, crml)
  ynms <- names(yldmerge)
  # yldred <- data.table(round(do.call(cbind, lapply(mannms, function(x) {
    # yldmerge[, rowMeans(.SD), .SDcols = grep(paste0(x, "[0-9]"), ynms)]
  # }))))
  # calculate 
  for(i in mannms) {
    onms <- list(paste0("YLD", i), grep(paste0(i, "[0-9]"), ynms))
    # yldmerge[, c(onms[[1]]) := round(rowMeans(.SD)), .SDcols = onms[[2]]]
    # Selecting rowMax here
    yldmerge[, c(onms[[1]]) := apply(.SD, 1, max), .SDcols = onms[[2]]]
    yldmerge[, c(onms[[2]]) := NULL]
  }

  # Internal check on yield means
  if(length(crml) == 8) crind <- list("S" = 1:4, "R" = 5:8, "I" = 9:12) 
  if(length(crml) == 6) crind <- list("S" = 1:3, "R" = 4:6, "I" = 7:9) 
  for(ii in sample(1:nrow(yldmerge), size = 20)) {
    for(jj in 1:length(mannms)) {  # jj = 2
      a <- yldmerge[ii, ]
      b <- sapply(crind[[jj]], function(x) {
        crml[[x]][mid == a$mid, ][[ncol(crml[[x]])]]
      })
      #if(a[[paste0("YLD", mannms[jj])]] != round(mean(b))) {
      if(a[[paste0("YLD", mannms[jj])]] != round(max(b))) {
        print(paste("Yield mismatch at", ii, mannms[jj]))
      }
    }
  }

  # swap in averaged yields
  # yldmerge[, mid := NULL]
  # yldmerge <- cbind(yldmerge, yldred)
  yldmerge
})
names(ylds) <- crops

soy_ylds <- fread(fp(p_yield, "DSSAT/soy_ylds.csv"))
# soy_ylds[, year := as.numeric(substr(PDAT, 1, 4))]
soy_ylds[, YR := 1:31, by = .(XNAME, TRNO)]  # unique year
soy_ylds[, c("RUNNO", "PROF", "PDAT", "MDAT", "PRCM") := NULL]
setnames(soy_ylds, "SOIL_ID...", "SOL")

# calculate mean and max as YLDS and YLDR
soy_ylds2 <- soy_ylds[,  round(mean(HWAH)), by = .(XNAME, x, y, WTH, SOL, YR)]
setnames(soy_ylds2, "V1", "YLDS")
soy_ylds2[, YLDR := soy_ylds[,  round(max(HWAH)), by = .(XNAME, YR)][, V1]]
soy_ylds2[, XNAME := NULL]

ylds$maize[, c("ID", "Zone", "zoneid") := NULL]
soy_ylds2 <- cbind(c("mid" = 1:nrow(soy_ylds2)), soy_ylds2)
setnames(soy_ylds2, "V1", "mid")

ylds <- list(ylds$maize, soy_ylds2)
names(ylds) <- c("maize", "soybean")

```

## Soil
### Profile data
Note: Using an updated version of `read_dssat_prof.R` from the rCropMod library. It needs to be rebuilt into that library.  Here it is being sourced before runnign the next section. 
```{r, eval = FALSE}
pts <- IDr_ss[, .(x, y)]
# pts <- cbind.data.frame("x" = 27.4738917, "y" = -16.2783358)  # in Zambia
coordinates(pts) <- ~x+y
soilid <- extract(soilgrid, pts)  # grid id corresponding to point
profs <- soilshare[CELL5M %in% soilid]  # profiles at that point
profdat <- read_fullSOL_prof(solfile = full_path(p_soil, "WI.SOL"), 
                             profiles = profs$SoilProfile) 
solvars <- unique(gsub("[[:digit:]]", "", names(profdat)))
lnms <- names(profdat)

# max depth in profile
mdepth <- rowMaxs(as.matrix(profdat[, .SD, .SDcols = grep("SLB", lnms)]), 
                  na.rm = TRUE)

# use depths as weights to calculate mean values of soil parameters to 1 m
# but need to fill to stop at 1 m, and also correct for soils that don't get to 
# 1 m.  Assume values for deepest horizon where deepest < 1 m are the values to 
# 1 m. Truncate horizon that spans 1 m at 1 m, makes this deepest horizon in 
# calc. Horizon depths become weights for weighted mean. 
tst <- profdat[, .SD , .SDcols = grep("SLB", lnms)]
dtest <- function(x) {
  o <- x
  o[is.na(o)] <- -10000
  if(all(o < 100)) {
    o[which.max(o)] <- 100
  } else if(any(o > 100)) {
    o[which(o > 100)[1]] <- 100
  }
  o[o < 0] <- NA
  o
}
depvec <- t(apply(as.matrix(tst), 1, dtest))  # fix depths

infill <- function(x, ind) {
  o <- x
  o0 <- which(o < 0)
  if(length(o0) > 0) {
    for(i in 1:length(o0)) {
      if(o0[i] < max(ind) & o0[i] > min(ind)) {
        o[o0[i]] <- mean(o[o0[i] - 1], o[o0[i] + 1])
      } else if(o0[i] == max(ind)) {
        o[o0[i]] <- o[o0[i] - 1]
      } else if(o0[i] == min(ind)) {
        o[o0[i]] <- o[o0[i] + 1]
      }
    }
  }  
  o
}

# calculate mean values to 1 m depth  
solvarsr <- solvars[solvars %in% 
                     c("SBDM", "SLOC", "SLCL", "SLSI", "SLHW")]
solvals <- do.call(cbind, lapply(solvarsr, function(i) {  # i <- solvarsr[2]
  mat <- as.matrix(profdat[, .SD , .SDcols = grep(i, lnms)])
  oval <- sapply(1:nrow(mat), function(j) {  # j <- 6
    ind <- which(!is.na(depvec[j, ]))
    matv <- mat[j, ind]
    matv <- infill(x = matv, ind)  # fill in NA values within profiles
    w <- depvec[j, ind] - c(0, depvec[j, ind[-length(ind)]])  # depth weights
    # weighted.mean(matv, depvec[j, ind], na.rm = TRUE)  # this was wrong
    weighted.mean(matv, w, na.rm = TRUE)
  })
  oval
}))
colnames(solvals) <- solvarsr

# finalize
sol1m <- cbind("prof" = profdat[, prof], data.table(solvals))
# unique(ylds$soybean$SOL) %in% sol1m$prof
```

### ISRIC grids
```{r, eval = FALSE}
p_spreds <- fp(p_root, "external/input_devel/soil")
mgrid <- raster("agroEcoTradeoff/external/data/ZA/ZA-mask.tif")
ext <- bbox(mgrid)[1:4]  # extent   
crpshp <- as(extent(mgrid), "SpatialPolygons")
crpshp@proj4string <- mgrid@crs

depths <- c(0, 5, 15, 30, 60, 100) 
thickness <- depths[-1] - depths[-length(depths)]

p_soc <- fp(p_sgrids, "carbon/SoilGrids1km")
fnms <- dir(p_soc, pattern = "sd[1-5]", full.names = TRUE)
socr <- stack(lapply(fnms, function(x) {
  r <- raster(x)
  r <- crop(r, spTransform(crpshp, r@crs)) / 10 # crop, convert permille to %
}))
# plot(socr[[1]])

# wmu <- function(x, na.rm = na.rm) weighted.mean(x, w = thickness, na.rm = na.rm) 
# soc1m <- calc(socr, fun = wmu, na.rm = TRUE)
soc1m <- weighted.mean(socr, w = thickness)
soc1m <- writeRaster(soc1m, filename = fp(p_spreds, "soctemp.tif"), 
                     overwrite = TRUE)
# soc1m <- raster(fp(p_spreds, "soctemp.tif"))
# weighted.mean(socr[1], thickness) == soc1m[1]

onm <- fp(p_spreds, "soctempalb.tif")
gdalwarp(srcfile = soc1m@file@name, t_srs = projection(mgrid), 
         dstfile = onm, r = "average", ot = "Float32", te = ext, 
         srcnodata = -32768, dstnodata = -99, tr = c(1000, 1000), 
         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
soc1m <- raster(onm)

# clay grids
p_cl <- fp(p_sgrids, "clay/SoilGrids1km")
fnms <- dir(p_cl, pattern = "sd[1-5]", full.names = TRUE)
sclr <- stack(lapply(fnms, function(x) {
  r <- raster(x)
  r <- crop(r, spTransform(crpshp, r@crs))
  # r[r > 100] <- 100
  r
}))
scl1m <- weighted.mean(sclr, w = thickness)
scl1m <- writeRaster(scl1m, filename = fp(p_spreds, "claytemp.tif"), 
                     overwrite = TRUE)
# scl1m <- raster(fp(p_spreds, "claytemp.tif"))
# weighted.mean(sclr[2], thickness) == scl1m[2]
onm <- fp(p_spreds, "claytempalb.tif")
gdalwarp(srcfile = scl1m@file@name, t_srs = projection(mgrid), 
         dstfile = onm, r = "average", ot = "Float32", te = ext, 
         srcnodata = -32768, dstnodata = -99, tr = c(1000, 1000), 
         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
scl1m <- raster(onm)

# ph
p_ph <- fp(p_sgrids, "pH/SoilGrids1km")
fnms <- dir(p_ph, pattern = "sd[1-5]", full.names = TRUE)
sphr <- stack(lapply(fnms, function(x) {
  r <- raster(x)
  r <- crop(r, spTransform(crpshp, r@crs)) / 10
  r
}))
sph1m <- weighted.mean(sphr, w = thickness)
sph1m <- writeRaster(sph1m, filename = fp(p_spreds, "phtemp.tif"), 
                     overwrite = TRUE)
# sph1m <- raster(fp(p_spreds, "phtemp.tif"))
# weighted.mean(sphr[2], thickness) == sph1m[2]

onm <- fp(p_spreds, "phtempalb.tif")
gdalwarp(srcfile = sph1m@file@name, t_srs = projection(mgrid), 
         dstfile = onm, r = "average", ot = "Float32", te = ext, 
         srcnodata = -32768, dstnodata = -99, tr = c(1000, 1000), 
         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
sph1m <- raster(onm)
# plot(sph1m)

soilgrids <- brick(stack(list("SLOC" = soc1m, "SLCL" = scl1m, "SLHW" = sph1m)))
soilgrids <- writeRaster(soilgrids, filename = fp(p_spreds, "soilgrids.tif"), 
                         overwrite = TRUE)
# plot(soilgrids)
```

## Climate
### Indices
```{r, eval = FALSE}
sdate <- "1948/01/01"
edate <- "2010/12/31"
dts <- seq.Date(as.Date(sdate), as.Date(edate), "days")
yrs <- as.numeric(substr(dts, 1, 4))
ano <- unique(yrs)
mos <- as.numeric(substr(dts, 6, 7))
dys <- as.numeric(substr(dts, 9, 10))

# variabes to derive: MAP, N raindays, rainfall depth, n rain days, 
# n dry days in Jan and Feb, mean max temp, N days > 30, annual mean temp, 
# GDD > 30, average daily SRAD
# set up indices for selecting dates
ystrt <- which(mos == 7 & dys == 15) 
ystrt <- ystrt[-length(ystrt)]
yend <- which(mos == 7 & dys == 14)[-1]
# set up seasonal index
sind <- rep(0, length(dts))  # whole year spanning growing season
for(i in 1:length(ystrt)) sind[ystrt[i]:yend[i]] <- i
sindu <- unique(sind)[-1]
janind <- rep(0, length(dts))  # january of each year
for(i in sindu) janind[which(sind == i & mos == 1)] <- i
febind <- rep(0, length(dts))  # feb of each year
for(i in sindu) febind[which(sind == i & mos == 2)] <- i
jfmind <- rep(0, length(dts))  # djf of each growing season
for(i in sindu) jfmind[which(sind == i & mos %in% c(1, 2, 3))] <- i
ndjfind <- rep(0, length(dts))  # djf of each growing season
for(i in sindu) ndjfind[which(sind == i & mos %in% c(1, 2, 11, 12))] <- i
octind <- rep(0, length(dts))  # feb of each year
for(i in sindu) octind[which(sind == i & mos == 10)] <- i

sindu2 <- sindu[sindu %in% which(ano %in% 1979:2010)]

```

### Rainfall
```{r, eval=FALSE}
prec <- brick(fp(p_pgf, "prec.Zambia.nc"))

# functions for raindays
nrday <- function(x) length(which(x > 0))  # raindays
ndry <- function(x) length(which(x == 0))
sdepth <- function(x) mean(x[x > 0])  # storm depth

# storm depth in NDJF
# sdeps <- lapply(sindu2, function(x) { # x <- sindu2[1]
#   print(x)
#   ryr <- prec[[which(ndjfind == x)]]
#   r <- calc(ryr, sdepth)
#   names(r) <- as.character(ano[x])
#   r
# })
# sdeps <- stack(sdeps)
# 
# # storm depth in NDJF
# rdays <- lapply(sindu2, function(x) { # x <- sindu2[1]
#   print(x)
#   ryr <- prec[[which(djfind == x)]]
#   r <- calc(ryr, nrday)
#   names(r) <- as.character(ano[x])
#   r
# })
# rdays <- stack(rdays)
# 
# # dry days in January
# drydaysj <- lapply(sindu2, function(x) { # x <- sindu2[1]
#   print(x)
#   ryr <- prec[[which(janind == x)]]
#   r <- calc(ryr, ndry)
#   names(r) <- as.character(ano[x])
#   r
# })
# drydaysj <- stack(drydaysj)
# 
# # dry days in February
# drydaysf <- lapply(sindu2, function(x) { # x <- sindu2[1]
#   print(x)
#   ryr <- prec[[which(febind == x)]]
#   r <- calc(ryr, ndry)
#   names(r) <- as.character(ano[x])
#   r
# })
# drydaysf <- stack(drydaysf)

registerDoMC(7)

# total season rainfall
rftot <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(sind == sindu2[i])
  calc(prec[[ind]], sum)
})
writeRaster(rftot, filename = fp(p_clim, "rftot.tif"), overwrite = TRUE)

# total season rainfall
ndjftot <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(ndjfind == sindu2[i])
  calc(prec[[ind]], sum)
})
writeRaster(ndjftot, filename = fp(p_clim, "ndjftot.tif"), overwrite = TRUE)

# JFM season rainfall
jfmtot <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(jfmind == sindu2[i])
  calc(prec[[ind]], sum)
})
writeRaster(jfmtot, filename = fp(p_clim, "jfmtot.tif"), overwrite = TRUE)

# january rainfall
jantot <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(janind == sindu2[i])
  calc(prec[[ind]], sum)
})
writeRaster(jantot, filename = fp(p_clim, "jantot.tif"), overwrite = TRUE)

# february total
febtot <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(febind == sindu2[i])
  calc(prec[[ind]], sum)
})
writeRaster(febtot, filename = fp(p_clim, "febtot.tif"), overwrite = TRUE)

# february total
octtot <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(octind == sindu2[i])
  calc(prec[[ind]], sum)
})
writeRaster(octtot, filename = fp(p_clim, "octtot.tif"), overwrite = TRUE)

# par(mfrow = c(6, 6), mar = c(0, 0, 0, 0))
# for(i in 1:nlayers(octtot)) {
#   plot(octtot[[i]], axes = FALSE, box = FALSE)
#   plot(prov, add = TRUE, zlim = c(0, 1000))
# }
# which(1980:2010 %in% ) 
# pre2005 <- calc(octtot[[1:26]], mean)
# post2005 <- calc(octtot[[27:31]], mean)
# 
# par(mfrow = c(1, 2), mar = c(0, 0, 0, 0))
# plot(pre2005, axes = FALSE, box = FALSE, zlim = c(0, 400))
# plot(prov, add = TRUE)
# plot(post2005, axes = FALSE, box = FALSE, zlim = c(0, 400))
# plot(prov, add = TRUE)
# 
# par(mfrow = c(1, 1), mar = c(0, 0, 0, 0))
# plot((pre2005 - post2005) / pre2005 * 100, zlim = c(-100, 100))
# plot(prov, add = TRUE)

# plot(rftot[[1]] - ndjftot[[1]])
# plot(ndjftot[[1]]); plot(rftot[[1]])

# plot out to show artifacts at 1 degree boundaries
# p_prov <- paste0("/Users/lestes/Dropbox/data/zari/infrastructure/zambia/", 
#                  "rda_dump/Final Shape files RDA/")
# prov <- readOGR(fp(p_prov, "prov.shp"), layer = "prov")
# plot(rftot[[1]])
# plot(calc((ryr > 0), sum))
# plot(sdeps[[1]])
# plot(rdays[[23]])
```

### Temperature
```{r, eval = FALSE}
tmax <- brick(fp(p_pgf, "Tmax.Zambia.nc")) 
tmin <- brick(fp(p_pgf, "Tmin.Zambia.nc"))

# Create tmean variable
tmu <- foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(sind == sindu2[i])
  tmu <- (tmax[[ind]] + tmin[[ind]]) / 2 - 273.15
}

# Functions
# GDD between 8-30 degrees
gdd830 <- function(x) {
  a <- x[x < 30] - 8
  sum(ifelse(a < 0, 0, a))
}
# x <- sample(4:40, size = 100, replace = T)
# gdd830(x)

# killing degree days
kdd <- function(x) {
  a <- x - 30
  sum(ifelse(a < 0, 0, a))
}
# plot(calc(tmu[[1]], kdd))

# frequency of days with T > 30
ftmax30 <- function(x) length(x[x > 30])

# growing degree day and killing degree days
kdd <- stack(kdy <- foreach(i = 1:length(tmu)) %dopar% calc(tmu[[i]], kdd))
writeRaster(kdd, filename = fp(p_clim, "kdd.tif"), overwrite = TRUE)
gdd <- stack(gdy <- foreach(i = 1:length(tmu)) %dopar% calc(tmu[[i]], gdd830))
writeRaster(gdd, filename = fp(p_clim, "gdd.tif"), overwrite = TRUE)

# tmax greater than 30, Jan and Feb
nkilljan <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(janind == sindu2[i])
  calc(tmax[[ind]] - 273.15, ftmax30)
})
writeRaster(nkilljan, filename = fp(p_clim, "nkilljan.tif"), overwrite = TRUE)
nkillfeb <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(febind == sindu2[i])
  calc(tmax[[ind]] - 273.15, ftmax30)
})
writeRaster(nkillfeb, filename = fp(p_clim, "nkillfeb.tif"), overwrite = TRUE)

# mean tmax in NDJF
tmax_gs <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(ndjfind == sindu2[i])
  calc(tmax[[ind]] - 273.15, mean)
})
writeRaster(tmax_gs, filename = fp(p_clim, "tmax_gs.tif"), overwrite = TRUE)

# mean tmax whole year
tmax_ano <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(sind == sindu2[i])
  calc(tmax[[ind]] - 273.15, mean)
})
writeRaster(tmax_ano, filename = fp(p_clim, "tmax_ano.tif"), overwrite = TRUE)

```

### Solar radiation
```{r, eval = FALSE}
srad <- brick(fp(p_pgf, "Rs.Zambia.nc"))

# convert to MJ m^2 
mjday <- function(x) x / 1000000 * (3600 * 24)

# total and mean solar radiation for the growing season
srtot_gs <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(ndjfind == sindu2[i])
  mj <- mjday(srad[[ind]])
  calc(mj, sum)
})
writeRaster(srtot_gs, filename = fp(p_clim, "srtot_gs.tif"), overwrite = TRUE)
srmu_gs <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(ndjfind == sindu2[i])
  mj <- mjday(srad[[ind]])
  calc(mj, mean)
})
writeRaster(srmu_gs, filename = fp(p_clim, "srmu_gs.tif"), overwrite = TRUE)

# total and mean gs for whole year
srtot_ano <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(sind == sindu2[i])
  mj <- mjday(srad[[ind]])
  calc(mj, sum)
})
writeRaster(srtot_ano, filename = fp(p_clim, "srtot_ano.tif"), 
            overwrite = TRUE)
srmu_ano <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(sind == sindu2[i])
  mj <- mjday(srad[[ind]])
  calc(mj, mean)
})
writeRaster(srmu_ano, filename = fp(p_clim, "srmu_ano.tif"), overwrite = TRUE)
```

## Assemble datasets

1. Predictors: 
    + Weather/climate: 
        +`rftot`, `ndjftot`, `jantot`, `febtot` for rainfall; `kdays`, `gdd`, `f30jan`, `f30feb`, `tmax_gs`, `tmax_ano` for temperature; `srtot_gs`, `srmu_gs`, `srtot_ano`, and `srmu_ano` for solar radiation
    + Soil: mean soil properties within top 1 m
    + Management variables: Codes S, R, I
2. Predictands: mean yields across cultivars within each management type

```{r, eval = FALSE}
# Organize rainfall
climnms <- c("rftot", "ndjftot", "jfmtot",  "jantot", "febtot", "octtot", "kdd",
             "gdd", "nkilljan", "nkillfeb", "tmax_ano", "tmax_gs", "srtot_ano",
             "srmu_ano", "srtot_gs", "srmu_gs")
bnms <- sapply(climnms, function(x) dir(p_clim, pattern = x, full.names = TRUE))

blist <- lapply(bnms, function(x) {
  b <- brick(x)
  names(b) <- paste0("Y", 1:nlayers(b))
  b
})

onms <- c("MAP", "MSP", "JFMP", "JP", "FP", "OP", "KDD", "GDD", "NKJ", "NKF", 
          "ATMX", "STMX", "SSRT", "SSRM", "ASRT", "ASRM")
climl <- lapply(1:length(blist), function(x) {
  r <- blist[[x]]
  names(r) <- paste0(onms[x], 1:nlayers(r))
  r
})
names(climl) <- onms

# melt predictors
xy <- unique(ylds$maize[, .(x, y)])
# xy[, ind := 1:.N]
cols <- c("x", "y")
xy[, (cols) := lapply(.SD, function(x) round(x, 5)), .SDcols = cols]

# climate predictors
climvals <- lapply(1:length(climl), function(i) extract(x = climl[[i]], xy))
climxy <- lapply(climvals, function(x) {  # x <- rvals[[1]]
  rmelt <- melt(cbind(xy, x), id = 1:2)
  setnames(rmelt, "value", unique(gsub("[0-9]", "", rmelt$variable)))
  rmelt[, variable := gsub("[[:alpha:]]", "", variable)]
  setnames(rmelt, "variable", "YR")
  rmelt[, ind := 1:.N]
  setkeyv(rmelt, c("ind", "x", "y", "YR"))
  rmelt
})
climpreds <- Reduce(mymerge, climxy)
climpreds[, YR := as.integer(YR)]
setkeyv(climpreds, c("x", "y", "YR"))

# # rainfall
# rvals <- lapply(1:length(rfl), function(i) extract(x = rfl[[i]], xy))
# rvxy <- lapply(rvals, function(x) {  # x <- rvals[[1]]
#   rmelt <- melt(cbind(xy, x), id = 1:2)
#   setnames(rmelt, "value", unique(gsub("[0-9]", "", rmelt$variable)))
#   rmelt[, variable := gsub("[[:alpha:]]", "", variable)]
#   setnames(rmelt, "variable", "YR")
#   # rmelt[, (cols) := lapply(.SD, as.character), .SDcols = cols]
#   rmelt[, ind := 1:.N]
#   setkeyv(rmelt, c("ind", "x", "y", "YR"))
#   rmelt
# })
# rfpreds <- Reduce(mymerge, rvxy)
# 
# # temperature
# tvals <- lapply(1:length(tfl), function(i) extract(x = tfl[[i]], xy))
# tvxy <- lapply(tvals, function(x) {  # x <- tvals[[3]]
#   rmelt <- melt(cbind(xy, x), id = 1:2)
#   setnames(rmelt, "value", unique(gsub("[0-9]", "", rmelt$variable)))
#   rmelt[, variable := gsub("[[:alpha:]]", "", variable)]
#   setnames(rmelt, "variable", "YR")
#   # rmelt[, (cols) := lapply(.SD, as.character), .SDcols = cols]
#   rmelt[, ind := 1:.N]
#   setkeyv(rmelt, c("ind", "x", "y", "YR"))
#   rmelt
# })
# tfpreds <- Reduce(mymerge, tvxy)
# 
# 
# # srad
# svals <- lapply(1:length(sfl), function(i) extract(x = sfl[[i]], xy))
# svxy <- lapply(svals, function(x) {  # x <- rvals[[1]]
#   rmelt <- melt(cbind(xy, x), id = 1:2)
#   setnames(rmelt, "value", unique(gsub("[0-9]", "", rmelt$variable)))
#   rmelt[, variable := gsub("[[:alpha:]]", "", variable)]
#   setnames(rmelt, "variable", "YR")
#   # rmelt[, (cols) := lapply(.SD, as.character), .SDcols = cols]
#   rmelt[, ind := 1:.N]
#   setkeyv(rmelt, c("ind", "x", "y", "YR"))
#   rmelt
# })
# sfpreds <- Reduce(mymerge, svxy)

# Join all climate predictors
# climpreds <- Reduce(mymerge, list(rfpreds, tfpreds, sfpreds))
# climpreds[, YR := as.integer(YR)]
# setkeyv(climpreds, c("x", "y", "YR"))

setnames(sol1m, "prof", "SOL")
setkey(sol1m, "SOL")
sols1mu <- unique(sol1m)

# merge together all the data
regdat <- lapply(ylds, function(i) {  # i <- ylds[[1]]
  i[, (cols) := lapply(.SD, function(x) round(x, 5)), .SDcols = cols]
  setkeyv(i, c("x", "y", "SOL"))
  m1 <- merge(i, sols1mu)
  m1[, YR := as.integer(YR)]
  setkeyv(m1, c("x", "y", "YR"))
  m2 <- merge(m1, climpreds)
})

all(regdat$maize$SOL == regdat$soybean$SOL)
all(regdat$maize$x == regdat$soybean$x)
all(regdat$maize$y == regdat$soybean$y)
all(regdat$maize$OP == regdat$soybean$OP)
all(regdat$maize$SLOC == regdat$soybean$SLOC)

# Checking merges (disabled at present)
# set.seed(234)
# samp <- sample(1:nrow(regdat$maize), size = 20)
# for(ii in samp) {  # ii <- samp[5]
#   a <- regdat$maize[ii, ]; b <- regdat$soybean[ii, ]
#   d <- ID[WTH == a$WTH & zoneid == a$zoneid & SOL == a$SOL]
#   e <- ylds$maize[WTH == b$WTH & zoneid == b$zoneid & SOL == b$SOL[1] & 
#                   YR == b$YR][1, ]
#   f <- ylds$soybean[WTH == b$WTH & zoneid == b$zoneid & SOL == b$SOL[1] & 
#                     YR == b$YR][1, ]
#   # SOIL climate match
#   for(i in list(b, d, e, f)) {
#     # i <- d
#     for(j in c("zoneid", "WTH", "SOL")) {
#       # print(j)
#       if(a[[j]] != i[[j]]) print(paste("mismatch at", j, "row", ii))
#     }
#     for(j in c("x", "y")) {
#       # print(j)
#       if(round(a[[j]], 5) != round(i[[j]], 5)) {  # round because IDr wasn't
#         print(paste("mismatch at", j, "row", ii))
#       }
#     }
#   }
#   for(i in c("YLDS", "YLDR")) { # i <- "YLDS"
#     if(a[[i]] != e[[i]]) print(paste("maize mismatch at", i, "row", ii))
#     if(b[[i]] != f[[i]]) print(paste("maize mismatch at", i, "row", ii))
#   }
# }

# ylds[[1]][ID == 6968 & zoneid == 1]
regdat <- lapply(regdat, function(i) {
  setcolorder(i, c(5, 1:4, 6:ncol(i)))
  setnames(i, "mid", "midold")
  i <- cbind(data.table("mid" = 1:nrow(i)), i)
  i <- setkey(i, "mid")
  i
})

all(regdat$maize$SOL == regdat$soybean$SOL)
all(regdat$maize$x == regdat$soybean$x)
all(regdat$maize$y == regdat$soybean$y)
all(regdat$maize$OP == regdat$soybean$OP)
all(regdat$maize$SLOC == regdat$soybean$SLOC)

regdt <- rbindlist(lapply(names(regdat), function(x) {
  CROP <- x
  cbind(CROP, regdat[[x]])
}))
save(regdt, file = fp(p_yield, "empirical/regdat.rda"))

```