### 2016-02-10 Rmd/yield-predictors.rmd
Deleted prior code for collecting yields from gridded yield datasets and merging with predictors

---
title: "Yield Mapping"
author: "Lyndon Estes"
date: "January 31, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Develop predictors for two-stage yield model

```{r, eval = FALSE}
library(agroEcoTradeoff)
library(gdalUtils)
library(rCropMod)
library(matrixStats)
library(doMC)

setwd("~/Dropbox/publications/zambiaToff/private/zambiaToff/")
p_root <- lmisc::set_base_path()
p_crop <- "~/Dropbox/data/agricultural/zambia/dssat/"
p_cropmod <- "~/Dropbox/projects/rCropMod/external/ext_data/"
p_soil <- "~/DSSAT45/SOIL/"
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_clim <- fp(p_root, "external/input_devel/climate")
p_climo <- "~/Dropbox/data/climate/rainfall/zambia/precip/"
p_yield <- fp(p_root, "external/input_devel/yield")
p_pgf <- "~/Dropbox/data/climate/afdm/pgf_bc/zambia/"
p_sgrids <- "~/Dropbox/data/soil/"
p_dssat <- fp(p_yield, "DSSAT")

# Read in Zambia grid
mgrid <- raster(full_path(p_basedat, "ZA-grid.tif"))

# Zambia provinces
p_prov <- paste0("/Users/lestes/Dropbox/data/zari/infrastructure/zambia/", 
                 "rda_dump/Final Shape files RDA/")
prov <- readOGR(fp(p_prov, "prov.shp"), layer = "prov")

# Read in original soil-climate data from rCropMod
soilgrid <- raster(fp(p_cropmod, "af_soil_grid.tif"))  # soil grid
soilshare <- fread(fp(p_cropmod, "af_soil_share.csv"))  # share of profiles

# Yield data selection points
gcs <- "+proj=longlat +datum=WGS84"
load(fp(p_yield, "empirical/stations.rda"))
zmetf@proj4string <- CRS(gcs)

zmetxy <- data.table(data.frame(zmetf)[, 1:3])
wthnms <- substr(zmetxy$nms, 1, 4)
ginds <- grep("g", wthnms)
combis <- expand.grid(data.frame(cbind(LETTERS, LETTERS, LETTERS)))[, 3:1]
combis <- paste0(combis[, 1], combis[, 2], combis[, 3])
wthnms[ginds] <- paste0("G", combis[1:length(ginds)])
zmetxy <- cbind(data.table("WTH" = wthnms), zmetxy)

```

# Extract subset of DSSAT yields, soils, climate data
## Yield
```{r, eval=FALSE}
# mymerge <- function(x, y) merge(x, y, all = TRUE)
crops <- c("maize", "soy")
ylds <- lapply(crops, function(X) {  # X <- crops[2]
  fnms <- dir(p_dssat, pattern = paste0(X, "_"), full.names = TRUE)
  cr <- fread(fnms, stringsAsFactors = FALSE)
  cr[, YR := RUNNO - min(RUNNO) + 1, by = .(XNAME, TRNO)]  # unique year  
  cr[, c("SOIL_ID...", "RUNNO") := NULL]

  # Yields. S becomes mean across cultivars, R becomes max of cultivars
  cr2 <- cr[, round(mean(HWAH)), by = .(XNAME, x, y, WTH, PROF, YR)]
  setnames(cr2, c("V1", "PROF"), c("YLDS", "SOL"))
  cr2[, YLDR := cr[, round(max(HWAH)), by = .(XNAME, YR)][, V1]]
  cr2
})
names(ylds) <- crops

```

## Soil
### Profile data
Note: Using an updated version of `read_dssat_prof.R` from the rCropMod library. It needs to be rebuilt into that library.  Here it is being sourced before runnign the next section. 
```{r, eval = FALSE}
soilid <- extract(soilgrid, zmetf)  # grid id corresponding to point
profs <- soilshare[CELL5M %in% soilid]  # profiles at that point
profdat <- read_fullSOL_prof(solfile = full_path(p_soil, "WI.SOL"), 
                             profiles = profs$SoilProfile)
setnames(profdat, "prof", "SOL")
solvars <- unique(gsub("[[:digit:]]", "", names(profdat)))
lnms <- names(profdat)

# max depth in profile
mdepth <- rowMaxs(as.matrix(profdat[, .SD, .SDcols = grep("SLB", lnms)]), 
                  na.rm = TRUE)

# use depths as weights to calculate mean values of soil parameters to 1 m
# but need to fill to stop at 1 m, and also correct for soils that don't get to 
# 1 m.  Assume values for deepest horizon where deepest < 1 m are the values to 
# 1 m. Truncate horizon that spans 1 m at 1 m, makes this deepest horizon in 
# calc. Horizon depths become weights for weighted mean. 
tst <- profdat[, .SD , .SDcols = grep("SLB", lnms)]
dtest <- function(x) {
  o <- x
  o[is.na(o)] <- -10000
  if(all(o < 100)) {
    o[which.max(o)] <- 100
  } else if(any(o > 100)) {
    o[which(o > 100)[1]] <- 100
  }
  o[o < 0] <- NA
  o
}
depvec <- t(apply(as.matrix(tst), 1, dtest))  # fix depths

infill <- function(x, ind) {
  o <- x
  o0 <- which(o < 0)
  if(length(o0) > 0) {
    for(i in 1:length(o0)) {
      if(o0[i] < max(ind) & o0[i] > min(ind)) {
        o[o0[i]] <- mean(o[o0[i] - 1], o[o0[i] + 1])
      } else if(o0[i] == max(ind)) {
        o[o0[i]] <- o[o0[i] - 1]
      } else if(o0[i] == min(ind)) {
        o[o0[i]] <- o[o0[i] + 1]
      }
    }
  }  
  o
}

# calculate mean values to 1 m depth  
solvarsr <- solvars[solvars %in% 
                     c("SBDM", "SLOC", "SLCL", "SLSI", "SLHW")]
solvals <- do.call(cbind, lapply(solvarsr, function(i) {  # i <- solvarsr[2]
  mat <- as.matrix(profdat[, .SD , .SDcols = grep(i, lnms)])
  oval <- sapply(1:nrow(mat), function(j) {  # j <- 6
    ind <- which(!is.na(depvec[j, ]))
    matv <- mat[j, ind]
    matv <- infill(x = matv, ind)  # fill in NA values within profiles
    w <- depvec[j, ind] - c(0, depvec[j, ind[-length(ind)]])  # depth weights
    # weighted.mean(matv, depvec[j, ind], na.rm = TRUE)  # this was wrong
    weighted.mean(matv, w, na.rm = TRUE)
  })
  oval
}))
colnames(solvals) <- solvarsr

# finalize
sol1m <- cbind("SOL" = profdat[, SOL], data.table(solvals))
# unique(ylds$soybean$SOL) %in% sol1m$prof
```

### ISRIC grids
```{r, eval = FALSE}
p_spreds <- fp(p_root, "external/input_devel/soil")
mgrid <- raster("agroEcoTradeoff/external/data/ZA/ZA-mask.tif")
ext <- bbox(mgrid)[1:4]  # extent   
crpshp <- as(extent(mgrid), "SpatialPolygons")
crpshp@proj4string <- mgrid@crs

depths <- c(0, 5, 15, 30, 60, 100) 
thickness <- depths[-1] - depths[-length(depths)]

p_soc <- fp(p_sgrids, "carbon/SoilGrids1km")
fnms <- dir(p_soc, pattern = "sd[1-5]", full.names = TRUE)
socr <- stack(lapply(fnms, function(x) {
  r <- raster(x)
  r <- crop(r, spTransform(crpshp, r@crs)) / 10 # crop, convert permille to %
}))
# plot(socr[[1]])

# wmu <- function(x, na.rm = na.rm) weighted.mean(x, w = thickness, na.rm = na.rm) 
# soc1m <- calc(socr, fun = wmu, na.rm = TRUE)
soc1m <- weighted.mean(socr, w = thickness)
soc1m <- writeRaster(soc1m, filename = fp(p_spreds, "soctemp.tif"), 
                     overwrite = TRUE)
# soc1m <- raster(fp(p_spreds, "soctemp.tif"))
# weighted.mean(socr[1], thickness) == soc1m[1]

onm <- fp(p_spreds, "soctempalb.tif")
gdalwarp(srcfile = soc1m@file@name, t_srs = projection(mgrid), 
         dstfile = onm, r = "average", ot = "Float32", te = ext, 
         srcnodata = -32768, dstnodata = -99, tr = c(1000, 1000), 
         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
soc1m <- raster(onm)

# clay grids
p_cl <- fp(p_sgrids, "clay/SoilGrids1km")
fnms <- dir(p_cl, pattern = "sd[1-5]", full.names = TRUE)
sclr <- stack(lapply(fnms, function(x) {
  r <- raster(x)
  r <- crop(r, spTransform(crpshp, r@crs))
  # r[r > 100] <- 100
  r
}))
scl1m <- weighted.mean(sclr, w = thickness)
scl1m <- writeRaster(scl1m, filename = fp(p_spreds, "claytemp.tif"), 
                     overwrite = TRUE)
# scl1m <- raster(fp(p_spreds, "claytemp.tif"))
# weighted.mean(sclr[2], thickness) == scl1m[2]
onm <- fp(p_spreds, "claytempalb.tif")
gdalwarp(srcfile = scl1m@file@name, t_srs = projection(mgrid), 
         dstfile = onm, r = "average", ot = "Float32", te = ext, 
         srcnodata = -32768, dstnodata = -99, tr = c(1000, 1000), 
         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
scl1m <- raster(onm)

# ph
p_ph <- fp(p_sgrids, "pH/SoilGrids1km")
fnms <- dir(p_ph, pattern = "sd[1-5]", full.names = TRUE)
sphr <- stack(lapply(fnms, function(x) {
  r <- raster(x)
  r <- crop(r, spTransform(crpshp, r@crs)) / 10
  r
}))
sph1m <- weighted.mean(sphr, w = thickness)
sph1m <- writeRaster(sph1m, filename = fp(p_spreds, "phtemp.tif"), 
                     overwrite = TRUE)
# sph1m <- raster(fp(p_spreds, "phtemp.tif"))
# weighted.mean(sphr[2], thickness) == sph1m[2]

onm <- fp(p_spreds, "phtempalb.tif")
gdalwarp(srcfile = sph1m@file@name, t_srs = projection(mgrid), 
         dstfile = onm, r = "average", ot = "Float32", te = ext, 
         srcnodata = -32768, dstnodata = -99, tr = c(1000, 1000), 
         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
sph1m <- raster(onm)
# plot(sph1m)

soilgrids <- brick(stack(list("SLOC" = soc1m, "SLCL" = scl1m, "SLHW" = sph1m)))
soilgrids <- writeRaster(soilgrids, filename = fp(p_spreds, "soilgrids.tif"), 
                         overwrite = TRUE)
# plot(soilgrids)
```

## Climate
### Indices
```{r, eval = FALSE}
sdate <- "1948/01/01"
edate <- "2010/12/31"
dts <- seq.Date(as.Date(sdate), as.Date(edate), "days")
yrs <- as.numeric(substr(dts, 1, 4))
ano <- unique(yrs)
mos <- as.numeric(substr(dts, 6, 7))
dys <- as.numeric(substr(dts, 9, 10))

# variabes to derive: MAP, N raindays, rainfall depth, n rain days, 
# n dry days in Jan and Feb, mean max temp, N days > 30, annual mean temp, 
# GDD > 30, average daily SRAD
# set up indices for selecting dates
ystrt <- which(mos == 7 & dys == 15) 
ystrt <- ystrt[-length(ystrt)]
yend <- which(mos == 7 & dys == 14)[-1]
# set up seasonal index
sind <- rep(0, length(dts))  # whole year spanning growing season
for(i in 1:length(ystrt)) sind[ystrt[i]:yend[i]] <- i
sindu <- unique(sind)[-1]
janind <- rep(0, length(dts))  # january of each year
for(i in sindu) janind[which(sind == i & mos == 1)] <- i
febind <- rep(0, length(dts))  # feb of each year
for(i in sindu) febind[which(sind == i & mos == 2)] <- i
jfmind <- rep(0, length(dts))  # djf of each growing season
for(i in sindu) jfmind[which(sind == i & mos %in% c(1, 2, 3))] <- i
ndjfind <- rep(0, length(dts))  # djf of each growing season
for(i in sindu) ndjfind[which(sind == i & mos %in% c(1, 2, 11, 12))] <- i
octind <- rep(0, length(dts))  # feb of each year
for(i in sindu) octind[which(sind == i & mos == 10)] <- i
djfmind <- rep(0, length(dts))  # feb of each year
for(i in sindu) djfmind[which(sind == i & mos  %in% c(12, 1:3))] <- i
# cbind(yrs[sind == 1], mos[sind == 1], djfmind[sind == 1])

sindu2 <- sindu[sindu %in% which(ano %in% 1979:2010)]

```

### Rainfall
```{r, eval=FALSE}
prec <- brick(fp(p_pgf, "prec.Zambia.nc"))

# functions for raindays
nrday <- function(x) length(which(x > 0))  # raindays
ndry <- function(x) length(which(x == 0))
sdepth <- function(x) mean(x[x > 0])  # storm depth

# storm depth in NDJF
# sdeps <- lapply(sindu2, function(x) { # x <- sindu2[1]
#   print(x)
#   ryr <- prec[[which(ndjfind == x)]]
#   r <- calc(ryr, sdepth)
#   names(r) <- as.character(ano[x])
#   r
# })
# sdeps <- stack(sdeps)
#

registerDoMC(7)

# total season rainfall
rftot <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(sind == sindu2[i])
  calc(prec[[ind]], sum)
})
writeRaster(rftot, filename = fp(p_clim, "rftot.tif"), overwrite = TRUE)

# total season rainfall
ndjftot <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(ndjfind == sindu2[i])
  calc(prec[[ind]], sum)
})
writeRaster(ndjftot, filename = fp(p_clim, "ndjftot.tif"), overwrite = TRUE)

# DJFM season rainfall
djfmtot <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(djfmind == sindu2[i])
  calc(prec[[ind]], sum)
})
writeRaster(djfmtot, filename = fp(p_clim, "djfmtot.tif"), overwrite = TRUE)

# JFM season rainfall
jfmtot <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(jfmind == sindu2[i])
  calc(prec[[ind]], sum)
})
writeRaster(jfmtot, filename = fp(p_clim, "jfmtot.tif"), overwrite = TRUE)

# january rainfall
jantot <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(janind == sindu2[i])
  calc(prec[[ind]], sum)
})
writeRaster(jantot, filename = fp(p_clim, "jantot.tif"), overwrite = TRUE)

# february total
febtot <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(febind == sindu2[i])
  calc(prec[[ind]], sum)
})
writeRaster(febtot, filename = fp(p_clim, "febtot.tif"), overwrite = TRUE)

# february total
octtot <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(octind == sindu2[i])
  calc(prec[[ind]], sum)
})
writeRaster(octtot, filename = fp(p_clim, "octtot.tif"), overwrite = TRUE)

# plot(rftot[[1]] - ndjftot[[1]])
# plot(ndjftot[[1]]); plot(rftot[[1]])

# plot out to show artifacts at 1 degree boundaries
# p_prov <- paste0("/Users/lestes/Dropbox/data/zari/infrastructure/zambia/", 
#                  "rda_dump/Final Shape files RDA/")
# prov <- readOGR(fp(p_prov, "prov.shp"), layer = "prov")
# plot(rftot[[1]])
# plot(calc((ryr > 0), sum))
# plot(sdeps[[1]])
# plot(rdays[[23]])
```

### Temperature
```{r, eval = FALSE}
tmax <- brick(fp(p_pgf, "Tmax.Zambia.nc")) 
tmin <- brick(fp(p_pgf, "Tmin.Zambia.nc"))

# Create tmean variable
tmu <- foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(sind == sindu2[i])
  tmu <- (tmax[[ind]] + tmin[[ind]]) / 2 - 273.15
}
writeRaster(stack(tmu), filename = fp(p_clim, "tmu.tif"), overwrite = TRUE)

# Functions
gddf <- function(x) {
  a <- x - 8
  sum(ifelse(a < 0, 0, a))
}

# GDD between 8-30 degrees
gdd830 <- function(x) {
  a <- x[x < 30] - 8
  sum(ifelse(a < 0, 0, a))
}
# x <- sample(4:40, size = 100, replace = T)
# gdd830(x)

# killing degree days
kdd <- function(x) {
  a <- x - 30
  sum(ifelse(a < 0, 0, a))
}
# plot(calc(tmu[[1]], kdd))

# frequency of days with T > 30
ftmax30 <- function(x) length(x[x > 30])

# growing degree day and killing degree days
kdd <- stack(kdy <- foreach(i = 1:length(tmu)) %dopar% calc(tmu[[i]], kdd))
writeRaster(kdd, filename = fp(p_clim, "kdd.tif"), overwrite = TRUE)
gdd30 <- stack(gdy <- foreach(i = 1:length(tmu)) %dopar% calc(tmu[[i]], gdd830))
writeRaster(gdd30, filename = fp(p_clim, "gdd30.tif"), overwrite = TRUE)
gdd <- stack(gdy <- foreach(i = 1:length(tmu)) %dopar% calc(tmu[[i]], gddf))
writeRaster(gdd30, filename = fp(p_clim, "gdd.tif"), overwrite = TRUE)

# tmax greater than 30, Jan and Feb
nkilljan <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(janind == sindu2[i])
  calc(tmax[[ind]] - 273.15, ftmax30)
})
writeRaster(nkilljan, filename = fp(p_clim, "nkilljan.tif"), overwrite = TRUE)
nkillfeb <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(febind == sindu2[i])
  calc(tmax[[ind]] - 273.15, ftmax30)
})
writeRaster(nkillfeb, filename = fp(p_clim, "nkillfeb.tif"), overwrite = TRUE)

# mean tmax in NDJF
tmax_gs <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(ndjfind == sindu2[i])
  calc(tmax[[ind]] - 273.15, mean)
})
writeRaster(tmax_gs, filename = fp(p_clim, "tmax_gs.tif"), overwrite = TRUE)

# mean tmax whole year
tmax_ano <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(sind == sindu2[i])
  calc(tmax[[ind]] - 273.15, mean)
})
writeRaster(tmax_ano, filename = fp(p_clim, "tmax_ano.tif"), overwrite = TRUE)

```

### Solar radiation
```{r, eval = FALSE}
srad <- brick(fp(p_pgf, "Rs.Zambia.nc"))

# convert to MJ m^2 
mjday <- function(x) x / 1000000 * (3600 * 24)

# total and mean solar radiation for the growing season
srtot_gs <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(ndjfind == sindu2[i])
  mj <- mjday(srad[[ind]])
  calc(mj, sum)
})
writeRaster(srtot_gs, filename = fp(p_clim, "srtot_gs.tif"), overwrite = TRUE)
srmu_gs1 <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(ndjfind == sindu2[i])
  mj <- mjday(srad[[ind]])
  calc(mj, mean)
})
writeRaster(srmu_gs1, filename = fp(p_clim, "srmu_gs1.tif"), overwrite = TRUE)

# DJFM mean solar radiation
srmu_gs2 <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(djfmind == sindu2[i])
  mj <- mjday(srad[[ind]])
  calc(mj, mean)
})
writeRaster(srmu_gs2, filename = fp(p_clim, "srmu_gs2.tif"), overwrite = TRUE)

# JFM mean solar radiation
srmu_gs3 <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(jfmind == sindu2[i])
  mj <- mjday(srad[[ind]])
  calc(mj, mean)
})
writeRaster(srmu_gs3, filename = fp(p_clim, "srmu_gs3.tif"), overwrite = TRUE)

# total and mean gs for whole year
srtot_ano <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(sind == sindu2[i])
  mj <- mjday(srad[[ind]])
  calc(mj, sum)
})
writeRaster(srtot_ano, filename = fp(p_clim, "srtot_ano.tif"), 
            overwrite = TRUE)
srmu_ano <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(sind == sindu2[i])
  mj <- mjday(srad[[ind]])
  calc(mj, mean)
})
writeRaster(srmu_ano, filename = fp(p_clim, "srmu_ano.tif"), overwrite = TRUE)
```

## Assemble datasets

1. Predictors: 
    + Weather/climate: 
        +`rftot`, `ndjftot`, `jantot`, `febtot` for rainfall; `kdays`, `gdd`, `f30jan`, `f30feb`, `tmax_gs`, `tmax_ano` for temperature; `srtot_gs`, `srmu_gs`, `srtot_ano`, and `srmu_ano` for solar radiation (plus others added on the way)
    + Soil: mean soil properties within top 1 m
    + Management parameters: Codes S, R, 
2. Predictands: mean yields across cultivars within each management type

```{r, eval = FALSE}
# Organize rainfall
climnms <- c("rftot", "ndjftot", "^djfmtot", "^jfmtot", "jantot", "febtot",
             "octtot", "kdd", "gdd\\.", "gdd30", "nkilljan", "nkillfeb",
             "tmax_ano", "tmax_gs", "srtot_ano", "srmu_ano", "srtot_gs",
             "srmu_gs1", "srmu_gs2", "srmu_gs3")
bnms <- sapply(climnms, function(x) dir(p_clim, pattern = x, full.names = TRUE))

blist <- lapply(bnms, function(x) {
  b <- brick(x)
  names(b) <- paste0("Y", 1:nlayers(b))
  b
})

onms <- c("MAP", "MSP", "DJFMP", "JFMP", "JP", "FP", "OP", "KDD", "GDD", 
          "GDDL", "NKJ", "NKF", "ATMX", "STMX", "ASRT", "ASRM", "SSRT", "SSRMI",
          "SSRMII", "SSRMIII")
climl <- lapply(1:length(blist), function(x) {
  r <- blist[[x]]
  names(r) <- paste0(onms[x], 1:nlayers(r))
  r
})
names(climl) <- onms

# melt predictors
xy <- unique(ylds$maize[, .(x, y)])

# climate predictors
mymerge <- function(x, y) merge(x, y, all = TRUE)
climvals <- lapply(1:length(climl), function(i) extract(x = climl[[i]], xy))
climxy <- lapply(climvals, function(x) {  # x <- climvals[[18]]
  rmelt <- melt(cbind(xy, x), id = 1:2)
  setnames(rmelt, "value", unique(gsub("[0-9]", "", rmelt$variable)))
  rmelt[, variable := gsub("[[:alpha:]]", "", variable)]
  setnames(rmelt, "variable", "YR")
  rmelt[, ind := 1:.N]
  setkeyv(rmelt, c("ind", "x", "y", "YR"))
  rmelt
})
climpreds <- Reduce(mymerge, climxy)
climpreds[, YR := as.integer(YR)]
setkeyv(climpreds, c("x", "y", "YR"))

# soil predictors
setkey(sol1m, "SOL")
sols1mu <- unique(sol1m)

# merge together all the data
cols <- c("x", "y")
regdat <- lapply(ylds, function(i) {  # i <- ylds[[1]]
  i[, (cols) := lapply(.SD, function(x) round(x, 5)), .SDcols = cols]
  setkeyv(i, c("x", "y", "SOL"))
  m1 <- merge(i, sols1mu)
  m1[, YR := as.integer(YR)]
  setkeyv(m1, c("x", "y", "YR"))
  m2 <- merge(m1, climpreds)
})

all(regdat$maize$SOL == regdat$soy$SOL)
all(regdat$maize$x == regdat$soy$x)
all(regdat$maize$y == regdat$soy$y)
all(regdat$maize$OP == regdat$soy$OP)
all(regdat$maize$SLOC == regdat$soy$SLOC)

regdt <- rbindlist(lapply(names(regdat), function(x) {
  CROP <- x
  cbind(CROP, regdat[[x]])
}))
save(regdt, file = fp(p_yield, "empirical/regdat.rda"))

```