### 2016-02-14 Rmd/empirical-yields.Rmd

---
title: "Stage 2 Yield Prediction"
author: "Lyndon Estes"
date: "February 5, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Empirical prediction of yield

A yield transfer function for DSSAT yields

```{r, eval = FALSE}
library(agroEcoTradeoff)
library(matrixStats)
library(doMC)
library(mgcv)
library(rworldxtra)

setwd("~/Dropbox/publications/zambiaToff/private/zambiaToff/")
p_root <- lmisc::set_base_path()
p_clim <- fp(p_root, "external/input_devel/climate")
p_yield <- fp(p_root, "external/input_devel/yield")
p_sgrids <- fp(p_root, "external/input_devel/soil") 
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_cropland <- full_path(p_root, "external/input_devel/cropland")
p_figures <- full_path(p_root, "paper/figures/")

# Read in Zambia grid
mgrid <- raster(full_path(p_basedat, "ZA-grid.tif"))

# Zambia provinces
p_prov <- paste0("/Users/lestes/Dropbox/data/zari/infrastructure/zambia/", 
                 "rda_dump/Final Shape files RDA/")
prov <- readOGR(fp(p_prov, "prov.shp"), layer = "prov")

# Regression variables
load(fp(p_yield, "empirical/regdat.rda"))

```

# Fit model
## Exploratory plots
```{r, eval = FALSE}
plot(prov)
regdt[CROP == "maize", plot(DJFMP, YLDR)]
regdt[CROP == "maize", plot(DJFMP, YLDS)]
regdt[CROP == "soy", plot(DJFMP, YLDR)]
regdt[CROP == "soy", plot(DJFMP, YLDS)]
regdt[CROP == "maize", mean(YLDR)]
regdt[CROP == "maize", mean(YLDS)]
regdt[CROP == "soy", mean(YLDR)]
regdt[CROP == "soy", mean(YLDS)]

regnew <- copy(regdt)
# regdt[CROP == "maize", points(x, y, pch = 16, cex = 0.5)]
# regdt[CROP == "maize" & YLDS > 15000, 
#       points(x, y, pch = 16, cex = 0.5, col = "red")]
# regdt[CROP == "soy", plot(MAP, YLDR)]
# regdt[CROP == "soy" & YLDS > 1700, 
#       points(x, y, pch = 16, cex = 0.5, col = "blue")]

# Prune out yields > 8000 maize > 1700 soy in subsistence case
# soyfilt <- regdt[CROP == "soy" & YLDS > 1700, unique(.SD), 
#                 .SDcols = c("x", "y", "SOL", "WTH", "Zone")]
# mzfilt <- regdt[CROP == "maize" & YLDS > 9000, unique(.SD), 
#                 .SDcols = c("x", "y", "SOL", "WTH")]
# # regnew <- regdt[!x %in% soyfilt$x | !y %in% soyfilt$y]
# regnew <- regdt[!x %in% mzfilt$x | !y %in% mzfilt$y]
# regnew[CROP == "maize", points(x, y, pch = 16, cex = 0.5, col = "green")]

# regnew[CROP == "maize", hist(YLDS)]
# regnew[CROP == "maize", hist(YLDR)]

# summary(lm(YLDS ~ poly(MAP, 2), data = regnew[CROP == "maize", ]))
# summary(lm(YLDR ~ poly(MSP, 2), data = regnew[CROP == "maize", ]))

reg_plotter <- function(dep, pred) {
  lm1 <- lm(dep ~ pred)
  lm2 <- lm(dep ~ poly(pred, 2))
  plot(pred, dep, pch = 16, cex = 0.1)
  abline(lm1)
  prng <- c(floor(min(pred)), ceiling(max(pred)))
  dum <- data.frame(seq(prng[1], prng[2], by = diff(prng) / 100))
  # dum <- data.frame(floor(min(pred)):ceiling(max(pred)))
  colnames(dum) <- "pred"
  preds <- predict(lm2, newdata = dum)
  lines(dum[, 1], preds, col = "red")
  mtext(round(summary(lm1)$adj.r.squared, 4), side = 3, adj = 0)
  mtext(round(summary(lm2)$adj.r.squared, 4), side = 3, adj = 0.5)
}

# Rainfall
regnew[CROP == "maize", reg_plotter(YLDS, MAP)] ##
regnew[CROP == "maize", reg_plotter(YLDS, MSP)] #
regnew[CROP == "maize", reg_plotter(YLDS, DJFMP)] ###
regnew[CROP == "maize", reg_plotter(YLDS, JFMP)] #
regnew[CROP == "maize", reg_plotter(YLDS, FP)]  
regnew[CROP == "maize", reg_plotter(YLDS, OP)]

# soil
regnew[CROP == "maize", reg_plotter(YLDS, SLHW)]  ###
regnew[CROP == "maize", reg_plotter(YLDS, SLOC)]  # 
regnew[CROP == "maize", reg_plotter(YLDS, SLCL)]  ## 
regnew[CROP == "maize", reg_plotter(YLDS, SBDM)]  ##

# TEMP
regnew[CROP == "maize", reg_plotter(YLDS, TMU)]  ## 
regnew[CROP == "maize", reg_plotter(YLDS, STMX)]  # 
regnew[CROP == "maize", reg_plotter(YLDS, ATMX)]  #
regnew[CROP == "maize", reg_plotter(YLDS, GDD)]  # 
regnew[CROP == "maize", reg_plotter(YLDS, KDD)]  
regnew[CROP == "maize", reg_plotter(YLDS, NKJ)]  ##
regnew[CROP == "maize", reg_plotter(YLDS, NKF)]  ##

# srad
regnew[CROP == "maize", reg_plotter(YLDS, ASRT)]  
regnew[CROP == "maize", reg_plotter(YLDS, ASRM)]  
regnew[CROP == "maize", reg_plotter(YLDS, SSRMI)]  ##
regnew[CROP == "maize", reg_plotter(YLDS, SSRMII)]  # 
regnew[CROP == "maize", reg_plotter(YLDS, SSRMIII)] #

```

## Reduce variables and fit models

Reduce down to means across years, and select 10% and 90% percentile years

### Maize
```{r, eval = FALSE}
regred <- copy(regnew[CROP == "maize"])
regred[, mid2 := 1:.N]

# regred[, c("Y10S", "Y90S") := as.integer(0)]
# regred[CROP == "maize" & SOL == "HC_GEN0010" & WTH == "KALA", mean(GDD)]
regmu <- regred[CROP == "maize", lapply(.SD, mean), 
                by = list(CROP, x, y, WTH, SOL), 
                .SDcols = names(regred)[c(8:ncol(regred))]]
regmu[, YST := "mu"]
# regnew[CROP == "maize" & Zone == "Zone2a" & WTH == "AABS" & 
#          SOL == "WI_CMZR003", mean(YLDS)]
qs <- function(x, y, ptile) {
  DF <- cbind(1:length(x), x, y)
  DF <- DF[order(DF[, 2]), ]
  DF[round(nrow(DF) * ptile), 3]
}

y <- 30:60
DF <- cbind(1:31, sample(2000:3000, size = 31), y)
DF <- DF[order(DF[, 2]), ]
DF[round(nrow(DF) * 0.9), 3]
sort(DF[, 2])

# a <- regnew[CROP == "maize" & WTH == "AEQA" & SOL == "WI_RGTZ054", YLDS]
# b <- regnew[CROP == "maize" & WTH == "AEQA" & SOL == "WI_RGTZ054", mid]
# qs(a, b, 0.9)
# sort(a)

# Select out quantile values
m10 <- regred[,
  qs(YLDS, mid2, 0.1), by = list(CROP, x, y, WTH, SOL)
][, V1]
m90 <- regred[,
  qs(YLDS, mid2, 0.9), by = list(CROP, x, y, WTH, SOL)
][, V1]

# 
y10 <- regred[mid2 %in% m10]
y90 <- regred[mid2 %in% m90]
y10[, YST := "10"]
y90[, YST := "90"]

y1090 <- rbind(y10, y90)
cnames <- colnames(y1090)[colnames(y1090) %in% colnames(regmu)]
setcolorder(regmu, cnames)
regred2 <- rbind(regmu, y1090[, cnames, with = FALSE])

par(mfrow = c(2, 2))
for(i in c("10", "mu", "90")) {
  regred2[CROP == "maize" & YST == i, hist(YLDS)]
}
regred2[CROP == "maize", mean(YLDS), by = YST]

# Fit gams
regred2[, round(cor(.SD), 2), 
        .SDcols = c("MAP", "MSP", "DJFMP", "JFMP", "JP", "FP", "OP")]  # 0.7
#        MAP  MSP DJFMP  JFMP    JP    FP    OP
# MAP   1.00 0.91  0.91  0.88  0.59  0.57  0.23
# MSP   0.91 1.00  0.90  0.83  0.71  0.63  0.06
# DJFMP 0.91 0.90  1.00  0.96  0.73  0.66 -0.02
# JFMP  0.88 0.83  0.96  1.00  0.73  0.73 -0.03
# JP    0.59 0.71  0.73  0.73  1.00  0.31 -0.07
# FP    0.57 0.63  0.66  0.73  0.31  1.00 -0.09
# OP    0.23 0.06 -0.02 -0.03 -0.07 -0.09  1.00
regred2[, round(cor(.SD), 2),  
        .SDcols = c("TMU", "GDD", "KDD", "NKF", "NKJ", "STMX", "ATMX")]
#       TMU  GDD  KDD  NKF  NKJ STMX ATMX
# TMU  1.00 0.87 0.45 0.80 0.87 0.95 0.90
# GDD  0.87 1.00 0.27 0.65 0.71 0.80 0.89
# KDD  0.45 0.27 1.00 0.31 0.48 0.42 0.34
# NKF  0.80 0.65 0.31 1.00 0.87 0.81 0.78
# NKJ  0.87 0.71 0.48 0.87 1.00 0.86 0.80
# STMX 0.95 0.80 0.42 0.81 0.86 1.00 0.90
# ATMX 0.90 0.89 0.34 0.78 0.80 0.90 1.00
regred2[, round(cor(.SD), 2),  
        .SDcols = c("TMU", "SSRMI", "SSRMIII", "MAP", "DJFMP", "JFMP", "GDD")]
#           TMU SSRMI SSRMIII   MAP DJFMP  JFMP   GDD
# TMU      1.00  0.70    0.69 -0.65 -0.62 -0.60  0.87
# SSRMI    0.70  1.00    0.91 -0.69 -0.60 -0.58  0.40
# SSRMIII  0.69  0.91    1.00 -0.64 -0.59 -0.61  0.45
# MAP     -0.65 -0.69   -0.64  1.00  0.91  0.88 -0.45
# DJFMP   -0.62 -0.60   -0.59  0.91  1.00  0.96 -0.47
# JFMP    -0.60 -0.58   -0.61  0.88  0.96  1.00 -0.43
# GDD      0.87  0.40    0.45 -0.45 -0.47 -0.43  1.00

regred2[, round(cor(.SD), 2), 
        .SDcols = c("SLOC", "SBDM", "SLCL", "SLHW", "SLSI")]  
#       SLOC  SBDM  SLCL  SLHW  SLSI
# SLOC  1.00 -0.67  0.53 -0.30  0.03
# SBDM -0.67  1.00 -0.77  0.34 -0.26
# SLCL  0.53 -0.77  1.00 -0.64 -0.09
# SLHW -0.30  0.34 -0.64  1.00  0.36
# SLSI  0.03 -0.26 -0.09  0.36  1.00

gam1 <- gam(YLDS ~ s(MAP, k = 3)  + s(TMU, k = 3) + s(SLHW, k = 3) + 
              s(SLCL, k = 3) + s(SLOC, k = 3) + 
              s(SSRMI, k = 3), family = gaussian(link = "log"), method = "ML", 
            data = regred2[CROP == "maize"])

gam2 <- gam(YLDS ~ s(MAP, k = 3) + s(TMU, k = 3) + s(SLHW, k = 3) + 
              s(SLCL, k = 3) + s(SLOC, k = 3), 
            family = gaussian(link = "log"), method = "ML", 
            data = regred2[CROP == "maize"])

gam3 <- gam(YLDS ~ poly(MAP, 2) + s(TMU, k = 3) + s(SLHW, k = 3) + 
              s(SLCL, k = 3) + s(SLOC, k = 3), 
            family = gaussian(link = "log"), method = "ML", 
            data = regred2[CROP == "maize"])
gam4 <- gam(YLDS ~ poly(MAP, 2) + poly(TMU, 2) + s(SLHW, k = 3) + 
              s(SLCL, k = 3) + s(SLOC, k = 3), 
            family = gaussian(link = "log"), method = "ML", 
            data = regred2[CROP == "maize"])
gam5 <- gam(YLDS ~ poly(MAP, 2) + poly(TMU, 2) + s(SLHW, k = 3) + 
              s(SLCL, k = 3), 
            family = gaussian(link = "log"), method = "ML", 
            data = regred2[CROP == "maize"])
gam6 <- gam(YLDS ~ poly(DJFMP, 2) + poly(TMU, 2) + s(SLHW, k = 3) + 
              s(SLCL, k = 3), 
            family = gaussian(link = "log"), method = "ML", 
            data = regred2[CROP == "maize"])
gam7 <- gam(YLDS ~ poly(DJFMP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
              s(SLCL, k = 3), 
            family = gaussian(link = "log"), method = "ML", 
            data = regred2[CROP == "maize"])
gam8 <- gam(YLDS ~ poly(DJFMP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
              poly(SLCL, 2), 
            family = gaussian(link = "log"), method = "ML", 
            data = regred2[CROP == "maize"])
gam9 <- gam(YLDS ~ poly(MSP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
              poly(SLCL, 2), 
            family = gaussian(link = "log"), method = "ML", 
            data = regred2[CROP == "maize"])
gam10 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
              poly(SLCL, 2), family = gaussian(link = "log"), method = "ML", 
            data = regred2[CROP == "maize"])
gam11 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
              poly(SLCL, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "ML", 
            data = regred2[CROP == "maize"])
gam12 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
              poly(SLCL, 2) + poly(SLOC, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "ML", 
            data = regred2[CROP == "maize"])
gam13 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2) + poly(SSRMI, 2) + 
               poly(SLHW, 2) + poly(SLCL, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "ML", 
            data = regred2[CROP == "maize"])
gam14 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2) + poly(SSRMI, 2) + 
               poly(SLHW, 2) + poly(SLCL, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "ML", 
            data = regnew[CROP == "maize"])
gam15 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2) + poly(SSRMI, 2) + 
               poly(SLHW, 2) + poly(SLCL, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "ML", 
            data = regred2[YST == "mu"])
gam16 <- gam(YLDS ~ s(JFMP, k = 3)  + s(TMU, k = 3) + s(SLHW, k = 3) + 
               s(SLCL, k = 3) + s(SSRMI, k = 3), 
             family = gaussian(link = "log"), method = "ML", 
             data = regred2[CROP == "maize"])
gam17 <- gam(YLDS ~ poly(JFMP, 2) + poly(GDD, 2) + poly(SSRMI, 2) + 
               poly(SLHW, 2) + poly(SLCL, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "ML", 
            data = regred2[CROP == "maize"])
gam18 <- gam(YLDS ~ poly(JFMP, 2) + poly(GDD, 2) + poly(SSRMI, 2) + 
               poly(SLHW, 2) + poly(SLCL, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "ML", 
            data = regnew[CROP == "maize"])
gam19 <- gam(YLDS ~ poly(DJFMP, 2) + poly(TMU, 2) + poly(SSRMI, 2) + 
               poly(SLHW, 2) + poly(SLCL, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam20 <- gam(YLDS ~ poly(DJFMP, 2) + poly(GDD, 2) + poly(SSRMI, 2) + 
               poly(SLHW, 2) + poly(SLCL, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regred2)
gam21 <- gam(YLDS ~ poly(DJFMP, 2) + poly(GDD, 2) + poly(SSRMI, 2) + 
               poly(SLHW, 2) + poly(SLCL, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regnew[CROP == "maize"])
gam22 <- gam(YLDS ~ poly(MAP, 2) + poly(GDD, 2) + KDD + poly(SSRMI, 2) + 
               poly(SLHW, 2) + poly(SLCL, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regnew[CROP == "maize"])
gam23 <- gam(YLDS ~ poly(MAP, 2) + poly(GDD, 2) + poly(SSRMI, 2) + 
               poly(SLHW, 2) + poly(SLCL, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regnew[CROP == "maize"])

summary(gam1)
plot.gam(gam1, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam2)
plot.gam(gam2, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam3)
plot.gam(gam3, residuals = FALSE, all.terms = TRUE, pages = 1)
summary(gam4)
plot.gam(gam4, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam5)  #
plot.gam(gam5, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam6)  #+  MAP -> DJMF
plot.gam(gam6, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam7)
plot.gam(gam7, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam8)
plot.gam(gam8, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam9)  #- DJFM -> MSP
plot.gam(gam9, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam10) #+ MSP -> JFM ~ DJMF
plot.gam(gam10, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam11)  
plot.gam(gam11, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam12)  #<-> SLOC
plot.gam(gam12, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam13)  #+ SLOC out SSRMI in
plot.gam(gam13, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam14)  #+ SLOC out SSRMI in, full dataset
plot.gam(gam14, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam15)  #+ SLOC out SSRMI in, means only
plot.gam(gam15, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam16)  #- Smooth version of 13, explains less 
plot.gam(gam16, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam17)  # GDD Subs in exactly for TMU, less correlated with SSRMI
plot.gam(gam17, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam18)  # 18 on full dataset
plot.gam(gam18, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam19)  #+ DJFMP + TMU 
plot.gam(gam19, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam20)  #+ DJFMP GDD
plot.gam(gam20, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam21)  #+ 20 with full dataset
plot.gam(gam21, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam22)  #+ MAP and KDD full dataset
plot.gam(gam22, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam23)  #+ MAP full dataset
plot.gam(gam23, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)

par(mfrow = c(2, 2))
regred2[CROP == "maize", plot(YLDS, predict(gam13, type = "response"))]
lines(c(2000, 12000), c(2000, 12000), col = "red")
regnew[CROP == "maize", plot(YLDS, predict(gam14, type = "response"))]
lines(c(2000, 12000), c(2000, 12000), col = "red")
regred2[YST == "mu", plot(YLDS, predict(gam15, type = "response"))]
lines(c(2000, 12000), c(2000, 12000), col = "red")
regred2[, plot(YLDS, predict(gam20, type = "response"))]
lines(c(2000, 12000), c(2000, 12000), col = "red")
regnew[CROP == "maize", plot(YLDS, predict(gam21, type = "response"))]
lines(c(2000, 12000), c(2000, 12000), col = "red")

# Predict on 17 and 18

```

### soy

```{r, eval = FALSE}
regreds <- copy(regnew[CROP == "soy"])
regreds[, mid2 := 1:.N]

regmus <- regreds[, lapply(.SD, mean), by = list(CROP, x, y, WTH, SOL), 
                  .SDcols = names(regred)[c(8:ncol(regred))]]
regmus[, YST := "mu"]

# Select out quantile values
m10 <- regreds[,
  qs(YLDR, mid2, 0.1), by = list(CROP, x, y, WTH, SOL)
][, V1]
m90 <- regreds[,
  qs(YLDR, mid2, 0.9), by = list(CROP, x, y, WTH, SOL)
][, V1]

# Select out 10th and 90th percentile years
y10s <- regreds[mid2 %in% m10]
y90s <- regreds[mid2 %in% m90]
y10s[, YST := "10"]
y90s[, YST := "90"]

y1090s <- rbind(y10s, y90s)
cnames <- colnames(y1090s)[colnames(y1090s) %in% colnames(regmus)]
setcolorder(regmus, cnames)
regreds2 <- rbind(regmus, y1090s[, cnames, with = FALSE])

par(mfrow = c(2, 2))
for(i in c("10", "mu", "90")) {
  regreds2[YST == i, hist(YLDR)]
}
regreds2[ , mean(YLDR), by = YST]

# Rainfall
regnew[CROP == "soy", reg_plotter(YLDS, MAP)] ###
regnew[CROP == "soy", reg_plotter(YLDS, MSP)] ##
regnew[CROP == "soy", reg_plotter(YLDS, DJFMP)] ###
regnew[CROP == "soy", reg_plotter(YLDS, JFMP)] ###
regnew[CROP == "soy", reg_plotter(YLDS, JP)]   
regnew[CROP == "soy", reg_plotter(YLDS, FP)] #  
regnew[CROP == "soy", reg_plotter(YLDS, OP)]

# soil
regnew[CROP == "soy", reg_plotter(YLDS, SLHW)]  ###
regnew[CROP == "soy", reg_plotter(YLDS, SLOC)]  ## 
regnew[CROP == "soy", reg_plotter(YLDS, SLCL)]  ### 
regnew[CROP == "soy", reg_plotter(YLDS, SBDM)]  ##

# TEMP
regnew[CROP == "soy", reg_plotter(YLDS, TMU)]  ### 
regnew[CROP == "soy", reg_plotter(YLDS, STMX)]  ### 
regnew[CROP == "soy", reg_plotter(YLDS, ATMX)]  ##
regnew[CROP == "soy", reg_plotter(YLDS, GDD)]  # 
regnew[CROP == "soy", reg_plotter(YLDS, KDD)]  
regnew[CROP == "soy", reg_plotter(YLDS, NKJ)]  #L
regnew[CROP == "soy", reg_plotter(YLDS, NKF)]  ##

# srad
regnew[CROP == "soy", reg_plotter(YLDS, ASRT)]  ##
regnew[CROP == "soy", reg_plotter(YLDS, ASRM)]  ##
regnew[CROP == "soy", reg_plotter(YLDS, SSRMI)]  ## L
regnew[CROP == "soy", reg_plotter(YLDS, SSRMII)]  ## L 
regnew[CROP == "soy", reg_plotter(YLDS, SSRMIII)] ### L

sgam1 <- gam(YLDS ~ s(MAP, k = 3) + s(TMU, k = 3)  + s(SLHW, k = 3) + 
               s(SLCL, k = 3) + s(SLOC, k = 3) + s(SSRMIII, k = 3), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam2 <- gam(YLDS ~ poly(MAP, 2) + poly(TMU, 2)  + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SLOC, 2) + poly(SSRMIII, 2), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam3 <- gam(YLDS ~ poly(DJFMP, 2) + poly(TMU, 2)  + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SLOC, 2) + poly(SSRMIII, 2), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam4 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2)  + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SLOC, 2) + poly(SSRMIII, 2), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam5 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2)  + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SLOC, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam6 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2)  + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam7 <- gam(YLDS ~ poly(JFMP, 2) + poly(GDD, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam8 <- gam(YLDS ~ poly(JFMP, 2) + poly(STMX, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam9 <- gam(YLDS ~ poly(JFMP, 2) + poly(NKJ, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam10 <- gam(YLDS ~ poly(JFMP, 2) + poly(NKF, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam11 <- gam(YLDS ~ poly(JFMP, 2) + poly(ATMX, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam12 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam13 <- gam(YLDS ~ poly(JFMP, 2) + poly(GDD, 2) + KDD + poly(SLHW, 2) + 
               poly(SLCL, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam14 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
# sgam14a <- gam(YLDS ~ poly(MAP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
#                poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
#              family = gaussian(link = "log"), method = "REML", 
#             data = regreds2[YST == "mu"])
# summary(sgam14a)
# plot.gam(sgam14a, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
sgam15 <- gam(YLDS ~ poly(JFMP, 2) + poly(GDD, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regnew[CROP == "soy"])
sgam16 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
            data = regnew[CROP == "soy"])

summary(sgam1)
plot.gam(sgam1, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam2)  #+ poly version
plot.gam(sgam2, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam3) #+ DJFMP 
plot.gam(sgam3, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam4) #+ Weakly higher for JFMP
plot.gam(sgam4, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam5) #+ s(x,y)
plot.gam(sgam5, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam6) #- dropping SLOC doesn't lose too much, wasn't sensible
plot.gam(sgam6, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam7) # <-> swapping in GDD doesn't change much
plot.gam(sgam7, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam8) #+ STMX 
plot.gam(sgam8, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam9) #- NKJ loses a bit, but looks more sensible
plot.gam(sgam9, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam10) #<-> NKF loses a bit, but wrong direction
plot.gam(sgam10, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam11) #<-> Annual TMx, same deal
plot.gam(sgam11, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam12) #<-> TMU still increasing, even with SSRMIII dropped
plot.gam(sgam12, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam13) #<-> KDD added, going in wrong direction
plot.gam(sgam13, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam14a) #<-> same as 6
plot.gam(sgam14a, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam15) #<-> Full data version of 7 (with GDD), becomes significant
plot.gam(sgam15, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam16) #<-> Full data version of 14 (Tmu)
plot.gam(sgam16, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)

par(mfrow = c(2, 2))
plot(regreds2[, YLDS], predict(sgam7, type = "response"))
lines(c(0, 8000), c(0, 8000), col = "red")
plot(regnew[CROP == "soy", YLDS], predict(sgam15, type = "response"))
lines(c(0, 8000), c(0, 8000), col = "red")
plot(regreds2[, YLDS], predict(sgam14, type = "response"))
lines(c(0, 8000), c(0, 8000), col = "red")

# gams to keep 
save(gam13, gam14, gam20, gam21, gam23, sgam7, sgam14, sgam15, sgam16, 
     file = fp(p_yield, "empirical/gams1.rda"))
# load(fp(p_yield, "empirical/gams1.rda"))
summary(gam13)
summary(sgam14)

```

## Load in grid predictors
```{r, eval = FALSE}
soilgrids <- brick(fp(p_sgrids, "soilgrids.tif"))
climvars <- c("rftot", "djfmtot", "^jfmtot", "gdd\\.", "tmu", "srmu_gs1", 
              "srmu_gs3")
climgrids <- sapply(climvars, function(x) {  # x <- climvars[2]
  print(x)
  brick(dir(p_clim, full.names = TRUE)[grep(x, dir(p_clim))])
})
climgrids_mu <- sapply(climgrids, function(x) calc(x, mean))
climst <- stack(climgrids_mu)
names(climst) <- c("MAP", "DJFMP", "JFMP", "GDD", "TMU", "SSRMI", "SSRMIII")

# disaggregate down to 1 km
climstbase <- disaggregate(climst[[1]], fact = 30)
climststep <- disaggregate(climst, fact = 20)
climst1km <- resample(climst, climstbase)

# need to warp soilgrids back to that also
ext <- bbox(climst1km)[1:4]  # extent   
onm <- fp(p_sgrids, "soilgridsgcs.tif")
gdalwarp(srcfile = soilgrids@file@name, t_srs = projection(climst1km), 
         dstfile = onm, r = "average", ot = "Float32", te = ext, 
         srcnodata = -32768, dstnodata = -99, tr = res(climst1km), 
         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
soilgcs <- brick(onm)
names(soilgcs) <- c("SLOC", "SLCL", "SLHW")
plot(soilgcs)

# need also a grid of x,y values
x <- soilgcs[[1]]
x[] <- xFromCell(x, 1:ncell(x))
y <- soilgcs[[1]]
y[] <- yFromCell(y, 1:ncell(y))
plot(y)
xy <- stack(x, y)
names(xy) <- c("x", "y")
predst <- stack(xy, climst1km, soilgcs)
# plot(predst[["TMU"]])

```

## Predicted yields surfaces
```{r, eval = FALSE}
mzred <- raster::predict(predst, gam20, type = "response")
mzred2 <- raster::predict(predst, gam13, type = "response")
mzfull <- raster::predict(predst, gam21, type = "response")
mzfull2 <- raster::predict(predst, gam14, type = "response")
mzfull3 <- raster::predict(predst, gam23, type = "response")
soyred <- raster::predict(predst, sgam7, type = "response")
soyred2 <- raster::predict(predst, sgam14, type = "response")
soyfull <- raster::predict(predst, sgam15, type = "response")
soyfull2 <- raster::predict(predst, sgam16, type = "response")

# Not unplausible looking, maize not sure
par(mfrow = c(2, 2), mar = c(1, 1, 1, 1))
plot(mzred)
plot(mzred2)
plot(mzfull)
plot(mzfull2)
plot(mzfull3)
plot(soyred)
plot(soyred2)
plot(soyfull)
plot(soyfull2)
plot(prov, add = TRUE)

# prov@proj4string <- soyfull@crs
# soymsk <- crop(soyfull, prov)
# soymsk2 <- crop(soyfull2, prov)
# soymsk3 <- crop(soyfull3, prov)

# Save bricks and models
yldrs <- brick(stack(list(mzred, mzred2, mzfull, mzfull2, mzfull3, soyred, 
                          soyred2, soyfull, soyfull2)))
writeRaster(yldrs, filename = fp(p_yield, "empirical/yield_surfaces2.tif"), 
            overwrite = TRUE)
yldrs <- brick(fp(p_yield, "empirical/yield_surfaces2.tif"))

onm <- fp(p_yield, "empirical/yield_surfaces2_alb.tif")
gdalwarp(srcfile = yldrs@file@name, t_srs = projection(mgrid), 
         dstfile = onm, r = "average", ot = "Float32", te = bbox(mgrid)[1:4], 
         srcnodata = 255, dstnodata = 255, tr = c(1000, 1000), 
         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
yldralb <- brick(onm)
names(yldralb) <- c("maizer1", "maizer2", "maizef1", "maizef2", "maizef3",
                    "soyr1", "soyr2", "soyf1", "soyf2")

yldralbmsk <- raster::mask(yldralb, mgrid)
writeRaster(yldralbmsk, 
            filename = fp(p_yield, "empirical/yield_surfaces2_albmsk.tif"))
# yldralbmsk <- brick(fp(p_yield, "empirical/yield_surfaces2_albmsk.tif"))
# names(yldralbmsk) <- c("maizer1", "maizer2", "maizef1", "maizef2", "maizef3",
#                        "soyr1", "soyr2", "soyf1", "soyf2")

prov@proj4string <- raster()@crs
png(fp(p_yield, "empirical/yield_maps.png"), height = 800, width = 800)
par(mfrow = c(3, 3), mar = rep(2, 4), oma = c(0, 0, 0, 2))
for(i in 1:9) {
  plot(yldralbmsk[[i]], axes = FALSE, box = FALSE)
  plot(spTransform(prov, mgrid@crs), add = TRUE)
  mtext(names(yldralb)[[i]], side = 3, line = 0)
}
dev.off()

```

## Adjust to FAO yields

Per `maize-soy-grid.Rmd`
```{r, eval = FALSE}
load(full_path(p_cropland, "currprod.rda"))
load(full_path(p_cropland, "usda-yields.rda"))
load(full_path(p_cropland, "fao-yields.rda"))
cols <- names(usda)[4:9]
usda_ylds <- usda[var == "prod", cols, with = FALSE] / 
  usda[var == "harvarea", cols, with = FALSE]
usda_mu <- round(rowMeans(usda_ylds), 1)

muprod <- lapply(c("Maize", "Soybeans"), function(x) {
  fao[Crop == x & var == "prod" & Year %in% c(2008:2015), sum(Value), 
      by = Year]
})
names(muprod) <- c("maize", "soy")
mupa <- lapply(c("Maize", "Soybeans"), function(x) {
  fao[Crop == x & var == "pa" & Year %in% c(2008:2015), sum(Value), 
      by = Year]
})
names(mupa) <- c("maize", "soy")
fao_mu <- c("maize" = mean(muprod$maize[, V1] / mupa$maize[, V1]), 
            "soy" = mean(muprod$soy[, V1] / mupa$soy[, V1]))

# current yields, mean of USDA and FAO
curr_mu <- round(colMeans(rbind(usda_mu, fao_mu)), 1)  # 2.3 maize, 2.0 soy

# Calculate the mean yields for the country in each management scenario (across 
# cultivar types)
cropland <- raster(fp(p_cropland, "cropland-1k.tif"))
plot(cropland)
crops <- c("maize", "soy")
ydiffs <- lapply(crops, function(x) { # x <- "maize"
  rl <- yldralbmsk[[grep(x, names(yldralbmsk))]]
  yr <- lapply(1:nlayers(rl), function(y) {  # y <- rl[[1]]
    ysum <- cellStats(rl[[y]] * cropland, sum)
    clsum <- cellStats(cropland, sum)
    wmu <- ysum / clsum
    rmu <- cellStats(rl[[y]], mean)
    rvar <- rl[[y]] / rmu
    list("ratio" = rvar, "mu" = rmu, "croplandmu" = wmu)
  })
  names(yr) <- names(rl)
  yr
})
names(ydiffs) <- crops

# Take the DSSAT means and figure out how much they differ relative to their 
# overall mean across management types - use as management specific ratio
dssat_mus <- sapply(ydiffs, function(x) sapply(x, function(y) y$mu))
dssat_cmus <- sapply(ydiffs, function(x) sapply(x, function(y) y$croplandmu))
# dssat_adj <- sweep(dssat_mus, MARGIN = 2,  colMeans(dssat_mus), FUN = `/`)

# Bruinsma 2007-2050 growth
bruinsma <- c(maize = 3224 / 1674, soy = 2124 / 1257)
fao <- fread(fp(p_yield, "empirical/zambia-maize-soy-yield.csv"))
fao2005 <- fao[year > 2004,]
yrates <- sapply(c("Maize", "Soybeans"), function(i) {
  fao[crop == i, coef(lm(yield / yield[1] ~ year))[2]]
})
ymult <- 1 + yrates * (2050-2014)  # fao yield multipliers

# take average of Bruinsma and FAO projection to account for maize
ymults <- c("maize" = sum(ymult[1] * 0.25, bruinsma[1] * 0.75), bruinsma[2])

# adjusted DSSAT factors
potmus <- curr_mu * c(ymults)  # 4.43 maize, 3.4 soy
potmu_mz <- potmus[1] / (dssat_cmus[[1]] / 1000)  # maize on cropland mean
potmu_soy <- potmus[2] / (dssat_mus[[2]] / 1000)  # soy on overall mean

# adjust yields
mzfao <- lapply(1:length(potmu_mz), function(x) {
  (yldralbmsk[[x]] * potmu_mz[x]) / 1000
})
names(mzfao) <- names(yldralbmsk)[1:5]

soyfao <- lapply(1:length(potmu_soy), function(x) {
  (yldralbmsk[[x + 5]] * potmu_soy[x]) / 1000
})
names(soyfao) <- names(yldralbmsk)[6:9]

# plot(soyfao[[1]])
mzfao <- brick(stack(mzfao))
soyfao <- brick(stack(soyfao))
writeRaster(mzfao, filename = fp(p_yield, "empirical/maize-faoadj.tif"), 
            overwrite = TRUE)
writeRaster(soyfao, filename = fp(p_yield, "empirical/soy-faoadj.tif"), 
            overwrite = TRUE)
# mzfao <- brick(fp(p_yield, "empirical/maize-faoadj.tif"))
# soyfao <- brick(fp(p_yield, "empirical/soy-faoadj.tif"))

data("countriesHigh")
zambia <- countriesHigh[countriesHigh@data$SOVEREIGNT == "Zambia", ]
zambia <- spTransform(zambia, mzfao@crs)

png(fp(p_figures, "yields.png"), width = 800, height = 400)
par(mfrow = c(1, 2), mar = c(0, 0, 0, 4))
plot(zambia, col = "grey80", border = "transparent")
plot(mzfao[[1]], add = TRUE)
plot(zambia, col = "grey80", border = "transparent")
plot(soyfao[[1]], add = TRUE)
dev.off()
```