### 2016-02-14 Rmd/main-analysis.Rmd

---
title: "Main Analyses"
author: "Lyndon Estes"
date: "February 11, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Main Analyses

```{r, warning=FALSE, message=FALSE}
library(agroEcoTradeoff)
library(readxl)
library(gdalUtils)
library(gisutils)
setwd("~/Dropbox/publications/zambiaToff/private/zambiaToff/")
p_root <- getwd()
p_demand <- fp(p_root, "external/input_devel/demand/")
p_mod <- fp(p_root, "agroEcoTradeoff")
p_moddat <- fp(p_mod, "external/data/ZA")
p_figures <- full_path(p_root, "paper/figures/")
p_roads <- fp(p_root, "external/input_devel/roads")
p_roads2 <- paste0("/Users/lestes/Dropbox/data/zari/infrastructure/zambia/", 
                   "rda_dump/Final Shape files RDA/")
p_out <- fp(p_mod, "external/output")
```

```{r, eval = FALSE}
# roads
mgrid <- raster(fp(p_moddat, "ZA-mask.tif"))
load(fp(p_mod, "external/data/ZA/parks_roads.rda"))
zamroads <- readOGR(fp(p_roads, "osmroads/zamroadsfp.shp"), 
                    layer = "zamroadsfp")
roads <- zamroads
save(pas, roads, zambia, 
     file = full_path(p_mod, "external/data/ZA/parks_roads.rda"))

```

## Demand
Following the example of Gasparri et al (2015), who report demand and area planted trends for soybean in Southern Africa (9 countries) between 2000-2013, we use the production trends for both maize and soybean for the periods for 2000-2014 to calculate scenario for production increases by extrapolating linearly from recent trends in production for each crop between 2014 and 2050. We took an average of both the 2000-2014 as well as the 2005-2014 trends, which show larger rates of gain, which may reflect the more recent agricultural development interests (cite).

```{r, eval = FALSE}
# fao stats for maize area harvested
ha <- data.table(read_excel(fp(p_demand, "fao-saf-maize-soybean-ha.xlsx")))
prod <- data.table(read_excel(fp(p_demand, "fao-saf-maize-soybean-prod.xlsx")))

# countries included in Gasparri et al (2015)
gasparri <- unique(ha$AreaName)[c(1:2, 7:8, 13, 15, 17, 19, 22)]
har <- ha[AreaName %in% gasparri]
setkey(har, "AreaName")
prodr <- prod[AreaName %in% gasparri]
setkey(prodr, "AreaName")
# prodr[AreaName == "Zambia" & Year %in% 2012:2014, mean(Value), by = ItemName]
# har[, unique(AreaName)]; prodr[, unique(AreaName)]

haregion <- har[, mean(Value), by = .(ItemName, Year)]
ha_rates <- sapply(c("Maize", "Soybeans"), function(i) {
  haregion[ItemName == i, coef(lm(V1 / V1[1] ~ Year))[2]]
})
prodregion <- prodr[, mean(Value), by = .(ItemName, Year)]
prod_rates <- sapply(c("Maize", "Soybeans"), function(i) {
  prodregion[ItemName == i, coef(lm(V1 / V1[1] ~ Year))[2]]
})
# prodregion[ItemName == "Maize", plot(Year, V1)]
# prodregion[ItemName == "Soybeans", plot(Year, V1)]
# haregion[ItemName == "Maize", plot(Year, V1)]
# haregion[ItemName == "Soybeans", plot(Year, V1)]

ha_rates05 <- sapply(c("Maize", "Soybeans"), function(i) {
  haregion[ItemName == i & Year > 2004, coef(lm(V1 / V1[1] ~ Year))[2]]
})
prod_rates05 <- sapply(c("Maize", "Soybeans"), function(i) {
  prodregion[ItemName == i & Year > 2004, coef(lm(V1 / V1[1] ~ Year))[2]]
})
# (1 + prod_rates05[2] * 13) * prodregion[ItemName == "Soybeans", V1]

allrates <- rbind(1 + ha_rates * (2050-2014), 1 + prod_rates * (2050-2014), 
                  1 + ha_rates05 * (2050-2014), 1 + prod_rates05 * (2050-2014))

# a <- 2
# b <- 10
# (a + sum((b - a) / (2010 - 2000) * rep(1, 10))) / a
# ((b - a) / (2010 - 2000))
# 5 * 100
# targets(c("maize" = 2), 10, 2)

round(colMeans(allrates[c(2, 4), ]), 1)  ## by production
round(colMeans(allrates[c(1, 3), ]), 1)  ## by area harvested
round(colMeans(allrates), 1)  ## averaged across production and area harvested
```

### Yield gap closure on existing land

We use the production based target as our potential scenario for Zambia.  Next is to factor in an assumption for how much demand will be met by closing current yield gaps
```{r}
setwd(p_mod)
path <- "external/data/ZA"
# load(fp(path, "currprod.rda"))
# load("~/Dropbox/projects/agroEcoTradeoff/external/data/ZA/currprod.rda")
# save(currprod, file = fp(path, "currprod.rda"))
cp <- c("maize" = 2912004, "soy" = 203732)

# gap closing function
gapper <- function(cp, gaps, pct) ((((cp * gaps) - cp) * pct) + cp) / cp

# Current gaps
mzgap <- 4.43 / 2.3 
soygap <- 3.4 / 2

# assume 50% closed on current land
g <- gapper(cp, c(mzgap, soygap), 0.25)

# targets given g
targ <- c("maize" = 3, "soy" = 9)
# targ <- c("maize" = 3.2, "soy" = 8.7)
targs <- targets(targ, cp, g)

```

## Tradeoff runs

### Best areas to meet targets for each crop

Running the model individually for each crop. This currently entails updating and the `currprod.rda` file, and then reinvoking `input_handler`  
```{r, eval = FALSE}
# currprod_switch <- function(cropind, path = p_moddat) {
#   master_prod <- c("maize" = 2912004, "soy" = 203732)
#   currprod <- master_prod[cropind]
#   save(currprod, file = fp(path, "currprod.rda"))
#   return(currprod)
# }

# Separate soy and maize impacts
cb <- c("Y" = 1, "C" = 0, "BD" = 0, "COST" = 0)
mzto <- tradeoff_mod(prod_targ = targ["maize"], cbetas = cb, currprodmod = g[1])
soyto <- tradeoff_mod(prod_targ = targ["soy"], cbetas = cb, currprodmod = g[2])
mzsoyto <- tradeoff_mod(prod_targ = targ, cbetas = cb, currprodmod = g)

# mzto$conv[, sum(maize)]
# soyto$conv[, sum(soy)]

CRSobj <- CRS(mzto$inputs$sp$crs)
mztor <- dt_to_raster(mzto$conv, CRSobj)
soytor <- dt_to_raster(soyto$conv, CRSobj)
mzsoytor <- dt_to_raster(mzsoyto$conv, CRSobj)  # converges on same results

# rasters
mzsoyr <- stack(list(mztor, soytor * 2))
mzsoyr <- calc(mzsoyr, sum)
plot(mzsoyr)
# plot(roads)

# Plot
i1 <- rbind(mzto$impacts, soyto$impacts)
crops <- c("maize", "soy", "maize/soy")
# plot(1:10, pch = 20, cex = 4, col = "gold2")
# plot(mzsoyr, col = c("grey", "goldenrod", "green4", "orange"))
imp_plot2(mzsoyr, i1, p_figures, "ideal_yield2", "", crops, TRUE, 
         c("gold2", "green3", "orange"))

```

### Best Areas for Other Three Factors
```{r, eval = FALSE}
basedt <- soyto$inputs$mask
ybl <- list(yb1 <- c(1, 1))
step <- 0.05
cnames <- c("C", "BD", "COST")
ps <- round(do.call(rbind, pareto_steps(cnames, step)), 2); head(ps)
Yv = "conv_area"; Cv = "tot_C"; BDv = "int_mean"; COSTv = "mu_cost"
Yst = sum; Cst = sum; BDst = mean; COSTst = mean
crops <- list("maize", "soy", c("maize", "soy"))
croplnms <- list("maize", "soy", "both")
# lapply(crops, function(x) targ[x])
# lapply(crops, function(x) g[x])
cbdc <- lapply(crops, function(x) {
  a <- Sys.time()
  print(paste("Starting run for", x, "at", a))
  otab <- pareto(cnames, step = step, prod_targ = targ[x], 
                 yblist = ybl, currprodmod = g[x], input_key = "ZA", 
                 Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
                 Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
                 silent = FALSE, ncl = 7)
  load(fp(p_out, "bcode.rda"))
  heatr <- heat(fp(p_out, bcode), basedt, todisk = TRUE) 
  print(Sys.time() - a)
  c("bcode" = bcode, otab, "heat" = heatr)
})
names(cbdc) <- croplnms

# check for datasets where certain weights were removed from contention
clength <- t(sapply(cbdc, function(x) c(nrow(x$params), nrow(x$optitab))))
cbdcmz <- heat2(cbdc$maize, fp(p_out, cbdc$maize$bcode), basedt, todisk = TRUE)
cbdc$maize$heat <- cbdcmz

# load(fp(p_out, "ZA_20160213_172518_13217_37779/impacts_tab.rda"))
# save(cbdc, file = fp(p_out, "results.rda"))
# par(mfrow = c(2, 2), mar = c(0, 0, 0, 4))
# plot(cbdc$both$heat, axes = FALSE, box = FALSE)
# plot(cbdc$maize$heat, axes = FALSE, box = FALSE)
# plot(cbdc$soy$heat, axes = FALSE, box = FALSE)
# a <- Sys.time()
# print(a)
# cbdc_mz <- pareto(cnames, step = step, prod_targ = targ["maize"], 
#                   yblist = ybl, currprodmod = g["maize"], input_key = "ZA", 
#                   Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
#                   Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
#                   silent = FALSE, ncl = 8)
# Sys.time() - a
# load(fp(p_out, "bcode.rda"))
# bcodes <- bcode
# a <- Sys.time()
# cbdc_soy <- pareto(cnames, step = step, prod_targ = targ["soy"], 
#                    yblist = ybl, currprodmod = g["soy"], input_key = "ZA", 
#                    Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
#                    Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
#                    silent = FALSE, ncl = 8)
# Sys.time() - a
# load(fp(p_out, "bcode.rda"))
# bcodes[2] <- bcode
# implist <- list(cbdc_mz, cbdc_soy)
# names(implist) <- bcodes

# tstmp <- function() {
#   paste0(gsub(" ", "", gsub(":|-", "", as.character(Sys.time()))))
# }
# save(implist, file = fp(p_out, paste0("imptabs_", tstmp(), ".rda")))
# load(fp(p_out, "imptabs_20160212165418.rda"))
# implist

# Create overlap maps
# CRSobj <- CRS(soyto$inputs$sp$crs)
# cbdc_mz_ht <- heat_func(fp(p_out, names(implist)[1]), basedt) 
# cbdc_soy_ht <- heat_func(fp(p_out, names(implist)[2]), basedt) 
# plot(cbdc_soy_ht + cbdc_mz_ht)
# plot(cbdc_soy_ht)
# Might need to run it for both crops simultaneously

# Do RGB plot comparing overlaps between two

```

### Conservation interests versus agricultural interests
```{r, eval = FALSE}
ybl <- list(yb1 <- c(1, 1))
step <- 0.02
cnames <- c("Y", "COST")
ps <- round(do.call(rbind, pareto_steps(cnames, step)), 2)
# ps[which(rowSums(ps[, c(1, 3)]) == 0), ]

ycost <- lapply(crops, function(x) {
  a <- Sys.time()
  print(paste("Starting run for", x, "at", a))
  otab <- pareto(cnames, step = step, prod_targ = targ[x], 
                  yblist = ybl, currprodmod = g[x], input_key = "ZA", 
                  Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
                  Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
                  silent = FALSE, ncl = 7)
  load(fp(p_out, "bcode.rda"))
  heatr <- heat(fp(p_out, bcode), basedt, todisk = TRUE) 
  print(Sys.time() - a)
  c("bcode" = bcode, otab, "heat" = heatr)
})
names(ycost) <- croplnms

clength <- t(sapply(ycost, function(x) c(nrow(x$params), nrow(x$optitab))))

cnames <- c("C", "BD")
bdc <- lapply(crops, function(x) {
  a <- Sys.time()
  print(paste("Starting run for", x, "at", a))
  otab <- pareto(cnames, step = step, prod_targ = targ[x], 
                  yblist = ybl, currprodmod = g[x], input_key = "ZA", 
                  Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
                  Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
                  silent = FALSE, ncl = 7)
  load(fp(p_out, "bcode.rda"))
  heatr <- heat(fp(p_out, bcode), basedt, todisk = TRUE) 
  print(Sys.time() - a)
  c("bcode" = bcode, otab, "heat" = heatr)
})
names(bdc) <- croplnms
# bdc3 <- c(bdc, bdc2)  # vestiges to show I forgot to do maize soy case
# bdc <- bdc3

bdcfix <- lapply(bdc, function(x) {  # x <- bdc$maize
  if(nrow(x$params) > nrow(x$optitab)) {
    print(paste("fixing heat map for", x$bcode))
    heatfix <- heat2(x, fp(p_out, x$bcode), basedt, todisk = TRUE)
  } else {
    heatfix <- NULL
  }
  heatfix
})
for(i in 1:length(bdc)) bdc[[i]]$heat <- bdcfix[[i]]

save(cbdc, ycost, bdc, file = fp(p_out, "mainresults1302.rda"))
```

### All 4 together
```{r, eval = FALSE}
step <- 0.05
cnames <- c("Y", "C", "BD", "COST")
ps <- round(do.call(rbind, pareto_steps(cnames, step)), 2)
ybdc <- lapply(crops, function(x) {
  a <- Sys.time()
  print(paste("Starting run for", x, "at", a))
  otab <- pareto(cnames, step = step, prod_targ = targ[x], 
                  yblist = ybl, currprodmod = g[x], input_key = "ZA", 
                  Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
                  Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
                  silent = FALSE, ncl = 7)
  load(fp(p_out, "bcode.rda"))
  # heatr <- heat(fp(p_out, bcode), basedt, todisk = TRUE) 
  print(Sys.time() - a)
  c("bcode" = bcode, otab) #, "heat" = heatr)
})
names(ybdc) <- croplnms

# ran because of fail on iteration for both (croplnms not properly compiled)
# bnames <- c("ZA_20160213_231906_17404_28678", "ZA_20160213_234959_17404_39818")
# ybdc <- lapply(bnames, function(x) {
#   bpath <- fp(p_out, x)
#   load(fp(bpath, "optitab.rda"))
#   load(fp(bpath, "parms.rda"))
#   list("bcode" = x, "optitab" = optimal, "params" = parms)
# })
# names(ybdc) <- croplnms[1:2]

# ybdc2 <- lapply(crops[3], function(x) {
#   a <- Sys.time()
#   print(paste("Starting run for", x, "at", a))
#   otab <- pareto(cnames, step = step, prod_targ = targ[x], 
#                   yblist = ybl, currprodmod = g[x], input_key = "ZA", 
#                   Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
#                   Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
#                   silent = FALSE, ncl = 7)
#   load(fp(p_out, "bcode.rda"))
#   # heatr <- heat(fp(p_out, bcode), basedt, todisk = TRUE) 
#   print(Sys.time() - a)
#   c("bcode" = bcode, otab) #, "heat" = heatr)
# })
# names(ybdc2) <- croplnms[[3]]
# ybdc3 <- c(ybdc, ybdc2)  # vestiges to show I forgot to do maize soy case
# ybdc <- ybdc3

# heat maps
ybdcheat <- lapply(ybdc, function(x) {  # x <- bdc$maize
  print(paste("N params =", nrow(x$params), ": N optimal =",  nrow(x$optitab)))
  heatfix <- heat2(x, fp(p_out, x$bcode), basedt, todisk = TRUE, 
                   overwrite = TRUE)
  heatfix
})
for(i in 1:length(ybdc)) ybdc[[i]]$heat <- ybdcheat[[i]]

plot(ybdc$maize$heat)
par(mfrow = c(1, 2))
plot(ycost$maize$heat)
plot(ycost$soy$heat)
plot(ycost$both$heat)
plot(bdc$both$heat)

```