### 2016-02-18 Rmd/carbon.Rmd
Redid Carbon stocks to correct fact that I was using veg biomass, not C, adding below ground, and re-ran through most of main-analysis up until first plots, using updated model to account for faulty grid ordering.

---
title: "Carbon datasets, updated"
author: "Lyndon Estes"
date: "January 18, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# New carbon layer
```{r, eval = FALSE}
library(agroEcoTradeoff)
library(GSIF)

p_proj <- "/Users/lestes/Dropbox/publications/zambiaToff/private/zambiaToff/"
p_dat <- "/Users/lestes/Dropbox/data/"
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
# p_findat <- "~/Dropbox/projects/agroEcoTradeoff/external/data/dt/"
p_c <- full_path(p_dat, "carbon/baccini/africa/")
p_devel <- full_path(p_proj, "external/input_devel/carbon")
p_soc <- full_path(p_dat, "soil/carbon/SoilGrids1km")
p_bd <- full_path(p_dat, "soil/bulkdens/SoilGrids1km")

# load datasets
mgrid <- raster(full_path(p_basedat, "ZA-grid.tif"))
vcarb <- raster(full_path(p_c, "biomass-C.tif"))
ocarb <- brick(full_path(p_basedat, "ZA-carbon.tif"))  # original carbon
names(ocarb) <- c("veg", "soil", "total")
# plot(ocarb < 30)

```

## Vegetation carbon
```{r, eval=FALSE}
onm <- full_path(p_devel, "veg-biomass.tif")
ext <- bbox(mgrid)[1:4]  # extent   
gdalwarp(srcfile = vcarb@file@name, t_srs = projection(mgrid), 
         dstfile = onm, r = "average", ot = "Float32", te = ext, 
         srcnodata = -32768, dstnodata = -99, tr = c(1000, 1000), 
         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
zvbiom <- raster(onm)  # vegetation carbon

# convert to carbon and then above and below ground carbon
zvcarb <- zvbiom * 0.47  # IPCC 2006

# mean root shoot ratio, average of Mokany et al (2006) for savannas (Table 2)
# and Mugasha et al (2013), Tanzanian miombo root-shoot ratio
rooshoo <- 1 + mean(c(0.642, 0.4))
zvcarbf <- zvcarb * rooshoo
onm2 <- fp(p_devel, "veg-carbon.tif")
writeRaster(zvcarbf, filename = onm2, overwrite = TRUE)
# plot(zvcarbf - ocarb[[1]])

```

## Soil carbon

ISRIC WISE grids, from [here](http://soilgrids1km.isric.org), with units and ancillary information from [here](http://www.isric.org/content/faq-soilgrids#How_much_does_it_cost). Conversion to t/ha from [here](http://www.soilquality.org.au/factsheets/organic-carbon) and [here](https://www.agric.wa.gov.au/climate-change/what-soil-organic-carbon?page=0%2C1).
```{r, eval = FALSE}
crpshp <- as(extent(mgrid), "SpatialPolygons")
crpshp@proj4string <- mgrid@crs

# organic carbon 
socr <- lapply(dir(p_soc, full.names = TRUE), function(x) {
  r <- raster(x)
  r <- crop(r, spTransform(crpshp, r@crs))
  r / 10  # convert from permille to percent
})
# plot(socr[[1]])

# bulk density converted kg / m3 (or g / cm3)
bd <- lapply(dir(p_bd, full.names = TRUE), function(x) {
  r <- raster(x)
  r <- crop(r, spTransform(crpshp, r@crs))
  r  / 1000  # convert from Mg/m3 to kg/m3
})
# plot(bd[[1]] * 1000)

# layer depth
depths <- c(0, 5, 15, 30, 60, 100, 200) 
thickness <- depths[-1] - depths[-length(depths)]

# plot(round(bd[[1]], 1) == 1.2)
# tst <- (socr[[1]] == 1.3 & round(bd[[1]], 1) == 1.2)
# 1.2 * 1.2 * 10
# 10000 * 0.1
# depths / 100

# calculate carbon in each layer
soct <- lapply(1:length(thickness), function(x) {
  socr[[x]] * bd[[x]] * thickness[x]
})

# then total for top 1 m
soct_1m <- calc(stack(soct[1:5]), sum)
soct_1ma <- projectRaster(soct_1m, mgrid)
writeRaster(soct_1ma, filename = full_path(p_devel, "soil-C-1m.tif"), 
            overwrite = TRUE)

# plot(mask(soct_1ma, mgrid))
# plot(ocarb$soil)
# plot(mask(soct_1ma, mgrid) - ocarb$soil)  # same order of magnitude

zamcarb <- brick(stack(zvcarbf, soct_1ma), 
                 filename = full_path(p_devel, "ZA-carbon.tif"), 
                 overwrite = TRUE)

```

## Resulting model inputs

`ZA-carbon.tif` - grid of carbon densities, for impact and prioritization