### 2016-02-18 Rmd/main-analysis.Rmd

---
title: "Main Analyses"
author: "Lyndon Estes"
date: "February 11, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Main Analyses

```{r, warning=FALSE, message=FALSE}
library(agroEcoTradeoff)
library(readxl)
library(gdalUtils)
library(gisutils)

setwd("~/Dropbox/publications/zambiaToff/private/zambiaToff/")
p_root <- getwd()
p_demand <- fp(p_root, "external/input_devel/demand/")
p_mod <- fp(p_root, "agroEcoTradeoff")
p_moddat <- fp(p_mod, "external/data/ZA")
p_figures <- full_path(p_root, "paper/figures/")
p_roads <- fp(p_root, "external/input_devel/roads")
p_roads2 <- paste0("/Users/lestes/Dropbox/data/zari/infrastructure/zambia/", 
                   "rda_dump/Final Shape files RDA/")
p_out <- fp(p_mod, "external/output")
```

```{r, eval = FALSE}
# roads
mgrid <- raster(fp(p_moddat, "ZA-mask.tif"))
load(fp(p_mod, "external/data/ZA/parks_roads.rda"))
zamroads <- readOGR(fp(p_roads, "osmroads/zamroadsfp.shp"), 
                    layer = "zamroadsfp")
roads <- zamroads
save(pas, roads, zambia, 
     file = full_path(p_mod, "external/data/ZA/parks_roads.rda"))

```

## Demand
Following the example of Gasparri et al (2015), who report demand and area planted trends for soybean in Southern Africa (9 countries) between 2000-2013, we use the production trends for both maize and soybean for the periods for 2000-2014 to calculate scenario for production increases by extrapolating linearly from recent trends in production for each crop between 2014 and 2050. We took an average of both the 2000-2014 as well as the 2005-2014 trends, which show larger rates of gain, which may reflect the more recent agricultural development interests (cite).

```{r, eval = FALSE}
# fao stats for maize area harvested
ha <- data.table(read_excel(fp(p_demand, "fao-saf-maize-soybean-ha.xlsx")))
prod <- data.table(read_excel(fp(p_demand, "fao-saf-maize-soybean-prod.xlsx")))

# countries included in Gasparri et al (2015)
gasparri <- unique(ha$AreaName)[c(1:2, 7:8, 13, 15, 17, 19, 22)]
har <- ha[AreaName %in% gasparri]
setkey(har, "AreaName")
prodr <- prod[AreaName %in% gasparri]
setkey(prodr, "AreaName")
# prodr[AreaName == "Zambia" & Year %in% 2012:2014, mean(Value), by = ItemName]
# har[, unique(AreaName)]; prodr[, unique(AreaName)]

haregion <- har[, mean(Value), by = .(ItemName, Year)]
ha_rates <- sapply(c("Maize", "Soybeans"), function(i) {
  haregion[ItemName == i, coef(lm(V1 / V1[1] ~ Year))[2]]
})
prodregion <- prodr[, mean(Value), by = .(ItemName, Year)]
prod_rates <- sapply(c("Maize", "Soybeans"), function(i) {
  prodregion[ItemName == i, coef(lm(V1 / V1[1] ~ Year))[2]]
})
# prodregion[ItemName == "Maize", plot(Year, V1)]
# prodregion[ItemName == "Soybeans", plot(Year, V1)]
# haregion[ItemName == "Maize", plot(Year, V1)]
# haregion[ItemName == "Soybeans", plot(Year, V1)]

ha_rates05 <- sapply(c("Maize", "Soybeans"), function(i) {
  haregion[ItemName == i & Year > 2004, coef(lm(V1 / V1[1] ~ Year))[2]]
})
prod_rates05 <- sapply(c("Maize", "Soybeans"), function(i) {
  prodregion[ItemName == i & Year > 2004, coef(lm(V1 / V1[1] ~ Year))[2]]
})
# (1 + prod_rates05[2] * 13) * prodregion[ItemName == "Soybeans", V1]

allrates <- rbind(1 + ha_rates * (2050-2014), 1 + prod_rates * (2050-2014), 
                  1 + ha_rates05 * (2050-2014), 1 + prod_rates05 * (2050-2014))

# a <- 2
# b <- 10
# (a + sum((b - a) / (2010 - 2000) * rep(1, 10))) / a
# ((b - a) / (2010 - 2000))
# 5 * 100
# targets(c("maize" = 2), 10, 2)

round(colMeans(allrates[c(2, 4), ]), 1)  ## by production
round(colMeans(allrates[c(1, 3), ]), 1)  ## by area harvested
round(colMeans(allrates), 1)  ## averaged across production and area harvested
```

### Yield gap closure on existing land

We use the production based target as our potential scenario for Zambia.  Next is to factor in an assumption for how much demand will be met by closing current yield gaps
```{r}
setwd(p_mod)
path <- "external/data/ZA"
# load(fp(path, "currprod.rda"))
# load("~/Dropbox/projects/agroEcoTradeoff/external/data/ZA/currprod.rda")
# save(currprod, file = fp(path, "currprod.rda"))
cp <- c("maize" = 2912004, "soy" = 203732)

# gap closing function
gapper <- function(cp, gaps, pct) ((((cp * gaps) - cp) * pct) + cp) / cp

# Current gaps
mzgap <- 4.43 / 2.3 
soygap <- 3.4 / 2

# assume 50% closed on current land
g <- gapper(cp, c(mzgap, soygap), 0.25)

# targets given g
targ <- c("maize" = 3, "soy" = 9)
# targ <- c("maize" = 3.2, "soy" = 8.7)
targs <- targets(targ, cp, g)

```

## Tradeoff runs

### Best areas to meet targets for each crop

Running the model individually for each crop. This currently entails updating and the `currprod.rda` file, and then reinvoking `input_handler`  
```{r, eval = FALSE}
# currprod_switch <- function(cropind, path = p_moddat) {
#   master_prod <- c("maize" = 2912004, "soy" = 203732)
#   currprod <- master_prod[cropind]
#   save(currprod, file = fp(path, "currprod.rda"))
#   return(currprod)
# }

# Separate soy and maize impacts
cb <- c("Y" = 1, "C" = 0, "BD" = 0, "COST" = 0)
mzto <- tradeoff_mod(prod_targ = targ["maize"], cbetas = cb, currprodmod = g[1])
soyto <- tradeoff_mod(prod_targ = targ["soy"], cbetas = cb, currprodmod = g[2])
mzsoyto <- tradeoff_mod(prod_targ = targ, cbetas = cb, currprodmod = g)

# mzto$conv[, sum(maize)]
# soyto$conv[, sum(soy)]

CRSobj <- CRS(mzto$inputs$sp$crs)
mztor <- dt_to_raster(mzto$conv, CRSobj)
soytor <- dt_to_raster(soyto$conv, CRSobj)
mzsoytor <- dt_to_raster(mzsoyto$conv, CRSobj)  # converges on same results

# rasters
mzsoyr <- stack(list(mztor, soytor * 2))
mzsoyr <- calc(mzsoyr, sum)
# plot(mzsoyr)
# plot(roads)

# Plot
i1 <- rbind(mzto$impacts, soyto$impacts)
crops <- c("maize", "soy", "maize/soy")
# plot(1:10, pch = 20, cex = 4, col = "gold2")
# plot(mzsoyr, col = c("grey", "goldenrod", "green4", "orange"))
imp_plot2(mzsoyr, i1, p_figures, "ideal_yield", "", crops, 1, TRUE, 
         c("gold2", "green3", "orange"))

basedt <- soyto$inputs$mask
```

### Best Areas for Other Three Factors

Switch off for now
```{r, eval = FALSE}
# ybl <- list(yb1 <- c(1, 1))
# step <- 0.05
# cnames <- c("C", "BD", "COST")
# ps <- round(do.call(rbind, pareto_steps(cnames, step)), 2); head(ps)
# Yv = "conv_area"; Cv = "tot_C"; #BDv = "int_prior"; COSTv = "mu_cost"
# BDv = c(Yv, "int_prior"); COSTv = c(Yv, "mu_cost")
# wmu <- function(x, na.rm = na.rm) {
#   weighted.mean(x = x[, 2], w = x[, 1], na.rm = na.rm)
# }
# # Yst = sum; Cst = sum; BDst = mean; COSTst = mean
# Yst = sum; Cst = sum; BDst = wmu; COSTst = wmu
# 
# # crops <- list("maize", "soy", c("maize", "soy"))
# crops <- list(c("maize", "soy"))
# # croplnms <- list("maize", "soy", "both")
# croplnms <- list("both")
# # lapply(crops, function(x) targ[x])
# # lapply(crops, function(x) g[x])
# cbdc <- lapply(crops, function(x) {
#   a <- Sys.time()
#   print(paste("Starting run for", x, "at", a))
#   otab <- pareto(cnames, step = step, prod_targ = targ[x], 
#                  yblist = ybl, currprodmod = g[x], input_key = "ZA", 
#                  Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
#                  Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
#                  silent = FALSE, ncl = 7)
#   load(fp(p_out, "bcode.rda"))
#   heatr <- heat(fp(p_out, bcode), basedt, todisk = TRUE) 
#   print(Sys.time() - a)
#   c("bcode" = bcode, otab, "heat" = heatr)
# })
# names(cbdc) <- croplnms
# 
# # cbdc <- lapply("ZA_20160215_071436_1558_16149", function(x) {
# #   bpath <- fp(p_out, x)
# #   load(fp(bpath, "impacts_tab.rda"))
# #   inlist <- lapply(unique(outtab$iter), function(x) {
# #     outtab[outtab$iter == x, ]
# #   })
# #   # load(fp(bpath, "optitab.rda"))
# #   optimal <- batch_stat(inlist, Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, 
# #                         Yst = Yst, Cst = Cst, BDst = BDst, COSTst = COSTst)
# #   save(optimal, file = fp(bpath, "optitab.rda"))
# #   load(fp(bpath, "parms.rda"))
# #   heatr <- heat2(fp(p_out, x), basedt, todisk = TRUE, overwrite = TRUE) 
# #   list("bcode" = x, "optimal" = optimal, "params" = parms, "heat" = heatr)
# # })
# # names(cbdc) <- croplnms
# 
# # check for datasets where certain weights were removed from contention
# # clength <- t(sapply(cbdc, function(x) c(nrow(x$params), nrow(x$optitab))))
# # cbdcmz <- heat2(cbdc$maize, fp(p_out, cbdc$maize$bcode), basedt, todisk = TRUE)
# # cbdc$maize$heat <- cbdcmz

```

### Conservation interests versus agricultural interests
```{r, eval = FALSE}
ybl <- list(yb1 <- c(1, 1))
step <- 0.02
cnames <- c("Y", "COST")
ps <- round(do.call(rbind, pareto_steps(cnames, step)), 2)
# ps[which(rowSums(ps[, c(1, 3)]) == 0), ]
Yv = "conv_area"; Cv = "tot_C"
BDv = c(Yv, "int_prior"); COSTv = c(Yv, "mu_cost")
wmu <- function(x, na.rm = na.rm) {
  weighted.mean(x = x[, 2], w = x[, 1], na.rm = na.rm)
}
Yst = sum; Cst = sum; BDst = wmu; COSTst = wmu

crops <- list(c("maize", "soy"))
croplnms <- list("both")

ycost <- lapply(crops, function(x) {
  a <- Sys.time()
  print(paste("Starting run for", x, "at", a))
  otab <- pareto(cnames, step = step, prod_targ = targ[x], 
                  yblist = ybl, currprodmod = g[x], input_key = "ZA", 
                  Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
                  Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
                  silent = FALSE, ncl = 7)
  load(fp(p_out, "bcode.rda"))
  heatr <- heat2(fp(p_out, bcode), basedt, todisk = TRUE) 
  print(Sys.time() - a)
  c("bcode" = bcode, otab, "heat" = heatr)
})
names(ycost) <- croplnms
# plot(ycost$both$optitab$Y, ycost$both$optitab$COST)

# fix because ran with heat2 not yet adjusted for improperly ordered csv files
ycostfix <- lapply(ycost, function(x) {  # x <- ycost$both
  if(nrow(x$params) > nrow(x$optitab)) {
    print(paste("fixing heat map for", x$bcode))
    heatfix <- heat2(fp(p_out, x$bcode), basedt, TRUE, TRUE)
  } else {
    heatfix <- NULL
  }
  heatfix
})
for(i in 1:length(ycost)) ycost[[i]]$heat <- ycostfix[[i]]

# ycost <- lapply("ZA_20160215_072118_1558_04246", function(x) {
#   bpath <- fp(p_out, x)
#   load(fp(bpath, "impacts_tab.rda"))
#   inlist <- lapply(unique(outtab$iter), function(x) {
#     outtab[outtab$iter == x, ]
#   })
#   optimal <- batch_stat(inlist, Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, 
#                         Yst = Yst, Cst = Cst, BDst = BDst, COSTst = COSTst)
#   save(optimal, file = fp(bpath, "optitab.rda"))
#   load(fp(bpath, "parms.rda"))
#   heatr <- heat2(fp(p_out, x), basedt, todisk = TRUE, overwrite = TRUE) 
#   list("bcode" = x, "optimal" = optimal, "params" = parms, "heat" = heatr)
# })
# names(ycost) <- croplnms
# clength <- t(sapply(ycost, function(x) c(nrow(x$params), nrow(x$optitab))))

cnames <- c("C", "BD")
bdc <- lapply(crops, function(x) {
  a <- Sys.time()
  print(paste("Starting run for", x, "at", a))
  otab <- pareto(cnames, step = step, prod_targ = targ[x], 
                  yblist = ybl, currprodmod = g[x], input_key = "ZA", 
                  Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
                  Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
                  silent = FALSE, ncl = 7)
  load(fp(p_out, "bcode.rda"))
  # heatr <- heat2(fp(p_out, bcode), basedt, todisk = TRUE) 
  print(Sys.time() - a)
  c("bcode" = bcode, otab, "heat" = heatr)
})
names(bdc) <- croplnms

# fix because ran with heat2 not yet adjusted for improperly ordered csv files
bdcfix <- lapply(bdc, function(x) {  # x <- ycost$both
  if(nrow(x$params) > nrow(x$optitab)) {
    print(paste("fixing heat map for", x$bcode))
    heatfix <- heat2(fp(p_out, x$bcode), basedt, TRUE, TRUE)
  } else {
    heatfix <- NULL
  }
  heatfix
})
for(i in 1:length(bdc)) bdc[[i]]$heat <- bdcfix[[i]]

# plot(bdc$both$optitab$C, bdc$both$optitab$BD)
# bdc3 <- c(bdc, bdc2)  # vestiges to show I forgot to do maize soy case
# bdc <- bdc3

# # Fix pareto tab
# load(fp(fp(p_out, bdc$both$bcode), "impacts_tab.rda"))
# inlist <- lapply(unique(outtab$iter), function(x) outtab[outtab$iter == x, ])
# outtabf <- batch_stat(inlist, Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, 
#                       Yst = Yst, Cst = Cst, BDst = BDst, COSTst = COSTst)

# bdcfix <- lapply(bdc, function(x) {  # x <- bdc$maize
#   if(nrow(x$params) > nrow(x$optitab)) {
#     print(paste("fixing heat map for", x$bcode))
#     heatfix <- heat2(x, fp(p_out, x$bcode), basedt, todisk = TRUE)
#   } else {
#     heatfix <- NULL
#   }
#   heatfix
# })
# for(i in 1:length(bdc)) bdc[[i]]$heat <- bdcfix[[i]]

# bdc <- lapply("ZA_20160215_072332_1558_54465", function(x) {
#   bpath <- fp(p_out, x)
#   load(fp(bpath, "impacts_tab.rda"))
#   inlist <- lapply(unique(outtab$iter), function(x) {
#     outtab[outtab$iter == x, ]
#   })
#   # load(fp(bpath, "optitab.rda"))
#   optimal <- batch_stat(inlist, Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, 
#                         Yst = Yst, Cst = Cst, BDst = BDst, COSTst = COSTst)
#   save(optimal, file = fp(bpath, "optitab.rda"))
#   load(fp(bpath, "parms.rda"))
#   heatr <- heat2(fp(p_out, x), basedt, todisk = TRUE, overwrite = TRUE) 
#   list("bcode" = x, "optimal" = optimal, "params" = parms, "heat" = heatr)
# })
# names(bdc) <- croplnms

# save(cbdc, ycost, bdc, file = fp(p_out, "mainresults1402.rda"))
save(ycost, bdc, file = fp(p_out, "mainresults1702.rda"))
```

### All 4 together
```{r, eval = FALSE}
step <- 0.05
cnames <- c("Y", "C", "BD", "COST")
ps <- round(do.call(rbind, pareto_steps(cnames, step)), 2)
ybdc <- lapply(crops, function(x) {
  a <- Sys.time()
  print(paste("Starting run for", x, "at", a))
  otab <- pareto(cnames, step = step, prod_targ = targ[x], 
                  yblist = ybl, currprodmod = g[x], input_key = "ZA", 
                  Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
                  Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
                  silent = FALSE, ncl = 7)
  load(fp(p_out, "bcode.rda"))
  # heatr <- heat(fp(p_out, bcode), basedt, todisk = TRUE) 
  print(Sys.time() - a)
  c("bcode" = bcode, otab) #, "heat" = heatr)
})
names(ybdc) <- croplnms

# fix because ran with heat2 not yet adjusted for improperly ordered csv files
ybdcfix <- lapply(ybdc, function(x) {  # x <- ycost$both
  if(nrow(x$params) > nrow(x$optitab)) {
    print(paste("fixing heat map for", x$bcode))
    heatfix <- heat2(fp(p_out, x$bcode), basedt, TRUE, TRUE)
  } else {
    heatfix <- NULL
  }
  heatfix
})
for(i in 1:length(ybdc)) ybdc[[i]]$heat <- ybdcfix[[i]]

# ran because of fail on iteration for both (croplnms not properly compiled)
# bnames <- c("ZA_20160213_231906_17404_28678", "ZA_20160213_234959_17404_39818")
# ybdc <- lapply("ZA_20160215_072541_1558_17131", function(x) {
#   bpath <- fp(p_out, x)
#   load(fp(bpath, "impacts_tab.rda"))
#   inlist <- lapply(unique(outtab$iter), function(x) {
#     outtab[outtab$iter == x, ]
#   })
#   # load(fp(bpath, "optitab.rda"))
#   optimal <- batch_stat(inlist, Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, 
#                         Yst = Yst, Cst = Cst, BDst = BDst, COSTst = COSTst)
#   save(optimal, file = fp(bpath, "optitab.rda"))
#   load(fp(bpath, "parms.rda"))
#   heatr <- heat2(fp(p_out, x), basedt, todisk = TRUE, overwrite = TRUE) 
#   list("bcode" = x, "optimal" = optimal, "params" = parms, "heat" = heatr)
# })
# names(ybdc) <- croplnms

# ybdc2 <- lapply(crops[3], function(x) {
#   a <- Sys.time()
#   print(paste("Starting run for", x, "at", a))
#   otab <- pareto(cnames, step = step, prod_targ = targ[x], 
#                   yblist = ybl, currprodmod = g[x], input_key = "ZA", 
#                   Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
#                   Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
#                   silent = FALSE, ncl = 7)
#   load(fp(p_out, "bcode.rda"))
#   # heatr <- heat(fp(p_out, bcode), basedt, todisk = TRUE) 
#   print(Sys.time() - a)
#   c("bcode" = bcode, otab) #, "heat" = heatr)
# })
# names(ybdc2) <- croplnms[[3]]
# ybdc3 <- c(ybdc, ybdc2)  # vestiges to show I forgot to do maize soy case
# ybdc <- ybdc3

# heat maps
# ybdcheat <- lapply(ybdc, function(x) {  # x <- bdc$maize
#   print(paste("N params =", nrow(x$params), ": N optimal =",  nrow(x$optitab)))
#   heatfix <- heat2(x, fp(p_out, x$bcode), basedt, todisk = TRUE, 
#                    overwrite = TRUE)
#   heatfix
# })
# for(i in 1:length(ybdc)) ybdc[[i]]$heat <- ybdcheat[[i]]

# plot(ybdc$maize$heat)
# par(mfrow = c(1, 2))
# plot(ycost$maize$heat)
# plot(ycost$soy$heat)
# plot(ycost$both$heat)
# plot(bdc$both$heat)
# save(ybdc, cbdc, ycost, bdc, file = fp(p_out, "mainresults1502.rda"))
save(ybdc, ycost, bdc, file = fp(p_out, "mainresults1702.rda"))
# load(fp(p_out, "mainresults1402.rda"))
```

## Figures

How reconcilable are the different land use interests, and what would be the impacts of each?

I am using the two crops combined approach because there is little overlap between the two.

### Figure 1
```{r, eval = FALSE}
# select out results for "pure" cases (weight = 1)
pure <- sapply(cnames, function(x) which(round(ybdc$both$params[, x], 4) == 1))

fnames <- imp_sorter(fp(p_out, ybdc$both$bcode), inds = pure, fullnames = TRUE)
pure_r <- lapply(fnames, function(x) {  # x <- pure[1]
  DT <- cbind(fread(x), basedt[, .(x, y, convertible)])
  DTr <- dt_to_raster(DT, CRSobj)
})
# sapply(1:4, function(x) which(ps[, x] == 1))
# sum across crops and multiply by convertible areas
rs <- lapply(pure_r, function(x) calc(x[[1:2]], sum) * x[[3]])  # actual area
rs2 <- lapply(pure_r, function(x) calc(x[[1:2]], sum))  # conv pixels
# check
# sapply(rs, function(x) cellStats(x, sum))
# ybdc$both$optitab[ind %in% pure, Y]  # same, but C optimized out
# sapply(rs2, function(x) cellStats(x, sum))
# plot(pure_r[[3]])

# figure out overlaps between different pure objectives
pms <- expand.grid(rep(list(0:1), 4))[-1, ]  # logic grid
rownames(pms) <- 1:15
rs_over <- lapply(1:nrow(pms), function(x) {  # apply logic tests
  (rs2[[1]] == pms[x, 1]) & (rs2[[2]] == pms[x, 2]) & (rs2[[3]] == pms[x, 3]) &     (rs2[[4]] == pms[x, 4])
})
rs_overs <- stack(rs_over)

# legend to show class of overlap
pms$leg <- c("Y", "C", "YC", "Bd", "YBd", "BdC", "YBdC", "Ct", 
             "YCt", "CCt", "YCCt", "BdCt", "YBdCt", "CBdCt", 
             "YCBdCt")

# frequency of overlaps in different categories
# sum(pms[c(1, 3, 5, 7, 9, 11, 13, 15), "hist"])
pms$hist <- sapply(1:nlayers(rs_overs), function(x) {
  round(cellStats(rs_overs[[x]] * pure_r[[1]]$convertible, sum), 1)
})

# legend color
pms$col <- c("dodgerblue3", "goldenrod2", "tomato1", "yellow", 
             rep("tomato1", 2), "red3", "deepskyblue1", rep("tomato1", 2),
             rep("red3", 5))

# recode with unique number for overlap class, collapse to RasterLayer, 
# simplify classes for plotting
rs_over_rc <- stack(lapply(1:length(rs_over), function(x) rs_over[[x]] * x))
rs_over_rc <- calc(rs_over_rc, sum)
# plot(rs_over_rc == 6)
rclmat <- cbind(0:15, c(0, 1, 3, 5, 4, 5, 5, 6, 2, 5, 5, 6, 5, 6, 6, 6))
leg2 <- c("Yield", "Cost", "Carbon", "Biodiversity", "2 overlap", "3+ overlap")
rs_over_rc2 <- raster::reclassify(rs_over_rc, rclmat)

# plot(1:2, pch = 20, cex = 3, col = c("yellow", "yellow2"))

# For Panel B - impact statistics 
bpath <- fp(p_out, ybdc$both$bcode)
load(fp(bpath, "impacts_tab.rda"))  # get impacts table

# select out corresponding impact statistics, summarize
pure2 <- unlist(lapply(pure, function(x) c(x, x+1)))
vars <- c("iter", "conv_area", "tot_C", "fpa_conv", "mu_cost", "rarity",
          "priority",  "int_prior", "int_mean")
conv_stats <- selected_impacts(pure, fp(p_out, ybdc$both$bcode))
# conv_stats <- outtab[outtab$iter %in% pure, vars]
# conv_stats <- data.frame(t(sapply(pure, function(x) {  # x <- pure[1]
#   DF <- conv_stats[conv_stats$iter == x, ]
#   c(colSums(DF[, 2:4]), 
#     sapply(5:ncol(DF), function(y) weighted.mean(DF[, y], DF$conv_area)))
# }))) 
# colnames(conv_stats) <- vars[-1]

# rescale for common plot
svec <- c(100000, 10000000, 1000, 1, 0.1, 0.1, 0.1, 0.1)  # rescale factor
conv_stats_n <- sweep(as.matrix(conv_stats)[, -1], MARGIN = 2, svec, "/")
conv_stats_n <- conv_stats_n[, c(1, 4, 2, 6:ncol(conv_stats_n), 3)]

# Plots
# Panel A
# cols <- RColorBrewer::brewer.pal(11, "RdYlBu")
# cols <- colorspace::diverge_hsv(n = 11)
# colsi <- sapply(c(0.5, 0.75, 1), function(x) rgb(1, 1, 0.9, alpha = x))
cols <- c("dodgerblue3", "deepskyblue1", "goldenrod2", "yellow", "tomato1",
          "red3")
plot(1:6, pch = 20, cex = 3, col = cols)

tiff(fp(p_figures, "single_priorities.tif"), height = 3, width = 8, 
     units = "in", res = 300)
par(mfrow = c(1, 3))
imp_plot2(dtr = rs_over_rc2, impacts = i1, fpath = p_figures, 
          fnm = "pure_overlap", cexscale = 0.6, ftitle = "", crops = leg2, 
          stats = FALSE, cols = cols, hold = TRUE, marg = c(0, 0, 0, 0))

# Panel B
# totf <- cellStats(rs_over_rc > 0, sum)
par(mar = c(7, 4, 2, 2))
bp <- barplot(pms$hist / sum(pms$hist) * 100, ylim = c(0, 25), las = 2, 
              col = pms$col, mgp = c(2, 0.25, 0), tcl = -0.2, 
              ylab = "% converted area")
axis(side = 1, at = bp, labels = pms$leg, las = 2, mgp = c(2, 0.25, 0), 
     tcl = -0.2)
# TeachingDemos::subplot( 
#   bp <- barplot(pms$hist / sum(pms$hist) * 100, ylim = c(0, 25), las = 2, 
#                 col = pms$col, mgp = c(2, 0.25, 0), tcl = -0.2, 
#                 ylab = "% converted pixels", names.arg = pms$leg), 
#   x=grconvertX(c(0.1,0.45), from='npc'),
#   y=grconvertY(c(0.6,1), from='npc'),
#   type='fig', pars=list(mar=c(3,3,0,0) + 0.1))

# Panel C
# barplot(conv_stats_n)
blabs <- c("Area", "Time", "Carbon", "Priority", "Priority/intactness", 
           "Intactness", "Forest reserves")
# blabs <- sub(':', '\n', blabs)
tlabs <- c(expression(10^5~"km"^2), "hours", expression(10^7~"tons"), 
           rep("X10", 3), expression(10^3~"km"^2))
bp <- barplot(conv_stats_n, beside = TRUE, ylim = c(0, 14), las = 2, 
              col = pms$col[c(1, 2, 4, 8)], names.arg = rep("", length(blabs)), 
              ylab = "Impact*", mgp = c(2, 0.25, 0), tcl = -0.2)
text(x = bp[2, ] - 0.4, y = rep(-0.3, length(bp[1, ])), adj = 1, 
     labels = blabs, xpd = NA, srt = 90)
text(x = bp[3, ] + 0.4, y = rep(-0.3, length(bp[3, ])), adj = 1, 
     labels = tlabs, xpd = NA, srt = 90, cex = 0.8)
dev.off()

#### Need to redo pareto to pass in weighted mean function for all average 
#### costs (DONE)

```

## Figure 2 & 3 (probably supplemental)
```{r, eval = FALSE}
# yc_bdc_st <- as.matrix(stack(ycost$both$heat, bdc$both$heat) / 100)
# yc_bdc_st <- stack(ycost$both$heat, ycost$both$heat * 0, bdc$both$heat) / 100
# plot(ycost$both$heat > 1)
# plotRGB(yc_bdc_st, r = 1, g = 2, b = 3, colNA = "transparent", scale = 1)
# rgb(red = yc_bdc_st[, 1], rep(0, nrow(yc_bdc_st)), yc_bdc_st[, 2])
# plot(ycost$both$heat)

# individual maps
png(full_path(p_figures, "all-4-maps.png"), height = 7, width = 7, 
     units = "in", res = 300)
par(mfrow = c(2, 2))#, oma = c(0, 0, 0, 2))
ftitles <- c("Yield", "Carbon", "Biodiversity", "Cost")
legp <- c(TRUE, rep(FALSE, 3))
for(i in 1:4) {
  dtr <- calc(pure_r[[i]][[1:2]] * c(1, 2), sum)
  imp_plot2(dtr, impacts = i1, fpath = p_figures, fnm = "all-4-maps", 
            cexscale = 0.6, ftitle = ftitles[i], crops = c("maize", "soy"), 
            stats = FALSE, hold = TRUE, legend = legp[i], marg = c(0, 0, 2, 0))
}
dev.off()

# heat maps
implist <- list(ycost$both, bdc$both)
load(full_path(set_base_path(), paste0("external/data/ZA/parks_roads.rda")))
pacols <- c("transparent", "grey85", "grey75")
png(full_path(p_figures, "ycost-bdc-heat.png"), height = 400, width = 900)
par(mfrow = c(1, 2), oma = c(0, 0, 0, 2))
for(i in 1:2) {
  par(mar = c(1, 1, 1, 4))
  plot(zambia, col = "grey90", border = "transparent")
  ir <- implist[[i]]$heat
  ir[ir == 0] <- NA
  plot(ir, col = rev(heat.colors(20)), add = TRUE, )
  plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2],
       border = FALSE)
  plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3],
       border = FALSE)
  plot(roads, add = TRUE, lwd = 0.1, col = "grey")
}
dev.off()

```

## Figure 4

Value of just a bit of compromise

```{r, eval = FALSE}
p <- data.frame(ybdc$both$params)
outtab[1, ]
p[, c("Y", "C", "BD" ,"COST")] <- round(p[, c("Y", "C", "BD" ,"COST")], 2)
# scens <- list(c(0.9, 0.05, 0.05, 0), 
#               c(0.85, 0.05, 0.05, 0.05), 
#               c(0.1, 0.45, 0.45, 0.0), 
#               c(0.05, 0.45, 0.45, 0.05))
scens <- lapply(c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9), function(x) {
  c(1 - x, rep(x / 2, 2), 0)
})
# scens <- lapply(c(0, 0.15, 0.3, 0.45, 0.6, 0.75), function(x) {
#   c(1 - x, rep(x / 3, 3))
# })
pind <- sapply(scens, function(x) {
  x <- round(x, 2)
  which(p$Y == x[1] & p$C == x[2] & p$BD == x[3] & p$COST == x[4])
})

# select out impacts corresponding to those runs
cstats <- selected_impacts(pind, p_out, ybdc$both$bcode)

# check on selection
# il <- input_handler(p_moddat, crops[[1]])
# chk <- tradeoff_mod(targ, ybdc$both$params[pind[5], 7:10], 
#                     currprodmod = g, exist_list = il)
# DF <- data.frame(chk$impacts)
# outtab[outtab$iter == pind[5], -2]
# round(c(colSums(DF[, 2:4]),
#         sapply(5:ncol(DF), function(z) weighted.mean(DF[, z], DF$conv_area))), 
#       2)
# cstats[5, ]
outtab[outtab$iter == pind[5], ]
outtab[, 1:4][1:100,]

fnames <- dir(bpath, pattern = "csv")
fnames[pind[5]]

# calculate max impacts across both crops, and figure out the impact range, 
# and what percent that is relative to the worst case
allstats <- selected_impacts(1:nrow(p), p_out, ybdc$both$bcode)
minmaxstats <- allstats[, lapply(.SD, range)]
imp_range <- minmaxstats[, lapply(.SD, diff)]  # difference between best/worst
pct_of_tot <- minmaxstats[, lapply(.SD, function(x) diff(x) / max(x))]

dfunc <- function(x, y) 100 - y / x * 100
stnms <- names(cstats)[-1]
imp_pct <- cstats[, lapply(stnms, function(x) {
  dfunc(minmaxstats[[x]][2], get(x))
})]
setnames(imp_pct, stnms)

plot(imp_pct$tot_C, imp_pct$conv_area, type = "l", xlab = "", ylab = "", 
     axes = FALSE)        # xaxs = "r", yaxs = "r")

sctab <- do.call(rbind, scens) * 100
cx <- 0.7
lcols <- c("black", "red", "blue")
imp_pct[, {
  png(fp(p_figures, "weight_impact.png"), height = 4, width = 4, 
      units = "in", res = 300)
  par(mgp = c(2, 0.75, 0), mar = c(3, 3, 1, 1))
  plot(conv_area, tot_C, type = "l", xlab = "", ylab = "", #xaxt = "n", 
       axes = FALSE, ylim = c(0, 65), xlim = c(2, 12))#, xaxs = "r", yaxs = "r")
  lines(conv_area, int_prior, col = "red")
  lines(conv_area, mu_cost, col = "blue")
  axis(1, at = seq(2, 12, 1), cex = cx, cex.axis = cx, line = -0.25, 
       tcl = -0.2, mgp = c(1, 0.25, 0))
  axis(2, at = seq(0, 60, 5), las = 2, cex = cx, cex.axis = cx, line = -0.25, 
       tcl = -0.15, mgp = c(1, 0.25, 0))
  mtext("% land savings", side = 1, line = 1, cex = cx)
  mtext("% Carbon/Biodiv/Cost savings", side = 2, line = 1, cex = cx)
  inms <- c("tot_C", "int_prior", "mu_cost")
  for(i in 1:length(inms)) {
    text(conv_area[-1], get(inms[i])[-1], labels = sctab[-1, i + 1], 
         cex = cx * 0.8, pos = 3, offset = 0.2, col = lcols[i])
    text(conv_area[1] + 0.1, i, labels = 0, cex = cx * 0.8, pos = 3, 
         offset = i * 0.15 - 0.2, col = lcols[i])
  }
  legend("bottomleft", legend = c("Carbon", "Biodiv", "Cost"), cex = cx,
         text.col = lcols, col = lcols, bty = "n", lty = 1)
  dev.off()
}]

scens_r <- lapply(pind, function(x) {  # x <- pind[[1]]
  print(x)
  inm <- paste0(outtab[outtab$iter == x, "rc"][1], ".csv")
  fpath <- fp(fp(p_out, ybdc$both$bcode), inm)
  # fname <- dir(bpath, pattern = "csv", full.names = TRUE)[x]
  DT <- cbind(fread(fpath), basedt[, .(x, y, convertible)])
  DTr <- dt_to_raster(DT, CRSobj)
})

scens_rs2 <- lapply(scens_r, function(x) calc(stack(x[[1:2]]), sum))
# freq((rs2$Y + scens_rs2[[1]]) * scens_r[[1]]$convertible
# plot(scens_rs2[[1]] + scens_rs2[[2]])

# round(sapply(scens_rs2, function(x) {
#   cellStats(x * scens_r[[1]]$convertible * 100, sum)
# }))
# cstats$conv_area

# pms <- list(c(1, 2), c(1, 3), c(1, 4), c(1, 5))
# pms <- expand.grid(rep(list(0:1), 5))[-1, ]  # logic grid

# rasters showing overlap as weight is progressively put on C and BD
scens_over <- lapply(2:6, function(x) scens_rs2[[1]] + scens_rs2[[x]] * 2)

# overlap areas
scens_over_a <- sapply(1:length(scens_over), function(x) { # x <- 4
  sapply(1:3, function(y) {  # y <- 3
    r <- scens_over[[x]] == y
    # plot(r)
    round(cellStats(r * scens_r[[1]]$convertible, sum), 1)
  })
})
sum(scens_over_a[c(2, 3), 4])
# over_cols <- sapply(seq(0, 0.4, 0.1), function(x) rgb(1 - x, 0, x))
# par(mar = c(0, 0, 0, 0))
# plot(1:3, pch = 20, col = rgb(1, 0, 1))
# plot(zambia)

cols <- c("grey", c(rgb(0, 1, 0), rgb(0, 0, 1), rgb(1, 0, 1)))
ftits <- c("Y=0.9")
par(mfrow = c(1, 4))
for(i in 1:4) {
  plot(zambia)
  imp_plot2(scens_over[[i]], impacts = i1, fpath = p_figures, fnm = "bob", 
            col = cols, add = TRUE, 
            legend = FALSE)
}
plot(scens_over[[1]], col = c("transparent", over_cols[[5]]))
plot(scens_rs2[[1]])
plot(scens_over[[4]])
plot(calc(stack(scens_over), sum))

plot(calc(stack(scens_rs2[2:4]), sum))

ybdc$both$optimal[, {
  par(mfrow = c(2, 2), mar = rep(3, 4))
  hist(.SD, axes = FALSE, labels = FALSE)
}, .SDcol = 1:4]

frange <- function(x) diff(range(x)) / max(x)
fqtile <- function(x) diff(quantile(x, probs = c(0.025, 0.975))) / max(x) 
ybdc$both$optimal[, lapply(.SD, frange), .SDcol = 1:4]
ybdc$both$optimal[, lapply(.SD, fqtile), .SDcol = 1:4]

hist(ybdc$both$optimal$Y)

```