### 2016-02-23 Rmd/figures.Rmd
Separated figures onto new Rmd, and completely analysis up through weight selection as function of range in data relative to worst case.

---
title: "Main Analyses"
author: "Lyndon Estes"
date: "February 11, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Main Analyses

```{r, warning=FALSE, message=FALSE}
library(agroEcoTradeoff)
library(readxl)
library(gdalUtils)
library(gisutils)
# URL <- paste0("https://cran.r-project.org/src/contrib/Archive/spatstat/",
#               "spatstat_1.42-2.tar.gz")
# install.packages(URL, repos = NULL, type="source", dependencies = TRUE)
library(spatstat)

setwd("~/Dropbox/publications/zambiaToff/private/zambiaToff/")
p_root <- getwd()
p_demand <- fp(p_root, "external/input_devel/demand/")
p_mod <- fp(p_root, "agroEcoTradeoff")
p_moddat <- fp(p_mod, "external/data/ZA")
p_figures <- full_path(p_root, "paper/figures/")
p_other <- fp(p_root, "external/input_devel/other/")
# p_roads <- fp(p_root, "external/input_devel/roads")
# p_roads2 <- paste0("/Users/lestes/Dropbox/data/zari/infrastructure/zambia/", 
#                    "rda_dump/Final Shape files RDA/")
p_out <- fp(p_mod, "external/output")
```

```{r, eval = FALSE}
# roads
mgrid <- raster(fp(p_moddat, "ZA-mask.tif"))
load(fp(p_mod, "external/data/ZA/parks_roads.rda"))

# farmblocks
fbp <- "/Users/lestes/Dropbox/data/zari/misc/NEW4GEO/Fblocks.sqlite"
# fb <- readOGR(fbp, layer = "fblocks")
# fbalb <- spTransform(fb, mgrid@crs)
# fbalb$ID <- 1:nrow(fbalb)
# fbalb@data <- fbalb@data[, "ID", drop = FALSE]
fnm <- fp(p_other, "fblocks_alb.sqlite")
# writeOGR(fbalb, dsn = fnm, layer = "fblocks_alb", driver = "SQLite")

# read in farmblocks
fbalb <- readOGR(fnm, layer = "fblocks_alb")
fbr <- rasterize(fbalb, mgrid)
fbr[is.na(fbr)] <- 0
fbr[fbr > 0] <- 1
fbr <- raster::mask(fbr, mgrid)
# plot(fbr)

```

<a href="#top">Back to top</a>

## Tradeoff runs

### Best areas to meet targets for each crop

Running the model individually for each crop and for both crops.
```{r, eval = FALSE}
setwd(p_mod)
path <- "external/data/ZA"

# gaps 
cp <- c("maize" = 2912004, "soy" = 203732)  # current production
gapper <- function(cp, gaps, pct) ((((cp * gaps) - cp) * pct) + cp) / cp
mzgap <- 4.43 / 2.3 
soygap <- 3.4 / 2
g <- gapper(cp, c(mzgap, soygap), 0.25)  # assume 50% closed on current land
targ <- c("maize" = 3, "soy" = 9)  # targets given g
targs <- targets(targ, cp, g)

# Separate soy and maize impacts
cb <- c("Y" = 1, "C" = 0, "BD" = 0, "COST" = 0)
mzto <- tradeoff_mod(prod_targ = targ["maize"], cbetas = cb, currprodmod = g[1])
soyto <- tradeoff_mod(prod_targ = targ["soy"], cbetas = cb, currprodmod = g[2])
mzsoyto <- tradeoff_mod(prod_targ = targ, cbetas = cb, currprodmod = g)

# rasters
CRSobj <- CRS(mzto$inputs$sp$crs)
mztor <- dt_to_raster(mzto$conv, CRSobj)
soytor <- dt_to_raster(soyto$conv, CRSobj)
mzsoytor <- dt_to_raster(mzsoyto$conv, CRSobj)  # converges on same results

# rasters
mzsoyr <- stack(list(mztor, soytor * 2))
mzsoyr <- calc(mzsoyr, sum)
# plot(mzsoyr)
# plot(roads)

# Plot
i1 <- rbind(mzto$impacts, soyto$impacts)
crops <- c("maize", "soy", "maize/soy")
# plot(1:10, pch = 20, cex = 4, col = "gold2")
# plot(mzsoyr, col = c("grey", "goldenrod", "green4", "orange"))
imp_plot2(mzsoyr, i1, p_figures, "ideal_yield", "", crops, 1, TRUE, 
         c("gold2", "green3", "orange"))

basedt <- soyto$inputs$mask
crops <- list(c("maize", "soy"))

```
<a href="#top">Back to top</a>

```{r, eval = FALSE}
load(fp(p_out, "mainresults1702.rda"))
```

<a href="#top">Back to top</a>

## Figures

How reconcilable are the different land use interests, and what would be the impacts of each?

I am using the two crops combined approach because there is little overlap between the two.

### Figure 1
```{r, eval = FALSE}
# select out results for "pure" cases (weight = 1)
cnames <- c("Y", "C", "BD", "COST")
pure <- sapply(cnames, function(x) which(round(ybdc$both$params[, x], 4) == 1))

fnames <- imp_sorter(fp(p_out, ybdc$both$bcode), inds = pure, fullnames = TRUE)
pure_r <- lapply(fnames, function(x) {  # x <- pure[1]
  DT <- cbind(fread(x), basedt[, .(x, y, convertible)])
  DTr <- dt_to_raster(DT, CRSobj)
})

# sum across crops and multiply by convertible areas
rs <- lapply(pure_r, function(x) calc(x[[1:2]], sum) * x[[3]])  # actual area
rs2 <- lapply(pure_r, function(x) calc(x[[1:2]], sum))  # conv pixels
names(rs2) <- cnames
# check
# sapply(rs, function(x) cellStats(x, sum))
# ybdc$both$optitab[ind %in% pure, Y]  # same, but C optimized out
# sapply(rs2, function(x) cellStats(x, sum))
# plot(pure_r[[3]])

# figure out overlaps between different pure objectives
pms2a <- expand.grid(rep(list(0:1), 4))[-1, ] 
pms2b <- pms2a[which(rowSums(pms2a) == 2), ]
pms2b$cl <- c("YC", "YBd", "CBd", "YCt", "CCt", "BdCt")
rownames(pms2b) <- 1:6

# pairwise overlaps
rs_over <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
  calc(stack(rs2[which(pms2b[x, ] == 1)]), sum)
})
names(rs_over) <- pms2b$cl

# Jaccard Index of overlaps
rs_over_a <- sapply(1:length(rs_over), function(x) { # x <- 4
  sapply(1:2, function(y) {  # y <- 3
    r <- rs_over[[x]] == y
    round(cellStats(r * pure_r[[1]]$convertible, sum), 1)
  })
})
pms2b$JS <- round(rs_over_a[2, ] / colSums(rs_over_a), 3)

# calculate nearest neighbor distance between each version, using spatstat 
# nncross
zam_owin <- owin(xrange = bbox(zambia)[1, ], yrange = bbox(zambia)[2, ],
                 unitname = "meter")
which1 <- function(x) x == 1  # selector function for r to xy
rsover_nn <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests; x = 2
  ind <- which(pms2b[x, 1:4] == 1)
  rl <- rs2[ind]
  rxy <- lapply(rl, function(x) rasterToPoints(x, fun = which1))
  pps <- lapply(rxy, function(x) {
    ppp(x = x[, 1], y = x[, 2], window = zam_owin)
  })
  # reorder to find which has most points, to put that in as X
  reord <- sort(sapply(pps, function(x) x$n), decreasing = TRUE)
  ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
})
pms2b$NND <- round(sapply(rsover_nn, function(x) mean(x$dist))) / 1000

cols_tr <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0, 0.75, 1), rgb(1, 0.75, 0))
plot(1:4, cex = 3, pch = 20, col = cols_tr)

# impact table

# For Panel B - impact statistics 
# bpath <- fp(p_out, ybdc$both$bcode)
# load(fp(bpath, "impacts_tab.rda"))  # get impacts table

# select out corresponding impact statistics, summarize
conv_stats <- selected_impacts(pure, fp(p_out, ybdc$both$bcode))
impsr <- lapply(c(2, 5, 3, 7:9, 4), function(x) conv_stats[[x]])

# Plot setup
m1 <- do.call(cbind, lapply(1:4, function(x) matrix(rep(x, 35), 5, 7)))
m2 <- matrix(rep(5, 20), 4, 5)
m3 <- do.call(cbind, lapply(6:12, function(x) matrix(rep(x, 12), 4, 3)))
lmat <- rbind(m1, cbind(rep(5, 4), m2, m3, rep(13, 4)))

# plot(1:10, ylab = ylabs[2])
# ylabs <- c("Jaccard Index", 
ylabs <- c(expression(paste("Converted area (", 10^5, " ", km^2, ")")), 
           "Hours", "C loss (Mt)", "Priority", "Intactness/priority",
           "Intactness", expression(paste("FR converted (ha"^2, ")")))
# yscale <- c(1, 10^5, 1, 100000, 1, 1, 1, 1)
yscale <- c(10^5, 1, 100000, 1, 1, 1, 1)
ftits <- list("Y 100%", "C 100%", "Biod 100%", "Cost 100%")
leg <- c("Converted")

yaxes <- list(seq(0, 14, 2), seq(0, 7, 0.5),               
              seq(0, 1100, 100), seq(0, 0.12, 0.02), seq(0, 0.6, 0.1), 
              seq(0, 1, 0.1), seq(0, 6500, 500))
bcols <- c(list(c(cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2], 
                  cols_tr[3])), lapply(1:7, function(x) rep("black", 4)))

# plot
tiff(fp(p_figures, "single_priorities2.tif"), height = 3.5, width = 7, 
     units = "in", res = 300)
l <- layout(mat = lmat)
# layout.show(l)

# map plots
for(i in 1:4) { # i <- 1
  imp_plot2(rs2[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
            crops = leg, ftitle = ftits[[i]], col = cols_tr[[i]], 
            legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.4, 
            marg = rep(0, 4))
}

# impact plots
mgp <- c(1.6, 0.25, 0)
par(mar = c(4, 3, 0, 0))
with(pms2b, {
  plot(NND, JS, col = cols_tr[c(2, 3, 3, 4, 4, 4)], pch = 16, cex = 1.5, 
       axes = FALSE, xlab = "", ylab = "", ylim = c(0, 0.3))
  points(NND, JS, col = cols_tr[c(1, 1, 2, 1, 2, 3)], pch = 20, cex = 1)
  axis(1, at = seq(0, 230, 20), cex.axis = 0.6, las = 2, mgp = mgp, 
       tcl = -0.2)
  axis(2, at = seq(0, 0.3, 0.05), cex.axis = 0.6, las = 2, mgp = mgp, 
       tcl = -0.2)
  mtext(text = "NN Distance (km)", side = 1, line = 1, cex = 0.45)
  mtext(text = "Jaccard Index", side = 2, line = 1.25, cex = 0.45)
})
mgp <- c(1, 0.25, 0)
for(i in 1:7) {
  v <- impsr[[i]] / yscale[i]
  b <- barplot(impsr[[i]] / yscale[i], las = 2,  col = cols_tr, 
               mgp = c(1.5, 0.25, 0), tcl = -0.2, yaxt = "n", 
               ylim = range(yaxes[[i]]), ylab = ylabs[i], cex.axis = 0.5,
               cex.lab = 0.7)
  axis(2, at = yaxes[[i]], cex.axis = 0.6, las = 2, mgp = mgp, 
       tcl = -0.2)
}
dev.off()

```

<a href="#top">Back to top</a>

### Figures 2 & 3 (probably supplemental)
```{r, eval = FALSE}
# individual maps
png(full_path(p_figures, "all-4-maps.png"), height = 7, width = 7, 
     units = "in", res = 300)
par(mfrow = c(2, 2))#, oma = c(0, 0, 0, 2))
ftitles <- c("Yield", "Carbon", "Biodiversity", "Cost")
legp <- c(TRUE, rep(FALSE, 3))
for(i in 1:4) {
  dtr <- calc(pure_r[[i]][[1:2]] * c(1, 2), sum)
  imp_plot2(dtr, impacts = i1, fpath = p_figures, fnm = "all-4-maps", 
            cexscale = 0.6, ftitle = ftitles[i], crops = c("maize", "soy"), 
            stats = FALSE, hold = TRUE, legend = legp[i], marg = c(0, 0, 2, 0))
}
dev.off()

# heat maps
implist <- list(ycost$both, bdc$both)
load(full_path(set_base_path(), paste0("external/data/ZA/parks_roads.rda")))
pacols <- c("transparent", "grey85", "grey75")
png(full_path(p_figures, "ycost-bdc-heat.png"), height = 400, width = 900)
par(mfrow = c(1, 2), oma = c(0, 0, 0, 2))
for(i in 1:2) {
  par(mar = c(1, 1, 1, 4))
  plot(zambia, col = "grey90", border = "transparent")
  ir <- implist[[i]]$heat
  ir[ir == 0] <- NA
  plot(ir, col = rev(heat.colors(20)), add = TRUE, )
  plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2],
       border = FALSE)
  plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3],
       border = FALSE)
  plot(roads, add = TRUE, lwd = 0.1, col = "grey")
}
dev.off()

```

<a href="#top">Back to top</a>

### Figures 4 (main) and 5 (probably supplemental)

What happens when weight is progressively added

```{r, eval = FALSE}
p <- data.frame(ybdc$both$params)
p[, c("Y", "C", "BD" ,"COST")] <- round(p[, c("Y", "C", "BD" ,"COST")], 2)
scens <- lapply(c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9), function(x) {
  c(1 - x, rep(x / 2, 2), 0)
})
pind <- sapply(scens, function(x) {
  x <- round(x, 2)
  which(p$Y == x[1] & p$C == x[2] & p$BD == x[3] & p$COST == x[4])
})

# select out impacts corresponding to those runs
cstats <- selected_impacts(pind, fp(p_out, ybdc$both$bcode))

# check on selection
# il <- input_handler(p_moddat, crops[[1]])
# chk <- tradeoff_mod(prod_targ = targ, cbetas = ybdc$both$params[pind[5], 7:10], 
#                     currprodmod = g, exist_list = il)
# DF <- data.frame(chk$impacts)
# round(c(colSums(DF[, 1:2]),
#         sapply(4:ncol(DF), function(z) weighted.mean(DF[, z], DF$conv_area))),2)
# cstats[5, ]
# fnames <- dir(bpath, pattern = "csv")
# fnames[pind[5]]

# calculate max impacts across both crops, and figure out the impact range, 
# and what percent that is relative to the worst case
allstats <- selected_impacts(1:nrow(p), fp(p_out, ybdc$both$bcode))
minmaxstats <- allstats[, lapply(.SD, range)]
imp_range <- minmaxstats[, lapply(.SD, diff)]  # difference between best/worst
pct_of_tot <- minmaxstats[, lapply(.SD, function(x) diff(x) / max(x))]

dfunc <- function(x, y) 100 - y / x * 100
stnms <- names(cstats)[-1]
imp_pct <- cstats[, lapply(stnms, function(x) {
  dfunc(minmaxstats[[x]][2], get(x))
})]
setnames(imp_pct, stnms)

# plot(imp_pct$tot_C, imp_pct$conv_area, type = "l", xlab = "", ylab = "", 
#      axes = FALSE)        # xaxs = "r", yaxs = "r")

sctab <- do.call(rbind, scens) * 100
cx <- 0.7
lcols <- c("black", "red", "blue")
imp_pct[, {
  png(fp(p_figures, "weight_impact2.png"), height = 4, width = 4, 
      units = "in", res = 300)
  par(mgp = c(2, 0.75, 0), mar = c(3, 3, 1, 1))
  plot(conv_area, tot_C, type = "l", xlab = "", ylab = "", #xaxt = "n", 
       axes = FALSE, ylim = c(0, 65), xlim = c(2, 12))#, xaxs = "r", yaxs = "r")
  lines(conv_area, int_prior, col = "red")
  lines(conv_area, mu_cost, col = "blue")
  axis(1, at = seq(2, 12, 1), cex = cx, cex.axis = cx, line = -0.25, 
       tcl = -0.2, mgp = c(1, 0.25, 0))
  axis(2, at = seq(0, 60, 5), las = 2, cex = cx, cex.axis = cx, line = -0.25, 
       tcl = -0.15, mgp = c(1, 0.25, 0))
  mtext("% land savings", side = 1, line = 1, cex = cx)
  mtext("% Carbon/Biodiv/Cost savings", side = 2, line = 1, cex = cx)
  inms <- c("tot_C", "int_prior", "mu_cost")
  for(i in 1:length(inms)) {
    text(conv_area[-1], get(inms[i])[-1], labels = sctab[-1, i + 1], 
         cex = cx * 0.8, pos = 3, offset = 0.2, col = lcols[i])
    text(conv_area[1] + 0.1, i, labels = 0, cex = cx * 0.8, pos = 3, 
         offset = i * 0.15 - 0.2, col = lcols[i])
  }
  legend("bottomleft", legend = c("Carbon", "Biodiv", "Cost"), cex = cx,
         text.col = lcols, col = lcols, bty = "n", lty = 1)
  dev.off()
}]

fnames <- fp(fp(p_out, ybdc$both$bcode), 
             imp_sorter(fp(p_out, ybdc$both$bcode), inds = pind))
scens_r <- lapply(fnames, function(x) {  # x <- pind[[1]]
  print(x)
  DT <- cbind(fread(x), basedt[, .(x, y, convertible)])
  DTr <- dt_to_raster(DT, CRSobj)
})

scens_rs2 <- lapply(scens_r, function(x) calc(stack(x[[1:2]]), sum))
names(scens_rs2) <- paste0("w", 1:length(scens_rs2))

pms2b <- do.call(rbind.data.frame, lapply(2:6, function(x) {
  a <- rep(0, 6)
  a[c(1, x)] <- 1
  a
}))
colnames(pms2b) <- paste0("Var", 1:ncol(pms2b))
pms2b$cl <- c("W010", "W020", "W030", "W040", "W050")
rownames(pms2b) <- 1:5

# pairwise overlaps
scens_over <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
  calc(stack(scens_rs2[which(pms2b[x, ] == 1)]), sum)
})
names(scens_over) <- pms2b$cl

# Jaccard Index of overlaps
scens_over_a <- sapply(1:length(scens_over), function(x) { # x <- 4
  sapply(1:2, function(y) {  # y <- 3
    r <- scens_over[[x]] == y
    round(cellStats(r * pure_r[[1]]$convertible, sum), 1)
  })
})
pms2b$JS <- round(scens_over_a[2, ] / colSums(scens_over_a), 3)

# calculate nearest neighbor distance between each version, using spatstat 
# nncross
zam_owin <- owin(xrange = bbox(zambia)[1, ], yrange = bbox(zambia)[2, ],
                 unitname = "meter")
which1 <- function(x) x == 1  # selector function for r to xy
scenover_nn <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests; x = 4
  print(x)
  ind <- which(pms2b[x, 1:6] == 1)
  rl <- scens_rs2[ind]
  rxy <- lapply(rl, function(x) rasterToPoints(x, fun = which1))
  pps <- lapply(rxy, function(x) {
    ppp(x = x[, 1], y = x[, 2], window = zam_owin)
  })
  # reorder to find which has most points, to put that in as X
  reord <- sort(sapply(pps, function(x) x$n), decreasing = TRUE)
  ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
})
pms2b$NND <- round(sapply(scenover_nn, function(x) mean(x$dist))) / 1000

cols_tr <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0, 0.75, 1), rgb(1, 0.75, 0))
plot(1:4, cex = 3, pch = 20, col = cols_tr)

# plot(scens_rs2[[1]] + scens_rs2[[2]])
# check
# round(sapply(scens_rs2, function(x) {
#   cellStats(x * scens_r[[1]]$convertible * 100, sum)
# }))
# cstats$conv_area

# pms <- list(c(1, 2), c(1, 3), c(1, 4), c(1, 5))
# pms <- expand.grid(rep(list(0:1), 5))[-1, ]  # logic grid

# rasters showing overlap as weight is progressively put on C and BD
scens_over2 <- lapply(2:6, function(x) scens_rs2[[1]] + scens_rs2[[x]] * 2)

# # overlap areas
# scens_over_a <- sapply(1:length(scens_over), function(x) { # x <- 4
#   sapply(1:3, function(y) {  # y <- 3
#     r <- scens_over[[x]] == y
#     # plot(r)
#     round(cellStats(r * scens_r[[1]]$convertible, sum), 1)
#   })
# })
# colSums(scens_over_a[c(2, 3), ])
# over_cols <- sapply(seq(0, 0.4, 0.1), function(x) rgb(1 - x, 0, x))
# par(mar = c(0, 0, 0, 0))
# plot(1:3, pch = 20, col = rgb(1, 0, 1))
# plot(zambia)

# cols <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0.75, 0, 0.75))
# cols <- c("red", "blue", "gold")
bluecols <- sapply(c(1, 0.75, 0.5, 0.25, 0), function(x) rgb(0, x, 1))
bluecols[5] <- rgb(0.4, 0, 1)
cols <- sapply(1:length(bluecols), function(x) c("red", bluecols[x], "gold"))
plot(1:ncol(cols), pch = 20, col = cols[2, ], cex = 20)
ftits <- list("Y 90% C 5% Biod 5% Cost 0%",
              "Y 80% C 10% Biod 10% Cost 0%",
              "Y 70% C 15% Biod 15% Cost 0%",
              "Y 60% C 20% Biod 20% Cost 0%", 
              "Y 50% C 25% Biod 25% Cost 0%")
leg <- c("Y", "YCBd", "Y + YCBd")

tiff(fp(p_figures, "changing_weights2.tif"), height = 3.5, width = 7, 
     units = "in", res = 300)
par(mfrow = c(2, 3), oma = c(0, 0, 0, 1))
for(i in 1:5) { # i <- 1
  # map
  imp_plot2(scens_over2[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
            crops = leg, ftitle = ftits[[i]], col = cols[, i], 
            legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.4, 
            marg = c(0, 0, 1, 0), titline = 0)
}
mgp <- c(1, 0.25, 0)
par(mar = c(3, 3, 1, 0))
with(pms2b, {
     plot(NND, JS, col = cols[2, ], pch = 16, cex = 1.5, 
          axes = FALSE, xlab = "", ylab = "", ylim = c(0, 0.4), 
          xlim = c(0, 180))
     points(NND, JS, col = cols[[1]][1], pch = 20, cex = 1)
     axis(1, at = seq(0, 180, 20), cex.axis = 0.6, las = 2, mgp = mgp, 
          tcl = -0.2)
     axis(2, at = seq(0, 1, 0.1), cex.axis = 0.6, las = 2, mgp = mgp, 
          tcl = -0.2)
     mtext(text = "NN Distance (km)", side = 1, line = 1, cex = 0.45)
     mtext(text = "Jaccard Index", side = 2, line = 1, cex = 0.45)
})
dev.off()

#   # barplot
#   par(mar = c(5, 4, 3, 3))
#   overl <- scens_over_a[1:3, i][c(1, 3, 2)]
#   bp <- barplot(overl, las = 2, ylim = c(0, 12000),  
#                 col = cols[c(1, 3, 2)], mgp = c(1.5, 0.25, 0), tcl = -0.2, 
#                 ylab = expression(km^2), cex.axis = 0.6, cex.lab = 0.7)
#   axis(side = 1, at = bp, labels = leg, las = 2, 
#        mgp = c(1, 0.25, 0), tcl = -0.2, cex.axis = 0.6)
# }
# dev.off()

```
<a href="#top">Back to top</a>

### Figure 6

Objective weighting of outcomes, based on range of potential impacts

```{r, eval = FALSE}
# objective weights
obj_wgts <- round(pct_of_tot[, {
  c(conv_area, tot_C, int_mean, mu_cost) / 
    sum(c(conv_area, tot_C, int_mean, mu_cost))
}], 2)
# round(obj_wgts, 2)
names(obj_wgts) <- cnames

# Objective weights
toobj <- tradeoff_mod(targ, cbetas = obj_wgts, currprodmod = g, it = "OBJ")

# ybdc$both$params[1, ]
# yp <- ybdc$both$params[, 7:10]
# which(round(yp[, 1], 2) == 0.05 & round(yp[, 2], 2) == 0.30 & 
#         round(yp[, 3], 2) == 0.25 & round(yp[, 4], 2) == 0.4)

# objective weights on ag only 
# considering that these might be two sides of the same coin
ag_wgts <- round(obj_wgts[c(1, 4)] / sum(obj_wgts[c(1, 4)]), 2)
ag_wgts <- c(ag_wgts[1], "C" = 0, "BD" = 0, ag_wgts[2])
ag_wgts2 <- ag_wgts[c(4, 2:3, 1)]
names(ag_wgts2) <- names(ag_wgts)

toag <- tradeoff_mod(targ, cbetas = ag_wgts, currprodmod = g, it = "AGOBJ")
toag2 <- tradeoff_mod(targ, cbetas = ag_wgts2, currprodmod = g, it = "AGOBJ2")

# subjective weights, where we want to treat everyone equally
subj_wgts <- rep(0.25, 4)
names(subj_wgts) <- cnames  # rename to make sure model runs properly
tosubj <- tradeoff_mod(targ, cbetas = subj_wgts, currprodmod = g, it = "SUBJ")

# extract impacts of these, and add in BAU case
tores <- lapply(list(toag, toag2, toobj, tosubj), function(x) {  # x <- toobj
  DTsum <- data.table(x$conv[, rowSums(.SD), .SDcols = c(crops[[1]])])
  DTspat <- cbind(basedt[, .(x, y)], DTsum)
  DTr <- dt_to_raster(DTspat, CRSobj)
  DTr <- DTr + (fbr * 2)  # overlap with farmblocks, for output
  list("map" = DTr, "imp" = x$impacts)
})
names(tores) <- c("ag", "ag2", "obj", "subj")

# Calculate area of overlaps between farmblocks and 
yfbr <- calc(mzsoytor, sum) + fbr * 2
plot(yfbr)
fb_over_a <- sapply(1:length(tores), function(x) { # x <- 4
  sapply(1:3, function(y) {  # y <- 3
    r <- tores[[x]]$map == y
    round(cellStats(r * scens_r[[1]]$convertible, sum), 1)
  })
})
fb_sim <- fb_over_a[3, ] / colSums(fb_over_a)

# grid for calculating overlaps
pms2a <- expand.grid(rep(list(0:1), 4))[-1, ] 
pms2b <- pms2a[which(rowSums(pms2a) == 2), ]
pms2b$cl <- c("AgAg2", "AgObj", "Ag2Obj", "AgSubj", "Ag2Subj", "ObjSubj")

toresr <- lapply(tores, function(x) {
  r <- x$map
  r[r == 2] <- 0
  r[r == 3] <- 1
  r
})

# tores_over_r <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
#   (toresr[[1]] == pms2b[x, 1]) & (toresr[[2]] == pms2b[x, 2]) & 
#     (toresr[[3]] == pms2b[x, 3]) & (toresr[[4]] == pms2b[x, 4])
# })
# names(tores_over_r) <- pms2b$cl
# tores_overs <- stack(tores_over)
tores_over_r <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
  calc(stack(toresr[which(pms2b[x, ] == 1)]), sum)
})
names(tores_over_r) <- pms2b$cl

# jaccard's similarity, per https://gist.github.com/jlehtoma/3369793
tores_over_a <- sapply(1:length(tores_over_r), function(x) { # x <- 4
  sapply(1:2, function(y) {  # y <- 3
    r <- tores_over_r[[x]] == y
    round(cellStats(r * scens_r[[1]]$convertible, sum), 1)
  })
})
pms2b$JS <- round(tores_over_a[2, ] / colSums(tores_over_a), 3)

# calculate nearest neighbor distance between each version, using spatstat 
# nncross
# plot(zam_owin)
zam_owin <- owin(xrange = bbox(zambia)[1, ], yrange = bbox(zambia)[2, ],
                 unitname = "meter")
which1 <- function(x) x == 1  # selector function for r to xy
tores_nn <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests; x = 2
  ind <- which(pms2b[x, 1:4] == 1)
  rl <- toresr[ind]
  rxy <- lapply(rl, function(x) rasterToPoints(x, fun = which1))
  pps <- lapply(rxy, function(x) {
    ppp(x = x[, 1], y = x[, 2], window = zam_owin)
  })
  # reorder to find which has most points, to put that in as X
  reord <- sort(sapply(pps, function(x) x$n), decreasing = TRUE)
  ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
})
pms2b$NND <- round(sapply(tores_nn, function(x) mean(x$dist))) / 1000
# with(pms2b, plot(NND, JS))

# do sum factoring in overlaps 
# sapply(tores_over_r, function(x) {
#   round(cellStats(x * scens_r[[1]]$convertible, sum), 1)
# })

cols_tr <- c(rgb(1, 0, 0), rgb(1, 0.75, 0), rgb(0, 0, 1), rgb(0, 0.75, 1))
plot(1:4, cex = 3, pch = 20, col = cols_tr)

# impact table
imp_list <- list(toag$impacts, toag2$impacts, toobj$impacts, tosubj$impacts)
imps <- rbindlist(lapply(imp_list, function(x) selected_impacts2(x)))
impsr <- lapply(c(1, 4, 2, 6:8, 3), function(x) imps[[x]])

# layout matrix
m1 <- do.call(cbind, lapply(1:4, function(x) matrix(rep(x, 35), 5, 7)))
m2 <- matrix(rep(5, 20), 4, 5)
m3 <- do.call(cbind, lapply(6:12, function(x) matrix(rep(x, 12), 4, 3)))
lmat <- rbind(m1, cbind(rep(5, 4), m2, m3, rep(13, 4)))

# plot(1:10, ylab = ylabs[2])
# ylabs <- c("Jaccard Index", 
ylabs <- c(expression(paste("Converted area (", 10^5, " ", km^2, ")")), 
           "Hours", "C loss (Mt)", "Priority", "Intactness/priority",
           "Intactness", expression(paste("FR converted (ha"^2, ")")))
# yscale <- c(1, 10^5, 1, 100000, 1, 1, 1, 1)
yscale <- c(10^5, 1, 100000, 1, 1, 1, 1)
ftits <- list("Y 11% C 0% Biod 0% Cost 89%",
              "Y 89% C 0% Biod 0% Cost 11%",
              "Y 5% C 29% Biod 25% Cost 41%",
              "Y 25% C 25% Biod 25% Cost 25%")
leg <- c("Converted")

# yaxes <- list(seq(0, 0.8, 0.1), seq(0, 14, 2), seq(0, 2.5, 0.5), 
yaxes <- list(seq(0, 14, 2), seq(0, 2.5, 0.5),               
              seq(0, 1000, 200), seq(0, 0.2, 0.05), seq(0, 0.5, 0.1), 
              seq(0, 0.8, 0.2), seq(0, 3500, 500))
# bcols <- c(list(c(cols_tr[1], cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2],
bcols <- c(list(c(cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2], 
                cols_tr[3])), lapply(1:7, function(x) rep("black", 4)))
# lcols <- c(list(c(cols_tr[2], cols_tr[3], cols_tr[3], cols_tr[4], cols_tr[4] 
# lcols <- lapply(1:7, function(x) cols_tr)

tiff(fp(p_figures, "objective_weights.tif"), height = 3.5, width = 7, 
     units = "in", res = 300)
l <- layout(mat = lmat)
# layout.show(l)

# map plots
for(i in 1:4) { # i <- 1
  # map
  # plot(zambia)
  imp_plot2(toresr[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
            crops = leg, ftitle = ftits[[i]], col = cols_tr[[i]], 
            legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.4, 
            marg = rep(0, 4))
}

# impact plots
mgp <- c(1, 0.25, 0)
par(mar = c(4, 3, 0, 0))
with(pms2b, {
     plot(NND, JS, col = cols_tr[c(2, 3, 3, 4, 4, 4)], pch = 16, cex = 1.5, 
          axes = FALSE, xlab = "", ylab = "", ylim = c(0, 1))
     points(NND, JS, col = cols_tr[c(1, 1, 2, 1, 2, 3)], pch = 20, cex = 1)
     axis(1, at = seq(0, 130, 10), cex.axis = 0.6, las = 2, mgp = mgp, 
          tcl = -0.2)
     axis(2, at = seq(0, 1, 0.1), cex.axis = 0.6, las = 2, mgp = mgp, 
          tcl = -0.2)
     mtext(text = "NN Distance (km)", side = 1, line = 1, cex = 0.45)
     mtext(text = "Jaccard Index", side = 2, line = 1, cex = 0.45)
})
for(i in 1:7) {
  v <- impsr[[i]] / yscale[i]
  b <- barplot(impsr[[i]] / yscale[i], las = 2,  col = cols_tr, 
               mgp = c(1.5, 0.25, 0), tcl = -0.2, yaxt = "n", 
               ylim = range(yaxes[[i]]), ylab = ylabs[i], cex.axis = 0.5,
               cex.lab = 0.7)
  axis(2, at = yaxes[[i]], cex.axis = 0.6, las = 2, mgp = mgp, 
       tcl = -0.2)
}
dev.off()

# ybdc$both$optitab[, {
#   par(mfrow = c(2, 2), mar = rep(3, 4))
#   hist(.SD, axes = FALSE, labels = FALSE)
# }, .SDcol = 1:4]
# 
# frange <- function(x) diff(range(x)) / max(x)
# fqtile <- function(x) diff(quantile(x, probs = c(0.025, 0.975))) / max(x) 
# ybdc$both$optimal[, lapply(.SD, frange), .SDcol = 1:4]
# ybdc$both$optimal[, lapply(.SD, fqtile), .SDcol = 1:4]
# 
# hist(ybdc$both$optimal$Y)

```