### 2016-02-24 Rmd/figures.Rmd
Updated results with new figure 3. Fixed small mistake in Figure 2, which was to select impact range from optimized results, not all results (very tiny difference).

---
title: "Main Analyses"
author: "Lyndon Estes"
date: "February 11, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Main Analyses

```{r, warning=FALSE, message=FALSE}
library(agroEcoTradeoff)
library(readxl)
library(gdalUtils)
library(gisutils)
# URL <- paste0("https://cran.r-project.org/src/contrib/Archive/spatstat/",
#               "spatstat_1.42-2.tar.gz")
# install.packages(URL, repos = NULL, type="source", dependencies = TRUE)
library(spatstat)

setwd("~/Dropbox/publications/zambiaToff/private/zambiaToff/")
p_root <- getwd()
p_demand <- fp(p_root, "external/input_devel/demand/")
p_mod <- fp(p_root, "agroEcoTradeoff")
p_moddat <- fp(p_mod, "external/data/ZA")
p_figures <- full_path(p_root, "paper/figures/")
p_other <- fp(p_root, "external/input_devel/other/")
# p_roads <- fp(p_root, "external/input_devel/roads")
# p_roads2 <- paste0("/Users/lestes/Dropbox/data/zari/infrastructure/zambia/", 
#                    "rda_dump/Final Shape files RDA/")
p_out <- fp(p_mod, "external/output")
```

```{r, eval = FALSE}
# roads
mgrid <- raster(fp(p_moddat, "ZA-mask.tif"))
load(fp(p_mod, "external/data/ZA/parks_roads.rda"))

# farmblocks
fbp <- "/Users/lestes/Dropbox/data/zari/misc/NEW4GEO/Fblocks.sqlite"
# fb <- readOGR(fbp, layer = "fblocks")
# fbalb <- spTransform(fb, mgrid@crs)
# fbalb$ID <- 1:nrow(fbalb)
# fbalb@data <- fbalb@data[, "ID", drop = FALSE]
fnm <- fp(p_other, "fblocks_alb.sqlite")
# writeOGR(fbalb, dsn = fnm, layer = "fblocks_alb", driver = "SQLite")

# read in farmblocks
fbalb <- readOGR(fnm, layer = "fblocks_alb")
fbr <- rasterize(fbalb, mgrid)
fbr[is.na(fbr)] <- 0
fbr[fbr > 0] <- 1
fbr <- raster::mask(fbr, mgrid)
# plot(fbr)

```

<a href="#top">Back to top</a>

## Tradeoff runs

### Best areas to meet targets for each crop

Running the model individually for each crop and for both crops.
```{r, eval = FALSE}
setwd(p_mod)
path <- "external/data/ZA"

# gaps 
cp <- c("maize" = 2912004, "soy" = 203732)  # current production
gapper <- function(cp, gaps, pct) ((((cp * gaps) - cp) * pct) + cp) / cp
mzgap <- 4.43 / 2.3 
soygap <- 3.4 / 2
g <- gapper(cp, c(mzgap, soygap), 0.25)  # assume 50% closed on current land
targ <- c("maize" = 3, "soy" = 9)  # targets given g
targs <- targets(targ, cp, g)

# Separate soy and maize impacts
cb <- c("Y" = 1, "C" = 0, "BD" = 0, "COST" = 0)
mzto <- tradeoff_mod(prod_targ = targ["maize"], cbetas = cb, currprodmod = g[1])
soyto <- tradeoff_mod(prod_targ = targ["soy"], cbetas = cb, currprodmod = g[2])
mzsoyto <- tradeoff_mod(prod_targ = targ, cbetas = cb, currprodmod = g)

# rasters
CRSobj <- CRS(mzto$inputs$sp$crs)
mztor <- dt_to_raster(mzto$conv, CRSobj)
soytor <- dt_to_raster(soyto$conv, CRSobj)
mzsoytor <- dt_to_raster(mzsoyto$conv, CRSobj)  # converges on same results

# rasters
mzsoyr <- stack(list(mztor, soytor * 2))
mzsoyr <- calc(mzsoyr, sum)
# plot(mzsoyr)
# plot(roads)

# Plot
i1 <- rbind(mzto$impacts, soyto$impacts)
crops <- c("maize", "soy", "maize/soy")
# plot(1:10, pch = 20, cex = 4, col = "gold2")
# plot(mzsoyr, col = c("grey", "goldenrod", "green4", "orange"))
imp_plot2(mzsoyr, i1, p_figures, "ideal_yield", "", crops, 1, TRUE, 
         c("gold2", "green3", "orange"))

basedt <- soyto$inputs$mask
crops <- list(c("maize", "soy"))

```
<a href="#top">Back to top</a>

```{r, eval = FALSE}
load(fp(p_out, "mainresults1702.rda"))
```

<a href="#top">Back to top</a>

## Figures

How reconcilable are the different land use interests, and what would be the impacts of each?

I am using the two crops combined approach because there is little overlap between the two.

### Figure 1
```{r, eval = FALSE}
# select out results for "pure" cases (weight = 1)
cnames <- c("Y", "C", "BD", "COST")
pure <- sapply(cnames, function(x) which(round(ybdc$both$params[, x], 4) == 1))

fnames <- imp_sorter(fp(p_out, ybdc$both$bcode), inds = pure, fullnames = TRUE)
pure_r <- lapply(fnames, function(x) {  # x <- pure[1]
  DT <- cbind(fread(x), basedt[, .(x, y, convertible)])
  DTr <- dt_to_raster(DT, CRSobj)
})

# sum across crops and multiply by convertible areas
rs <- lapply(pure_r, function(x) calc(x[[1:2]], sum) * x[[3]])  # actual area
rs2 <- lapply(pure_r, function(x) calc(x[[1:2]], sum))  # conv pixels
names(rs2) <- cnames
# check
# sapply(rs, function(x) cellStats(x, sum))
# ybdc$both$optitab[ind %in% pure, Y]  # same, but C optimized out
# sapply(rs2, function(x) cellStats(x, sum))
# plot(pure_r[[3]])

# figure out overlaps between different pure objectives
pms2a <- expand.grid(rep(list(0:1), 4))[-1, ] 
pms2b <- pms2a[which(rowSums(pms2a) == 2), ]
pms2b$cl <- c("YC", "YBd", "CBd", "YCt", "CCt", "BdCt")
rownames(pms2b) <- 1:6

# pairwise overlaps
rs_over <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
  calc(stack(rs2[which(pms2b[x, ] == 1)]), sum)
})
names(rs_over) <- pms2b$cl

# Jaccard Index of overlaps
rs_over_a <- sapply(1:length(rs_over), function(x) { # x <- 4
  sapply(1:2, function(y) {  # y <- 3
    r <- rs_over[[x]] == y
    round(cellStats(r * pure_r[[1]]$convertible, sum), 1)
  })
})
pms2b$JS <- round(rs_over_a[2, ] / colSums(rs_over_a), 3)

# calculate nearest neighbor distance between each version, using spatstat 
# nncross
zam_owin <- owin(xrange = bbox(zambia)[1, ], yrange = bbox(zambia)[2, ],
                 unitname = "meter")
which1 <- function(x) x == 1  # selector function for r to xy
rsover_nn <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests; x = 2
  ind <- which(pms2b[x, 1:4] == 1)
  rl <- rs2[ind]
  rxy <- lapply(rl, function(x) rasterToPoints(x, fun = which1))
  pps <- lapply(rxy, function(x) {
    ppp(x = x[, 1], y = x[, 2], window = zam_owin)
  })
  # reorder to find which has most points, to put that in as X
  reord <- sort(sapply(pps, function(x) x$n), decreasing = TRUE)
  ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
})
pms2b$NND <- round(sapply(rsover_nn, function(x) mean(x$dist))) / 1000

cols_tr <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0, 0.75, 1), rgb(1, 0.75, 0))
plot(1:4, cex = 3, pch = 20, col = cols_tr)

# impact table

# For Panel B - impact statistics 
# bpath <- fp(p_out, ybdc$both$bcode)
# load(fp(bpath, "impacts_tab.rda"))  # get impacts table

# select out corresponding impact statistics, summarize
conv_stats <- selected_impacts(pure, fp(p_out, ybdc$both$bcode))
impsr <- lapply(c(2, 5, 3, 7:9, 4), function(x) conv_stats[[x]])

# Plot setup
m1 <- do.call(cbind, lapply(1:4, function(x) matrix(rep(x, 35), 5, 7)))
m2 <- matrix(rep(5, 20), 4, 5)
m3 <- do.call(cbind, lapply(6:12, function(x) matrix(rep(x, 12), 4, 3)))
lmat <- rbind(m1, cbind(rep(5, 4), m2, m3, rep(13, 4)))

# plot(1:10, ylab = ylabs[2])
# ylabs <- c("Jaccard Index", 
ylabs <- c(expression(paste("Converted area (", 10^5, " ", km^2, ")")), 
           "Hours", "C loss (Mt)", "Priority", "Intactness/priority",
           "Intactness", expression(paste("FR converted (ha"^2, ")")))
# yscale <- c(1, 10^5, 1, 100000, 1, 1, 1, 1)
yscale <- c(10^5, 1, 100000, 1, 1, 1, 1)
ftits <- list("Y 100%", "C 100%", "Biod 100%", "Cost 100%")
leg <- c("Converted")

yaxes <- list(seq(0, 14, 2), seq(0, 7, 0.5),               
              seq(0, 1100, 100), seq(0, 0.12, 0.02), seq(0, 0.6, 0.1), 
              seq(0, 1, 0.1), seq(0, 6500, 500))
bcols <- c(list(c(cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2], 
                  cols_tr[3])), lapply(1:7, function(x) rep("black", 4)))

# plot
tiff(fp(p_figures, "single_priorities2.tif"), height = 3.5, width = 7, 
     units = "in", res = 300)
l <- layout(mat = lmat)
# layout.show(l)

# map plots
for(i in 1:4) { # i <- 1
  imp_plot2(rs2[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
            crops = leg, ftitle = ftits[[i]], col = cols_tr[[i]], 
            legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.4, 
            marg = rep(0, 4))
}

# impact plots
mgp <- c(1.6, 0.25, 0)
par(mar = c(4, 3, 0, 0))
with(pms2b, {
  plot(NND, JS, col = cols_tr[c(2, 3, 3, 4, 4, 4)], pch = 16, cex = 1.5, 
       axes = FALSE, xlab = "", ylab = "", ylim = c(0, 0.3))
  points(NND, JS, col = cols_tr[c(1, 1, 2, 1, 2, 3)], pch = 20, cex = 1)
  axis(1, at = seq(0, 230, 20), cex.axis = 0.6, las = 2, mgp = mgp, 
       tcl = -0.2)
  axis(2, at = seq(0, 0.3, 0.05), cex.axis = 0.6, las = 2, mgp = mgp, 
       tcl = -0.2)
  mtext(text = "NN Distance (km)", side = 1, line = 1, cex = 0.45)
  mtext(text = "Jaccard Index", side = 2, line = 1.25, cex = 0.45)
})
mgp <- c(1, 0.25, 0)
for(i in 1:7) {
  v <- impsr[[i]] / yscale[i]
  b <- barplot(impsr[[i]] / yscale[i], las = 2,  col = cols_tr, 
               mgp = c(1.5, 0.25, 0), tcl = -0.2, yaxt = "n", 
               ylim = range(yaxes[[i]]), ylab = ylabs[i], cex.axis = 0.5,
               cex.lab = 0.7)
  axis(2, at = yaxes[[i]], cex.axis = 0.6, las = 2, mgp = mgp, 
       tcl = -0.2)
}
dev.off()

```

<a href="#top">Back to top</a>

### Figures 2 & 3 (probably supplemental)
```{r, eval = FALSE}
# individual maps
png(full_path(p_figures, "all-4-maps.png"), height = 7, width = 7, 
     units = "in", res = 300)
par(mfrow = c(2, 2))#, oma = c(0, 0, 0, 2))
ftitles <- c("Yield", "Carbon", "Biodiversity", "Cost")
legp <- c(TRUE, rep(FALSE, 3))
for(i in 1:4) {
  dtr <- calc(pure_r[[i]][[1:2]] * c(1, 2), sum)
  imp_plot2(dtr, impacts = i1, fpath = p_figures, fnm = "all-4-maps", 
            cexscale = 0.6, ftitle = ftitles[i], crops = c("maize", "soy"), 
            stats = FALSE, hold = TRUE, legend = legp[i], marg = c(0, 0, 2, 0))
}
dev.off()

# heat maps
implist <- list(ycost$both, bdc$both)
load(full_path(set_base_path(), paste0("external/data/ZA/parks_roads.rda")))
pacols <- c("transparent", "grey85", "grey75")
png(full_path(p_figures, "ycost-bdc-heat.png"), height = 400, width = 900)
par(mfrow = c(1, 2), oma = c(0, 0, 0, 2))
for(i in 1:2) {
  par(mar = c(1, 1, 1, 4))
  plot(zambia, col = "grey90", border = "transparent")
  ir <- implist[[i]]$heat
  ir[ir == 0] <- NA
  plot(ir, col = rev(heat.colors(20)), add = TRUE, )
  plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2],
       border = FALSE)
  plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3],
       border = FALSE)
  plot(roads, add = TRUE, lwd = 0.1, col = "grey")
}
dev.off()

```

<a href="#top">Back to top</a>

### Figures 4 (main) and 5 (probably supplemental)

What happens when weight is progressively added

```{r, eval = FALSE}
p <- data.frame(ybdc$both$params)
pare <- ybdc$both$optitab
p[, c("Y", "C", "BD" ,"COST")] <- round(p[, c("Y", "C", "BD" ,"COST")], 2)
scens <- lapply(c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9), function(x) {
  c(1 - x, rep(x / 2, 2), 0)
})
pind <- sapply(scens, function(x) {
  x <- round(x, 2)
  which(p$Y == x[1] & p$C == x[2] & p$BD == x[3] & p$COST == x[4])
})

# select out impacts corresponding to those runs
cstats <- selected_impacts(pind, fp(p_out, ybdc$both$bcode))

# check on selection
# il <- input_handler(p_moddat, crops[[1]])
# chk <- tradeoff_mod(prod_targ = targ, cbetas = ybdc$both$params[pind[5], 7:10], 
#                     currprodmod = g, exist_list = il)
# DF <- data.frame(chk$impacts)
# round(c(colSums(DF[, 1:2]),
#         sapply(4:ncol(DF), function(z) weighted.mean(DF[, z], DF$conv_area))),2)
# cstats[5, ]
# fnames <- dir(bpath, pattern = "csv")
# fnames[pind[5]]

# calculate max impacts across both crops, and figure out the impact range, 
# and what percent that is relative to the worst case
allstats <- selected_impacts(pare$ind, fp(p_out, ybdc$both$bcode))
minmaxstats <- allstats[, lapply(.SD, range)]
imp_range <- minmaxstats[, lapply(.SD, diff)]  # difference between best/worst
pct_of_tot <- minmaxstats[, lapply(.SD, function(x) diff(x) / max(x))]

dfunc <- function(x, y) 100 - y / x * 100
stnms <- names(cstats)[-1]
imp_pct <- cstats[, lapply(stnms, function(x) {
  dfunc(minmaxstats[[x]][2], get(x))
})]
setnames(imp_pct, stnms)

# plot(imp_pct$tot_C, imp_pct$conv_area, type = "l", xlab = "", ylab = "", 
#      axes = FALSE)        # xaxs = "r", yaxs = "r")

sctab <- do.call(rbind, scens) * 100
cx <- 0.7
lcols <- c("black", "red", "blue")
imp_pct[, {
  png(fp(p_figures, "weight_impact.png"), height = 4, width = 4, 
      units = "in", res = 300)
  par(mgp = c(2, 0.75, 0), mar = c(3, 3, 1, 1))
  plot(conv_area, tot_C, type = "l", xlab = "", ylab = "", #xaxt = "n", 
       axes = FALSE, ylim = c(0, 65), xlim = c(2, 12))#, xaxs = "r", yaxs = "r")
  lines(conv_area, int_prior, col = "red")
  lines(conv_area, mu_cost, col = "blue")
  axis(1, at = seq(2, 12, 1), cex = cx, cex.axis = cx, line = -0.25, 
       tcl = -0.2, mgp = c(1, 0.25, 0))
  axis(2, at = seq(0, 60, 5), las = 2, cex = cx, cex.axis = cx, line = -0.25, 
       tcl = -0.15, mgp = c(1, 0.25, 0))
  mtext("% land savings", side = 1, line = 1, cex = cx)
  mtext("% Carbon/Biodiv/Cost savings", side = 2, line = 1, cex = cx)
  inms <- c("tot_C", "int_prior", "mu_cost")
  for(i in 1:length(inms)) {
    text(conv_area[-1], get(inms[i])[-1], labels = sctab[-1, i + 1], 
         cex = cx * 0.8, pos = 3, offset = 0.2, col = lcols[i])
    text(conv_area[1] + 0.1, i, labels = 0, cex = cx * 0.8, pos = 3, 
         offset = i * 0.15 - 0.2, col = lcols[i])
  }
  legend("bottomleft", legend = c("Carbon", "Biodiv", "Cost"), cex = cx,
         text.col = lcols, col = lcols, bty = "n", lty = 1)
  dev.off()
}]

fnames <- fp(fp(p_out, ybdc$both$bcode), 
             imp_sorter(fp(p_out, ybdc$both$bcode), inds = pind))
scens_r <- lapply(fnames, function(x) {  # x <- pind[[1]]
  print(x)
  DT <- cbind(fread(x), basedt[, .(x, y, convertible)])
  DTr <- dt_to_raster(DT, CRSobj)
})

scens_rs2 <- lapply(scens_r, function(x) calc(stack(x[[1:2]]), sum))
names(scens_rs2) <- paste0("w", 1:length(scens_rs2))

pms2b <- do.call(rbind.data.frame, lapply(2:6, function(x) {
  a <- rep(0, 6)
  a[c(1, x)] <- 1
  a
}))
colnames(pms2b) <- paste0("Var", 1:ncol(pms2b))
pms2b$cl <- c("W010", "W020", "W030", "W040", "W050")
rownames(pms2b) <- 1:5

# pairwise overlaps
scens_over <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
  calc(stack(scens_rs2[which(pms2b[x, ] == 1)]), sum)
})
names(scens_over) <- pms2b$cl

# Jaccard Index of overlaps
scens_over_a <- sapply(1:length(scens_over), function(x) { # x <- 4
  sapply(1:2, function(y) {  # y <- 3
    r <- scens_over[[x]] == y
    round(cellStats(r * pure_r[[1]]$convertible, sum), 1)
  })
})
pms2b$JS <- round(scens_over_a[2, ] / colSums(scens_over_a), 3)

# calculate nearest neighbor distance between each version, using spatstat 
# nncross
zam_owin <- owin(xrange = bbox(zambia)[1, ], yrange = bbox(zambia)[2, ],
                 unitname = "meter")
which1 <- function(x) x == 1  # selector function for r to xy
scenover_nn <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests; x = 4
  print(x)
  ind <- which(pms2b[x, 1:6] == 1)
  rl <- scens_rs2[ind]
  rxy <- lapply(rl, function(x) rasterToPoints(x, fun = which1))
  pps <- lapply(rxy, function(x) {
    ppp(x = x[, 1], y = x[, 2], window = zam_owin)
  })
  # reorder to find which has most points, to put that in as X
  reord <- sort(sapply(pps, function(x) x$n), decreasing = TRUE)
  ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
})
pms2b$NND <- round(sapply(scenover_nn, function(x) mean(x$dist))) / 1000

cols_tr <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0, 0.75, 1), rgb(1, 0.75, 0))
plot(1:4, cex = 3, pch = 20, col = cols_tr)

# plot(scens_rs2[[1]] + scens_rs2[[2]])
# check
# round(sapply(scens_rs2, function(x) {
#   cellStats(x * scens_r[[1]]$convertible * 100, sum)
# }))
# cstats$conv_area

# pms <- list(c(1, 2), c(1, 3), c(1, 4), c(1, 5))
# pms <- expand.grid(rep(list(0:1), 5))[-1, ]  # logic grid

# rasters showing overlap as weight is progressively put on C and BD
scens_over2 <- lapply(2:6, function(x) scens_rs2[[1]] + scens_rs2[[x]] * 2)

# # overlap areas
# scens_over_a <- sapply(1:length(scens_over), function(x) { # x <- 4
#   sapply(1:3, function(y) {  # y <- 3
#     r <- scens_over[[x]] == y
#     # plot(r)
#     round(cellStats(r * scens_r[[1]]$convertible, sum), 1)
#   })
# })
# colSums(scens_over_a[c(2, 3), ])
# over_cols <- sapply(seq(0, 0.4, 0.1), function(x) rgb(1 - x, 0, x))
# par(mar = c(0, 0, 0, 0))
# plot(1:3, pch = 20, col = rgb(1, 0, 1))
# plot(zambia)

# cols <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0.75, 0, 0.75))
# cols <- c("red", "blue", "gold")
bluecols <- sapply(c(1, 0.75, 0.5, 0.25, 0), function(x) rgb(0, x, 1))
bluecols[5] <- rgb(0.4, 0, 1)
cols <- sapply(1:length(bluecols), function(x) c("red", bluecols[x], "gold"))
plot(1:ncol(cols), pch = 20, col = cols[2, ], cex = 20)
ftits <- list("Y 90% C 5% Biod 5% Cost 0%",
              "Y 80% C 10% Biod 10% Cost 0%",
              "Y 70% C 15% Biod 15% Cost 0%",
              "Y 60% C 20% Biod 20% Cost 0%", 
              "Y 50% C 25% Biod 25% Cost 0%")
leg <- c("Y", "YCBd", "Y + YCBd")

tiff(fp(p_figures, "changing_weights2.tif"), height = 3.5, width = 7, 
     units = "in", res = 300)
par(mfrow = c(2, 3), oma = c(0, 0, 0, 1))
for(i in 1:5) { # i <- 1
  # map
  imp_plot2(scens_over2[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
            crops = leg, ftitle = ftits[[i]], col = cols[, i], 
            legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.4, 
            marg = c(0, 0, 1, 0), titline = 0)
}
mgp <- c(1, 0.25, 0)
par(mar = c(3, 3, 1, 0))
with(pms2b, {
     plot(NND, JS, col = cols[2, ], pch = 16, cex = 1.5, 
          axes = FALSE, xlab = "", ylab = "", ylim = c(0, 0.4), 
          xlim = c(0, 180))
     points(NND, JS, col = cols[[1]][1], pch = 20, cex = 1)
     axis(1, at = seq(0, 180, 20), cex.axis = 0.6, las = 2, mgp = mgp, 
          tcl = -0.2)
     axis(2, at = seq(0, 1, 0.1), cex.axis = 0.6, las = 2, mgp = mgp, 
          tcl = -0.2)
     mtext(text = "NN Distance (km)", side = 1, line = 1, cex = 0.45)
     mtext(text = "Jaccard Index", side = 2, line = 1, cex = 0.45)
})
dev.off()

#   # barplot
#   par(mar = c(5, 4, 3, 3))
#   overl <- scens_over_a[1:3, i][c(1, 3, 2)]
#   bp <- barplot(overl, las = 2, ylim = c(0, 12000),  
#                 col = cols[c(1, 3, 2)], mgp = c(1.5, 0.25, 0), tcl = -0.2, 
#                 ylab = expression(km^2), cex.axis = 0.6, cex.lab = 0.7)
#   axis(side = 1, at = bp, labels = leg, las = 2, 
#        mgp = c(1, 0.25, 0), tcl = -0.2, cex.axis = 0.6)
# }
# dev.off()

```
<a href="#top">Back to top</a>

### Figure 6

"Objective" subjective weighting of outcomes, based on different preferences. The examples here are intended to understand the impacts that will happen when different actors with different land use interests are willing to make some compromise, e.g. they are willing to give up 5-10% of the value of their best case cost, whereas the other parties must be willing to pay more. Do that for each of the 4, and then compare to the case where each gets equal weight.

```{r, eval = FALSE}
# use minmax stats from optimal solution set
# pare <- ybdc$both$optitab
optirange <- pare[, lapply(.SD, range), .SDcols = 1:4]
# optirange[1, lapply(.SD, function(x) x * 1.05)]

# y_comp <- sweep(as.matrix(pare)[, 1:4], MARGIN = 2, swvec, `<=`)
cvar <- "C"
compromise <- function(pare, optirange, cvar, cost1 = 1.05, cost2) {  
  sval <- optirange[1, (eval(parse(text = cvar))) * cost1]
  pare_ss <- as.matrix(pare[(eval(parse(text = cvar))) <= sval])
  vars <- names(pare)[!names(pare) %in% c(cvar, "ind")]
  swvec <- as.matrix(optirange[1, vars, with = FALSE])[1, ]
  ocosts <- sweep(pare_ss[, vars], MARGIN = 2, swvec, "/")
  # cost2 <- seq(1, 2, 0.05)
  # pare_ss[2, vars][3] / swvec[3]
  # ocosts[2, ]

  chks <- sapply(cost2, function(x) {  # x <- cost2[13]
    chkvals <- ifelse(ocosts < x, 1, 0)
    chkvec <- which(rowSums(chkvals) == 3)
    if(length(chkvec) == 1) {
      ochks <- chkvec
    } else if(length(chkvec) > 1) {
      ochks <- chkvec[which.min(rowSums(ocosts[chkvec, ]))]  # select minimum
    } else {
      ochks <- 0
    }
    ochks
  })
  # pare_ss[269, ]
  # ocosts[269, ]
  # pare_ss[269, "COST"] / optirange[1, COST]
  # optirange[1, Y] * 1.05
  c("cost" = cost2[chks > 0][1], pare_ss[chks[chks > 0][1], "ind"])
}

# indices of costs in other three variables
cost2 <- seq(1, 5, 0.05)
y_comp <- compromise(pare, optirange, cvar = "Y", cost1 = 1.05, cost2 = cost2)
c_comp <- compromise(pare, optirange, cvar = "C", cost1 = 1.05, cost2 = cost2)
bd_comp <- compromise(pare, optirange, cvar = "BD", cost1 = 1.05, cost2 = cost2)
ct_comp <- compromise(pare, optirange, cvar = "COST", cost1 = 1.05, cost2=cost2)
# pare[ind == y_comp["ind"], .(Y, C, BD, COST)] / optirange[1, ]
# pare[ind == c_comp["ind"], .(Y, C, BD, COST)] / optirange[1, ]
# pare[ind == bd_comp["ind"], .(Y, C, BD, COST)] / optirange[1, ]
# pare[ind == ct_comp["ind"], .(Y, C, BD, COST)] / optirange[1, ]

# select out those values from the results tables and grab conversion areas
parms <- round(ybdc$both$params[, 7:10], 2)
eqind <- which(apply(parms, 1, function(x) all(x == 0.25))) # iter of eq wt run
# pare[ind == eqind]

pind <- c(y_comp["ind"], c_comp["ind"], bd_comp["ind"], ct_comp["ind"], eqind)
# round(ybdc$both$params[pind, 7:10], 2)
fnames <- fp(fp(p_out, ybdc$both$bcode), 
             imp_sorter(fp(p_out, ybdc$both$bcode), inds = pind))
objsubj_r <- lapply(fnames, function(x) {  # x <- pind[[1]]
  print(x)
  DT <- cbind(fread(x), basedt[, .(x, y, convertible)])
  DTr <- dt_to_raster(DT, CRSobj)
})
names(objsubj_r) <- c("ycomp", "ccomp", "bdcomp", "ctcomp", "eqwt")

# run for subjective weights (could have pull this out also) as check
# run for subjective weights, where we want to treat everyone equally
# eq_wgts <- rep(0.25, 4)
# names(eq_wgts) <- cnames  # rename to make sure model runs properly
# tosubj <- tradeoff_mod(targ, cbetas = eq_wgts, currprodmod = g, it = "SUBJ")
# tosubj_imp <- selected_impacts2(tosubj$impacts)

# tosubj_imp <- lapply(c(1, 4, 2, 6:8, 3), function(x) tosubj_imp[[x]])

# pull out the impact tables (brief check shows they are the same values)
all_stats <- selected_impacts(unname(pind), fp(p_out, ybdc$both$bcode))
impsr <- lapply(c(2, 5, 3, 7:9, 4), function(x) all_stats[[x]])
# pare[ind == y_comp["ind"],]; pare[ind == c_comp["ind"],]; 
# pare[ind == bd_comp["ind"],]; pare[ind == ct_comp["ind"],]
# tosubj_imp

# grid for calculating overlaps
pms2a <- expand.grid(rep(list(0:1), 5))[-1, ] 
pms2b <- pms2a[which(rowSums(pms2a) == 2), ]
pms2b$cl <- c("YC", "YBD", "CBd", "YCt", "CCt", "BDCt", "YEq", "CEq", 
              "BdEq", "CtEq")

toresr <- lapply(objsubj_r, function(x) calc(x[[1:2]], sum))
tores_over_r <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
  calc(stack(toresr[which(pms2b[x, ] == 1)]), sum)
})
names(tores_over_r) <- pms2b$cl

# jaccard's similarity, per https://gist.github.com/jlehtoma/3369793
tores_over_a <- sapply(1:length(tores_over_r), function(x) { # x <- 4
  sapply(1:2, function(y) {  # y <- 3
    r <- tores_over_r[[x]] == y
    round(cellStats(r * scens_r[[1]]$convertible, sum), 1)
  })
})
pms2b$JS <- round(tores_over_a[2, ] / colSums(tores_over_a), 3)

# calculate nearest neighbor distance between each version, using spatstat 
# nncross
# plot(zam_owin)
zam_owin <- owin(xrange = bbox(zambia)[1, ], yrange = bbox(zambia)[2, ],
                 unitname = "meter")
which1 <- function(x) x == 1  # selector function for r to xy
tores_nn <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests; x = 2
  ind <- which(pms2b[x, 1:5] == 1)  ### check col number
  rl <- toresr[ind]
  rxy <- lapply(rl, function(x) rasterToPoints(x, fun = which1))
  pps <- lapply(rxy, function(x) {
    ppp(x = x[, 1], y = x[, 2], window = zam_owin)
  })
  # reorder to find which has most points, to put that in as X
  reord <- sort(sapply(pps, function(x) x$n), decreasing = TRUE)
  ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
})
pms2b$NND <- round(sapply(tores_nn, function(x) mean(x$dist))) / 1000
# with(pms2b, plot(NND, JS))

# do sum factoring in overlaps 
# sapply(tores_over_r, function(x) {
#   round(cellStats(x * scens_r[[1]]$convertible, sum), 1)
# })

# cols_tr <- c(rgb(1, 0, 0), rgb(1, 0.75, 0), rgb(0, 0, 1), rgb(0, 0.75, 1))
cols_tr <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0, 0.75, 1), rgb(1, 0.5, 0), 
             "gold")#rgb(1, 1, 0))
plot(1:length(cols_tr), cex = 3, pch = 20, col = cols_tr)

# layout matrix
m1 <- do.call(cbind, lapply(1:5, function(x) matrix(rep(x, 45), 6, 9)))
m2 <- matrix(rep(6, 36), 4, 9)
m3 <- do.call(cbind, lapply(7:13, function(x) matrix(rep(x, 20), 4, 5)))
f1 <- lapply(14:15, function(x) cbind(rep(x, 4), rep(x, 4)))
lmat <- rbind(m1, cbind(m2, m3, rep(14, 4)))
# lmat <- rbind(m1, m3)
# cbind(m2, m3)

# plot(1:10, ylab = ylabs[2])
# ylabs <- c("Jaccard Index", 
ylabs <- c(expression(paste("Conv. area (", 10^5, " ", km^2, ")")), 
           "Hours", "C loss (Mt)", "Priority", "Intactness/priority",
           "Intactness", expression(paste("FR converted (ha"^2, ")")))
# yscale <- c(1, 10^5, 1, 100000, 1, 1, 1, 1)
yscale <- c(10^5, 1, 100000, 1, 1, 1, 1)

# figure titles
ftits <- lapply(pind, function(x) {
  paste(sapply(1:4, function(y) {
    paste(paste(cnames[y], paste0(parms[x, y] * 100, "%")))
  }), collapse = " ")
})
leg <- c("Converted")

# yaxes <- list(seq(0, 0.8, 0.1), seq(0, 14, 2), seq(0, 2.5, 0.5), 
maxes <- sapply(c(2, 5, 3, 7:9, 4), function(j) minmaxstats[[j]][2])
mins <- sapply(c(2, 5, 3, 7:9, 4), function(j) minmaxstats[[j]][1])
ymaxes <- maxes / yscale
ymins <- mins / yscale
yaxes <- list(seq(0, 15, 3), 0:7, seq(0, 1100, 100), seq(0, 0.15, 0.05), 
              seq(0, 0.6, 0.1), seq(0, 1, 0.1), seq(0, 7000, 1000))

# bcols <- c(list(c(cols_tr[1], cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2],
pms2b$icol <- cols_tr[c(1, 1, 2, 1, 2, 3, 1, 2, 3, 4)]  # circle inner
pms2b$ocol <- cols_tr[c(2, 3, 3, 4, 4, 4, 5, 5, 5, 5)]  # circle outer

tiff(fp(p_figures, "objective-subjective_weights.tif"), height = 2.5, width = 7, 
     units = "in", res = 300)
l <- layout(mat = lmat)
# layout.show(l)

# map plots
for(i in 1:length(toresr)) { # i <- 1
  # map
#   par(mar = rep(0, 4))
#   plot(zambia, col = cols_tr[i])
  imp_plot2(toresr[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
            crops = leg, ftitle = ftits[[i]], col = cols_tr[[i]], 
            legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.3, 
            marg = rep(0, 4), inset = c(0.1, 0.1, 0, 0))
}

# impact plots
mgp <- c(0.9, 0.25, 0)
par(mar = c(3, 3, 0, 3))
with(pms2b, {
     plot(NND, JS, col = pms2b$icol, pch = 16, cex = 0.8, 
          axes = FALSE, xlab = "", ylab = "", ylim = c(0, 0.6), 
          xlim = c(0, 25))
     points(NND, JS, col = pms2b$ocol, pch = 20, cex = 0.5)
     axis(1, at = seq(0, 25, 5), cex.axis = 0.5, las = 2, mgp = mgp, 
          tcl = -0.2)
     axis(2, at = seq(0, 0.6, 0.1), cex.axis = 0.5, las = 2, mgp = mgp, 
          tcl = -0.2)
     mtext(text = "NN Distance (km)", side = 1, line = 0.5, cex = 0.4)
     mtext(text = "Jaccard Index", side = 2, line = 1, cex = 0.4)
})
par(mar = c(3, 3, 0, 0))
mgp <- c(1.3, 0.25, 0)
for(i in 1:7) { # i <- 1
  v <- impsr[[i]] / yscale[i]
  b <- barplot(impsr[[i]] / yscale[i], las = 2,  col = cols_tr, 
               mgp = mgp, tcl = -0.2, yaxt = "n", 
               ylim = range(yaxes[[i]]), ylab = ylabs[i], cex.axis = 0.4,
               cex.lab = 0.55)
  lines(c(-1, 7), rep(ymins[i], 2), lwd = 1, col = "grey50")
  lines(c(-1, 7), rep(ymaxes[i], 2), lwd = 1)
  axis(2, at = yaxes[[i]], cex.axis = 0.5, las = 2, mgp = mgp, 
       tcl = -0.2)
}
dev.off()

```

# Earlier version of figure 3

```{r, eval = FALSE}
# objective weights
obj_wgts <- round(pct_of_tot[, {
  c(conv_area, tot_C, int_mean, mu_cost) / 
    sum(c(conv_area, tot_C, int_mean, mu_cost))
}], 2)
# round(obj_wgts, 2)
names(obj_wgts) <- cnames

# Objective weights
toobj <- tradeoff_mod(targ, cbetas = obj_wgts, currprodmod = g, it = "OBJ")

# ybdc$both$params[1, ]
# yp <- ybdc$both$params[, 7:10]
# which(round(yp[, 1], 2) == 0.05 & round(yp[, 2], 2) == 0.30 & 
#         round(yp[, 3], 2) == 0.25 & round(yp[, 4], 2) == 0.4)

# objective weights on ag only 
# considering that these might be two sides of the same coin
ag_wgts <- round(obj_wgts[c(1, 4)] / sum(obj_wgts[c(1, 4)]), 2)
ag_wgts <- c(ag_wgts[1], "C" = 0, "BD" = 0, ag_wgts[2])
ag_wgts2 <- ag_wgts[c(4, 2:3, 1)]
names(ag_wgts2) <- names(ag_wgts)

toag <- tradeoff_mod(targ, cbetas = ag_wgts, currprodmod = g, it = "AGOBJ")
toag2 <- tradeoff_mod(targ, cbetas = ag_wgts2, currprodmod = g, it = "AGOBJ2")

# subjective weights, where we want to treat everyone equally
subj_wgts <- rep(0.25, 4)
names(subj_wgts) <- cnames  # rename to make sure model runs properly
tosubj <- tradeoff_mod(targ, cbetas = subj_wgts, currprodmod = g, it = "SUBJ")

# extract impacts of these, and add in BAU case
tores <- lapply(list(toag, toag2, toobj, tosubj), function(x) {  # x <- toobj
  DTsum <- data.table(x$conv[, rowSums(.SD), .SDcols = c(crops[[1]])])
  DTspat <- cbind(basedt[, .(x, y)], DTsum)
  DTr <- dt_to_raster(DTspat, CRSobj)
  DTr <- DTr + (fbr * 2)  # overlap with farmblocks, for output
  list("map" = DTr, "imp" = x$impacts)
})
names(tores) <- c("ag", "ag2", "obj", "subj")

# Calculate area of overlaps between farmblocks and 
yfbr <- calc(mzsoytor, sum) + fbr * 2
plot(yfbr)
fb_over_a <- sapply(1:length(tores), function(x) { # x <- 4
  sapply(1:3, function(y) {  # y <- 3
    r <- tores[[x]]$map == y
    round(cellStats(r * scens_r[[1]]$convertible, sum), 1)
  })
})
fb_sim <- fb_over_a[3, ] / colSums(fb_over_a)

# grid for calculating overlaps
pms2a <- expand.grid(rep(list(0:1), 4))[-1, ] 
pms2b <- pms2a[which(rowSums(pms2a) == 2), ]
pms2b$cl <- c("AgAg2", "AgObj", "Ag2Obj", "AgSubj", "Ag2Subj", "ObjSubj")

toresr <- lapply(tores, function(x) {
  r <- x$map
  r[r == 2] <- 0
  r[r == 3] <- 1
  r
})

# tores_over_r <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
#   (toresr[[1]] == pms2b[x, 1]) & (toresr[[2]] == pms2b[x, 2]) & 
#     (toresr[[3]] == pms2b[x, 3]) & (toresr[[4]] == pms2b[x, 4])
# })
# names(tores_over_r) <- pms2b$cl
# tores_overs <- stack(tores_over)
tores_over_r <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
  calc(stack(toresr[which(pms2b[x, ] == 1)]), sum)
})
names(tores_over_r) <- pms2b$cl

# jaccard's similarity, per https://gist.github.com/jlehtoma/3369793
tores_over_a <- sapply(1:length(tores_over_r), function(x) { # x <- 4
  sapply(1:2, function(y) {  # y <- 3
    r <- tores_over_r[[x]] == y
    round(cellStats(r * scens_r[[1]]$convertible, sum), 1)
  })
})
pms2b$JS <- round(tores_over_a[2, ] / colSums(tores_over_a), 3)

# calculate nearest neighbor distance between each version, using spatstat 
# nncross
# plot(zam_owin)
zam_owin <- owin(xrange = bbox(zambia)[1, ], yrange = bbox(zambia)[2, ],
                 unitname = "meter")
which1 <- function(x) x == 1  # selector function for r to xy
tores_nn <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests; x = 2
  ind <- which(pms2b[x, 1:4] == 1)
  rl <- toresr[ind]
  rxy <- lapply(rl, function(x) rasterToPoints(x, fun = which1))
  pps <- lapply(rxy, function(x) {
    ppp(x = x[, 1], y = x[, 2], window = zam_owin)
  })
  # reorder to find which has most points, to put that in as X
  reord <- sort(sapply(pps, function(x) x$n), decreasing = TRUE)
  ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
})
pms2b$NND <- round(sapply(tores_nn, function(x) mean(x$dist))) / 1000
# with(pms2b, plot(NND, JS))

# do sum factoring in overlaps 
# sapply(tores_over_r, function(x) {
#   round(cellStats(x * scens_r[[1]]$convertible, sum), 1)
# })

cols_tr <- c(rgb(1, 0, 0), rgb(1, 0.75, 0), rgb(0, 0, 1), rgb(0, 0.75, 1))
plot(1:4, cex = 3, pch = 20, col = cols_tr)

# impact table
imp_list <- list(toag$impacts, toag2$impacts, toobj$impacts, tosubj$impacts)
imps <- rbindlist(lapply(imp_list, function(x) selected_impacts2(x)))
impsr <- lapply(c(1, 4, 2, 6:8, 3), function(x) imps[[x]])

# layout matrix
m1 <- do.call(cbind, lapply(1:4, function(x) matrix(rep(x, 35), 5, 7)))
m2 <- matrix(rep(5, 20), 4, 5)
m3 <- do.call(cbind, lapply(6:12, function(x) matrix(rep(x, 12), 4, 3)))
lmat <- rbind(m1, cbind(rep(5, 4), m2, m3, rep(13, 4)))

# plot(1:10, ylab = ylabs[2])
# ylabs <- c("Jaccard Index", 
ylabs <- c(expression(paste("Converted area (", 10^5, " ", km^2, ")")), 
           "Hours", "C loss (Mt)", "Priority", "Intactness/priority",
           "Intactness", expression(paste("FR converted (ha"^2, ")")))
# yscale <- c(1, 10^5, 1, 100000, 1, 1, 1, 1)
yscale <- c(10^5, 1, 100000, 1, 1, 1, 1)
ftits <- list("Y 11% C 0% Biod 0% Cost 89%",
              "Y 89% C 0% Biod 0% Cost 11%",
              "Y 5% C 29% Biod 25% Cost 41%",
              "Y 25% C 25% Biod 25% Cost 25%")
leg <- c("Converted")

# yaxes <- list(seq(0, 0.8, 0.1), seq(0, 14, 2), seq(0, 2.5, 0.5), 
yaxes <- list(seq(0, 14, 2), seq(0, 2.5, 0.5),               
              seq(0, 1000, 200), seq(0, 0.2, 0.05), seq(0, 0.5, 0.1), 
              seq(0, 0.8, 0.2), seq(0, 3500, 500))
# bcols <- c(list(c(cols_tr[1], cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2],
bcols <- c(list(c(cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2], 
                cols_tr[3])), lapply(1:7, function(x) rep("black", 4)))
# lcols <- c(list(c(cols_tr[2], cols_tr[3], cols_tr[3], cols_tr[4], cols_tr[4] 
# lcols <- lapply(1:7, function(x) cols_tr)

tiff(fp(p_figures, "objective_weights.tif"), height = 3.5, width = 7, 
     units = "in", res = 300)
l <- layout(mat = lmat)
# layout.show(l)

# map plots
for(i in 1:4) { # i <- 1
  # map
  # plot(zambia)
  imp_plot2(toresr[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
            crops = leg, ftitle = ftits[[i]], col = cols_tr[[i]], 
            legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.4, 
            marg = rep(0, 4))
}

# impact plots
mgp <- c(1, 0.25, 0)
par(mar = c(4, 3, 0, 0))
with(pms2b, {
     plot(NND, JS, col = cols_tr[c(2, 3, 3, 4, 4, 4)], pch = 16, cex = 1.5, 
          axes = FALSE, xlab = "", ylab = "", ylim = c(0, 1))
     points(NND, JS, col = cols_tr[c(1, 1, 2, 1, 2, 3)], pch = 20, cex = 1)
     axis(1, at = seq(0, 130, 10), cex.axis = 0.6, las = 2, mgp = mgp, 
          tcl = -0.2)
     axis(2, at = seq(0, 1, 0.1), cex.axis = 0.6, las = 2, mgp = mgp, 
          tcl = -0.2)
     mtext(text = "NN Distance (km)", side = 1, line = 1, cex = 0.45)
     mtext(text = "Jaccard Index", side = 2, line = 1, cex = 0.45)
})
for(i in 1:7) {
  v <- impsr[[i]] / yscale[i]
  b <- barplot(impsr[[i]] / yscale[i], las = 2,  col = cols_tr, 
               mgp = c(1.5, 0.25, 0), tcl = -0.2, yaxt = "n", 
               ylim = range(yaxes[[i]]), ylab = ylabs[i], cex.axis = 0.5,
               cex.lab = 0.7)
  axis(2, at = yaxes[[i]], cex.axis = 0.6, las = 2, mgp = mgp, 
       tcl = -0.2)
}
dev.off()

# ybdc$both$optitab[, {
#   par(mfrow = c(2, 2), mar = rep(3, 4))
#   hist(.SD, axes = FALSE, labels = FALSE)
# }, .SDcol = 1:4]
# 
# frange <- function(x) diff(range(x)) / max(x)
# fqtile <- function(x) diff(quantile(x, probs = c(0.025, 0.975))) / max(x) 
# ybdc$both$optimal[, lapply(.SD, frange), .SDcol = 1:4]
# ybdc$both$optimal[, lapply(.SD, fqtile), .SDcol = 1:4]
# 
# hist(ybdc$both$optimal$Y)

```