### 2016-02-26 Rmd/figures.Rmd
Near final figures for main text done after revision to Fig 1 and 3, and new Figure 4 made. Model rerun for 100% yield gap closure on existing land.

---
title: "Main Analyses"
author: "Lyndon Estes"
date: "February 11, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Figures

```{r, warning=FALSE, message=FALSE}
library(agroEcoTradeoff)
library(readxl)
library(gdalUtils)
library(gisutils)
# URL <- paste0("https://cran.r-project.org/src/contrib/Archive/spatstat/",
#               "spatstat_1.42-2.tar.gz")
# install.packages(URL, repos = NULL, type="source", dependencies = TRUE)
library(spatstat)

setwd("~/Dropbox/publications/zambiaToff/private/zambiaToff/")
p_root <- getwd()
p_demand <- fp(p_root, "external/input_devel/demand/")
p_mod <- fp(p_root, "agroEcoTradeoff")
p_moddat <- fp(p_mod, "external/data/ZA")
p_figures <- full_path(p_root, "paper/figures/")
p_other <- fp(p_root, "external/input_devel/other/")
# p_roads <- fp(p_root, "external/input_devel/roads")
# p_roads2 <- paste0("/Users/lestes/Dropbox/data/zari/infrastructure/zambia/", 
#                    "rda_dump/Final Shape files RDA/")
p_out <- fp(p_mod, "external/output")
```

```{r, eval = FALSE}
# roads
mgrid <- raster(fp(p_moddat, "ZA-mask.tif"))
load(fp(p_mod, "external/data/ZA/parks_roads.rda"))

# farmblocks
fbp <- "/Users/lestes/Dropbox/data/zari/misc/NEW4GEO/Fblocks.sqlite"
# fb <- readOGR(fbp, layer = "fblocks")
# fbalb <- spTransform(fb, mgrid@crs)
# fbalb$ID <- 1:nrow(fbalb)
# fbalb@data <- fbalb@data[, "ID", drop = FALSE]
fnm <- fp(p_other, "fblocks_alb.sqlite")
# writeOGR(fbalb, dsn = fnm, layer = "fblocks_alb", driver = "SQLite")

# read in farmblocks
fbalb <- readOGR(fnm, layer = "fblocks_alb")
fbr <- rasterize(fbalb, mgrid)
fbr[is.na(fbr)] <- 0
fbr[fbr > 0] <- 1
fbr <- raster::mask(fbr, mgrid)
# plot(fbr)

```

<a href="#top">Back to top</a>

## Tradeoff runs

### Best areas to meet targets for each crop

Running the model individually for each crop and for both crops.
```{r, eval = FALSE}
setwd(p_mod)
path <- "external/data/ZA"

# gaps 
cp <- c("maize" = 2912004, "soy" = 203732)  # current production
gapper <- function(cp, gaps, pct) ((((cp * gaps) - cp) * pct) + cp) / cp
mzgap <- 4.43 / 2.3 
soygap <- 3.4 / 2
g <- gapper(cp, c(mzgap, soygap), 1)  # assume 50% closed on current land
targ <- c("maize" = 4, "soy" = 10)  # targets given g
targs <- targets(targ, cp, g)

# Separate soy and maize impacts
cb <- c("Y" = 1, "C" = 0, "BD" = 0, "COST" = 0)
mzto <- tradeoff_mod(prod_targ = targ["maize"], cbetas = cb, currprodmod = g[1])
soyto <- tradeoff_mod(prod_targ = targ["soy"], cbetas = cb, currprodmod = g[2])
mzsoyto <- tradeoff_mod(prod_targ = targ, cbetas = cb, currprodmod = g)

# rasters
CRSobj <- CRS(mzto$inputs$sp$crs)
mztor <- dt_to_raster(mzto$conv, CRSobj)
soytor <- dt_to_raster(soyto$conv, CRSobj)
mzsoytor <- dt_to_raster(mzsoyto$conv, CRSobj)  # converges on same results

# rasters
mzsoyr <- stack(list(mztor, soytor * 2))
mzsoyr <- calc(mzsoyr, sum)
# plot(mzsoyr)
# plot(roads)

# Plot
i1 <- rbind(mzto$impacts, soyto$impacts)
crops <- c("maize", "soy", "maize/soy")
# plot(1:10, pch = 20, cex = 4, col = "gold2")
# plot(mzsoyr, col = c("grey", "goldenrod", "green4", "orange"))
imp_plot2(mzsoyr, i1, p_figures, "ideal_yield2", "", crops, 1.2, TRUE, 
         c("red", "blue1", "gold"))

basedt <- soyto$inputs$mask
crops <- list(c("maize", "soy"))

```
<a href="#top">Back to top</a>

<a href="#top">Back to top</a>

## Figures

How reconcilable are the different land use interests, and what would be the impacts of each?

I am using the two crops combined approach because there is little overlap between the two.

### Figure 1
```{r, eval = FALSE}
load(fp(p_out, "mainresults2502.rda"))

# select out results for "pure" cases (weight = 1)
cnames <- c("Y", "C", "BD", "COST")
pure <- sapply(cnames, function(x) which(round(ybdc$both$params[, x], 4) == 1))

# check on dominated status
# ybdc$both$optitab[ind %in% 20:22]  # pure C not in optimum table
sic <- selected_impacts(20:22, fp(p_out, ybdc$both$bcode))
# (1 - sic$tot_C[1] / sic$tot_C[2]) * 100  # C is .07 more (dominated)
# probably because of over-shooting target
# ctst <- tradeoff_mod(prod_targ = targ, 
#                      cbetas = c("Y" = 0, "C" = 1, "BD" = 0, "COST" = 0),
#                      currprodmod = g)  
# selected_impacts2(ctst$impacts)  # correct one selected
# ybdc$both$params[20:22, ]

fnames <- imp_sorter(fp(p_out, ybdc$both$bcode), inds = pure, fullnames = TRUE)
pure_r <- lapply(fnames, function(x) {  # x <- pure[1]
  DT <- cbind(fread(x), basedt[, .(x, y, convertible)])
  DTr <- dt_to_raster(DT, CRSobj)
})

# sum across crops and multiply by convertible areas
rs <- lapply(pure_r, function(x) calc(x[[1:2]], sum) * x[[3]])  # actual area
rs2 <- lapply(pure_r, function(x) calc(x[[1:2]], sum))  # conv pixels
names(rs2) <- cnames
# check
# sapply(rs, function(x) cellStats(x, sum))
# ybdc$both$optitab[ind %in% pure, Y]  # same, but C optimized out
# sapply(rs2, function(x) cellStats(x, sum))
# plot(pure_r[[3]])

# figure out overlaps between different pure objectives
pms2a <- expand.grid(rep(list(0:1), 4))[-1, ] 
pms2b <- pms2a[which(rowSums(pms2a) == 2), ]
pms2b$cl <- c("YC", "YBd", "CBd", "YCt", "CCt", "BdCt")
rownames(pms2b) <- 1:6

# pairwise overlaps
rs_over <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
  calc(stack(rs2[which(pms2b[x, ] == 1)]), sum)
})
names(rs_over) <- pms2b$cl

# Jaccard Index of overlaps
rs_over_a <- sapply(1:length(rs_over), function(x) { # x <- 4
  sapply(1:2, function(y) {  # y <- 3
    r <- rs_over[[x]] == y
    round(cellStats(r * pure_r[[1]]$convertible, sum), 1)
  })
})
pms2b$JS <- round(rs_over_a[2, ] / colSums(rs_over_a) * 100, 2)

# calculate nearest neighbor distance between each version, using spatstat 
# nncross
zam_owin <- owin(xrange = bbox(zambia)[1, ], yrange = bbox(zambia)[2, ],
                 unitname = "meter")
which1 <- function(x) x == 1  # selector function for r to xy
rsover_nn <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests; x = 2
  ind <- which(pms2b[x, 1:4] == 1)
  rl <- rs2[ind]
  rxy <- lapply(rl, function(x) rasterToPoints(x, fun = which1))
  pps <- lapply(rxy, function(x) {
    ppp(x = x[, 1], y = x[, 2], window = zam_owin)
  })
  # reorder to find which has most points, to put that in as X
  reord <- sort(sapply(pps, function(x) x$n), decreasing = TRUE)
  ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
})
pms2b$NND <- round(sapply(rsover_nn, function(x) mean(x$dist))) / 1000

cols_tr <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0, 0.75, 1), rgb(1, 0.75, 0))
plot(1:4, cex = 3, pch = 20, col = cols_tr)

# impact table

# For Panel B - impact statistics 
# bpath <- fp(p_out, ybdc$both$bcode)
# load(fp(bpath, "impacts_tab.rda"))  # get impacts table

# select out corresponding impact statistics, summarize
conv_stats <- selected_impacts(pure, fp(p_out, ybdc$both$bcode))
# impsr <- lapply(c(2, 5, 3, 7:9, 4), function(x) conv_stats[[x]])
impsr <- lapply(c(2, 3, 8, 5), function(x) conv_stats[[x]])

# Plot setup
m1 <- do.call(cbind, lapply(1:4, function(x) matrix(rep(x, 35), 5, 7)))
m2 <- do.call(cbind, lapply(c(5, 8, 11, 14), function(x) {
  a <- matrix(rep(x, 16), 4, 4)
  b <- matrix(rep(x + 1, 6), 2, 3)
  cbind(a, rbind(b, b + 1))
}))

# m2 <- matrix(rep(5, 20), 4, 5)
# m3 <- do.call(cbind, lapply(6:12, function(x) matrix(rep(x, 12), 4, 3)))
# lmat <- rbind(m1, cbind(rep(5, 4), m2, m3, rep(13, 4)))
lmat <- rbind(m1, m2)

# plot(1:10, ylab = ylabs[2])
# ylabs <- c("Jaccard Index", 
ylabs <- c(expression(paste("Converted area (", 10^5, " ", km^2, ")")), 
           "Carbon loss (Mt)", #"Priority", 
           "Biodiversity Index", "Travel time (hours)")#,
           # "Intactness", expression(paste("FR converted (ha"^2, ")")))
# yscale <- c(1, 10^5, 1, 100000, 1, 1, 1, 1)
yscale <- c(10^5, 100000, 1, 1)#, 1, 1, 1)
# ftits <- list("Y 100%", "C 100%", "Biod 100%", "Cost 100%")
ftits <- list("Maximize Yields", "Minimize Carbon Loss", 
              "Minimize Biodiversity Loss", "Minimize Travel Time")
# leg <- c("Converted")

yaxes <- list(seq(0, 16, 2), seq(0, 1200, 100), seq(0, 0.6, 0.1), 
              seq(0, 7, 1))#,               
              # seq(0, 1200, 100), #seq(0, 0.12, 0.02)), 
              #seq(0, 1, 0.1), seq(0, 6500, 500))
# bcols <- c(list(c(cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2], 
#                   cols_tr[3])), lapply(1:7, function(x) rep("black", 4)))

# plot
# tiff(fp(p_figures, "single_priorities2.tif"), height = 3.5, width = 7, 
png(fp(p_figures, "single_priorities2.png"), height = 3.5, width = 7, 
    units = "in", res = 300)
# postscript(fp(p_figures, "single_priorities2.eps"), horizontal = FALSE, 
#            height = 3.5, width = 7)
l <- layout(mat = lmat)
# layout.show(l)

# map plots
legtf <- c(FALSE, FALSE, FALSE, FALSE)
for(i in 1:4) { # i <- 1
  # par(mar = rep(0, 4))
  # plot(zambia, col = cols_tr[i])
  imp_plot2(rs2[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
            crops = "", ftitle = ftits[[i]], col = cols_tr[[i]], 
            legend = legtf[i], hold = TRUE, stats = FALSE, cexscale = 0.4, 
            marg = rep(0, 4))
}  
# impact plots
# mgp <- c(1.6, 0.25, 0)
# par(mar = c(4, 3, 0, 0))
# with(pms2b, {
#   plot(NND, JS, col = cols_tr[c(2, 3, 3, 4, 4, 4)], pch = 16, cex = 1.5, 
#        axes = FALSE, xlab = "", ylab = "", ylim = c(0, 0.3))
#   points(NND, JS, col = cols_tr[c(1, 1, 2, 1, 2, 3)], pch = 20, cex = 1)
#   axis(1, at = seq(0, 230, 20), cex.axis = 0.6, las = 2, mgp = mgp, 
#        tcl = -0.2)
#   axis(2, at = seq(0, 0.3, 0.05), cex.axis = 0.6, las = 2, mgp = mgp, 
#        tcl = -0.2)
#   mtext(text = "NN Distance (km)", side = 1, line = 1, cex = 0.45)
#   mtext(text = "Jaccard Index", side = 2, line = 1.25, cex = 0.45)
# })
mgp <- c(1, 0.25, 0)
# for(i in 1:7) {
cx <- 0.6
for(i in 1:4) {
  v <- impsr[[i]] / yscale[i]
  par(mar = c(2, 3.5, 2, 0))
  b <- barplot(impsr[[i]] / yscale[i], las = 2, col = cols_tr, 
               mgp = c(1, 0.25, 0), tcl = -0.2, yaxt = "n", 
               ylim = range(yaxes[[i]]), ylab = ylabs[i], cex.axis = cx,
               cex.lab = 0.7)
  axis(2, at = yaxes[[i]], cex.axis = cx, las = 2, mgp = mgp, tcl = -0.2)
  # mtext("Impacts", side = 3, cex = 1)
  ind <- which(pms2b[, paste0("Var", i)] == 1)
  with(pms2b[ind, ], {
    par(mar = c(0.5, 3, 2, 1), mgp = mgp)
    barplot(JS, col = cols_tr[-i], ylim = c(0, 30), yaxt = "n", 
            ylab = "% Overlap", cex.lab = cx)
    axis(2, at=seq(0, 30, 10), cex.axis = cx, las = 2, mgp = mgp, tcl = -0.2)
    # mtext("Promixity", side = 3, cex = 1)
    par(mar = c(2, 3, 0.5, 1), mgp = mgp)
    barplot(NND, col = cols_tr[-i], ylim = c(0, 250), las = 2, yaxt = "n", 
            ylab = "", cex.lab = cx)
    axis(2, at=seq(0, 250, 50), cex.axis = cx, las = 2, mgp = mgp, tcl = -0.2)
    mtext("Average distance", side = 2, line = 1.72, cex = 0.4)
    mtext("to neighbor (km)", side = 2, line = 1.2, cex = 0.4)
  })
}
par(xpd = NA)
legx <- grconvertX(0.4, from = "ndc", to = "user")
legy <- grconvertY(0.5, from = "ndc", to = "user")
legend(x = legx, y = legy, legend = "Protected Areas", pch = 15, col = "grey75", 
       bty = "n", pt.cex = 3 * 0.6, cex = 1.5 * 0.6)
dev.off()

```

<a href="#top">Back to top</a>

### Figures 2 & 3 (probably supplemental)
```{r, eval = FALSE}
# individual maps
png(full_path(p_figures, "all-4-maps.png"), height = 7, width = 7, 
     units = "in", res = 300)
par(mfrow = c(2, 2))#, oma = c(0, 0, 0, 2))
ftitles <- c("Yield", "Carbon", "Biodiversity", "Cost")
legp <- c(TRUE, rep(FALSE, 3))
for(i in 1:4) {
  dtr <- calc(pure_r[[i]][[1:2]] * c(1, 2), sum)
  imp_plot2(dtr, impacts = i1, fpath = p_figures, fnm = "all-4-maps", 
            cexscale = 0.6, ftitle = ftitles[i], crops = c("maize", "soy"), 
            stats = FALSE, hold = TRUE, legend = legp[i], marg = c(0, 0, 2, 0))
}
dev.off()

# heat maps
# implist <- list(ycost$both, bdc$both)
# load(full_path(set_base_path(), paste0("external/data/ZA/parks_roads.rda")))
# pacols <- c("transparent", "grey85", "grey75")
# png(full_path(p_figures, "ycost-bdc-heat.png"), height = 400, width = 900)
# par(mfrow = c(1, 2), oma = c(0, 0, 0, 2))
# for(i in 1:2) {
#   par(mar = c(1, 1, 1, 4))
#   plot(zambia, col = "grey90", border = "transparent")
#   ir <- implist[[i]]$heat
#   ir[ir == 0] <- NA
#   plot(ir, col = rev(heat.colors(20)), add = TRUE, )
#   plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2],
#        border = FALSE)
#   plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3],
#        border = FALSE)
#   plot(roads, add = TRUE, lwd = 0.1, col = "grey")
# }
# dev.off()

```

<a href="#top">Back to top</a>

### Figures 4 (main) and 5 (probably supplemental)

What happens when weight is progressively added

```{r, eval = FALSE}
p <- data.frame(ybdc$both$params)
pare <- ybdc$both$optitab
p[, c("Y", "C", "BD" ,"COST")] <- round(p[, c("Y", "C", "BD" ,"COST")], 2)
scens <- lapply(c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9), function(x) {
  c(1 - x, rep(x / 2, 2), 0)
})
pind <- sapply(scens, function(x) {
  x <- round(x, 2)
  which(p$Y == x[1] & p$C == x[2] & p$BD == x[3] & p$COST == x[4])
})

# select out impacts corresponding to those runs
cstats <- selected_impacts(pind, fp(p_out, ybdc$both$bcode))
cstats <- cstats[, .(iter, conv_area, tot_C, mu_cost, int_prior)]

# check on selection
# il <- input_handler(p_moddat, crops[[1]])
# chk <- tradeoff_mod(prod_targ = targ, cbetas = ybdc$both$params[pind[5], 7:10], 
#                     currprodmod = g, exist_list = il)
# DF <- data.frame(chk$impacts)
# round(c(colSums(DF[, 1:2]),
#         sapply(4:ncol(DF), function(z) weighted.mean(DF[, z], DF$conv_area))),2)
# cstats[5, ]
# fnames <- dir(bpath, pattern = "csv")
# fnames[pind[5]]

# calculate max impacts across both crops, and figure out the impact range, 
# and what percent that is relative to the worst case
allstats <- selected_impacts(pare$ind, fp(p_out, ybdc$both$bcode))
minmaxstats <- allstats[, lapply(.SD, range)]
imp_range <- minmaxstats[, lapply(.SD, diff)]  # difference between best/worst
pct_of_tot <- minmaxstats[, lapply(.SD, function(x) diff(x) / max(x))]

# cost is how much you pay relative to the best case
# savings is how much you gain relative to the worst case
dfunc <- function(x, y) (1 - y / x) * 100
# dfunc <- function(x, y) (x / y - 1) * 100
stnms <- names(cstats)[-1]
imp_pct <- cstats[, lapply(stnms, function(x) {
  mx <- minmaxstats[[x]][2]
  dfunc(mx, get(x))
})]
dfunc(minmaxstats[2, ], minmaxstats[1, ])  # maximum avoided cost

setnames(imp_pct, paste0(stnms, "p"))
imp_pct <- cbind("conv_km" = cstats[, conv_area] / 10^5, imp_pct)
ywgts <- do.call(rbind, scens)[, 1]
xwgts <- rowSums(do.call(rbind, scens)[, 2:3])
imp_pct <- cbind(cbind(ywgts, xwgts), cstats, imp_pct)

cols_tr <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0, 0.75, 1), rgb(1, 0.75, 0))
# sctab <- do.call(rbind, scens) * 100
cx <- 0.7
# lcols <- c("black", "red", "blue")
# imp_pct[, {
#   png(fp(p_figures, "weight_impact.png"), height = 4, width = 4, 
#       units = "in", res = 300)
#   par(mgp = c(2, 0.75, 0), mar = c(3, 3, 1, 1))
#   plot(conv_area, tot_C, type = "l", xlab = "", ylab = "", #xaxt = "n", 
#   lines(conv_area, int_prior, col = "red")
#   lines(conv_area, mu_cost, col = "blue")
#   axis(1, at = seq(2, 12, 1), cex = cx, cex.axis = cx, line = -0.25, 
#        tcl = -0.2, mgp = c(1, 0.25, 0))
#   axis(2, at = seq(0, 60, 5), las = 2, cex = cx, cex.axis = cx, line = -0.25, 
#        tcl = -0.15, mgp = c(1, 0.25, 0))
#   mtext("% land savings", side = 1, line = 1, cex = cx)
#   mtext("% Carbon/Biodiv/Cost savings", side = 2, line = 1, cex = cx)
#   inms <- c("tot_C", "int_prior", "mu_cost")
#   for(i in 1:length(inms)) {
#     text(conv_area[-1], get(inms[i])[-1], labels = sctab[-1, i + 1], 
#          cex = cx * 0.8, pos = 3, offset = 0.2, col = lcols[i])
#     text(conv_area[1] + 0.1, i, labels = 0, cex = cx * 0.8, pos = 3, 
#          offset = i * 0.15 - 0.2, col = lcols[i])
#   }
#   legend("bottomleft", legend = c("Carbon", "Biodiv", "Cost"), cex = cx,
#          text.col = lcols, col = lcols, bty = "n", lty = 1)
imp_pct[, {
  png(fp(p_figures, "weight_impact3.png"), height = 4, width = 4, 
      units = "in", res = 300)
  plot(xwgts, conv_areap, type = "l", xlab = "", ylab = "", axes = FALSE, 
       ylim = c(0, 55), col = cols_tr[1])
  lines(xwgts, tot_Cp, col = cols_tr[2])
  lines(xwgts, int_priorp, col = cols_tr[3])
  lines(xwgts, mu_costp, col = cols_tr[4])
  axis(1, at = xwgts, labels = seq(0, 90, 10), cex = cx, cex.axis = cx, 
       line = -0.25, tcl = -0.2, mgp = c(1, 0.25, 0))
  axis(2, at = seq(0, 55, 5), las = 2, cex = cx, cex.axis = cx, line = -0.25, 
       tcl = -0.15, mgp = c(1, 0.25, 0))
  mtext("% Weight taken from Yield and divided", side = 1, line = 1, cex = cx)
  mtext("equally between Carbon & Biodiversity", side = 1, line = 1.75, 
        cex = cx)
  mtext("% Avoided Cost", side = 2, line = 1, cex = cx)
  legend("bottomright", 
         legend = c("Yield", "Carbon", "Biodiversity", "Travel time"), 
         cex = cx, text.col = cols_tr, col = cols_tr, bty = "n", lty = 1, 
         inset = c(0, 0.25, 0, 0))
  dev.off()
}]

# fnames <- fp(fp(p_out, ybdc$both$bcode), 
#              imp_sorter(fp(p_out, ybdc$both$bcode), inds = pind))
# scens_r <- lapply(fnames, function(x) {  # x <- pind[[1]]
#   print(x)
#   DT <- cbind(fread(x), basedt[, .(x, y, convertible)])
#   DTr <- dt_to_raster(DT, CRSobj)
# })
# 
# scens_rs2 <- lapply(scens_r, function(x) calc(stack(x[[1:2]]), sum))
# names(scens_rs2) <- paste0("w", 1:length(scens_rs2))
# 
# pms2b <- do.call(rbind.data.frame, lapply(2:6, function(x) {
#   a <- rep(0, 6)
#   a[c(1, x)] <- 1
#   a
# }))
# colnames(pms2b) <- paste0("Var", 1:ncol(pms2b))
# pms2b$cl <- c("W010", "W020", "W030", "W040", "W050")
# rownames(pms2b) <- 1:5
# 
# # pairwise overlaps
# scens_over <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
#   calc(stack(scens_rs2[which(pms2b[x, ] == 1)]), sum)
# })
# names(scens_over) <- pms2b$cl
# 
# # Jaccard Index of overlaps
# scens_over_a <- sapply(1:length(scens_over), function(x) { # x <- 4
#   sapply(1:2, function(y) {  # y <- 3
#     r <- scens_over[[x]] == y
#     round(cellStats(r * pure_r[[1]]$convertible, sum), 1)
#   })
# })
# pms2b$JS <- round(scens_over_a[2, ] / colSums(scens_over_a), 3)
# 
# # calculate nearest neighbor distance between each version, using spatstat 
# # nncross
# zam_owin <- owin(xrange = bbox(zambia)[1, ], yrange = bbox(zambia)[2, ],
#                  unitname = "meter")
# which1 <- function(x) x == 1  # selector function for r to xy
# scenover_nn <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests; x = 4
#   print(x)
#   ind <- which(pms2b[x, 1:6] == 1)
#   rl <- scens_rs2[ind]
#   rxy <- lapply(rl, function(x) rasterToPoints(x, fun = which1))
#   pps <- lapply(rxy, function(x) {
#     ppp(x = x[, 1], y = x[, 2], window = zam_owin)
#   })
#   # reorder to find which has most points, to put that in as X
#   reord <- sort(sapply(pps, function(x) x$n), decreasing = TRUE)
#   ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
# })
# pms2b$NND <- round(sapply(scenover_nn, function(x) mean(x$dist))) / 1000
# 
# cols_tr <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0, 0.75, 1), rgb(1, 0.75, 0))
# plot(1:4, cex = 3, pch = 20, col = cols_tr)
# 
# # plot(scens_rs2[[1]] + scens_rs2[[2]])
# # check
# # round(sapply(scens_rs2, function(x) {
# #   cellStats(x * scens_r[[1]]$convertible * 100, sum)
# # }))
# # cstats$conv_area
# 
# # pms <- list(c(1, 2), c(1, 3), c(1, 4), c(1, 5))
# # pms <- expand.grid(rep(list(0:1), 5))[-1, ]  # logic grid
# 
# # rasters showing overlap as weight is progressively put on C and BD
# scens_over2 <- lapply(2:6, function(x) scens_rs2[[1]] + scens_rs2[[x]] * 2)
# 
# # # overlap areas
# # scens_over_a <- sapply(1:length(scens_over), function(x) { # x <- 4
# #   sapply(1:3, function(y) {  # y <- 3
# #     r <- scens_over[[x]] == y
# #     # plot(r)
# #     round(cellStats(r * scens_r[[1]]$convertible, sum), 1)
# #   })
# # })
# # colSums(scens_over_a[c(2, 3), ])
# # over_cols <- sapply(seq(0, 0.4, 0.1), function(x) rgb(1 - x, 0, x))
# # par(mar = c(0, 0, 0, 0))
# # plot(1:3, pch = 20, col = rgb(1, 0, 1))
# # plot(zambia)
# 
# # cols <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0.75, 0, 0.75))
# # cols <- c("red", "blue", "gold")
# bluecols <- sapply(c(1, 0.75, 0.5, 0.25, 0), function(x) rgb(0, x, 1))
# bluecols[5] <- rgb(0.4, 0, 1)
# cols <- sapply(1:length(bluecols), function(x) c("red", bluecols[x], "gold"))
# plot(1:ncol(cols), pch = 20, col = cols[2, ], cex = 20)
# ftits <- list("Y 90% C 5% Biod 5% Cost 0%",
#               "Y 80% C 10% Biod 10% Cost 0%",
#               "Y 70% C 15% Biod 15% Cost 0%",
#               "Y 60% C 20% Biod 20% Cost 0%", 
#               "Y 50% C 25% Biod 25% Cost 0%")
# leg <- c("Y", "YCBd", "Y + YCBd")
#               
# tiff(fp(p_figures, "changing_weights2.tif"), height = 3.5, width = 7, 
#      units = "in", res = 300)
# par(mfrow = c(2, 3), oma = c(0, 0, 0, 1))
# for(i in 1:5) { # i <- 1
#   # map
#   imp_plot2(scens_over2[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
#             crops = leg, ftitle = ftits[[i]], col = cols[, i], 
#             legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.4, 
#             marg = c(0, 0, 1, 0), titline = 0)
# }
# mgp <- c(1, 0.25, 0)
# par(mar = c(3, 3, 1, 0))
# with(pms2b, {
#      plot(NND, JS, col = cols[2, ], pch = 16, cex = 1.5, 
#           axes = FALSE, xlab = "", ylab = "", ylim = c(0, 0.4), 
#           xlim = c(0, 180))
#      points(NND, JS, col = cols[[1]][1], pch = 20, cex = 1)
#      axis(1, at = seq(0, 180, 20), cex.axis = 0.6, las = 2, mgp = mgp, 
#           tcl = -0.2)
#      axis(2, at = seq(0, 1, 0.1), cex.axis = 0.6, las = 2, mgp = mgp, 
#           tcl = -0.2)
#      mtext(text = "NN Distance (km)", side = 1, line = 1, cex = 0.45)
#      mtext(text = "Jaccard Index", side = 2, line = 1, cex = 0.45)
# })
# dev.off()
# 
# 
# #   # barplot
# #   par(mar = c(5, 4, 3, 3))
# #   overl <- scens_over_a[1:3, i][c(1, 3, 2)]
# #   bp <- barplot(overl, las = 2, ylim = c(0, 12000),  
# #                 col = cols[c(1, 3, 2)], mgp = c(1.5, 0.25, 0), tcl = -0.2, 
# #                 ylab = expression(km^2), cex.axis = 0.6, cex.lab = 0.7)
# #   axis(side = 1, at = bp, labels = leg, las = 2, 
# #        mgp = c(1, 0.25, 0), tcl = -0.2, cex.axis = 0.6)
# # }
# # dev.off()
#

```
<a href="#top">Back to top</a>

### Figure 6

"Objective" subjective weighting of outcomes, based on different preferences. The examples here are intended to understand the impacts that will happen when different actors with different land use interests are willing to make some compromise, e.g. they are willing to give up 5-10% of the value of their best case cost, whereas the other parties must be willing to pay more. Do that for each of the 4, and then compare to the case where each gets equal weight.

```{r, eval = FALSE}
# use minmax stats from optimal solution set
# pare <- ybdc$both$optitab
optirange <- pare[, lapply(.SD, range), .SDcols = 1:4]
optirange[1, lapply(.SD, function(x) x * 1.05)] - optirange[1, ]

# y_comp <- sweep(as.matrix(pare)[, 1:4], MARGIN = 2, swvec, `<=`)
cvar <- "C"
compromise <- function(pare, optirange, cvar, cost1 = 1.05, cost2) {  
  sval <- optirange[1, (eval(parse(text = cvar))) * cost1]
  pare_ss <- as.matrix(pare[(eval(parse(text = cvar))) <= sval])
  vars <- names(pare)[!names(pare) %in% c(cvar, "ind")]
  swvec <- as.matrix(optirange[1, vars, with = FALSE])[1, ]
  ocosts <- sweep(pare_ss[, vars], MARGIN = 2, swvec, "/")
  # cost2 <- seq(1, 2, 0.05)
  # pare_ss[2, vars][3] / swvec[3]
  # ocosts[2, ]

  chks <- sapply(cost2, function(x) {  # x <- cost2[13]
    chkvals <- ifelse(ocosts < x, 1, 0)
    chkvec <- which(rowSums(chkvals) == 3)
    if(length(chkvec) == 1) {
      ochks <- chkvec
    } else if(length(chkvec) > 1) {
      ochks <- chkvec[which.min(rowSums(ocosts[chkvec, ]))]  # select minimum
    } else {
      ochks <- 0
    }
    ochks
  })
  # pare_ss[269, ]
  # ocosts[269, ]
  # pare_ss[269, "COST"] / optirange[1, COST]
  # optirange[1, Y] * 1.05
  c("cost" = cost2[chks > 0][1], pare_ss[chks[chks > 0][1], "ind"])
}

# indices of costs in other three variables
cost2 <- seq(1, 5, 0.05)
y_comp <- compromise(pare, optirange, cvar = "Y", cost1 = 1.05, cost2 = cost2)
c_comp <- compromise(pare, optirange, cvar = "C", cost1 = 1.05, cost2 = cost2)
bd_comp <- compromise(pare, optirange, cvar = "BD", cost1 = 1.05, cost2 = cost2)
ct_comp <- compromise(pare, optirange, cvar = "COST", cost1 = 1.05, cost2=cost2)
# pare[ind == y_comp["ind"], .(Y, C, BD, COST)] / optirange[1, ]
# pare[ind == c_comp["ind"], .(Y, C, BD, COST)] / optirange[1, ]
# pare[ind == bd_comp["ind"], .(Y, C, BD, COST)] / optirange[1, ]
# pare[ind == ct_comp["ind"], .(Y, C, BD, COST)] / optirange[1, ]

# select out those values from the results tables and grab conversion areas
parms <- round(ybdc$both$params[, 7:10], 2)
eqind <- which(apply(parms, 1, function(x) all(x == 0.25))) # iter of eq wt run
# pare[ind == eqind]

pind <- c(y_comp["ind"], c_comp["ind"], bd_comp["ind"], ct_comp["ind"], eqind)
# round(ybdc$both$params[pind, 7:10], 2)
fnames <- fp(fp(p_out, ybdc$both$bcode), 
             imp_sorter(fp(p_out, ybdc$both$bcode), inds = pind))
objsubj_r <- lapply(fnames, function(x) {  # x <- pind[[1]]
  print(x)
  DT <- cbind(fread(x), basedt[, .(x, y, convertible)])
  DTr <- dt_to_raster(DT, CRSobj)
})
names(objsubj_r) <- c("ycomp", "ccomp", "bdcomp", "ctcomp", "eqwt")

# run for subjective weights (could have pull this out also) as check
# run for subjective weights, where we want to treat everyone equally
# eq_wgts <- rep(0.25, 4)
# names(eq_wgts) <- cnames  # rename to make sure model runs properly
# tosubj <- tradeoff_mod(targ, cbetas = eq_wgts, currprodmod = g, it = "SUBJ")
# tosubj_imp <- selected_impacts2(tosubj$impacts)

# tosubj_imp <- lapply(c(1, 4, 2, 6:8, 3), function(x) tosubj_imp[[x]])

# pull out the impact tables (brief check shows they are the same values)
all_stats <- selected_impacts(unname(pind), fp(p_out, ybdc$both$bcode))
# impsr <- lapply(c(2, 5, 3, 7:9, 4), function(x) all_stats[[x]])
impsr <- lapply(c(2, 3, 8, 5), function(x) all_stats[[x]])

# pare[ind == y_comp["ind"],]; pare[ind == c_comp["ind"],]; 
# pare[ind == bd_comp["ind"],]; pare[ind == ct_comp["ind"],]
# tosubj_imp

# grid for calculating overlaps
pms2a <- expand.grid(rep(list(0:1), 5))[-1, ] 
pms2b <- pms2a[which(rowSums(pms2a) == 2), ]
pms2b$cl <- c("YC", "YBD", "CBd", "YCt", "CCt", "BDCt", "YEq", "CEq", 
              "BdEq", "CtEq")

toresr <- lapply(objsubj_r, function(x) calc(x[[1:2]], sum))
tores_over_r <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
  calc(stack(toresr[which(pms2b[x, ] == 1)]), sum)
})
names(tores_over_r) <- pms2b$cl

# jaccard's similarity, per https://gist.github.com/jlehtoma/3369793
tores_over_a <- sapply(1:length(tores_over_r), function(x) { # x <- 4
  sapply(1:2, function(y) {  # y <- 3
    r <- tores_over_r[[x]] == y
    round(cellStats(r * pure_r[[1]]$convertible, sum), 1)
  })
})
pms2b$JS <- round(tores_over_a[2, ] / colSums(tores_over_a) * 100, 2)

# calculate nearest neighbor distance between each version, using spatstat 
# nncross
# plot(zam_owin)
zam_owin <- owin(xrange = bbox(zambia)[1, ], yrange = bbox(zambia)[2, ],
                 unitname = "meter")
which1 <- function(x) x == 1  # selector function for r to xy
tores_nn <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests; x = 2
  ind <- which(pms2b[x, 1:5] == 1)  ### check col number
  rl <- toresr[ind]
  rxy <- lapply(rl, function(x) rasterToPoints(x, fun = which1))
  pps <- lapply(rxy, function(x) {
    ppp(x = x[, 1], y = x[, 2], window = zam_owin)
  })
  # reorder to find which has most points, to put that in as X
  reord <- sort(sapply(pps, function(x) x$n), decreasing = TRUE)
  ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
})
pms2b$NND <- round(sapply(tores_nn, function(x) mean(x$dist))) / 1000
pms2bb <- pms2b[pms2b$Var5 != 1, -5]
# with(pms2b, plot(NND, JS))

# do sum factoring in overlaps 
# sapply(tores_over_r, function(x) {
#   round(cellStats(x * scens_r[[1]]$convertible, sum), 1)
# })

# cols_tr <- c(rgb(1, 0, 0), rgb(1, 0.75, 0), rgb(0, 0, 1), rgb(0, 0.75, 1))
# cols_tr <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0, 0.75, 1), rgb(1, 0.5, 0), 
#              "gold")#rgb(1, 1, 0))
# plot(1:length(cols_tr), cex = 3, pch = 20, col = cols_tr)

# layout matrix
# m1 <- do.call(cbind, lapply(1:5, function(x) matrix(rep(x, 45), 6, 9)))
# m2 <- matrix(rep(6, 36), 4, 9)
# m3 <- do.call(cbind, lapply(7:13, function(x) matrix(rep(x, 20), 4, 5)))
# f1 <- lapply(14:15, function(x) cbind(rep(x, 4), rep(x, 4)))
# lmat <- rbind(m1, cbind(m2, m3, rep(14, 4)))
# lmat <- rbind(m1, m3)
# cbind(m2, m3)

# # plot(1:10, ylab = ylabs[2])
# # ylabs <- c("Jaccard Index", 
# ylabs <- c(expression(paste("Conv. area (", 10^5, " ", km^2, ")")), 
#            "Hours", "C loss (Mt)", "Priority", "Intactness/priority",
#            "Intactness", expression(paste("FR converted (ha"^2, ")")))
# # yscale <- c(1, 10^5, 1, 100000, 1, 1, 1, 1)
# yscale <- c(10^5, 1, 100000, 1, 1, 1, 1)

# # figure titles
# ftits <- lapply(pind, function(x) {
#   paste(sapply(1:4, function(y) {
#     paste(paste(cnames[y], paste0(parms[x, y] * 100, "%")))
#   }), collapse = " ")
# })
# leg <- c("Converted")

# # yaxes <- list(seq(0, 0.8, 0.1), seq(0, 14, 2), seq(0, 2.5, 0.5), 
# maxes <- sapply(c(2, 5, 3, 7:9, 4), function(j) minmaxstats[[j]][2])
# mins <- sapply(c(2, 5, 3, 7:9, 4), function(j) minmaxstats[[j]][1])
# ymaxes <- maxes / yscale
# ymins <- mins / yscale
# yaxes <- list(seq(0, 15, 3), 0:7, seq(0, 1100, 100), seq(0, 0.15, 0.05), 
#               seq(0, 0.6, 0.1), seq(0, 1, 0.1), seq(0, 7000, 1000))
#

maxes <- sapply(c(2:3, 8, 5), function(j) minmaxstats[[j]][2])
mins <- sapply(c(2:3, 8, 5), function(j) minmaxstats[[j]][1])
ymaxes <- maxes / yscale
ymins <- mins / yscale
pctlines <- rbind(ymins, 
                  t(sapply(c(0.25, 0.5, 0.75), function(x) {
                    (ymaxes - ymins) * x + ymins
                  })), ymaxes)

# # bcols <- c(list(c(cols_tr[1], cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2],
# pms2b$icol <- cols_tr[c(1, 1, 2, 1, 2, 3, 1, 2, 3, 4)]  # circle inner
# pms2b$ocol <- cols_tr[c(2, 3, 3, 4, 4, 4, 5, 5, 5, 5)]  # circle outer
# 
# tiff(fp(p_figures, "objective-subjective_weights.tif"), height = 2.5, width = 7,      units = "in", res = 300)
# l <- layout(mat = lmat)
# layout.show(l)

# # map plots
# for(i in 1:length(toresr)) { # i <- 1
#   # map
# #   par(mar = rep(0, 4))
# #   plot(zambia, col = cols_tr[i])
#   imp_plot2(toresr[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
#             crops = leg, ftitle = ftits[[i]], col = cols_tr[[i]], 
#             legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.3, 
#             marg = rep(0, 4), inset = c(0.1, 0.1, 0, 0))
# }  
# 
# # impact plots
# mgp <- c(0.9, 0.25, 0)
# par(mar = c(3, 3, 0, 3))
# with(pms2b, {
#      plot(NND, JS, col = pms2b$icol, pch = 16, cex = 0.8, 
#           axes = FALSE, xlab = "", ylab = "", ylim = c(0, 0.6), 
#           xlim = c(0, 25))
#      points(NND, JS, col = pms2b$ocol, pch = 20, cex = 0.5)
#      axis(1, at = seq(0, 25, 5), cex.axis = 0.5, las = 2, mgp = mgp, 
#           tcl = -0.2)
#      axis(2, at = seq(0, 0.6, 0.1), cex.axis = 0.5, las = 2, mgp = mgp, 
#           tcl = -0.2)
#      mtext(text = "NN Distance (km)", side = 1, line = 0.5, cex = 0.4)
#      mtext(text = "Jaccard Index", side = 2, line = 1, cex = 0.4)
# })
# par(mar = c(3, 3, 0, 0))
# mgp <- c(1.3, 0.25, 0)
# for(i in 1:7) { # i <- 1
#   v <- impsr[[i]] / yscale[i]
#   b <- barplot(impsr[[i]] / yscale[i], las = 2,  col = cols_tr, 
#                mgp = mgp, tcl = -0.2, yaxt = "n", 
#                ylim = range(yaxes[[i]]), ylab = ylabs[i], cex.axis = 0.4,
#                cex.lab = 0.55)
#   lines(c(-1, 7), rep(ymins[i], 2), lwd = 1, col = "grey50")
#   lines(c(-1, 7), rep(ymaxes[i], 2), lwd = 1)
#   axis(2, at = yaxes[[i]], cex.axis = 0.5, las = 2, mgp = mgp, 
#        tcl = -0.2)
# }
# dev.off()

# Plot setup
ftits <- list("Compromise 5% on Yield", "Compromise 5% on Carbon", 
              "Compromise 5% on Biodiversity", "Compromise 5% on Travel Time")
yaxes <- list(seq(0, 16, 4), seq(0, 1200, 100), seq(0, 0.6, 0.1), 
              seq(0, 7, 1))
pcttxt <- c("best", "50%", "worst")

# plot
png(fp(p_figures, "compromise_5pct.png"), height = 3.5, 
    width = 7, units = "in", res = 300)
l <- layout(mat = lmat)
# layout.show(l)

# map plots
legtf <- c(FALSE, FALSE, FALSE, FALSE)
for(i in 1:4) { # i <- 1
  # plot(zambia)
  imp_plot2(toresr[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
            crops = "", ftitle = ftits[[i]], col = cols_tr[[i]], 
            legend = legtf[i], hold = TRUE, stats = FALSE, cexscale = 0.4, 
            marg = rep(0, 4))
}  
# impact plots
mgp <- c(0.8, 0.25, 0)
cx <- 0.6
for(i in 1:4) { 
  v <- impsr[[i]][1:4] / yscale[i]
  par(mar = c(2, 3.5, 2, 0))
  b <- barplot(v, las = 2, col = cols_tr, mgp = c(1, 0.25, 0), tcl = -0.2, 
               yaxt = "n", ylim = range(yaxes[[i]]), ylab = ylabs[i], 
               cex.axis = cx, cex.lab = 0.7)
  for(j in 1:3) {
    lines(c(-1, 7), rep(pctlines[j + (j - 1), i], 2), lwd = 1, col = "grey50")
    text(4.5, y = rep(pctlines[j + (j - 1), i], 2), labels = pcttxt[j], 
         col = "grey50", xpd = NA, pos = 4, cex = cx * 0.9)
  }
  axis(2, at = yaxes[[i]], cex.axis = cx, las = 2, mgp = mgp, tcl = -0.2)
  with(pms2bb[ind, ], {
    par(mar = c(0.5, 3, 2, 1), mgp = mgp)
    barplot(JS, col = cols_tr[-i], ylim = c(0, 70), yaxt = "n", 
            ylab = "% Overlap", cex.lab = cx)
    axis(2, at=seq(0, 70, 10), cex.axis = cx, las = 2, mgp = mgp, tcl = -0.2)
    par(mar = c(2, 3, 0.5, 1), mgp = mgp)
    barplot(NND, col = cols_tr[-i], ylim = c(0, 30), las = 2, yaxt = "n", 
            ylab = "", cex.lab = cx)
    axis(2, at=seq(0, 30, 10), cex.axis = cx, las = 2, mgp = mgp, tcl = -0.2)
    mtext("Average distance", side = 2, line = 1.3, cex = 0.4)
    mtext("to neighbor (km)", side = 2, line = 0.8, cex = 0.4)
  })
}
par(xpd = NA)
legx <- grconvertX(0.4, from = "ndc", to = "user")
legy <- grconvertY(0.5, from = "ndc", to = "user")
legend(x = legx, y = legy, legend = "Protected Areas", pch = 15, col = "grey75",        bty = "n", pt.cex = 3 * 0.6, cex = 1.5 * 0.6)
dev.off()

# 25-25-25-25 Compromise
m1 <- matrix(rep(1, 24), nrow = 4, ncol = 6)
m2 <- do.call(cbind, lapply(2:3, function(x) matrix(rep(x, 4), 2, 2)))
m3 <- do.call(cbind, lapply(4:5, function(x) matrix(rep(x, 4), 2, 2)))
lmat <- cbind(m1, rbind(m2, m3))

png(fp(p_figures, "compromise_25pct.png"), height = 2, width = 4, 
    units = "in", res = 300)
l <- layout(mat = lmat)
# layout.show(l)

par(oma = c(0, 0, 0, 1))
# plot(zambia)
imp_plot2(toresr[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
          crops = "", ftitle = "Compromise equally", col = "purple", 
          legend = FALSE, hold = TRUE, stats = FALSE, cexscale = 0.4, 
          marg = c(0, 0, 1, 0), titline = 0)
legend("bottomright", legend = "Protected Areas", pch = 15, col = "grey75",
       bty = "n", pt.cex = 3 * 0.5, cex = 1.5 * 0.5, inset = c(0, 0.1, 0, 0))

mgp <- c(1.25, 0.25, 0)
cx <- 0.5
for(i in 1:4) { 
  v <- impsr[[i]][c(5, 1:4)] / yscale[i]
  if(i %in% 1:2) par(mar = c(1, 3.5, 2, 0.2), lwd = 0.5)
  if(i %in% 3:4) par(mar = c(2, 3.5, 1, 0.2), lwd = 0.5)
  b <- barplot(v, las = 2, col = c("purple", cols_tr), mgp = c(1, 0.25, 0), 
               tcl = -0.2, yaxt = "n", ylim = range(yaxes[[i]]), 
               ylab = ylabs[i], cex.axis = cx, cex.lab = 0.5)
  for(j in 1:3) {
    lines(c(-1, 7), rep(pctlines[j + (j - 1), i], 2), lwd = 1, col = "grey50")
    text(5.5, y = rep(pctlines[j + (j - 1), i], 2), labels = pcttxt[j], 
         col = "grey50", xpd = NA, pos = 4, cex = cx * 0.7)
  }
  axis(2, at = yaxes[[i]], cex.axis = cx, las = 2, mgp = mgp, tcl = -0.2)
}
dev.off()

```

<a href="#top">Back to top</a>

### Overlap with Farm Blocks
```{r, eval = FALSE}
# Calculate area of overlaps between farmblocks and 
Y <- calc(mzsoytor, sum)
yfbr <- Y + fbr * 2
eqfbr <- toresr$eqwt + fbr * 2
fbrlist <- list(yfbr, eqfbr)

fb_over_a <- sapply(1:2, function(x) { # x <- 4
  sapply(1:3, function(y) {  # y <- 3
    r <- fbrlist[[x]] == y
    round(cellStats(r * pure_r[[1]]$convertible, sum), 1)
  })
})
fb_sim <- fb_over_a[3, ] / colSums(fb_over_a)

fbrna <- fbr > 0
fbrna[fbrna == 0] <- NA

# calculate nearest neighbor dist between each version, using spatstat nncross
inlist <- list("Y" = Y, "eqwt" = toresr$eqwt)
fb_nn <- lapply(1:length(inlist), function(x) {  # x <- 1
  rl <- list(inlist[[x]], fbrna)
  names(rl) <- c(names(inlist)[x], "fbr")
  rxy <- lapply(rl, function(x) rasterToPoints(x, fun = which1))
  pps <- lapply(rxy, function(x) {
    ppp(x = x[, 1], y = x[, 2], window = zam_owin)
  })
  # reorder to find which has most points, to put that in as X
  reord <- sort(sapply(pps, function(x) x$n), decreasing = TRUE)
  ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
})
fb_nnd <- round(sapply(fb_nn, function(x) mean(x$dist))) / 1000

ftits <- c("Maximize Yields", "Compromise Equally")
leg <- c("converted", "farmblock", "converted in farmblock")
inlist <- list("Y" = yfbr, "eqwt" = eqfbr)
png(fp(p_figures, "fb_overlaps.png"), height = 2.5, width = 5, units = "in", 
     res = 300)
par(mfrow = c(1, 2))
for(i in 1:2) {
  imp_plot2(inlist[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
            crops = leg, ftitle = ftits[i], 
            col = c("red", "orange", rgb(0, 1, 1)),
            legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.3, 
            marg = c(0, 0, 1, 0), inset = c(0, 0, 0, 0), titline = 0)
  mtext(paste("Jaccard Index =", round(fb_sim[i], 2)), side = 3, line = -2, 
        cex = 0.4, adj = 0.2)
  mtext(paste("NN Distance =", round(fb_nnd[i]), "km"), side = 3, line = -2.5, 
        cex = 0.4, adj = 0.2)
}
dev.off()

```

### Impact of closing yield gaps
```{r, eval = FALSE}
# Separate soy and maize impacts
cb <- c("Y" = 0.25, "C" = 0.25, "BD" = 0.25, "COST" = 0.25)
g <- gapper(cp, c(mzgap, soygap), 0.5)  # assume 50% closed on current land
lowgap <- tradeoff_mod(prod_targ = targ, cbetas = cb, currprodmod = g)
g <- gapper(cp, c(mzgap, soygap), 1.5)  # assume 50% closed on current land
highgap <- tradeoff_mod(prod_targ = targ, cbetas = cb, currprodmod = g)

lowi <- selected_impacts2(lowgap$impacts)
highi <- selected_impacts2(highgap$impacts)
gaptab <- rbind(lowi, all_stats[5, names(all_stats)[-1], with = FALSE], highi)
gaptab[1, ] / gaptab[2, ]
gaptab[, tot_C / 100000]

```

### Additional supplementals
```{r, eval = FALSE}
il <- mzsoyto$inputs
ylds <- dt_to_raster(cbind(basedt[, .(x, y)], il$p_yield), CRSobj)
carb <- dt_to_raster(cbind(basedt[, .(x, y)], il$carbon), CRSobj)
cost <- dt_to_raster(cbind(basedt[, .(x, y)], il$cost), CRSobj)
closs <- carb[[1]] + carb[[2]] * 0.25
pacols <- c("transparent", "grey85", "grey75")

png(fp(p_figures, "yields.png"), height = 3.5, width = 8, unit = "in", 
     res = 300)
par(mar = c(0, 0, 0, 4), mfrow = c(1, 2))
for(i in 1:2) {
  plot(zambia, col = "white", border = "transparent")
  plot(ylds[[i]], legend = TRUE, add = TRUE)
  plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2],
       border = pacols[2])
  plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3],
       border = pacols[3])
  legend("bottomright", legend = c("GMAs", "NatParks"),
         pch = 15, col = c(pacols[2:3]), bty = "n",
         pt.cex = 2, cex = 1, inset = c(0.1, 0.1, 0, 0))

}  
dev.off()

png(fp(p_figures, "carbon.png"), height = 3.5, width = 4.5, unit = "in", 
     res = 300)
par(mar = c(0, 0, 0, 4))
plot(zambia, col = "white", border = "transparent")
plot(closs, legend = TRUE, add = TRUE)
plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2],
     border = pacols[2])
plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3],
     border = pacols[3])
legend("bottomright", legend = c("GMAs", "NatParks"),
       pch = 15, col = c(pacols[2:3]), bty = "n",
       pt.cex = 2, cex = 1, inset = c(0.1, 0.1, 0, 0))
dev.off()

png(fp(p_figures, "cost.png"), height = 3.5, width = 4.5, unit = "in", 
     res = 300)
par(mar = c(0, 0, 0, 4))
plot(zambia, col = "white", border = "transparent")
plot(cost, legend = TRUE, add = TRUE)
plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2],
     border = pacols[2])
plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3],
     border = pacols[3])
legend("bottomright", legend = c("GMAs", "NatParks"),
       pch = 15, col = c(pacols[2:3]), bty = "n",
       pt.cex = 2, cex = 1, inset = c(0.1, 0.1, 0, 0))
dev.off()

```

<a href="#top">Back to top</a>

### Deprecated version of Figure 3

```{r, eval = FALSE}
# objective weights
# obj_wgts <- round(pct_of_tot[, {
#   c(conv_area, tot_C, int_mean, mu_cost) / 
#     sum(c(conv_area, tot_C, int_mean, mu_cost))
# }], 2)
# # round(obj_wgts, 2)
# names(obj_wgts) <- cnames
# 
# # Objective weights
# toobj <- tradeoff_mod(targ, cbetas = obj_wgts, currprodmod = g, it = "OBJ")
# 
# # ybdc$both$params[1, ]
# # yp <- ybdc$both$params[, 7:10]
# # which(round(yp[, 1], 2) == 0.05 & round(yp[, 2], 2) == 0.30 & 
# #         round(yp[, 3], 2) == 0.25 & round(yp[, 4], 2) == 0.4)
# 
# # objective weights on ag only 
# # considering that these might be two sides of the same coin
# ag_wgts <- round(obj_wgts[c(1, 4)] / sum(obj_wgts[c(1, 4)]), 2)
# ag_wgts <- c(ag_wgts[1], "C" = 0, "BD" = 0, ag_wgts[2])
# ag_wgts2 <- ag_wgts[c(4, 2:3, 1)]
# names(ag_wgts2) <- names(ag_wgts)
# 
# toag <- tradeoff_mod(targ, cbetas = ag_wgts, currprodmod = g, it = "AGOBJ")
# toag2 <- tradeoff_mod(targ, cbetas = ag_wgts2, currprodmod = g, it = "AGOBJ2")
# 
# # subjective weights, where we want to treat everyone equally
# subj_wgts <- rep(0.25, 4)
# names(subj_wgts) <- cnames  # rename to make sure model runs properly
# tosubj <- tradeoff_mod(targ, cbetas = subj_wgts, currprodmod = g, it = "SUBJ")
# 
# # extract impacts of these, and add in BAU case
# tores <- lapply(list(toag, toag2, toobj, tosubj), function(x) {  # x <- toobj
#   DTsum <- data.table(x$conv[, rowSums(.SD), .SDcols = c(crops[[1]])])
#   DTspat <- cbind(basedt[, .(x, y)], DTsum)
#   DTr <- dt_to_raster(DTspat, CRSobj)
#   DTr <- DTr + (fbr * 2)  # overlap with farmblocks, for output
#   list("map" = DTr, "imp" = x$impacts)
# })
# names(tores) <- c("ag", "ag2", "obj", "subj")
# 
# # grid for calculating overlaps
# pms2a <- expand.grid(rep(list(0:1), 4))[-1, ] 
# pms2b <- pms2a[which(rowSums(pms2a) == 2), ]
# pms2b$cl <- c("AgAg2", "AgObj", "Ag2Obj", "AgSubj", "Ag2Subj", "ObjSubj")
# 
# toresr <- lapply(tores, function(x) {
#   r <- x$map
#   r[r == 2] <- 0
#   r[r == 3] <- 1
#   r
# })
# 
# # tores_over_r <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
# #   (toresr[[1]] == pms2b[x, 1]) & (toresr[[2]] == pms2b[x, 2]) & 
# #     (toresr[[3]] == pms2b[x, 3]) & (toresr[[4]] == pms2b[x, 4])
# # })
# # names(tores_over_r) <- pms2b$cl
# # tores_overs <- stack(tores_over)
# tores_over_r <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
#   calc(stack(toresr[which(pms2b[x, ] == 1)]), sum)
# })
# names(tores_over_r) <- pms2b$cl
# 
# # jaccard's similarity, per https://gist.github.com/jlehtoma/3369793
# tores_over_a <- sapply(1:length(tores_over_r), function(x) { # x <- 4
#   sapply(1:2, function(y) {  # y <- 3
#     r <- tores_over_r[[x]] == y
#     round(cellStats(r * scens_r[[1]]$convertible, sum), 1)
#   })
# })
# pms2b$JS <- round(tores_over_a[2, ] / colSums(tores_over_a), 3)
# 
# # calculate nearest neighbor distance between each version, using spatstat 
# # nncross
# # plot(zam_owin)
# zam_owin <- owin(xrange = bbox(zambia)[1, ], yrange = bbox(zambia)[2, ],
#                  unitname = "meter")
# which1 <- function(x) x == 1  # selector function for r to xy
# tores_nn <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests; x = 2
#   ind <- which(pms2b[x, 1:4] == 1)
#   rl <- toresr[ind]
#   rxy <- lapply(rl, function(x) rasterToPoints(x, fun = which1))
#   pps <- lapply(rxy, function(x) {
#     ppp(x = x[, 1], y = x[, 2], window = zam_owin)
#   })
#   # reorder to find which has most points, to put that in as X
#   reord <- sort(sapply(pps, function(x) x$n), decreasing = TRUE)
#   ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
# })
# pms2b$NND <- round(sapply(tores_nn, function(x) mean(x$dist))) / 1000
# # with(pms2b, plot(NND, JS))
# 
# # do sum factoring in overlaps 
# # sapply(tores_over_r, function(x) {
# #   round(cellStats(x * scens_r[[1]]$convertible, sum), 1)
# # })
# 
# cols_tr <- c(rgb(1, 0, 0), rgb(1, 0.75, 0), rgb(0, 0, 1), rgb(0, 0.75, 1))
# plot(1:4, cex = 3, pch = 20, col = cols_tr)
# 
# # impact table
# imp_list <- list(toag$impacts, toag2$impacts, toobj$impacts, tosubj$impacts)
# imps <- rbindlist(lapply(imp_list, function(x) selected_impacts2(x)))
# impsr <- lapply(c(1, 4, 2, 6:8, 3), function(x) imps[[x]])
# 
# # layout matrix
# m1 <- do.call(cbind, lapply(1:4, function(x) matrix(rep(x, 35), 5, 7)))
# m2 <- matrix(rep(5, 20), 4, 5)
# m3 <- do.call(cbind, lapply(6:12, function(x) matrix(rep(x, 12), 4, 3)))
# lmat <- rbind(m1, cbind(rep(5, 4), m2, m3, rep(13, 4)))
# 
# # plot(1:10, ylab = ylabs[2])
# # ylabs <- c("Jaccard Index", 
# ylabs <- c(expression(paste("Converted area (", 10^5, " ", km^2, ")")), 
#            "Hours", "C loss (Mt)", "Priority", "Intactness/priority",
#            "Intactness", expression(paste("FR converted (ha"^2, ")")))
# # yscale <- c(1, 10^5, 1, 100000, 1, 1, 1, 1)
# yscale <- c(10^5, 1, 100000, 1, 1, 1, 1)
# ftits <- list("Y 11% C 0% Biod 0% Cost 89%",
#               "Y 89% C 0% Biod 0% Cost 11%",
#               "Y 5% C 29% Biod 25% Cost 41%",
#               "Y 25% C 25% Biod 25% Cost 25%")
# leg <- c("Converted")
# 
# # yaxes <- list(seq(0, 0.8, 0.1), seq(0, 14, 2), seq(0, 2.5, 0.5), 
# yaxes <- list(seq(0, 14, 2), seq(0, 2.5, 0.5),               
#               seq(0, 1000, 200), seq(0, 0.2, 0.05), seq(0, 0.5, 0.1), 
#               seq(0, 0.8, 0.2), seq(0, 3500, 500))
# # bcols <- c(list(c(cols_tr[1], cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2],
# bcols <- c(list(c(cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2], 
#                 cols_tr[3])), lapply(1:7, function(x) rep("black", 4)))
# # lcols <- c(list(c(cols_tr[2], cols_tr[3], cols_tr[3], cols_tr[4], cols_tr[4] 
# # lcols <- lapply(1:7, function(x) cols_tr)
# 
# tiff(fp(p_figures, "objective_weights.tif"), height = 3.5, width = 7, 
#      units = "in", res = 300)
# l <- layout(mat = lmat)
# # layout.show(l)
# 
# # map plots
# for(i in 1:4) { # i <- 1
#   # map
#   # plot(zambia)
#   imp_plot2(toresr[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
#             crops = leg, ftitle = ftits[[i]], col = cols_tr[[i]], 
#             legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.4, 
#             marg = rep(0, 4))
# }  
# 
# # impact plots
# mgp <- c(1, 0.25, 0)
# par(mar = c(4, 3, 0, 0))
# with(pms2b, {
#      plot(NND, JS, col = cols_tr[c(2, 3, 3, 4, 4, 4)], pch = 16, cex = 1.5, 
#           axes = FALSE, xlab = "", ylab = "", ylim = c(0, 1))
#      points(NND, JS, col = cols_tr[c(1, 1, 2, 1, 2, 3)], pch = 20, cex = 1)
#      axis(1, at = seq(0, 130, 10), cex.axis = 0.6, las = 2, mgp = mgp, 
#           tcl = -0.2)
#      axis(2, at = seq(0, 1, 0.1), cex.axis = 0.6, las = 2, mgp = mgp, 
#           tcl = -0.2)
#      mtext(text = "NN Distance (km)", side = 1, line = 1, cex = 0.45)
#      mtext(text = "Jaccard Index", side = 2, line = 1, cex = 0.45)
# })
# for(i in 1:7) {
#   v <- impsr[[i]] / yscale[i]
#   b <- barplot(impsr[[i]] / yscale[i], las = 2,  col = cols_tr, 
#                mgp = c(1.5, 0.25, 0), tcl = -0.2, yaxt = "n", 
#                ylim = range(yaxes[[i]]), ylab = ylabs[i], cex.axis = 0.5,
#                cex.lab = 0.7)
#   axis(2, at = yaxes[[i]], cex.axis = 0.6, las = 2, mgp = mgp, 
#        tcl = -0.2)
# }
# dev.off()
#

```

<a href="#top">Back to top</a>