### 2016-02-26 paper/main.tex

%%%% added from https://groups.google.com/forum/#!topic/latexusersgroup/CDlEjgNnF80

%\documentclass[onecolumn]{rsauthor}    
\documentclass[a4paper]{article}
\usepackage[affil-it]{authblk}

\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{endfloat}
\usepackage[numbers,sort,compress]{natbib}
%\bibpunct{[}{]}{,}{n}{,}{,}  % https://xianblog.wordpress.com/tag/natbib/ (allows natbib with PNAS)
\usepackage{endnotes}
\usepackage{setspace}
\usepackage{verbatim}
\usepackage[left=2.5cm, right=2.5cm, bottom=2cm, top=2cm]{geometry}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
%\usepackage{mathtime}
\usepackage{bm}
\usepackage{url}
%\usepackage{babel}
\usepackage{dcolumn}
\usepackage{multirow}
%%%%%%

\usepackage[
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,anchorcolor=blue,
  citecolor=blue,filecolor=blue,
  menucolor=blue,pagecolor=blue,
  urlcolor=blue]{hyperref}

% Let's add todonotes and comments:
%\usepackage[draft]{todonotes}
%
%% Select what to do with command \comment:  
%% \newcommand{\comment}[1]{}  %comment not showed
%\newcommand{\comment}[1]
%{{[\bfseries \color{blue} #1]}} %comment showed

\usepackage{lineno}
\usepackage{float}
\usepackage[anythingbreaks]{breakurl}

\title{Reconciling agriculture, carbon, and biodiversity in a savanna transformation frontier}
\author[1,2]{Estes, L.D.}
\author[2]{Searchinger, T.}
\author[2]{Spiegel, M.}
\author[2]{Tian, D.}
\author[3]{Kehoe, L.}
\author[3]{Kuemmerle, T.}
\author[1]{Caylor, K.K.}
\affil[1]{Civil and Environmental Engineering, Princeton University, Princeton, NJ, 08544 USA}
\affil[2]{Woodrow Wilson School, Princeton University, Princeton, NJ, 08544 USA}
\affil[3]{Humboldt University, Berlin}
\date{}

\begin{document}
\maketitle
%\begin{frontmatter}

\begin{abstract}
Blah blah
\end{abstract}
%% Text of abstract

%\begin{keyword}
%keywords
%\end{keyword}

%\end{frontmatter}

%\linenumbers

%% main text
\section*{\large Introduction}

   Meeting growing food demands while minimizing loss of native ecosystems and their carbon and biodiversity presents one of the world’s great challenges for food and the environment.  Many global sustainability studies and climate mitigation pathways call for eliminating or rapidly phasing out emissions from land use change.  CITATIONS IPCC AR V; UK SCIENCE OFFICE.  Unfortunately, these ambitions do not appear realistically attainable in sub-Saharan Africa due to a population likely to more than double to 2.1 billion by 2050. For example, even with healthy linear yield growth from 2006 to 2050 at 3\% to 5\% per year projected by the UN Food and Agriculture Organization (FAO), the region would likely need to add more than 100 million hectares of cropland by 2050.  Searchinger et al. 2013. Much and probably the great bulk of that new cropland is likely to occur in the ~500 million hectares of savannas and shrublands of the region that receive sufficient rainfall to permit crop production.  Searchinger et al. 2013.  This conversion would release tens of millions of tons of carbon and have large impacts on biodiversity.

[I would add in here something about the fact that this is a major challenge, it is one that needs careful planning to achieve, but that this planning has to be done at the country level as this is where most land use policy decisions are made.  Here we look at Zambia, which is perhaps best representative of the challenge.  [this ties us back to the call we made in the NCC paper for finer-level planning, and also inoculates us from the parochiality problem, as we are looking at a representative country].]

Zambia presents an acute part of the challenge.  According to the medium estimate of the United Nations, Zambia’s population is likely to grow 265\% from 16.2 million in 2015 to 43 million in 2050.  (United Nations Department of Economic and Social Affairs Population Division.  2015.  World Population Prospects:  The 2015 Revision, DVD Edition).   As per capita calorie food availability in Zambia is also low, hunger indices high (with child stunting rates at  40\%), https://www.wfp.org/stories/10-facts-about-hunger-zambia GET BETTER CITATION) , and the percentage of the population dependent on agriculture at X\%, Zambia has great need to boost its agricultural production.  Zambia has also been expanding its agricultural exports, and has agronomic and situational characteristics that put it in a strong position to expand exports to other parts of the region.[1]

At the same time, two thirds of Zambia’s 75 milloin hectares is dominated by Miombo Woodland considered dense to be categorized as forest by the UN REDD program, or X\%.  (Vinya, R., Syampungani, S., Kasumu, E.C., Monde, C. \& Kasubika, R. (2011). Preliminary Study on the Drivers of Deforestation and Potential for REDD+ in Zambia. A consultancy report prepared for Forestry Department and FAO under the national UN-REDD+ Programme Ministry of Lands \& Natural Resources. Lusaka, Zambia).  Using this measure of forest, Zambia’s deforestation rate has been estimated at 250,000 to 300,000 hectares per year, among the highest rates of deforestation of any country in the world. Ibid. Fifty percent of these X million hectares have carbon stocks in excess of X tC/h, and an additional X\% have carbon stocks in excess of X.

The biodiversity of these woody savannas is also high.  Overall, the wetter woodland-savannas and shrublands of Africa have a similar average richness for birds and mammals as the world’s wetter tropical forests, and 75\% of these habitats have higher overall vertebrate diversity than 75\% of the rest of the world.   (Searchiner et al. 2015).   Most of Zambia is covered by what the World Wildlife Fund has categorized as Central Zambian Miombo Woodland.  It contains mature miombo woodland trees typically 15 to 20 meters tall, around a quarter of which are evergreen, with a broadleaf shrub and grass understory beneath.[2]  It is also interspersed with a large quantity of wetlands, up to 30\% of many areas.   Its large expanse of habitat make it a center of large mammals, including elephant, black rhino and African buffalo.  Although endemism is low, birdlife is rich with X species, and reptile and amphibian endemism is high, with 19 endemic reptile species 13 endemic amphibians.  LYNDON, CAN WE ADD SOME NUMBERS OF PLANT SPECIES, AND VERTEBRATE SPECIES.   Zambia contains eight other ecoregions according to WWF classifications.[3]   Its extensive wetlands are particular centers of bird and animal life and under extensive pressures.  The value of this habitat is based in part on its remaining relative abundance, which makes it possible for wide-ranging mammals, including many predator species, to survive.

These multiple trends imply challenging trade-offs between agricultural development and environmental concerns.  They raise the question how could decision-makers best target agricultural expansion so that it balances economic, carbon and biodiversity concerns.  A related question is how could a system for targeting well illuminate the trade-offs to support decision-making and a proper role for influence by stakeholders.  And an important question is whether if so what benefits could be achieved by targeting land expansion in certain areas from the standpoint of achieving areas with a desirable mix of objectives.

[We need a somewhat clearer phrasing of our objectives here, or we need to state the objective below the next paragraph. What is that objective: two-fold I would say. 1) Is this sort of modeling approach useful and/or does it have realistic potential for finding compromises in a more general sense (i.e. can it be applied beyond just Zambia), 2) how much potential does Zambia appear to have for reducing the impacts of its expansion. Something along these lines.]

There is a long history of conservation planning that attempts to target conservation areas or objectives while minimizing economic costs subject or while limited land use budget for conservation \cite[e.g.][]{koh_spatially_2010, naidoo_integrating_2006,margules_systematic_2000}. Koh and Ghazoul \citep{koh_spatially_2010} provide one example of an effort to examine trade-offs among production, carbon and biodiversity for oil palm development in Indonesia by giving each land use priority equal weighting.  We adopt a similar approach here, using a newly developed model in which varying weights can be placed on the different objectives, thereby allowing optimal spatial allocations to be found that reflect different degrees of compromise between competing land use interests. To enable this functionality, the model is structured to evaluate each objective in terms of efficiency, or the ratio of cost to yield benefit that results from converting land to agriculture in order to achieve a production target.

We apply this model to answer a basic but important question: can agricultural expansion in Zambia be achieved for substantially lower environmental cost than a business as usual approach, where expansion occurs without consideration of its ecological consequences? This question is particularly important for tropical savanna shrubland, and grassland biomes, which are commonly assumed to provide a low environmental cost reserve for agricultural development \cite{searchinger_high_2015}.

\section*{\large Materials and Methods}
The model we developed, called \emph{AgroEcoTradeoff}, is designed to find the best land use configuration to satisfy the production targets for multiple crops while minimizing four costs: the total land area required, which is inversely related to productivity; the potential losses of carbon and biodiversity; and transportation costs, which are closely correlated with agricultural production costs (ref).  The model takes the basic form:

\begin{equation}
B_{ijk} = \frac{J_{jk}}{P_{ij}} \quad i \in m, \ j = 1, \ldots,n, \ k = 1, \ldots,q 
\end{equation}

Where J$_{k}$ is one of the four costs to be minimized, within a landscape composed of $n$ grid cells containing values for each of the $q$ cost objectives and potential productivity values $P_i$, for each of $m$ crops. $B_{ijk}$ therefore is an efficiency that describes the ecological cost per unit of production, which is normalized so that values are between 0 and 1, so that the different objectives can be compared in common units on a common scale. The normalized values are subtracted from 1 to equate high values with high probability of conversion. Each of these probabilities is then multiplied by a weight (range 0-1, constrained to sum to 1 across all four weights), and the four resulting products are summed, producing a scalarized single objective that represents the landscape conversion probability for each individual crop in each grid cell. An objective function seeks to maximize the sum of these conversion probabilities, based on the requirement that the sum of the potential production of the cells converted for each crop must meet that crop's production target. A second constraint restricts each cell so that it is converted to no more than one crop, as most crops are grown during the same season, and producing multiple crops on the same land would introduce further complexity in setting assumptions about the share of each crop allocated to shared pixels. We therefore formulated a greedy approximation algorithm, based on solving the knapsack problem  \citep{dantzig_discrete-variable_1957} and an algorithm for solving the generalized assignment problem \citep{cohen_efficient_2006}, which allocates each crop to its highest value cells.

This formulation allows quantitatively optimal solutions to be found across a wide range of weight combinations, which is valuable because it is rare to find a single solution that simultaneously obtains the lowest cost for each objective. By finding the optimal solutions across a range of weights, the model can identify the Pareto front, or the set of solutions where the cost in one objective cannot be further minimized without increasing the cost in another. This set then allows model users to choose the most appealing combination of weights based on their land use preferences, and their own assessments of what are and are not acceptable costs within each of the four competing objectives.

This weighting system thus has two basic components.  One reflects the pure value judgment for one output versus another, e.g., how much more a person cares about yield over carbon, or carbon over biodiversity.  The other component expresses each output in quantitative terms so that it reasonably expresses preferences for that output.  For example, units for the carbon, biodiversity and yield outputs are scaled so that someone seeking to weight yield at twice the value each of carbon and biodiversity can do so and have the model find an outcome that reasonably reflects this preference. This approach is necessary to support public decision-making processes, in which different individuals and stakeholders might assign different values to different interests, and where compromise solutions must be sought.

\subsection*{Model inputs} 
For conducting our analysis of potential land use tradeoffs in Zambia, we developed the following spatial datasets at a resolution of 1 km$^2$.

\subsubsection*{\emph{Agricultural potential and production scenarios}}
There are many ways of evaluating agricultural potential factoring in agronomic and economic characteristics.  Here we focused on two crops, maize and soybeans, because of their existing and potential significance for Zambia’s agricultural sector, and because they are considered to be the two dominant commodities that will be grown in Africa's broader savanna regions \citep{searchinger_high_2015}.  Maize is the largest crop by area and consumption in Zambia \citep{fao_fao_2012}, while soybeans are expanding rapidly as a commodity crop to meet rapidly growing regional and global demands for animal feed and direct human consumption \citep{gasparri_emerging_2015}.

We mapped potential yields for each crop using a three-step approach. In the first step, we used the decision support system for agrotechnology transfer \cite[DSSAT, ][]{jones_modelling_2003} to simulate maize and soybean yields at the sites of 40 weather station distributed across Zambia, which represent the points where a 31-year, gridded daily meteorological dataset has received the greatest amount of bias-correction and infilling by weather observations \citep{sheffield_development_2006,estes_changing_2014}. We used soil profiles corresponding to those locations drawn from the WISE v1.1 gridded soil profile database \citep{harvestchoice_converting_2010}, and simulated for each location maize yields over all 31 years under commercial input practices for three-four cultivars representative of short and medium season length according to the Zambian crop calendar \cite{grassini_how_2015}.  This step allowed us to characterize crop responses to Zambia’s inherent climatic variation, and to improved management practices, which was otherwise impossible given the coarse resolution of available observed yield data and the fact that much of Zambia’s current production occurs on low input, smallholder farms (cite). In the second step, we used an empirical model to map the DSSAT-modeled yields onto a 1 km grid, taking advantage of a finer resolution soil map that provides a common frame for estimating both yield potential and soil carbon stocks (see below), but has insufficient detail in several parameters (particularly effective rooting depth) required by DSSAT.  We used a generalized additive model \citep{wood_mgcv:_2001} to predict the values of the DSSAT-simulated yields, using best-fitting subsets of the following weather and soil variables: growing season precipitation; growing degree days; mean growing season shortwave solar radiation; percent organic carbon; pH; percent clay content. The soil predictors represented mean values in the top 1 m of soil, and were extracted from a new 1 km global database of soil properties \cite{hengl_soilgrids1km_2014}.  These first two steps were necessary to 1) capture crop responses to Zambia's climatic variability and potential management practices, and 2) to base potential yields on the same dataset used to estimate carbon costs (see below). This resulted in a map of potential yield under relatively high inputs and good management, and assuming no losses due to other limitations such as diseases. These yield values therefore overestimated what farmers are truly likely to achieve. To correct this, we therefore rescaled the yield maps to match the mean yields (4.5 t/ha for maize and 3.6 t/ha for soybean) that are projected to be achieved in Zambia by 2050, using FAO-projections to derive an annual growth factor, which we then applied to the 2009-2014 reported mean yields for each crop. Overall, the distribution of yields is most important to targeting land use, but the average yield determines the total quantity of land needed to meet production targets.

As a second measure related to agricultural potential, we estimated the travel time to the nearest city of 10,000 people or more (check this), which serves as a proxy for production costs, such as fertilizer prices and transport costs (citation), and is strongly correlated with the suitability of land for farming (cites from ConBio paper).  We used a cost-distance analysis drawing on spatial data for roads, rivers and water bodies, urban centers inside and neighboring Zambia, and national boundaries to calculate travel time in hours. Further details on the cost distance and yield estimation methods can be found in the online electronic supplement.

\subsubsection*{\emph{Carbon}} 
To estimate potential carbon costs from cropland conversions, we developed maps of vegetative and soil carbon stocks for Zambia. Vegetative carbon was calculated from a map of above-ground biomass developed by Baccini et al. \citep{baccini_estimated_2012}, from which we estimated below-ground biomass using savanna and miombo woodland-specific root-shoot ratios \cite{mokany_critical_2006,mugasha_allometric_2013}, and then converted the sum to carbon using a ratio of 0.47 (IPCC 2006 cite). We used the same soil carbon dataset used for crop yield prediction together with an accompanying map of bulk density to calculate soil carbon stocks in the top 1 m. The potential carbon loss due to agricultural conversion was calculated as 100\% of vegetative carbon and 25\% of soil carbon \cite{searchinger_high_2015}.

\subsubsection*{\emph{Biodiversity}}
There are many ways of evaluating biodiversity value and potential impacts, including species richness of different taxa in absolute terms or based on level of threat or endemism, and including factors designed to address the significance of contiguity or the impact of different types of development \citep[e.g.][]{newbold_global_2015,searchinger_high_2015, gasparri_emerging_2015,koh_spatially_2010}. Such measures should reflect both conservation priorities and the quality of available data.  For a fine-scaled, country-level analysis such as this one, the species range map data that form the basis for many impact assessments (Kehoe et al, 2015) have an effective resolution that is too coarse (Laura's cite). We therefore developed a composite measure of biodiversity value based on several datasets. First, we used the potential vegetation maps of East Africa \citep[][]{van_breugel_potential_2011}, a 2010 landcover dataset for Zambia (SERVIR cite), and a shapefile of Zambian protected areas to calculate how much of Zambia's 20 unique vegetation types remains after conversion to croplands and settlements, and what proportion of that remainder falls within national parks. From this we calculated a hybrid measure of rarity and threat, wherein we assigned a weight to each vegetation type based on its relative remaining area in proportion to Zambia's extent, log-transformed to account for a highly skewed distribution, and then re-weighted that weight by each vegetation types protected fraction, following \citep{van_breugel_environmental_2015}. We then used the remaining vegetation map to calculate a second measure, the proportion of untransformed vegetation within a 11X11 km neighborhood centered on each 1 km$^2$, providing a measure of vegetation intactness. By adding this measure to the first index and normalizing, we created a unitless biodiversity score (range 0-1) that gives equal weight to how rare or threatened a habitat is, and how undisturbed it is.

As a final modification, we masked out the areas of existing national parks and game management areas, on the assumption that they would not be used for expanding farmland, but retained a third protected category, forest reserves, as these tend to be far less stringently protected, having had 8.3\% of their area converted to cropland or settlements according to the landcover data. We assigned these areas a high biodiversity score of 0.917 (1 - the already converted proportion), indicating the value of keeping such areas protected, but recognizing that they have some chance of being converted, which can be estimated from their current degree of transformation.

\subsubsection*{\emph{Potential farmland and production targets}}
Our tradeoff analysis is concerned with cropland expansion into areas that are currently not farmed but could be. To identify such areas, we calculated the proportion of currently cropped and settled grid cells within each 1 km$^2$ grid cell, and used a 30 m digital elevation model to calculate slope percent over the entire country, and calculated the proportion of cells greater than 20\%. Slopes above this percent are considered unsuitable for farming (USDA cite), and slope-based measures are strong predictors of cropland presence or absence \citep{estes_using_2014}. Adding the two resulting maps together, subtracting their value from 1, and removing protected areas provided the fractional potential cropland in each cell. In addition to defining cropland availability, this map also provided the area term necessary for calculating production from potential yield.

To develop the production targets used by the model, we calculated the rate of change in annual production of maize for Zambia between 2000-2014, collected from the FAOStat database. For soybean we did the same, but used the mean production trend across the 9 southern African countries examined by Gasparri et al \citep{gasparri_emerging_2015}, given that soybean production is very underdeveloped in Zambia compared to countries such as South Africa that are currently experiencing rapid growth. Extrapolating linearly from 2015 until 2050, maize production will triple and soybean production will increase 9-fold. We used these linear extrapolations as the basis for our targets, adding an additional unit to each factor to account for the fact that Zambia, which has one of the highest agricultural potentials in sub-Saharan Africa, is likely to be subject to larger demands for exports to many of its neighbors, thus our final targets were a four-fold increase for maize (from 2009-2014 average production of 2,912,000 tons) and a 10-fold increase for soybean (from 203,700 tons).

Given that current Zambian maize and soybean yields are well below their current potentials (Waddington et al, 2010), it is an open question how much of future production needs will be satisfied by boosting production on existing croplands. However, it seems likely that some portion of this target will come from current land, given recent Zambian yield growth trends (FAOStat).  In our production scenario, we therefore assumed that an amount equal to a quarter of the gap (redo) between current and projected yields would be met on current farmland before significant expansion occurs.

\subsection*{Simulations}
Using the selected production targets, we ran the model across all combinations of weights for the four parameters, varying the weights in 5\% increments. This allowed us to evaluate the impact of different land use preferences on the location of conversion impacts, the range and magnitude of potential conversion impacts, including for each objective, and to test how sensitive costs in each objective were to changes in weights.

\section*{\large Results}
The first order determinant of land use conversion costs is unrelated to any weights applied to the four objectives, as it is based on how much of future production needs will be met on existing croplands, and therefore how much non-agricultural land will be spared from conversion. In our main scenario, we assumed that the gap between current and our projected future potential yields would be fully closed, but to illustrate the impact of this assumption on land conversion costs, we also ran the model (with equal weights across all objectives) for cases when 50\% and 150\% of the gap was closed. In the former case, the total land area required to meet the target increased by 18\% (14,634 to 17,202 km$^2$), average travel time increased 3\% (1.96 to 2.01 hours), carbon losses increased by 19\% (597 to 711 mt), and the average biodiversity index of converted areas increased by 2\% (0.297 to 0.303). In the latter case, the impacts were reversed, with impacts declining by almost exactly the same percentage.

To identify the minimum possible costs for each individual objective, and the degree to which their areas of new cropland correspond, we examined the simulations were 100\% of the weight was given to each objective. The yield maximizing objective (100\% weight to yield) describes both a cost in terms of the land area required, and a benefit that is based on the productivity gained from converting that land. Scenarios that involve the lowest land use costs also provide farmers the highest potential yields, and therefore have economic advantages.  An examination of overlaps shows two interesting results. First, converted areas that would maximize yield have little correspondence with the areas that minimize carbon costs, biodiversity impacts or travel time (Figure 1), as measured along two dimensions: actual shared area, as measured by the percentage overlap between the two maps, and their adjacency to one another, calculated as the average distance between each pixel of one map and its nearest neighbor in the other. Second, the areas selected for minimizing carbon, biodiversity, and transportation costs have high adjacency, albeit little overlap. All three criteria tend to identify lands along existing road networks and population and agricultural centers. The reason for this correlation is that carbon stocks and biodiversity values are usually lowest in areas of high human density, because of high levels of charcoal and wood harvesting (cite) and fragmentation caused by settlements, existing croplands, and road networks, which exist primarily to connect agricultural and urban centers.

\begin{figure}[!ht]
  \begin{center}
       \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{figures/single_priorities2.png}}
    %\includegraphics[scale=0.5]{fig2.eps}
    \caption{Areas converted to new maize and soybean croplands and their associated costs when each individual objective (yield maximization, and carbon cost, biodiversity cost, and transport time minimization) is given 100\% weight.  The impact categories are color-coded to the specific conversion maps.}
    \label{default}
  \end{center}
\end{figure}

One of the goals of this analysis is to determine whether opportunities exist to achieve a reasonable accommodation between different objectives, and here the results are also interesting. While the best case scenario for maximizing yield would convert  million hectares, the worst case scenario (which would minimize carbon costs) would convert Y million hectares, only a Z\% increase, which is far lower than the variability in potential crop yields (3.7-6.1 t ha$^{-1}$ for maize and 0.8 to 5.1 ha$^{-1}$ for soybean). This result reflects our model's use, and the value, of an efficiency approach. Because yield is in the denominator of each objective (e.g. carbon loss per ton of yield, or biodiversity value per ton of yield), land with better yield potential tend to be selected--even when the yield maximization priority is given no weight--so that the crop production can be achieved with the least cost.

By contrast to the limited differences for yields, the differences in transportation costs using existing transportation networks range from an average of 0.6 hours for each hectare to 6.9 hours, an increase of 1050\%.  Carbon costs from minimum to maximum vary by 188\%, and average biodiversity impacts vary by 100\%.  These differences suggest that there is high potential to reduce transportation, carbon and biodiversity costs for relatively sacrifice of yield potential.

This potential is illustrated in Figure 2, which shows how much cost is avoided in each objective when weights are transferred between them. Here avoided cost is calculated as the percent difference between the cost paid under a given weight combination and the largest cost that is incurred in all combinations (the worst case). For carbon, biodiversity, and travel time, the maximum cost occurs when yield is given 100\% weight, which is where the yield objective has the lowest possible cost, as measured by the percent of land saved relative to the largest area converted, because the cropland that is converted is the highest yielding. However, by sequentially adding weights in 5\% increments to both biodiversity and carbon conservation objectives, large reductions in not only those costs but also in average travel time (because of the correlations between these three measures). Weighting carbon and biodiversity by 15\% each reduces the average area of new lands (i.e. new crop area is increased) by just 3\%, in exchange for avoided costs of 34\% for biodiversity, 48\% for carbon, and 44\% for travel time, which respectively represent 68\%, 77\%, and 50\% of the maximum possible avoided cost for each objective.

\begin{figure}[!ht]
  \begin{center}
       \makebox[\textwidth][c]{\includegraphics[width=0.8\textwidth]{figures/weight_impact3.png}}
    %\includegraphics[scale=0.5]{fig2.eps}
    \caption{The impact of transferring weights from the yield maximization objective to the carbon and biodiversity objectives in 5\% increments, measured in terms of the percent cost that is avoided in each objective relative to the highest cost paid across all weight combinations.}
    \label{default}
  \end{center}
\end{figure}

%These substantial cost savings suggest that a user  achieved for relatively little transfer of weight between objectives indicates that a user primarily concerned with acquiring the highest-yielding land for agriculture would have to make relatively little sacrifice to give enormous gains to conservation interests, or to substantially reduce the potential costs of their farming operation. value of compromise in terms that users might be more likely

These results show that transferring relatively little weight between objectives can have large savings. However, people involved in making land use decisions are more likely to evaluate their willingness to compromise in actual cost units, and from the perspective of how much more they might be willing to pay to compromise with other interests. We therefore evaluated four further scenarios, in which a hypothetical land user with interests in a single objective willingly pays up to 5\% more than the lowest cost case (i.e. the 100\% weight outcome) in order to compromise with other interests. The yield maximizing user is thus willing to ``pay'' for an extra 68,400 km$^2$ of new cropland, the carbon and biodiversity conservationists would respectively sacrifice an additional 2,312,000 tons of carbon and 0.013 biodiversity index units, and the production cost-minded agricultural developer would pay just 4.1 minutes of travel time. Given these sacrifices, we identified how much more the other three interests would have to pay relative to their lowest cost (100\% weight) scenarios, treating each of the three equally, i.e. the amount each has to pay extra is the same percentage of their lowest cost case.

The impacts of this assessment show that in for all four cases, the carbon, biodiversity, and travel time objectives typically pay less than 50\% more than their lowest cost when another party is willing to pay 5\% more cost (Figure 3), and all four associated conversion maps show substantial convergence around road networks and population and agricultural centers.

\begin{figure}[!ht]
  \begin{center}
       \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{figures/compromise_5pct.png}}
    %\includegraphics[scale=0.5]{fig2.eps}
    \caption{The impact of transferring weights from the yield maximization objective to the carbon and biodiversity objectives in 5\% increments, measured in terms of the percent cost that is avoided in each objective relative to the highest cost paid across all weight combinations.}
    \label{default}
  \end{center}
\end{figure}

Figure 3 also maps and shows the results for the equal weighting option and how it contrasts with the minimum costs for each factor.  It shows increased savings for each factor of X to X for transportation costs, carbon and biodiversity compared to their maximum costs.  It also holds these costs to no more than X above their minimum costs.

\begin{figure}[!ht]
  \begin{center}
       \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{figures/compromise_25pct.png}}
    \caption{25\% compromise}
    \label{default}
  \end{center}
\end{figure}

Another interesting feature of these results is that there is little overlap between any targeting areas selected by any of the criteria we have calculated and designated farm development blocks.  In 2005, the Zambian Ministry of Agriculture designated nine farm blocks with roughly one million hectares for assisted agricultural development and spread around the country.  (Zambia Ministry of Agriculture. 2005. Farm Block Development Plan 2005-07).  Although these areas have in fact as a whole attracted little development, they remain “on the books,” and there are at least some reports of continued intent to develop them.  Undeveloped farm blocks worry Agriculture Minister, Zambia Daily Mall: https://www.daily-mail.co.zm/?p=29505

%\begin{equation}
%  \textrm{I} = \left(\frac{\textrm{A}}{\textrm{A + C}} + \frac{\textrm{D}}{\textrm{B + D}}\right)0.5
%\end{equation}

\section*{\large Discussion}

\section*{\large Data accessibility}
\section*{\large Authors' contributions}
\section*{\large Competing interests}
\section*{\large Funding}
\section*{\large Acknowledgements}
This work was supported by funds from: the Norwegian Aid Agency (NORAD) under the Agricultural Synergies Project; the Princeton Environmental Institute Grand Challenges program; the NASA New Investigator Program (NNX15AC64G), and the National Science Foundation (SES-1360463).

\section*{\large Supplementary materials}

%\label{}

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix

%% \section{}
%% \label{}

%% If you have bibdatabase file and want bibtex to generate the
%% bibitems, please use
%%
%\section*{\large References}
\bibliographystyle{prsb} 
\bibliography{/Users/lestes/Dropbox/publications/full}

%% else use the following coding to input the bibitems directly in the
%% TeX file.

%\begin{thebibliography}{00}
%
%%% \bibitem[Author(year)]{label}
%%% Text of bibliographic item
%
%\bibitem[ ()]{}
%
%\end{thebibliography}
\end{document}

\endinput
%%
%% End of file `elsarticle-template-harv.tex'.