### 2016-02-26 paper/supplementary_results.tex

\documentclass[a4paper]{article}
\usepackage[affil-it]{authblk}  
%\newcommand{\citetapos}[1]{\citeauthor{#1}'s \citeyearpar{#1}}  % apostrophe's in cites
\def\linenumberfont{\normalfont\small\sffamily}
\usepackage{graphicx}
\usepackage{epstopdf}
% From link below: next two lines allow caption to fill the full figure box, and caption formats to be changed (e.g. font)
% http://tex.stackexchange.com/questions/107350/caption-below-the-figure-and-aligned-with-left-side-of-figure
% http://ctan.mackichan.com/macros/latex/contrib/caption/caption-eng.pdf
\usepackage[left=2.5cm, right=2.5cm, bottom=2cm, top=2cm]{geometry}
\usepackage{caption}
\captionsetup[figure]{slc=off, font=footnotesize}
\captionsetup[table]{margin={4cm, 4cm}, slc=off, font=footnotesize}
\usepackage[numbers,sort,compress]{natbib}

\begin{document}
\section*{\Large SI Results}

\section*{\large Crop yields}

\begin{figure}[!ht]
       \begin{center}
       \makebox[\textwidth][c]{\includegraphics[width=0.8\textwidth]{figures/ideal_yield2.png}}
        \end{center}
      \caption{The distribution of converted areas when the model is run separately for maize and soybean when 100\% is placed on yield maximization. The map is color-coded according to which crop was allocated to each location (red for maize, blue for soybean, and yellow for the places where both crops were allocated).}
      \label{fig:default}
\end{figure}

\section*{\large References}
\bibliographystyle{prsb} 
\bibliography{/Users/lestes/Dropbox/publications/full}

\end{document}