### 2016-03-01 paper/main.tex

%%%% added from https://groups.google.com/forum/#!topic/latexusersgroup/CDlEjgNnF80

%\documentclass[onecolumn]{rsauthor}    
\documentclass[a4paper]{article}
\usepackage[affil-it]{authblk}

\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{endfloat}
\usepackage[numbers,sort,compress]{natbib}
%\bibpunct{[}{]}{,}{n}{,}{,}  % https://xianblog.wordpress.com/tag/natbib/ (allows natbib with PNAS)
\usepackage{endnotes}
\usepackage{setspace}
\usepackage{verbatim}
\usepackage[left=2.5cm, right=2.5cm, bottom=2cm, top=2cm]{geometry}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
%\usepackage{mathtime}
\usepackage{bm}
\usepackage{url}
%\usepackage{babel}
\usepackage{dcolumn}
\usepackage{multirow}
%%%%%%

\usepackage[
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,anchorcolor=blue,
  citecolor=blue,filecolor=blue,
  menucolor=blue,pagecolor=blue,
  urlcolor=blue]{hyperref}

% Let's add todonotes and comments:
%\usepackage[draft]{todonotes}
%
%% Select what to do with command \comment:  
%% \newcommand{\comment}[1]{}  %comment not showed
%\newcommand{\comment}[1]
%{{[\bfseries \color{blue} #1]}} %comment showed

\usepackage{lineno}
\usepackage{float}
\usepackage[anythingbreaks]{breakurl}

\title{Reconciling agriculture, carbon, and biodiversity in a savanna transformation frontier}
\author[1,2]{Estes, L.D.}
\author[2]{Searchinger, T.}
\author[1]{Spiegel, M.}
\author[1]{Tian, D.}
\author[3]{Sichinga, S.}
\author[3]{Mwale, M.}
\author[4]{Kehoe, L.}
\author[4]{Kuemmerle, T.}
\author[1]{Berven, A.}
\author[5]{Chaney, N.}
\author[1]{Sheffield, J.}
\author[1]{Wood, E.F.}
\author[1]{Caylor, K.K.}
\affil[1]{Civil and Environmental Engineering, Princeton University, Princeton, NJ, 08544 USA}
\affil[2]{Woodrow Wilson School, Princeton University, Princeton, NJ, 08544 USA}
\affil[3]{Zambia Agricultural Research Institute, Mt. Makhulu Research Station, Chilanga, Zambia}
\affil[4]{Geography Department, Humboldt University, 10099 Berlin, Germany}
\affil[5]{Program in Atmospheric and Ocean Sciences, Princeton University, Princeton, NJ USA}
\date{}

\begin{document}
\maketitle
%\begin{frontmatter}

\begin{abstract}
Rapidly rising populations and likely increases in incomes in sub-Saharan Africa make tens of millions of hectares of cropland expansion nearly inevitable, even with large increases in crop yields. Much of that expansion is likely to occur in higher rainfall savannas, with substantial costs to biodiversity and carbon storage. Zambia presents an acute example of this challenge, with an expected tripling of population by 2050, good potential to expand maize and soybean production, and large areas of relatively undisturbed miombo woodland and associated habitat types of high biodiversity value. Here we present a new model designed to explore the potential for targeting agricultural expansion in ways that achieve quantitatively optimal trade-offs between competing economic and environmental objectives: total converted land area (the reciprocal of potential yield); carbon loss, biodiversity loss, and transportation costs. To allow different interests to find potential compromises, users can apply varying weights to examine the effects of their subjective preferences on the spatial allocation of new croplands and its costs. We find that small compromises from the objective to convert the highest yielding areas permit large savings in transportation costs, and the carbon and biodiversity impacts resulting from savanna conversion. For example, transferring just 30\% of weight from a yield maximizing objective equally between carbon and biodiversity protection objectives would increase total cropland area by just 2.7\%, but result in avoided costs of 27\%-47\% for carbon, biodiversity, and transportation. Compromise solutions tend to focus agricultural expansion along existing transportation corridors and in already disturbed areas. Used appropriately, this type of model could help countries find agricultural expansion alternatives and related infrastructure and land use policies that help achieve production targets while helping to conserve Africa's rapidly transforming savannas.

%Rapidly rising populations and likely increases in incomes in sub-Saharan Africa make tens of millions of hectares of cropland expansion nearly inevitable even with large increases in crop yields.  Much of that expansion is likely to occur in higher rainfall wetter savannas, at substantial costs to biodiversity and carbon storage. Zambia presents an acute example of this challenge with an expected tripling of population by 2050, good agronomic potential to expand maize and soybean production, and large swaths of relatively undisturbed Miombo woodland and associated habitat types of high value to mammals and birds.  Here we present a new model designed to explore the potential to target agricultural expansion in ways that achieve reasonable trade-offs among different economic and environmental objectives.  To allow different interests to find potential compromises, the model allows users to apply different weights to different objectives and examine how they alter outcomes.  We find that modest compromises in yield targets permit large savings in transportation costs, and carbon and biodiversity impacts from savanna conversion. For example, transferring just 30\% of weight from a yield maximizing objective equally to carbon and biodiversity would increase total cropland area by just 2.7\%, but result in avoided cost of 27\%-47\% for the carbon, biodiversity, and transportation cost objectives. Compromise solutions tend to focus agricultural expansion along existing transportation corridors and in already disturbed areas.  Used appropriately, this type of model could offer a practical means by which countries could explore agricultural expansion alternatives and target transportation investments and land use policies to achieve production targets while helping to conserve the values of Africa's savannas.     
\end{abstract}
%% Text of abstract

%\begin{keyword}
%keywords
%\end{keyword}

%\end{frontmatter}

\linenumbers

%% main text
\section*{\large Introduction}
Meeting growing food demands while minimizing ecological losses presents one of the 21st Century's major challenges.  Many global sustainability studies and climate mitigation pathways call for eliminating or rapidly phasing out emissions from land use change \cite{searchinger_high_2015,field_summary_2014}.  Unfortunately, these ambitions do not appear realistically attainable in sub-Saharan Africa, due to a projected doubling of population by 2050 combined with rapidly growing  economies, which will combine to treble food demand \citep{tilman_global_2011,searchinger_high_2015}. For example, even with the substantial yield growth (3-5\% per year) projected by the UN Food and Agriculture Organization (FAO), the region would likely need to add more than 100 million hectares of cropland by 2050 \citep{searchinger_high_2015}. Much and probably the great bulk of that new cropland is likely to occur in the ~500 million hectares of savannas and shrublands of the region that receive sufficient rainfall to permit crop production \citep{searchinger_high_2015}.

Savannas sometimes receive limited respect in global land use assessments and have been a major focus of agricultural expansion, not only in Africa but in places such as the Cerrado of Brazil and northern Thailand \citep{morris_awakening_2009}.  Their carbon contents and biodiversity values may be completely ignored on the grounds that they are not forests \citep{searchinger_high_2015}. Yet, the wetter woodland-savannas and shrublands of Africa have a similar average richness, although not density, for birds and mammals as the world’s wetter tropical forests, and 75\% of these habitats have higher overall vertebrate diversity than 75\% of the rest of the world \citep{searchinger_high_2015}.  Their carbon contents are also substantial, particularly relative to their likely yields, so that their use for food crops is unlikely to result in fewer carbon emissions per ton of crop \citep{searchinger_high_2015}.  Conversion of over 100 million hectares of these lands would release tens of millions of tons of carbon.

The high likelihood of some cropland expansion in sub-Saharan Africa raises the prospect of whether this expansion could be located in ways that allow agricultural potential to be achieved, but for substantially lower loss of carbon and biodiversity than conventional agricultural development pathways.  Analyzing this question must be done at national or sub-national scales, because that is where key decisions are made, and it requires data that is both in resolution and quality to provide confidence for decision-makers.

In this paper we look at Zambia, which is a bellwether for this land use challenge in sub-Saharan Africa. On the one hand, Zambia has great need to boost its food production. Sixty percent of Zambia's population is rural {\citep{the_world_bank_world_2016}, per capita calorie food availability is just 2100 kcal \citep{fao_faostat:_2016}, and child stunting rates are 40\% \citep{haddad_global_2015}.  According to the medium estimate of the United Nations, Zambia’s population is projected to grow 265\% from 16.2 million in 2015 to 43 million in 2050 \citep{united_nations_world_2015}, which will combine with rapid economic growth \citep[$>$6\%,][]{the_world_bank_world_2016} to substantially increase per capita calorie demands \citep{tilman_global_2011}. Zambia has also been expanding its agricultural exports, and has agronomic and situational characteristics that put it in a strong position to continue to do so to other parts of the region.

At the same time, two thirds of Zambia’s 738,400 km$^2$ land area is dominated by woodland savannas, which have sufficiently dense tree cover to be classified as forest under either of the United Nations Framework Convention on Climate Change thresholds ($>$10 or $>$30\% tree cover) for defining this vegetation type \citep{sexton_conservation_2015}.  These ecosystems store substantial amounts of carbon, with 50\% of Zambia having a carbon density $\geq$130 t ha$^{-1}$ (vegetative carbon and soil carbon in the top 1 m), while an additional 40\% exceeds 80 t ha$^{-1}$. Meanwhile, Zambia’s deforestation rate of 2,500--3,000 km$^2$ year$^{-1}$ is one of the highest in the world \citep{vinya_preliminary_2011}.

The biodiversity of these woody savannas is also high.  Most of Zambia is covered by what the World Wildlife Fund has categorized as Central Zambian Miombo Woodland \citep{olson_terrestrial_2001,hogan_central_2004}. It contains mature miombo woodland trees typically 15 to 20 meters tall, around a quarter of which are evergreen, with a broadleaf shrub and grass understory beneath, and is interspersed with wetlands that comprise up to 30\% of the landscape \citep{hogan_central_2004}. In addition to hosting one of the world's three richest mammal assemblages \cite{olson_terrestrial_2001}, the large expanses of relatively intact habitat make these woodlands, and several of Zambia's eight other ecoregions \citep{fund_ecoregions_2011, estes_zambezian_????}, important centers of large mammal populations, including black rhino, African buffalo, and highly migratory or wide-ranging species, such as elephants and major predators \citep{hogan_central_2004}.  Zambia in general has a high bird diversity \citep{searchinger_high_2015}. Plant diversity in much of the country ranges between 3,000-5,000 species, a richness that is exceeded only in tropical forests \citep{kier_global_2005}.  Although there are relatively few endemic bird and mammal species, reptile and amphibian endemism is high, with 19 endemic reptile species and 13 endemic amphibians \citep{fund_ecoregions_2011}.

There is a long history of conservation planning that attempts to target conservation areas or objectives while minimizing economic costs or within a constrained budget of land area for conservation \cite[e.g.][]{koh_spatially_2010, naidoo_integrating_2006,margules_systematic_2000}. Koh and Ghazoul \citep{koh_spatially_2010} provide one example of an effort to examine trade-offs among production, carbon and biodiversity for oil palm development in Indonesia by giving each land use priority equal weighting.  One of the challenges for land use planning in practice is that decisions involve trade-offs among different objectives and values that most people find difficult to weight against each other \emph{ex ante}, because the values are often abstract, making them harder to evaluate without first viewing the consequential trade-offs.  Here we present a new model in which varying weights can be placed on the competing land use objectives. This set of options then allows model users to choose the most appealing combination of weights based on their land use preferences, and their own assessments of what are and are not acceptable costs within those objectives. The objective of this is to assist different stakeholders in finding compromises between competing land use interests while still meeting agricultural production targets. We apply this model to Zambia so that we can explore its strengths and limitations as a decision-making tool, and also to answer a basic but important question: is it possible to achieve a reasonable balance of different interests by reducing the carbon and biodiversity costs of expansion while still focusing on land use ares with high agricultural potential?

Answering this question is becoming increasingly important as savannas and other grassy biomes receive increasing agricultural development pressure, and may become increasingly targeted as an indirect result of the laudable efforts to conserve tropical forests \citep{searchinger_high_2015}.

\section*{\large Materials and Methods}

The model we developed seeks to find optimal land use configurations that satisfy the production targets for multiple crops while minimizing four costs: the total land area required, transportation costs, carbon released from land conversion, and impacts on biodiversity. Although total land area is treated as a cost, this cost is minimized by selecting areas with the highest likely yields. By minimizing total land required as well as transportation costs, the model finds areas that reflect two important indicators of agricultural potential. By minimizing carbon and biodiversity costs, the model finds areas that minimize two important dimensions of environmental impact.  The model chooses final areas depending on the weights that a user assigns to each of these four objectives.

To enable this functionality, and to identify optimal solutions, the model is structured to evaluate each objective in terms of production efficiency, or the ratio of cost per hectare to crop production per hectare (yield) that results from converting land to agriculture.  For land area costs, output is measured in terms of the number of km$^2$ required to meet the crop production target. For the other objectives, the indicators are transportation costs, carbon loss, or biodiversity impacts per ton of crop yield.

We focus our model initially on two crops, maize and soybeans, because of their existing and potential significance for Zambia's agricultural sector, and their likely dominance of in the future growth of agriculture in Africa's broader savanna regions \citep{searchinger_high_2015}.  Maize is the largest crop by area and consumption in Zambia \citep{fao_faostat:_2016}, while soybeans are expanding rapidly as a commodity crop to meet regional and global demands for animal feed and direct human consumption \citep{gasparri_emerging_2015}. Because the model is designed to meet production targets for multiple crops, we apply algorithms for allocating each crop to its highest value cells.

Further details on model structure and all model input calculations can be found in the online electronic supplement.

\subsection*{Model inputs}

Our model focuses on the costs of cropland expansion into areas that are currently not farmed but could be.  To identify such areas, we calculated the proportion of currently cropped and settled grid cells within each 1 km$^2$ grid cell.  We excluded the proportion of cells that are either protected or with slope greater than 20\% slope, which are both generally considered unsuitable for farming and strong actual predictors of cropland presence or absence \citep{estes_using_2014}.  We estimated the costs of converting each cell as described in the following sections.

\subsubsection*{\emph{Likely yields}}

We spatially estimated likely yields for maize and soybeans using a three-step approach. In the first step, we used the decision support system for agrotechnology transfer \cite[DSSAT, ][]{jones_modelling_2003} to simulate maize and soybean yields at the sites of 40 weather stations distributed across Zambia. They are the points where a 31-year, gridded daily meteorological dataset has received the greatest amount of bias-correction and infilling by weather observations \citep{sheffield_development_2006,chaney_spatial_2014, estes_changing_2014}. We used soil profiles corresponding to those locations drawn from the WISE v1.1 gridded soil profile database \citep{harvestchoice_converting_2010}, and simulated for each location maize yields over all 31 years under commercial input practices for three-four cultivars representative of short and medium season length according to the Zambian crop calendar \citep{grassini_how_2015}.

In the second step, we used an empirical model to map the DSSAT-modeled yields onto a 1 km grid, taking advantage of a finer resolution soil map that provides a common frame for estimating both yield potential and soil carbon stocks (see below), but has insufficient detail in several parameters (particularly effective rooting depth) required by DSSAT to provide a sound empirical basis for running the model directly.  We used a generalized additive model \citep{wood_mgcv:_2001} to predict the values of the DSSAT-simulated yields, using best-fitting subsets of the following weather and soil variables: growing season precipitation; growing degree days; mean growing season shortwave solar radiation; percent organic carbon; pH; percent clay content. The soil predictors represented mean values in the top 1 m of soil, and were extracted from a new 1 km global database of soil properties \cite{hengl_soilgrids1km_2014}.  This two step approach allowed us to more soundly capture crop responses to Zambia's climatic variability and potential management practices, and to base potential yields on the same dataset used to estimate carbon costs (see below).

The first two steps resulted in a map of potential yield under high inputs and good management, and assuming no losses due to other limitations such as diseases. These yield values therefore overestimate what farmers are truly likely to achieve. To reflect more realistic estimates of future yields, we therefore rescaled the yield maps to match the mean yields that are projected to be achieved in Zambia by 2050 (4.4 t ha$^{-1}$ for maize and 3.6 t ha$^{-1}$ for soybean).  We projected these yields by using FAO-projections for sub-Saharan Africa as a whole to derive an annual growth factor and applying that factor to the average 2009-2014 yields for each crop reported in Zambia.  Overall, the distribution of yields derived from our crop modeling is most important to targeting land use, while this choice of average yield determines the total quantity of land needed to meet production targets.

\subsubsection*{\emph{Transportation Costs}}

As a second measure related to agricultural potential, we estimated the travel time to the nearest major town. Travel time serves here as a proxy for differences in output prices and production costs, such as fertilizer prices and transport costs \cite{stifel_isolation_2008}, and, via the density of road networks, is strongly correlated with the suitability of land for farming \citep{laurance_estimating_2015,estes_using_2014}.  We used a cost-distance analysis drawing on spatial data for roads, rivers and water bodies, urban centers inside and neighboring Zambia, and national boundaries to calculate travel time in hours. Total travel costs for any land use scenario equal the total time needed to transport the tonnage of crops produced in each converted cell to the nearest town, which we divide by 20 to produce "truck hours" on the assumption of a 20-ton truck (although the division  of all cell results by this constant does not alter the mapping outcomes).

\subsubsection*{\emph{Carbon}}

To estimate potential carbon costs from cropland conversions, we developed maps of vegetative and soil carbon stocks for Zambia.  Vegetative carbon was calculated from a recent map of above-ground biomass developed by Baccini et al. \citep{baccini_estimated_2012}, from which we estimated below-ground biomass using savanna and miombo woodland-specific root-shoot ratios \cite{mokany_critical_2006, mugasha_allometric_2013}, and then converted the sum to carbon using a ratio of 0.47 \citep{ipcc_land_1996}. We used the same soil carbon dataset used for crop yield prediction together with an accompanying map of bulk density to calculate soil carbon stocks in the top 1 m. The potential carbon loss due to agricultural conversion was calculated as 100\% of vegetative carbon and 25\% of soil carbon \cite{searchinger_high_2015}.

\subsubsection*{\emph{Biodiversity}}

There are many ways of evaluating biodiversity value and potential impacts, including species richness of different taxa in absolute terms or based on level of threat or endemism, and including factors designed to address the significance of contiguity or the impact of different types of development \citep[e.g.][]{newbold_global_2015,searchinger_high_2015, gasparri_emerging_2015,koh_spatially_2010}.  The actual measures selected should reflect both conservation priorities and the quality of available data.  For a fine-scaled, country-level analysis such as this one, the species range map data that form the basis for broader impact assessments \citep[e.g.][]{kehoe_global_2015} have an effective resolution that is too coarse. We therefore developed a composite measure of biodiversity value based on several datasets.

First, we used the potential vegetation maps of East Africa \citep[][]{van_breugel_potential_2011}, a 2010 landcover dataset for Zambia\footnote{downloaded from http://apps.rcmrd.org/landcoverviewer/}, and a shapefile of protected areas to calculate how much of Zambia's 20 unique vegetation types remains after conversion to croplands and settlements, and what proportion of that remainder falls within national parks. From this estimate we calculated a hybrid measure of rarity and threat, wherein we assigned a weight to each vegetation type based on its relative remaining area in proportion to Zambia's extent, log-transformed to account for a highly skewed distribution.  We then re-weighted that weight by the protected fraction of each vegetation type's protected fraction, following \citep{van_breugel_environmental_2015}. We then used the remaining vegetation map to calculate a second measure, the proportion of untransformed vegetation within a 11X11 km neighborhood centered on each 1 km$^2$, providing a measure of vegetation intactness. By adding this measure to the first index and normalizing, we created a biodiversity score (range 0-1 per ha) that gives equal weight to how rare or threatened a habitat is, and how undisturbed its local area.  When the biodiversity score for each converted grid cell converted is summed, it generates a total cost of biodiversity impacts.

As a final modification, we masked out the areas of existing national parks and game management areas to respect the national judgment that they should not be used for expanding farmland.  We retained a third protected category, forest reserves, as these tend to be far less stringently protected, having already had 8.3\% of their area converted to cropland or settlements, but assigned these areas a high biodiversity score.

\subsubsection*{\emph{Potential farmland and production targets}}

Likely future demand for maize and soybeans produced in Zambia is difficult to estimate, and our model can easily be run with different production targets.  For this paper, we adopted targets of a four-fold increase for maize for 2050 compared to 2009-2014 average, and a 10-fold increase for soybeans.  These increases reflect the rate of growth in Zambia from 2000-2014, and a mean production trend for soybean growth in southern Africa estimated by Gasparri et al \citep{gasparri_emerging_2015}, with small additional adjustments to to account for the fact that Zambia has one of the highest agricultural potentials in sub-Saharan Africa and has high potential to expand exports.

\subsubsection*{\emph{Yields on current land}}

The amount of land conversion to meet a production target will depend on the growth of yields on existing croplands, which are well short of their current potentials \citep{waddington_getting_2010}.  For the principal scenarios we present in this paper, we assume that yields on current cropland will achieve the average likely yields for Zambia's \emph{potential} farmland.  However, to examine the possible significance of yield gains on existing cropland, we also examined scenarios in which the yield gains on existing cropland 1) only manage to close 50\% of the gap between current and our projected future yields, and 2) close 150\% of the gap.

\subsection*{Normalization and Weighting}

Weighting different objectives is to some extent based purely on preferences, but it is important that these be expressed in a numerical form and processed so that a quantitatively optimal solution for that set of preferences is identified. To achieve this result, we first express the different objectives as efficiencies (e.g. carbon loss per ton of crop yield, as described above), and then normalize the units of each objective based on the range of its scores. We then multiplying each normalized objective by its corresponding weight (which fall between 0 and 1 and must sum to 1, with values closer to 1 indicating stronger preferences) and then adding the resulting products. The result is a single score that ranks each grid cell according to its ability to meet the production target for the lowest cost, given those weights. The model then selects cells in descending rank order until each the cumulative production of converted cells reaches each crop's corresponding target (see online electronic supplement for details of land allocation between crops).

Although this weighting system treats choices as preferences, our model calculates the actual costs of each objective in absolute terms:  area of land converted (which is increasingly minimized as weight is added to the preference to maximize crop yields), total transportation costs from new croplands, total carbon released from conversions, and total impacts on biodiversity. After reviewing these costs, model users can then adjust the preferences to reflect desirable or politically acceptable trade-offs.

\section*{\large Results}

\subsection*{Significance of yield gains on existing cropland}

In our main scenario, we assumed that the gap between current and projected future yields would be fully closed on existing cropland, but we also ran the model (with equal weights across all objectives) for cases when only 50\% or 150\% of the gap was closed. In the former case, the total land area required to meet the target increased by 18\% (14,634 to 17,202 km$^2$), transport costs increased 22\% (5.52 X 10$^{-5}$ to 6.61 X 10$^{-5}$ truck hours), carbon losses increased by 19\% (597 to 711 Mt), and the total biodiversity index of converted areas increased by 20\% (5.53 X 10$^{-5}$ to 6.61 X 10$^{-5}$).   In the 150\% gap closure scenario, the impacts were reversed, with impacts declining by almost exactly the same percentage.

\subsection*{Overlap and adjacency from optimal individual solutions}

To identify the minimum possible costs for each individual objective, and the degree to which their areas of new cropland correspond, we examined simulations giving 100\% of the weight to each of the four different objectives. Figure 1 maps the results, and shows the costs for each objective in each simulation, as well as the extent of overlap among objectives.

The scenario is which the sole preference is to maximize yield (100\% weight) converts land with the most promising yield potential for farmers, which should have economic advantages.  These areas have little direct overlap with the areas picked by scenarios that minimize carbon costs, biodiversity impacts or travel time (Figure 1).  They also have little adjacency, as measured by the average distance between each pixel of one map and its nearest neighbor in the other (Figure 1).

However, the areas selected solely for minimizing carbon, biodiversity, or transportation costs also have little direct overlap, but the map reveals that they have high adjacency, confirmed by direct estimates of distance (Figure 1). All three criteria tend to identify lands along existing road networks and population and agricultural centers, with the carbon minimizing objective converting a larger area of land in the western half of the country. The reasons for this correlation are probably that carbon stocks and biodiversity values are lowest in areas of high human density, because of high levels of wood harvesting for charcoal and fuelwood, and because human uses similarly cause fragmentation that leads to low biodiversity scores.

\begin{figure}[!ht]
  \begin{center}
       \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{figures/single_priorities3.png}}
    \caption{Areas converted to new maize and soybean croplands and their associated costs when each individual objective (yield maximization, and carbon cost, biodiversity cost, and transport time minimization) is given 100\% weight. Cropland converted under each scenario is shown in the maps, colour-coded according to which objective receives 100\% weight. Large bar charts, colour-coded to the maps, display the absolute value of costs resulting from each scenario. Small bar charts show the percentage of overlap (top) and adjacency (bottom; nearest neighbor distance) between the map directly above it, and each of the other three maps.}
    \label{default}
  \end{center}
\end{figure}

\subsection*{Potential to harmonize different interests}

Although the ideal solutions to maximize each objective do not lead to much direct overlap, areas selected by less ideal but still good solutions for each objective may still overlap.  The reasons start with the different potential ranges of costs among the scenarios that maximize each objective.

While the best case scenario for maximizing yield would convert 13,668 km$^2$, the worst case scenario among our single-purpose optimization scenarios (the scenario minimizing carbon costs) would convert 15,302 km$^2$. This range in converted areas is only 12\%, which is far less than the range in yields across the country’s potential cropland (3.7-6.1 t/ha/y ha$^{-1}$ for maize and 0.8 to 5.1 t/ha/y ha$^{-1}$ for soybean). One reason is the significance of yield to minimizing each of the costs because of the need to meet a fixed production target; because yield is in the denominator of each objective (e.g. carbon loss per ton of yield), land with better yield potential tends to be selected even when the yield maximization priority is given no weight.  By contrast, the differences in travel times among single-objective scenarios range from a total of 3.4 X 10$^{-5}$ to 32.3 X 10$^{-5}$ truck hours, a difference of 872\%.  Carbon costs differ by 156\% (462 to 1184 Mt), and biodiversity of 68\% (4.96 X 10$^{-5}$ to 8.33 X 10$^{-5}$).  The low range in the yield-related cost (converted area), and the high ranges in the other costs, suggest that there is strong potential to reduce transportation, carbon and biodiversity costs for relatively little sacrifice of yield.

This potential is illustrated in Figure 2, which shows how much carbon, biodiversity, and transport cost can be avoided by transferring increasing amounts of weight to these objectives from the yield maximization objectives. Avoided cost is the percent difference between the cost paid under a given weight combination and the maximum cost found across all weighting scenarios (we ran all possible weight permutations varying in 5\% increments).  Sequentially adding weights in 5\% increments to both the biodiversity and carbon objectives results in little cost to yield, but rapid, large reductions in both biodiversity and carbon costs as well as transport costs, despite no weight given to transport costs because of its correlations with the other two objectives). For example, adding just 15\% weight each to carbon and biodiversity increases the area of new lands (i.e. reduces the average yield) by just 2.7\% (avoided area costs drop to 8\% from 10.7\%), but avoids costs equal to 27\% for biodiversity, 47\% for carbon, and 44\% for transportation costs. For comparison, in the best case (100\% weight) scenario for each objective, 61\%, 41\%, and 90\% of carbon, biodiversity, and transportation cost would be avoided.

\begin{figure}[!ht]
  \begin{center}
       \makebox[\textwidth][c]{\includegraphics[width=0.8\textwidth]{figures/weight_impact5.png}}
    %\includegraphics[scale=0.5]{fig2.eps}
    \caption{The impact of transferring weights from the yield maximization objective to the carbon and biodiversity objectives in 5\% increments, measured in terms of the percent cost that is avoided in each objective relative to the highest cost paid across all weight combinations. In these results, no weight was assigned to the transport cost objective.}
    \label{default}
  \end{center}
\end{figure}

Figure 3 shows the potential for compromise in which a hypothetical land user only interested in a single objective is willing to ``pay'' 5\% more costs than the lowest possible costs for that objective. For example, the yield maximizing user would be willing to pay for an extra 683 km$^2$ of new cropland, or the person interested in minimizing carbon costs would pay an additional 231 Mt of carbon.  Lines in the bar charts show the minimum and maximum costs across all possible weighting scenarios.  Each of the four different ``compromise'' scenarios, represented by a different colour, permits solutions that allow the other scenarios to save substantially relative to maximum potential costs. The carbon, biodiversity, and transport cost objectives in all but two cases avoid more than 50\% of their maximum potential costs when any of the other objectives is willing to pay 5\% more cost (Figure 3).

The location of the conversions associated with each of these compromises show substantial convergence along roads and population and agricultural centers. Overlap between the resulting cropland allocations exceeds 30\% between the yield and biodiversity and cost compromises, and 25\% between the biodiversity and carbon compromises. The carbon and cost compromises overlap least ($<$10\%), followed by carbon and yield ($\sim$15\%), but each pair of maps has high adjacency, with all compromise allocations having an average nearest neighbor distance of $<$20km.

\begin{figure}[!ht]
  \begin{center}
       \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{figures/compromise_5pct3.png}}
    \caption{The results of scenarios in which each land use objective is willing to pay 5\% more cost to compromise with the other three objectives. Cropland converted under each compromise scenario is shown in the maps, colour-coded according to which objective is making the 5\% compromise. Large bar charts, colour-coded to the maps, display the absolute value of costs resulting from each compromise scenario, together with the maximum, minimum, and median costs across all weighting permutations (grey horizontal lines in bar charts). Small bar charts show the percentage of overlap (top) and adjacency (bottom; nearest neighbor distance) between the map directly above it, and each of the other three maps, with the colour denoting which map is being compared.}
    \label{default}
  \end{center}
\end{figure}

The weight on the objective subject to the 5\% compromise is proportional to the size of its potential avoided cost (see Table 1 in supplementary results), thus yield requires a weight of only 40\% to pay 5\% or less of its lowest cost, whereas transportation cost and carbon, which have the largest cost ranges, require 85\% weight each, whereas biodiversity requires 50\% weight.

We also analyzed an ``equal compromise''  scenario in which each objective receives equal weight, as shown in Figure 4.  The bar charts show how this scenario, represented in purple, compares with each of the 5\% compromise scenarios shown in Figure 3, plus the best and worst case options among all weighting scenarios. Relative to the worst case, equal compromise avoids 80\% or more of the cost for carbon, biodiversity, and transportation, and 41\% of the cost for land area. With respect to the 5\% compromises, the equal weighting scenario is lower than, but compares favorably to, the $\sim$95\% savings achieved by carbon, biodiversity, and transportation costs within their individual cost categories, and to the 58\% savings by the yield objective.

As a general rule, the areas selected for conversion under equal compromise lie along existing major transportation corridors, and generally near areas of existing cropland.

 \begin{figure}[!ht]
  \begin{center}
       \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{figures/compromise_25pct2.png}}
    \caption{Areas converted to new maize and soybean croplands and their associated costs when when all objectives (yield maximization, and carbon, biodiversity, and transport cost minimization) compromise equally (i.e. receive 25\% weight). Cropland converted under this scenario is denoted by purple areas in the map, with its associated cost impacts shown alongside those from the 5\% compromise scenarios (Figure 3), together with the maximum, minimum, and median costs across all weighting permutations (grey horizontal lines in bar charts).}
    \label{default}
  \end{center}
\end{figure}

All of these scenarios also suggest the value of deliberate and optimized land use planning.  Zambia has designated a large number of farm blocks, and added nine more blocks with roughly one million hectares as recently as 2005 \citep{anonymous_farm_2005}. Although these areas have in fact as a whole attracted little development, they remain ``on the books'', and there are at least some reports of continued intent to develop them \citep{anonymous_undeveloped_2015}. But there is less than 2\% overlap between designated farm blocks and either the yield maximization scenario and the equal weighting scenario, and very little adjacency (see supplementary results).

\section*{\large Discussion}

Our results for Zambia suggest at least the potential to plan agricultural expansion such that it limits carbon and biodiversity costs with limited sacrifice of yield potential.  In fact, if agricultural projection goals alone are best represented by a combination of agronomic potential and existing road access, then solutions that provide a good balance of these objectives also correlate strongly with solutions that limit carbon and biodiversity impacts.

This finding demonstrates the importance of targeting land uses based on their ability to deliver a particular benefit for the lowest possible cost. The value of such efficiency-based approach, which we apply here to agricultural land use, is in line with findings from the field of conservation planning, where it has been shown that, by considering both the price of land and its ecological value, conservation objectives can be met for less than half the cost of approaches that target land solely based on ecological value \citep{naidoo_integrating_2006}.

Another result, if more obvious, is the potential benefits of increasing yields on existing cropland.  Actions that boost yields on existing cropland also have the potential to have feedback effects that lead to more local expansion, for example, by boosting the competitiveness and therefore quantity of exports \cite{ewers_increases_2009,angelsen_policies_2010}. Our analysis ignores those potential feedback effects because all options analyzed meet the same production targets, but our analyses quantities the large potential benefits if these yield gains are used to limit environmental impacts.

The fact that the most advantageous compromises occur along existing transportation corridors also appears to have several  advantages.  Road construction has been a primary driver of the location of agricultural expansion \citep{laurance_impacts_2009}. Zambia is already engaged in a program to improve existing major roads, but there is evidence that agriculture responds in particular to improvements in the quality of smaller, feeder rural roads to larger paved roadways \citep{bourguignon_rural_2008,blimpo_public_2013}. Improving smaller rural roads would be one way of focusing Zambia's agriculture development.  Because agricultural development tends to build on the private and public infrastructure that accompanies initial development, the focus on existing areas would also seem likely to be a more robust, long-term strategy for directing agricultural development \cite{laurance_estimating_2015}. A potential impediment to this, however, might be that land closer to roads is already ``spoken for'', and thus less likely to be available for new farms.

This model appears to be the right kind of tool for assisting the planning of agricultural development. Optimization models that require ex ante specification of preferences in mathematical terms work less well for decision-making that must weigh very different objectives in unclear ways.  Our model makes possible an iterative optimization process that allows people to realize their preferences in the face of real information about options.  It is particularly suitable to decisions that will inevitably have strong political elements, and to consensus-building among stakeholders with different preferences.

For this kind of tool to be truly legitimate and useful on a practical level, we believe several elements are required. First, it must be further developed and applied in an iterative way with the government and other stakeholders so that the impact of different weighting decisions, in terms of their costs and benefits, can be fully understood, thereby facilitating informed land use planning.

Second, development planning must reflect and incorporate other real world factors beyond those incorporated in our model. For example, large, multi-lane roads can lead to high rates of wildlife mortality, and serve as effective barriers to migration for large mammals, regardless of other land uses \citep{holdo_predicted_2011}.  Although our ``equal weight'' scenario identifies some target areas in the heavily wooded northeastern part of the country, these concerns would be one reason to explore whether areas closer to existing cropland concentrations in the south might be reasonable substitutes.  Such consolidation might also make it easier to support agricultural development with lower attendant infrastructure development costs.  And this biodiversity concern is only one example of other issues, many purely practical, which should, and inevitably will, influence where agriculture develops.

Third, data quality is critical. This kind of tool focuses on particular hectares, not broad averages, so data errors have little or no chance of averaging out through aggregation.  Crop modeling predictions and most of our other scores can be highly sensitive to a range of data parameters. These facts mean that long-term use of the tool should be accompanied by steady data improvement.  It also means that the tool should mostly serve as a guide to likely values. Before governments will make important decisions ``for keeps'', there should be site-specific ground-truthing of soil properties and yield potential, as well as confirmation of biodiversity and carbon characteristics.

Some elements of the model should also be further developed, depending on the ability to gather significant additional data. Analysis of agricultural potential could incorporate more economic factors, differentiate between smallholder and commercial farming practices, and incorporate irrigation potential. Additional information on land ownership and title would allow a finer delineation of potential agricultural land, while the carbon estimates would be improved by root-shoot specific to each vegetation type. Biodiversity impacts could be better assessed by incorporation data on endemism and the degree to which they are impacted by habitat fragmentation. One potentially promising approach would be to incorporate the biodiversity impact model developed by Newbold et al. \citep{newbold_global_2015}, provided the land use intensity maps on which it is conditioned can be developed for this resolution.

Overall, however, our results are promising, and suggest that efficiency-based land use planning approaches such as ours can help to substantially minimize the ecological cost of cropland expansion in sub-Saharan Africa's savannas, the 21st Century's emerging hotspot of agricultural land use change.

\section*{\large Data accessibility}
The model source and inputs data are available online at https://github.com/PrincetonUniversity/agroEcoTradeoff.

\section*{\large Authors' contributions}
L.E. and T.S. conceived and designed the research, together with K.C. L.E. and M.S. designed and coded the model. L.E. developed the input datasets, and ran and analyzed simulations, together with T.S. D.T. and L.E. developed the crop yield potential surfaces. L.K. and T.K. helped design the biodiversity impact metric. S.S. and M.M. collected and processed spatial datasets used in developing inputs. A.B. processed infrastructure data and helped develop the travel cost surface. J.S., N.C., and E.W. developed the bias-corrected weather data needed to estimate yield potential. L.E. and T.S. wrote the manuscript, with input from K.C., T.K., and L.K.

\section*{\large Competing interests}
We declare we have no competing interests.

\section*{\large Funding}
This work was supported by funds from: the Norwegian Aid Agency (NORAD) under the Agricultural Synergies Project; the Princeton Environmental Institute Grand Challenges program; the NASA New Investigator Program (NNX15AC64G), and the National Science Foundation (SES-1360463).

\section*{\large Acknowledgements}
We thank Paulo van Breugel for providing the potential vegetation map of East Africa, and Philip Thornton for advice regarding modeling of yield potential.

\section*{\large Supplementary materials}

%\section*{\large References}
\bibliographystyle{prsb} 
\bibliography{/Users/lestes/Dropbox/publications/full}

\end{document}

\endinput
%%
%% End of file `elsarticle-template-harv.tex'.