### 2016-05-05 paper/supplementary_results.tex

\documentclass[a4paper]{article}
\usepackage[affil-it]{authblk}  
%\newcommand{\citetapos}[1]{\citeauthor{#1}'s \citeyearpar{#1}}  % apostrophe's in cites
\def\linenumberfont{\normalfont\small\sffamily}
\usepackage{graphicx}
\usepackage{epstopdf}
% From link below: next two lines allow caption to fill the full figure box, and caption formats to be changed (e.g. font)
% http://tex.stackexchange.com/questions/107350/caption-below-the-figure-and-aligned-with-left-side-of-figure
% http://ctan.mackichan.com/macros/latex/contrib/caption/caption-eng.pdf
\usepackage[left=2.5cm, right=2.5cm, bottom=2cm, top=2cm]{geometry}
\usepackage{caption}
\captionsetup[figure]{slc=off, font=footnotesize}
\captionsetup[table]{margin={4cm, 4cm}, slc=off, font=footnotesize}
\usepackage[numbers,sort,compress]{natbib}
\usepackage[skip=2pt,font=normalsize]{caption}

\begin{document}
\section*{\Large SI Results}

\section*{\large Potential displacements caused by cropland allocation routine}
To assess the degree to which the model's cropland allocation routine impacts the areas selected for conversion for each crop (see SI Methods for details), we ran the model separately for each crop, using the same parameters applied in the main analysis for the yield maximization scenario. We then assessed the overlap in the resulting areas selected for conversion for each crop in the resulting output maps (Figure 1). The results indicated that the two crops were allocated to the same areas in $<$1\% of converted pixels.

\begin{figure}[!ht]
\captionsetup{width=0.95\linewidth, font=small}
       \begin{center}
       \makebox[\textwidth][c]{\includegraphics[width=0.8\textwidth]{figures/ideal_yield3.png}}
        \end{center}
      \caption{The distribution of converted areas when the model is run separately for maize and soybean when 100\% is placed on yield maximization. The map is color-coded according to which crop was allocated to each location (red for maize, blue for soybean, and yellow for the places where both crops were allocated).}
      \label{fig:default}
\end{figure}

\section*{\large Overlaps with farm blocks}

\begin{figure}[!ht]
\captionsetup{width=0.95\linewidth, font=small}
       \begin{center}
       \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{figures/fb_overlaps.png}}
        \end{center}
      \caption{Simulations run under the 100\% weight to yield maximization and the equal weights scenarios (25\% weight to all objectives), showing the areas of overlap with designated farm development blocks. }
      \label{fig:default}
\end{figure}

\clearpage
\section*{\large Weights associated with 5\% cost compromises}

\begin{table}[ht]
\begin{center}
\captionsetup{width=0.85\linewidth, font=small}
\caption{Weights corresponding to the scenarios in which each objective was a) willing to pay 5\% more than its lowest possible cost, and b) the costs for the other three objectives were minimized equitably.}
\begin{tabular}{ccccc}
\hline
& \multicolumn{4}{c}{Weights}\\
Objective making 5\% compromise & Yield & Carbon & Biodiversity & Transport Cost \\
\hline\hline
Yield & 40 & 10 & 5 & 45 \\
Carbon & 15 & 75 & 0 & 10 \\
Biodiversity & 15 & 5 & 50 & 30 \\
Travel time & 5 & 5 & 0 & 90 \\
\hline
\end{tabular}
\end{center}
\label{default}
\end{table}

%\section*{\large References}
\bibliographystyle{prsb} 
\bibliography{/Users/lestes/Dropbox/publications/full}

\end{document}